### 2021-12-07 draft.tex
Deploying to gh-pages from @ PoisotLab/ms_networkary_classifiers@f46d643cdd302ab05b4a42074527daf14f99c3b9 🚀

%!TEX TS-program = xelatex
\documentclass[11pt]{article}

\usepackage[english]{babel}

\usepackage{amsmath,amssymb,amsfonts}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{stix2}
\usepackage[scaled]{helvet}
\usepackage[scaled]{inconsolata}

\usepackage{lastpage}

\usepackage{setspace}

\usepackage{ccicons}

\usepackage[hang,flushmargin]{footmisc}

\usepackage{geometry}

\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}

\usepackage{fancyhdr}
\renewcommand{\headrulewidth}{0pt}\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}

\makeatletter
\newcounter{tableno}
\newenvironment{tablenos:no-prefix-table-caption}{
  \caption@ifcompatibility{}{
    \let\oldthetable\thetable
    \let\oldtheHtable\theHtable
    \renewcommand{\thetable}{tableno:\thetableno}
    \renewcommand{\theHtable}{tableno:\thetableno}
    \stepcounter{tableno}
    \captionsetup{labelformat=empty}
  }
}{
  \caption@ifcompatibility{}{
    \captionsetup{labelformat=default}
    \let\thetable\oldthetable
    \let\theHtable\oldtheHtable
    \addtocounter{table}{-1}
  }
}
\makeatother

\usepackage{array}
\newcommand{\PreserveBackslash}[1]{\let\temp=\\#1\let\\=\temp}
\let\PBS=\PreserveBackslash

\usepackage[breaklinks=true]{hyperref}
\hypersetup{colorlinks,%
citecolor=blue,%
filecolor=blue,%
linkcolor=blue,%
urlcolor=blue}
\usepackage{url}

\usepackage{caption}
\setcounter{secnumdepth}{0}
\usepackage{cleveref}

\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth
\else\Gin@nat@width\fi}
\makeatother
\let\Oldincludegraphics\includegraphics
\renewcommand{\includegraphics}[1]{\Oldincludegraphics[width=\maxwidth]{#1}}

\usepackage{longtable}
\usepackage{booktabs}

\usepackage{color}
\usepackage{fancyvrb}
\newcommand{\VerbBar}{|}
\newcommand{\VERB}{\Verb[commandchars=\\\{\}]}
\DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
% Add ',fontsize=\small' for more characters per line
\usepackage{framed}
\definecolor{shadecolor}{RGB}{248,248,248}
\newenvironment{Shaded}{\begin{snugshade}}{\end{snugshade}}
\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
\newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{#1}}
\newcommand{\DecValTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\FloatTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\CharTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\StringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\VerbatimStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\ImportTok}[1]{#1}
\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
\newcommand{\DocumentationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\AnnotationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{#1}}
\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\VariableTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
\newcommand{\OperatorTok}[1]{\textcolor[rgb]{0.81,0.36,0.00}{\textbf{#1}}}
\newcommand{\BuiltInTok}[1]{#1}
\newcommand{\ExtensionTok}[1]{#1}
\newcommand{\PreprocessorTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
\newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.77,0.63,0.00}{#1}}
\newcommand{\RegionMarkerTok}[1]{#1}
\newcommand{\InformationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\WarningTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\AlertTok}[1]{\textcolor[rgb]{0.94,0.16,0.16}{#1}}
\newcommand{\ErrorTok}[1]{\textcolor[rgb]{0.64,0.00,0.00}{\textbf{#1}}}
\newcommand{\NormalTok}[1]{#1}

\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newenvironment{CSLReferences}[3] % #1 hanging-ident, #2 entry spacing
 {% don't indent paragraphs
  \setlength{\parindent}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1 \everypar{\setlength{\hangindent}{\cslhangindent}}\ignorespaces\fi
  % set entry spacing
  \ifnum #2 > 0
  \setlength{\parskip}{#2\baselineskip}
  \fi
 }%
 {}
\usepackage{calc} % for \widthof, \maxof
\newcommand{\CSLBlock}[1]{#1\hfill\break}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\maxof{\widthof{#1}}{\csllabelwidth}}{#1}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth}{#1}}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}\geometry{verbose,letterpaper,tmargin=2.2cm,bmargin=2.2cm,lmargin=2.2cm,rmargin=2.2cm}

\usepackage{lineno}
\usepackage[nolists,noheads]{endfloat}

\pagestyle{plain}

\tolerance=1
\emergencystretch=\maxdimen
\hyphenpenalty=10000
\hbadness=10000

\doublespacing

\fancypagestyle{normal}
{
  \fancyhf{}
  \fancyfoot[R]{\footnotesize\sffamily\thepage\ of \pageref*{LastPage}}
}
\begin{document}
\raggedright
\thispagestyle{empty}
{\Large\bfseries\sffamily Template to prepare preprints and manuscripts
using markdown and github actions}
\vskip 5em

%
\href{https://orcid.org/0000-0002-0735-5184}{Timothée\,Poisot}%
%
\,\textsuperscript{1,2}

\textsuperscript{1}\,Université de
Montréal\quad \textsuperscript{2}\,Québec Centre for Biodiversity
Sciences

\textbf{Correspondance to:}\\
Timothée Poisot --- \texttt{timothee.poisot@umontreal.ca}\\

\vfill
This work is released by its authors under a CC-BY 4.0 license\hfill\ccby\\
Last revision: \emph{\today}

\clearpage
\thispagestyle{empty}

\vfill

\begin{enumerate}
    \itemThe prediction of species interaction networks is gaining
momentum as a way to circumvent limitations in data volume. Yet,
ecological networks are challenging to predict because they are
typically small and sparse. Dealing with extreme class imbalance is a
challenge for most binary classifiers, and there are currently no
guidelines as to how predictive models can be trained.%
    \itemUsing simple mathematical arguments and numerical experiments
in which a variety of classifiers (for supervised learning) are trained
on simulated networks, we develop a series of guidelines related to the
choice of measures to use for model selection, and the degree of
unbiasing to apply to the training dataset.%
    \itemClassifier accuracy and the ROC-AUC are not informative
measures for the performance of interaction prediction. PR-AUC is a
fairer assesment of performance. In some cases, even standard measures
can lead to selecting a more biased classifier because the effect of
connectance is strong. The amount of correction to apply to the training
dataset depends as a function of the classifier and the network
connectance.%
    \itemThese results reveal that training machines to predict networks
is a challenging task, and that in virtually all cases, the composition
of the training set needs to be experimented on before performing the
actual training. We discuss these consequences in the context of the low
volume of data.%
\end{enumerate}

\vfill

\clearpage
\linenumbers
\pagestyle{normal}

example on diagnostic test: rare events are hard to detect even with
really good models

summary of model challenges for networks

list of problems to solve - baseline values and response to bias -
effect of training set bias on performance - which models need the least
amount of interactions to work

summary of the results

\hypertarget{baseline-values}{%
\section{Baseline values}\label{baseline-values}}

\hypertarget{numerical-experiments}{%
\section{Numerical experiments}\label{numerical-experiments}}

\hypertarget{effect-of-training-set-on-performance}{%
\subsection{Effect of training set on
performance}\label{effect-of-training-set-on-performance}}

\hypertarget{required-amount-of-positives-to-get-the-best-performance}{%
\subsection{Required amount of positives to get the best
performance}\label{required-amount-of-positives-to-get-the-best-performance}}

\hypertarget{guidelines-for-prediction}{%
\section{Guidelines for prediction}\label{guidelines-for-prediction}}

\hypertarget{references}{%
\section{References}\label{references}}

\end{document}

### 2021-12-07 preprint.tex
Deploying to gh-pages from @ PoisotLab/ms_networkary_classifiers@f46d643cdd302ab05b4a42074527daf14f99c3b9 🚀

--- 2021-12-07 draft.tex
@@
 %!TEX TS-program = xelatex
-\documentclass[11pt]{article}
+\documentclass[10pt,oneside]{article}
+\usepackage[fontsize=9pt]{scrextend}
 
@@
 \newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth}{#1}}
-\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}\geometry{verbose,letterpaper,tmargin=2.2cm,bmargin=2.2cm,lmargin=2.2cm,rmargin=2.2cm}
+\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}\usepackage[table,dvipsnames]{xcolor}
 
-\usepackage{lineno}
-\usepackage[nolists,noheads]{endfloat}
+\geometry{includemp,
+            letterpaper,
+            top=2.4cm,
+            bottom=2.4cm,
+            left=1.0cm,
+            right=1.0cm,
+            marginparwidth=5cm,
+            marginparsep=1.0cm}
+
+\usepackage[singlelinecheck=off]{caption}
+
+\captionsetup{
+  font={small},
+  labelfont={bf},
+  format=plain,
+  labelsep=quad
+}
+
+\usepackage{floatrow}
+
+\floatsetup[figure]{margins=hangright,
+              facing=no,
+              capposition=beside,
+              capbesideposition={center,outside},
+              floatwidth=\textwidth}
+
+% \floatsetup[table]{margins=hangright,
+%              facing=no,
+%              capposition=beside,
+%              capbesideposition={center,outside},
+%              floatwidth=\textwidth}
 
@@
 
-\tolerance=1
-\emergencystretch=\maxdimen
-\hyphenpenalty=10000
-\hbadness=10000
+\setcounter{secnumdepth}{5}
 
-\doublespacing
+\usepackage{titlesec}
+
+\titleformat{\section}[block]
+{\normalfont\large\sffamily}
+{\thesection}{.5em}{\titlerule\\[.8ex]\bfseries}
+
+\titleformat{\subsection}[runin]
+{\normalfont\fontseries{b}\selectfont\filright\sffamily}
+{\thesubsection.}{.5em}{}
+
+\titleformat{\subsubsection}[runin]
+{\normalfont\itshape\rmfamily\bfseries}{\thesubsubsection}{1em}{}
+
+\fancypagestyle{firstpage}
+{
+   \fancyhf{}
+   \renewcommand{\headrulewidth}{0pt}
+   \fancyfoot[R]{\footnotesize\ccby}
+   \fancyfoot[L]{\footnotesize\sffamily\today}
+}
 
@@
 }
+
+\usepackage{tikz}
 \begin{document}
-\raggedright
-\thispagestyle{empty}
-{\Large\bfseries\sffamily Template to prepare preprints and manuscripts
-using markdown and github actions}
-\vskip 5em
+\tikz [remember picture, overlay] %
+\node [shift={(-0.6in,1.1cm)},scale=0.2,opacity=0.4] at (current page.south east)[anchor=south east]{\includegraphics{logo}};%
+\pagestyle{normal}
+\thispagestyle{firstpage}
 
+\newcommand{\colorRule}[3][black]{\textcolor[HTML]{#1}{\rule{#2}{#3}}}
+
+\noindent {\LARGE \textbf{\textsf{Template to prepare preprints and
+manuscripts using markdown and github actions}}}
+
+\medskip
+\begin{flushleft}
+{\small
 %
@@
 \,\textsuperscript{1,2}
-
-\textsuperscript{1}\,Université de
-Montréal\quad \textsuperscript{2}\,Québec Centre for Biodiversity
-Sciences
-
+\vskip 1em
+\textsuperscript{1}\,Université de Montréal; \textsuperscript{2}\,Québec
+Centre for Biodiversity Sciences\\
+\vskip 1em
 \textbf{Correspondance to:}\\
 Timothée Poisot --- \texttt{timothee.poisot@umontreal.ca}\\
+}
+\end{flushleft}
 
-\vfill
-This work is released by its authors under a CC-BY 4.0 license\hfill\ccby\\
-Last revision: \emph{\today}
+\vskip 2em
+\makebox[0pt][l]{\colorRule[CCCCCC]{2.0\textwidth}{0.5pt}}
+\vskip 2em
+\noindent
 
-\clearpage
-\thispagestyle{empty}
-
-\vfill
+\marginpar{\vskip 1em\flushright
+{\small{\bfseries Keywords}:\par
+species interaction networks\\binary classifiers\\machine
+learning\\regression\\supervised learning\\}
+}
 
@@
 
-\vfill
-
-\clearpage
-\linenumbers
-\pagestyle{normal}
+\vskip 2em
+\makebox[0pt][l]{\colorRule[CCCCCC]{2.0\textwidth}{0.5pt}}
+\vskip 2em
 

### 2021-12-07 draft.tex
Deploying to gh-pages from @ PoisotLab/ms_networkary_classifiers@c674d297723af7b596895362aa1bc58859555e33 🚀

@@
 \begin{enumerate}
-    \itemThe prediction of species interaction networks is gaining
+    \item The prediction of species interaction networks is gaining
 momentum as a way to circumvent limitations in data volume. Yet,
@@
 guidelines as to how predictive models can be trained.%
-    \itemUsing simple mathematical arguments and numerical experiments
+    \item Using simple mathematical arguments and numerical experiments
 in which a variety of classifiers (for supervised learning) are trained
@@
 unbiasing to apply to the training dataset.%
-    \itemClassifier accuracy and the ROC-AUC are not informative
+    \item Classifier accuracy and the ROC-AUC are not informative
 measures for the performance of interaction prediction. PR-AUC is a
@@
 connectance.%
-    \itemThese results reveal that training machines to predict networks
-is a challenging task, and that in virtually all cases, the composition
-of the training set needs to be experimented on before performing the
-actual training. We discuss these consequences in the context of the low
-volume of data.%
+    \item These results reveal that training machines to predict
+networks is a challenging task, and that in virtually all cases, the
+composition of the training set needs to be experimented on before
+performing the actual training. We discuss these consequences in the
+context of the low volume of data.%
 \end{enumerate}

### 2021-12-08 draft.tex
Deploying to gh-pages from @ PoisotLab/ms_networkary_classifiers@63549fd18f5af504ca6758ad8f0a0daf54daa184 🚀
Deploying to gh-pages from @ PoisotLab/ms_networkary_classifiers@66a85caf0e83f19b397eecd6090f5b69612a9004 🚀

@@
 \thispagestyle{empty}
-{\Large\bfseries\sffamily Template to prepare preprints and manuscripts
-using markdown and github actions}
+{\Large\bfseries\sffamily Guidelines for the supervised learning of
+species interactions}
 \vskip 5em
@@
 
+In this section, we will assume a network of connectance \(\rho\),
+\emph{i.e.} having \(\rho S^2\) interactions (where \(S\) is the species
+richness), and \((1-\rho) S^2\) non-interactions. Therefore, the vector
+describing the \emph{true} state of the network is
+\(\vec o = [\rho, (1-\rho)]\).
+
 \hypertarget{numerical-experiments}{%

### 2021-12-08 preprint.tex
Deploying to gh-pages from @ PoisotLab/ms_networkary_classifiers@63549fd18f5af504ca6758ad8f0a0daf54daa184 🚀
Deploying to gh-pages from @ PoisotLab/ms_networkary_classifiers@66a85caf0e83f19b397eecd6090f5b69612a9004 🚀

@@
 
-\noindent {\LARGE \textbf{\textsf{Template to prepare preprints and
-manuscripts using markdown and github actions}}}
+\noindent {\LARGE \textbf{\textsf{Guidelines for the supervised learning
+of species interactions}}}
 

### 2021-12-08 draft.tex
Deploying to gh-pages from @ PoisotLab/ms_networkary_classifiers@191f2799c8ed57ead307bc0ea76b62f84420343c 🚀

@@
 
+introduction to the confusion matrix
+
 list of problems to solve - baseline values and response to bias -
@@
 describing the \emph{true} state of the network is
-\(\vec o = [\rho, (1-\rho)]\).
+\(\mathbf{o}' = [\rho, (1-\rho)]\) (we can safely drop the \(S^2\)
+terms, as we will work on the confusion matrix, which ends up expressing
+\emph{relative} values).
+
+In order to write the values of the confusion matrix for a hypothetical
+classifier, we need to define two characteristics: its skill, and its
+bias. Skill, here, refers to the propensity of the classifier to get the
+correct answer (\emph{i.e.} to assign interactions where they are, and
+to not assign them where they are not). A no-skill classifier guesses at
+random, \emph{i.e.} it will guess interactions with a probability
+\(\rho\).
 

### 2021-12-08 draft.tex
Deploying to gh-pages from @ PoisotLab/ms_networkary_classifiers@de05b254eec882ac59dd902a5a1d22e9cd514b8d 🚀

@@
 richness), and \((1-\rho) S^2\) non-interactions. Therefore, the vector
-describing the \emph{true} state of the network is
-\(\mathbf{o}' = [\rho, (1-\rho)]\) (we can safely drop the \(S^2\)
+describing the \emph{true} state of the network is a column vector
+\(\mathbf{o}^T = [\rho (1-\rho)]\) (we can safely drop the \(S^2\)
 terms, as we will work on the confusion matrix, which ends up expressing
@@
 random, \emph{i.e.} it will guess interactions with a probability
-\(\rho\).
+\(\rho\). The predictions of a no-skill classifier can be expressed as a
+row vector \(\mathbf{p} = [\rho (1-\rho)]\). The confusion matrix
+\(\mathbf{M}\) for a no-skill classifier is given by the element-wise
+product of these vectors \(\mathbf{o} \odot \mathbf{p}\), \emph{i.e.}
+
+\[
+\mathbf{M} = \begin{pmatrix}
+    \rho^2 & \rho (1-\rho) \\
+    (1-\rho) \rho & (1-\rho)^2
+\end{pmatrix}
+\]
 

### 2021-12-08 draft.tex
Deploying to gh-pages from @ PoisotLab/ms_networkary_classifiers@ef77f3cb7cf4bb330d492cab4e350c00a27950b3 🚀

@@
     (1-\rho) \rho & (1-\rho)^2
+\end{pmatrix} \,.
+\]
+
+In order to regulate the skill of this classifier, we can define a skill
+matrix \(\mathbf{S}\) with diagonal elements equal to \(s\), and
+off-diagonal elements equal to \((1-s)\), and re-express the
+skill-adjusted confusion matrix as \(\mathbf{M} \odot \mathbf{S}\),
+\emph{i.e.}
+
+\[
+\begin{pmatrix}
+    \rho^2 & \rho (1-\rho) \\
+    (1-\rho) \rho & (1-\rho)^2
+\end{pmatrix} \odot \begin{pmatrix}
+    s & (1-s) \\
+    (1-s) & s
 \end{pmatrix}
@@
 
-\hypertarget{numerical-experiments}{%
-\section{Numerical experiments}\label{numerical-experiments}}
+Note that when \(s=0\), \(\mathbf{Tr}(\mathbf{M}) = 0\) (the classifier
+is \emph{always} wrong), when \(s=0.5\), the classifier is no-skill and
+guesses at random, and when \(s=1\), the classifier is perfect.
+
+The second element we can adjust in this hypothetical classifier is its
+bias, specifically its tendency to over-predict interactions. Like
+above, we can do so by defining a bias matrix \# Numerical experiments
 

### 2021-12-08 draft.tex
Deploying to gh-pages from @ PoisotLab/ms_networkary_classifiers@c337f170e85889437af63e80d0271b96ee61ca48 🚀

@@
     (1-s) & s
-\end{pmatrix}
+\end{pmatrix} \,.
 \]
@@
 bias, specifically its tendency to over-predict interactions. Like
-above, we can do so by defining a bias matrix \# Numerical experiments
+above, we can do so by defining a bias matrix \(\mathbf{B}\), where
+interactions are over-predicted with probability \(b\), and express the
+final classifier confusion matrix as
+\(\mathbf{M}\odot \mathbf{S}\odot \mathbf{B}\), \emph{i.e.}
+
+\[
+\begin{pmatrix}
+    \rho^2 & \rho (1-\rho) \\
+    (1-\rho) \rho & (1-\rho)^2
+\end{pmatrix} \odot \begin{pmatrix}
+    s & (1-s) \\
+    (1-s) & s
+\end{pmatrix} \odot \begin{pmatrix}
+    b & b \\
+    (1-b) & (1-b)
+\end{pmatrix}\,.
+\]
+
+The final expression for the confusion matrix in which we can regulate
+the skill and the bias is
+
+\[
+\mathbf{C} = \begin{pmatrix}
+    s\times b\times \rho^2 & (1-s)\times b\times \rho (1-\rho) \\
+    (1-s)\times (1-b)\times (1-\rho) \rho & s\times (1-b)\times (1-\rho)^2
+\end{pmatrix} \,.
+\]
+
+\hypertarget{numerical-experiments}{%
+\section{Numerical experiments}\label{numerical-experiments}}
 

### 2021-12-08 draft.tex
Deploying to gh-pages from @ PoisotLab/ms_networkary_classifiers@2aaa3526e1b06bdd5e6e0e7bf95b506fee905cc5 🚀

@@
 
-Note that when \(s=0\), \(\mathbf{Tr}(\mathbf{M}) = 0\) (the classifier
-is \emph{always} wrong), when \(s=0.5\), the classifier is no-skill and
+Note that when \(s=0\), \(\text{Tr}(\mathbf{M}) = 0\) (the classifier is
+\emph{always} wrong), when \(s=0.5\), the classifier is no-skill and
 guesses at random, and when \(s=1\), the classifier is perfect.

### 2021-12-08 draft.tex
Deploying to gh-pages from @ PoisotLab/ms_networkary_classifiers@0df28077f8bcc3a3e5912907b85311d7e7d64ee0 🚀

@@
 
-summary of model challenges for networks
+summary of model challenges for networks - Strydom et al. (2021)
+importance of drawing on traits + validation is challenging - Whalen et
+al. (2021) machine learning from genomics
 
@@
 \section{Baseline values}\label{baseline-values}}
+
+\hypertarget{confusion-matrix-with-skill-and-bias}{%
+\subsection{Confusion matrix with skill and
+bias}\label{confusion-matrix-with-skill-and-bias}}
 
@@
 
+\hypertarget{what-are-the-baseline-values-of-performance-measures}{%
+\subsection{What are the baseline values of performance
+measures?}\label{what-are-the-baseline-values-of-performance-measures}}
+
+In this section, we will change the values of \(b\) abd \(s\) for a
+given value of \(\rho\), and see how the values of common performance
+measures for binary classification are affected.
+
 \hypertarget{numerical-experiments}{%
@@
 \hypertarget{references}{%
-\section{References}\label{references}}
+\section*{References}\label{references}}
+\addcontentsline{toc}{section}{References}
+
+\hypertarget{refs}{}
+\begin{CSLReferences}{1}{0}
+\leavevmode\hypertarget{ref-Strydom2021RoaPre}{}%
+Strydom, T., Catchen, M. D., Banville, F., Caron, D., Dansereau, G.,
+Desjardins-Proulx, P., Forero-Muñoz, N. R., Higino, G., Mercier, B.,
+Gonzalez, A., Gravel, D., Pollock, L., \& Poisot, T. (2021). A roadmap
+towards predicting species interaction networks (across space and time).
+\emph{Philosophical Transactions of the Royal Society B: Biological
+Sciences}, \emph{376}(1837), 20210063.
+\url{https://doi.org/10.1098/rstb.2021.0063}
+
+\leavevmode\hypertarget{ref-Whalen2021NavPit}{}%
+Whalen, S., Schreiber, J., Noble, W. S., \& Pollard, K. S. (2021).
+Navigating the pitfalls of applying machine learning in genomics.
+\emph{Nature Reviews Genetics}, 1--13.
+\url{https://doi.org/10.1038/s41576-021-00434-9}
+
+\end{CSLReferences}
 

### 2021-12-08 draft.tex
Deploying to gh-pages from @ PoisotLab/ms_networkary_classifiers@77e05d04ed5f14a0bfd7e7d2f1748de2644bddfd 🚀

@@
 
+In the following section, we will generate random networks, and train
+four binary classifiers (as well as an ensemble model using the sum of
+the outputs) on 30\% of the interaction data. Networks are generated by
+picking random generality \(g\) and vulnerability \(v\) traits for
+\(S = 200\) species uniformly on the unit interval, and assigning an
+interaction from species \(i\) to species \(j\) if
+\(0.2g_i-\xi \le v_j \le 0.2g_i+\xi\), where \(\xi\) is a constant
+regulating the connectance of the networks, and varies uniformly in
+\([5\times 10^{-3}, 10^{-1}]\). This model gives fully interval networks
+that are close analogues to the niche model (Williams \& Martinez,
+2000), but has the benefit of only relying on two features
+(\(g_i, v_j\)), and having the exact same rule for all interactions. It
+is, therefore, a simple case which most classifiers should be able to
+learn.
+
 \hypertarget{effect-of-training-set-on-performance}{%
@@
 
+\leavevmode\hypertarget{ref-Williams2000SimRul}{}%
+Williams, R., \& Martinez, N. (2000). Simple rules yield complex food
+webs. \emph{Nature}, \emph{404}, 180--183.
+\href{http://userwww.sfsu.edu/\%20}{http://userwww.sfsu.edu/}
+
 \end{CSLReferences}

### 2021-12-08 draft.tex
Deploying to gh-pages from @ PoisotLab/ms_networkary_classifiers@a5943ded537de0e77810368ff8ebb79033739ccf 🚀

@@
 
-\hypertarget{effect-of-training-set-on-performance}{%
-\subsection{Effect of training set on
-performance}\label{effect-of-training-set-on-performance}}
+The training sample is composed of 30\% of the \(4\times 10^4\) possible
+entries in the network, \emph{i.e.} \(n=12000\). Out of these
+interactions, we pick a proportion \(\nu\) (the training set bias) to be
+positive, so that the training set has \(\nu n\) interactions, and
+\((1-\nu) n\) non-interactions. We vary \(\nu\) uniformly in \(]0,1[\).
+This allows to evaluate how the measures of binary classification
+performance respond to artificially rebalanced dataset for a given
+network connectance. Note that both \(\xi\) and \(\nu\) are sampled from
+a distribution rather than being picked on a grid; this is because there
+is no direct relationship between the value of \(\xi\) and the
+connectance of the simulated network, and therefore the precise value of
+\(\xi\) is not relevant for the analysis of the results.
+
+The dataset used for numerical experiments is composed of 20000 such
+\((\xi, \nu)\) pairs, on which four models are trained: a decision tree
+regressor, a boosted regression tree, a ridge regressor, and a random
+forest regressor. All models were taken from the \texttt{MLJ.jl} package
+(Blaom et al., 2020; Blaom \& Vollmer, 2020) in Julia 1.7 (Bezanson et
+al., 2017). The complete code to run the simulations is given as an
+appendix. \#\# Effect of training set on performance
 
@@
 \begin{CSLReferences}{1}{0}
+\leavevmode\hypertarget{ref-Bezanson2017JulFre}{}%
+Bezanson, J., Edelman, A., Karpinski, S., \& Shah, V. (2017). Julia: A
+Fresh Approach to Numerical Computing. \emph{SIAM Review}, \emph{59}(1),
+65--98. \url{https://doi.org/10.1137/141000671}
+
+\leavevmode\hypertarget{ref-Blaom2020MljJul}{}%
+Blaom, A. D., Kiraly, F., Lienart, T., Simillides, Y., Arenas, D., \&
+Vollmer, S. J. (2020). MLJ: A Julia package for composable machine
+learning. \emph{Journal of Open Source Software}, \emph{5}(55), 2704.
+\url{https://doi.org/10.21105/joss.02704}
+
+\leavevmode\hypertarget{ref-Blaom2020FleMod}{}%
+Blaom, A. D., \& Vollmer, S. J. (2020, December 31). \emph{Flexible
+model composition in machine learning and its implementation in MLJ}.
+\url{http://arxiv.org/abs/2012.15505}
+
 \leavevmode\hypertarget{ref-Strydom2021RoaPre}{}%

### 2021-12-08 draft.tex
Deploying to gh-pages from @ PoisotLab/ms_networkary_classifiers@63aca186d2ffaa114ff93cc23b677f40d26ffd9a 🚀

@@
 The dataset used for numerical experiments is composed of 20000 such
-\((\xi, \nu)\) pairs, on which four models are trained: a decision tree
-regressor, a boosted regression tree, a ridge regressor, and a random
-forest regressor. All models were taken from the \texttt{MLJ.jl} package
-(Blaom et al., 2020; Blaom \& Vollmer, 2020) in Julia 1.7 (Bezanson et
-al., 2017). The complete code to run the simulations is given as an
-appendix. \#\# Effect of training set on performance
+\((\xi, \nu)\) pairs, on which four learners are trained: a decision
+tree regressor, a boosted regression tree, a ridge regressor, and a
+random forest regressor. All models were taken from the \texttt{MLJ.jl}
+package (Blaom et al., 2020; Blaom \& Vollmer, 2020) in Julia 1.7
+(Bezanson et al., 2017). In order to pick the best adjacency matrix for
+a given learner, we performed a thresholding approach using 500 steps on
+predictions from the testing set, and picking the threshold that
+maximized Youden's informedness, which is usually the optimized target
+for imbalanced classification. During the thresholding step, we measured
+the area under the receiving-operator characteristic (ROC-AUC) and
+precision-recall (PR-AUC) curves, as measures of overall performance
+over the range of returned values. We report the ROC-AUC and PR-AUC, as
+well as a suite of other measures as introduced in the next section, for
+the best threshold. The ensemble model was generated by summing the
+predictions of all component models on the testing set (ranged in
+\([0,1]\)), then put through the same thresholding process. The complete
+code to run the simulations is given as an appendix.
+
+After the simulations were completed, we removed all runs (\emph{i.e.}
+pairs of \(\xi\) and \(\nu\)) for which at least one of the following
+conditions was met: the accuracy was 0, the true positive or true
+negative rates were 0, the connectance was larger than 0.2. This removes
+both the obviously failed model runs, and the networks that are more
+densely connected compared to the connectance of empirical food webs
+(and are therefore less difficult to predict, being less imbalanced).
+
+\hypertarget{effect-of-training-set-on-performance}{%
+\subsection{Effect of training set on
+performance}\label{effect-of-training-set-on-performance}}
 

### 2021-12-08 draft.tex
Deploying to gh-pages from @ PoisotLab/ms_networkary_classifiers@40755d0ee846afda73fd0e431f92d94208c395a1 🚀

@@
 
-introduction to the confusion matrix
+Binary classifiers are usually assessed by measuring properties of their
+confusion matrix, \emph{i.e.} the contingency table reporting true/false
+positive/negative hits. The same approach is used to evaluate
+\emph{e.g.} species occurrence models (Allouche et al., 2006). A
+confusion matrix is laid out as
+
+\[\begin{pmatrix}
+    \text{tp} & \text{fp} \\
+    \text{fn} & \text{tn}
+\end{pmatrix} \,,\]
+
+wherein \(\text{tp}\) is the number of interactions predicted as
+positive, \(\text{tn}\) is the number of non-interactions predicted as
+negative, \(\text{fp}\) is the number of non-interactions predicted as
+positive, and \(\text{fn}\) is the number of interactions predicted as
+negative. Almost all measures based on the confusion matrix express
+rates of error or success as proportions, and therefore the values of
+these components matter in a \emph{relative} way.
 
@@
 \begin{CSLReferences}{1}{0}
+\leavevmode\hypertarget{ref-Allouche2006AssAcc}{}%
+Allouche, O., Tsoar, A., \& Kadmon, R. (2006). Assessing the accuracy of
+species distribution models: Prevalence, kappa and the true skill
+statistic (TSS). \emph{Journal of Applied Ecology}, \emph{43}(6),
+1223--1232. \url{https://doi.org/10.1111/j.1365-2664.2006.01214.x}
+
 \leavevmode\hypertarget{ref-Bezanson2017JulFre}{}%

### 2021-12-08 draft.tex
Deploying to gh-pages from @ PoisotLab/ms_networkary_classifiers@5fff1ce67f2c5c2c640f8d6823454793f1667aba 🚀
Deploying to gh-pages from @ PoisotLab/ms_networkary_classifiers@cb9c3acc708fb9039a60bd90e98617b795b3246c 🚀

@@
 rates of error or success as proportions, and therefore the values of
-these components matter in a \emph{relative} way.
+these components matter in a \emph{relative} way. At a coarse scale, a
+classifier is \emph{accurate} when the trace of the matrix divided by
+the sum of the matrix is close to 1, with other measures focusing on
+different ways in which the classifer is wrong.
 
@@
 given value of \(\rho\), and see how the values of common performance
-measures for binary classification are affected.
+measures for binary classification are affected. Specifically, we will
+focus on four quantities: the accuracy
+(\((\text{tp}+\text{tn})/(\text{tp}+\text{tn}+\text{fp}+\text{fn})\)),
+the \emph{balanced} accuracy
+(\(\text{tp}/(2(\text{tp}+\text{fn}))+\text{tn}/(2(\text{tn}+\text{fp}))\)),
+Youden's J
+(\(\text{tp}/(\text{tp}+\text{fn})+\text{tn}/(\text{tn}+\text{fp})-1\)),
+and the \(F_1\) score (\(2\text{tp}/(2\text{tp}+\text{fp}+\text{fn})\)).
 

### 2021-12-08 draft.tex
Deploying to gh-pages from @ PoisotLab/ms_networkary_classifiers@eb8fed8c52c5e72640c6598edd3257ebc76d09d1 🚀

@@
 
+\textbf{Justification} of why these 4
+
+Assuming a no-skill unbiased classifier (\emph{i.e.}
+\(\mathbf{C}=\matbf{M}\)), the accuracy is \(\rho^2 + (1-\rho)^2\), the
+balanced accuracy is \(0.5\), Youden's J is \(0\), and \(F_1=\rho\). In
+other words, given a connectance \(\rho = 0.05\), we expect that a
+classifier guessing at random would still achieve an accuracy of
+\(0.905\); for a connectance of \(\rho = 0.01\), this accuracy
+\emph{increases} to over \(0.98\). In other words, networks with fewer
+interactions have inherently higher accuracy, because it is easy to
+predict the overwheling majority of non-interactions right.
+
+In order to examine how these values change w.r.t. the skill and bias,
+we performed a grid exploration of the values of \(s\) (from 0 to 1),
+and of \(\text{logit}(b)\), and visualize the result for a connectance
+of \textbf{TODO}.
+
 \hypertarget{numerical-experiments}{%

### 2021-12-08 draft.tex
Deploying to gh-pages from @ PoisotLab/ms_networkary_classifiers@3c7c22bca70e6bc71bb73bc363c08415f51deca3 🚀

@@
 Assuming a no-skill unbiased classifier (\emph{i.e.}
-\(\mathbf{C}=\matbf{M}\)), the accuracy is \(\rho^2 + (1-\rho)^2\), the
+\(\mathbf{C}=\mathbf{M}\)), the accuracy is \(\rho^2 + (1-\rho)^2\), the
 balanced accuracy is \(0.5\), Youden's J is \(0\), and \(F_1=\rho\). In
@@
 
-\hypertarget{effect-of-training-set-on-performance}{%
-\subsection{Effect of training set on
-performance}\label{effect-of-training-set-on-performance}}
+\hypertarget{effect-of-training-set-bias-on-performance}{%
+\subsection{Effect of training set bias on
+performance}\label{effect-of-training-set-bias-on-performance}}
 

### 2021-12-09 draft.tex
Deploying to gh-pages from @ PoisotLab/ms_networkary_classifiers@98b1970dc36f98ab33c18892639da1d76f72e072 🚀
Deploying to gh-pages from @ PoisotLab/ms_networkary_classifiers@441ee81db40c91c7be292d5e6cc0a92498cba381 🚀

@@
 
-Assuming a no-skill unbiased classifier (\emph{i.e.}
-\(\mathbf{C}=\mathbf{M}\)), the accuracy is \(\rho^2 + (1-\rho)^2\), the
-balanced accuracy is \(0.5\), Youden's J is \(0\), and \(F_1=\rho\). In
-other words, given a connectance \(\rho = 0.05\), we expect that a
-classifier guessing at random would still achieve an accuracy of
-\(0.905\); for a connectance of \(\rho = 0.01\), this accuracy
-\emph{increases} to over \(0.98\). In other words, networks with fewer
-interactions have inherently higher accuracy, because it is easy to
-predict the overwheling majority of non-interactions right.
+Assuming a no-skill unbiased classifier (\(s=0.5\), \emph{i.e.}
+\(\mathbf{C}=\mathbf{M}\) after normalization), the accuracy is
+\(\rho^2 + (1-\rho)^2\), the balanced accuracy is \(0.5\), Youden's J is
+\(0\), and \(F_1=\rho\). In other words, given a connectance
+\(\rho = 0.05\), we expect that a classifier guessing at random would
+still achieve an accuracy of \(0.905\); for a connectance of
+\(\rho = 0.01\), this accuracy \emph{increases} to over \(0.98\). In
+other words, networks with fewer interactions have inherently higher
+accuracy, because it is easy to predict the overwheling majority of
+non-interactions right.
 
@@
 of \textbf{TODO}.
+
+\hypertarget{what-are-the-baseline-values-of-thresholding-performance}{%
+\subsection{What are the baseline values of thresholding
+performance?}\label{what-are-the-baseline-values-of-thresholding-performance}}
+
+ROC-AUC v. PR-AUC, vary \(s\) to simulate a classifier ranging from
+no-skill to perfect
 

### 2021-12-09 code/confusionmatrix.jl
add code

@@ -1,0 +1,37 @@
+struct ConfusionMatrix
+    tp::Float64
+    tn::Float64
+    fp::Float64
+    fn::Float64
+    function ConfusionMatrix(a,b,c,d)
+        s = a+b+c+d
+        return new(a/s, b/s, c/s, d/s)
+    end
+end
+
+ConfusionMatrix(A::Matrix{Float64}) = ConfusionMatrix(A[1,1], A[2,2], A[2,1], A[1,2])
+
+tpr(M::ConfusionMatrix) = M.tp / (M.tp + M.fn)
+tnr(M::ConfusionMatrix) = M.tn / (M.tn + M.fp)
+ppv(M::ConfusionMatrix) = M.tp / (M.tp + M.fp)
+npv(M::ConfusionMatrix) = M.tn / (M.tn + M.fn)
+fnr(M::ConfusionMatrix) = M.fn / (M.fn + M.tp)
+fpr(M::ConfusionMatrix) = M.fp / (M.fp + M.tn)
+fdir(M::ConfusionMatrix) = M.fp / (M.fp + M.tp)
+fomr(M::ConfusionMatrix) = M.fn / (M.fn + M.tn)
+plr(M::ConfusionMatrix) = tpr(M) / fpr(M)
+nlr(M::ConfusionMatrix) = fnr(M) / tnr(M)
+pt(M::ConfusionMatrix) = sqrt(fpr(M)) / (sqrt(tpr(M)) + sqrt(fpr(M)))
+csi(M::ConfusionMatrix) = M.tp / (M.tp + M.fn + M.fp)
+prevalence(M::ConfusionMatrix) = (M.tp + M.fn) / (M.tp + M.fn + M.tn + M.fp)
+accuracy(M::ConfusionMatrix) = (M.tp + M.tn) / (M.tp + M.tn + M.fp + M.fn)
+balanced(M::ConfusionMatrix) = (tpr(M) + tnr(M)) * 0.5
+f1(M::ConfusionMatrix) = 2 * (ppv(M) * tpr(M)) / (ppv(M) + tpr(M))
+fm(M::ConfusionMatrix) = sqrt(ppv(M) * tpr(M))
+informedness(M::ConfusionMatrix) = tpr(M) + tnr(M) - 1.0
+markedness(M::ConfusionMatrix) = ppv(M) + npv(M) - 1.0
+dor(M::ConfusionMatrix) = plr(M) / nlr(M)
[…]

### 2021-12-09 code/connectance.jl

@@ -1,0 +1,120 @@
+using Statistics
+using StatsPlots
+using StatsFuns
+using StatsBase
+using Random
+using Statistics
+using MLJ
+
+# AUC
+function ∫(x::Array{T}, y::Array{T}) where {T<:Number}
+    S = zero(Float64)
+    for i in 2:length(x)
+        S += (x[i] - x[i - 1]) * (y[i] + y[i - 1]) * 0.5
+    end
+    return .-S
+end
+
+# Confusion matrix utilities
+include("confusionmatrix.jl")
+
+# Simple network-generating function with traits
+
+x = sort(rand(200))
+y = sort(rand(200))
+
+# Trait elements
+
+Tree = @load EvoTreeGaussian
+
+breadth = rand(250).*0.15
+
+𝐌 = Vector{ConfusionMatrix}(undef, length(breadth))
+ROCAUC = zeros(Float64, length(breadth))
+AUPRC = zeros(Float64, length(breadth))
+
+training_size = round(Int64, 0.3*length(𝐳))
+
+for (connectance_index, ρ) in enumerate(breadth)
+    
+    𝐱 = repeat(x; inner=length(y))
+    𝐲 = repeat(y; outer=length(x))
+    𝐳 = clamp.(𝐱 .- 0.8.*𝐱 .- ρ, 0, 1) .<= 𝐲 .<= clamp.(𝐱 .- 0.8.*𝐱 .+ ρ, 0, 1)
+    𝐗 = hcat(𝐱, 𝐲)
+    𝐘 = Float64.(𝐳)
+
+    n_positive = round(Int64, training_size*ρ)
+    idx_pos = sample(findall(iszero.(𝐘)), n_positive; replace=true)
+    idx_neg = sample(findall(isone.(𝐘)), training_size-n_positive; replace=true)
+
+    Iₚ = shuffle(vcat(idx_neg, idx_pos))
+    Iₒ = setdiff(eachindex(𝐘), Iₚ)
+    
+    for (midx,m) in enumerate(candidate_models)
+        this_machine = machine(m, table(𝐗), 𝐘)
+        fit!(this_machine, rows=Iₚ) 
+        prediction = mean.(MLJ.predict(this_machine, rows=Iₒ))
+        thresholds = LinRange(minimum(prediction), maximum(prediction), 500)
+        binobs = Bool.(𝐘[Iₒ])
+        M = Vector{ConfusionMatrix}(undef, length(thresholds))
[…]

### 2021-12-09 code/figures.jl

@@ -1,0 +1,67 @@
+using AlgebraOfGraphics, CairoMakie
+using DataFrames
+using CSV: CSV
+using DataFramesMeta
+using StatsBase
+using Statistics
+
+results = DataFrame(CSV.File("output.csv"))
+allowmissing!(results, :value)
+replace!(results.value, NaN => missing)
+dropmissing!(results)
+
+# Drop the models with obviously borked classifiers
+results.runid = hash.(results.connectance .* results.breadth .* results.bias)
+
+todrop = unique(vcat([
+    @subset(results, :measure .== "ACC", :value .== 0.0).runid,
+    @subset(results, :measure .== "TPR", :value .== 0.0).runid,
+    @subset(results, :measure .== "TNR", :value .== 0.0).runid,
+    @subset(results, :measure .== "FPR", :value .== 1.0).runid,
+    @subset(results, :measure .== "FNR", :value .== 1.0).runid,
+    @subset(results, :connectance .>= 0.2).runid
+]...))
+
+tokeep(f) = !(f in todrop)
+@subset!(results, tokeep.(:runid))
+
+# Make bins for connectance to make the plotting more efficient
+connectance_values = unique(select(results, [:runid, :connectance]))
+connectance_values.midpoint = zeros(Float64, size(connectance_values,1))
+n_connectance_bins = 4
+bins_connectance = LinRange(0.0, 0.2, n_connectance_bins+1)
+for i in 1:n_connectance_bins
+    _idx = findall(bins_connectance[i] .<= connectance_values.connectance .< bins_connectance[i+1])
+    connectance_values.midpoint[_idx] .= (bins_connectance[i+1] + bins_connectance[i])/2.0
+end
+
+# Join the two dataframes
+results = leftjoin(results, select(connectance_values, [:runid, :midpoint]), on=:runid)
+
+# Dataviz
+data(@subset(results, :measure .== "F1")) *
+    mapping(:bias, :value, row=:midpoint => nonnumeric, col=:model) *
+    (AlgebraOfGraphics.density() * visual(Contour, color=:grey, alpha=0.3) + smooth() * visual(linewidth=2.0)) |>
+    draw
+
+# Make bins for connectance to get the optimal bias
+connectance_values = unique(select(results, [:runid, :connectance]))
+connectance_values.conbin = zeros(Float64, size(connectance_values,1))
[…]

### 2021-12-09 code/main.jl

@@ -1,0 +1,35 @@
+using Statistics
+using AlgebraOfGraphics, CairoMakie
+using DataFrames
+using DataFramesMeta
+using CSV: CSV
+using StatsFuns
+
+# Confusion matrix utilities
+include("confusionmatrix.jl")
+
+function ConfusionMatrix(; s=0.0, b=0.5, ρ=0.05)
+    𝐂 = [ρ^2 (1 - ρ)*ρ; (1 - ρ)*ρ (1 - ρ)^2]
+    𝐒 = [s 1-s; 1-s s]
+    𝐁 = [b b; 1-b 1-b]
+    return ConfusionMatrix(𝐂 .* 𝐒 .* 𝐁)
+end
+
+# Calculations
+results = DataFrame(ρ = Float64[], s = Float64[], b = Float64[], measure = Symbol[], value = Float64[])
+ρ = 0.15
+for ls in LinRange(-10, 10, 9)
+    for lb in LinRange(-10, 10, 150)
+        𝐌 = ConfusionMatrix(ρ=ρ, s=logistic(ls), b=logistic(lb))
+        push!(results, (ρ, logistic(ls), logistic(lb), :Accuracy, accuracy(𝐌)))
+        push!(results, (ρ, logistic(ls), logistic(lb), Symbol("Positive Predictive Value"), ppv(𝐌)))
+        push!(results, (ρ, logistic(ls), logistic(lb), :F1, csi(𝐌)))
+        push!(results, (ρ, logistic(ls), logistic(lb), Symbol("κ"), κ(𝐌)))
+    end
+end
+
+data(results) *
+    mapping(:b => logit => "logit(bias)", :value => "Measure value", layout=:measure => nonnumeric, color=:s => logit => "logit(skill)", group=:s => nonnumeric) *
+    (visual(Lines, colormap=:tofino)) |> 
+    plt -> draw(plt, facet=(;linkyaxes = :none)) |>
+    plt -> save("analytics.png", plt, px_per_unit = 10)

### 2021-12-09 code/simulations.jl

@@ -1,0 +1,178 @@
+using Statistics
+using StatsFuns
+using StatsBase
+using DataFramesMeta
+using Random
+using Statistics
+using MLJ
+using DataFrames
+using CSV: CSV
+
+# AUC
+function ∫(x::Array{T}, y::Array{T}) where {T<:Number}
+    S = zero(Float64)
+    for i in 2:length(x)
+        S += (x[i] - x[i - 1]) * (y[i] + y[i - 1]) * 0.5
+    end
+    return .-S
+end
+
+# Confusion matrix utilities
+include("confusionmatrix.jl")
+
+# Network generation function
+function network(S, breadth)
+    # Get the network
+    x = sort(rand(S))
+    y = sort(rand(S))
+    𝐱 = repeat(x; inner=length(y))
+    𝐲 = repeat(y; outer=length(x))
+    𝐳 =
+        clamp.(𝐱 .- 0.8 .* 𝐱 .- breadth, 0, 1) .<=
+        𝐲 .<=
+        clamp.(𝐱 .- 0.8 .* 𝐱 .+ breadth, 0, 1)
+
+    # Features / labels matrix
+    return (table(hcat(𝐱, 𝐲)), Float64.(𝐳))
+end
+
+# Prepare the results
+results = [DataFrame(;
+    breadth=Float64[],
+    bias=Float64[],
+    connectance=Float64[],
+    model=Symbol[],
+    measure=Symbol[],
+    value=Float64[],
+) for _thr in 1:Threads.nthreads()]
+
+# Pick possible models based on a network
+_mock = network(10, 0.2)
+_mods = models() do model
+    matching(model, _mock...) &&
+        model.prediction_type == :deterministic &&
+        model.is_supervised &&
+        model.is_pure_julia
+end
+
+# these regression machines go brrr as f u c k
+DecisionTree = @load DecisionTreeRegressor pkg = DecisionTree
+RandomForest = @load RandomForestRegressor pkg = DecisionTree
+BoostedRegressor = @load EvoTreeRegressor pkg = EvoTrees
+RidgeRegressor = @load RidgeRegressor pkg = MLJLinearModels
+
+candidate_models = [
+    :DecTree => DecisionTree(),
+    :BRT => BoostedRegressor(),
+    :RF => RandomForest(),
+    :RR => RidgeRegressor()
+]
+
+S = 200
+
+_n_sims = 20000
+conditions_breadth = rand(_n_sims) .* 0.1 .+ 0.005
[…]

### 2021-12-09 draft.tex
Deploying to gh-pages from @ PoisotLab/ms_networkary_classifiers@9ef5c83b84870620d756aae345a87dc7b7f08d36 🚀
Deploying to gh-pages from @ PoisotLab/ms_networkary_classifiers@c3703997d21ea0b79a1ffa603d71af7891264633 🚀

@@
 
-\hypertarget{what-are-the-baseline-values-of-thresholding-performance}{%
-\subsection{What are the baseline values of thresholding
-performance?}\label{what-are-the-baseline-values-of-thresholding-performance}}
-
-ROC-AUC v. PR-AUC, vary \(s\) to simulate a classifier ranging from
-no-skill to perfect
+\begin{figure}
+\hypertarget{fig:analytics}{%
+\centering
+\includegraphics{figures/analytics.png}
+\caption{Consequences of changing the classifier skills (\(s\)) and bias
+(\(s\)) for a connectance \(\rho=0.15\), on accuracy, \(F_1\), postive
+predictive value, and \(\kappa\). \textbf{TODO} discussion of
+results}\label{fig:analytics}
+}
+\end{figure}
 

### 2021-12-10 draft.tex
Deploying to gh-pages from @ PoisotLab/ms_networkary_classifiers@9364d894c21b55736a0bd2d767725ff8e5523f1e 🚀

@@
 (\(s\)) for a connectance \(\rho=0.15\), on accuracy, \(F_1\), postive
-predictive value, and \(\kappa\). \textbf{TODO} discussion of
-results}\label{fig:analytics}
+predictive value, and \(\kappa\). Accuracy increases with skill, but
+also increases when the bias tends towards estimating \emph{fewer}
+interactions. The \(F_1\) score increases with skill but also increases
+when the bias tends towards estimating \emph{more} interactions; PPV
+behaves in the same way. Interestingly, \(\kappa\) responds as expected
+to skill (being negative whenever \(s < 0.5\)), and peaks for values of
+\(b \approx 0.5\); nevertheless, the value of bias for which \(\kappa\)
+is maximized in \emph{not} \(b=0.5\), but instead increases with
+classifier skill. In other words, at equal skill, maximizing \(\kappa\)
+would lead to select a \emph{more} bias
+classifier.}\label{fig:analytics}
 }

### 2021-12-10 draft.tex
Deploying to gh-pages from @ PoisotLab/ms_networkary_classifiers@c01929f4d4dbf64bab7adfa991b25d8cdd8d0e1e 🚀
Deploying to gh-pages from @ PoisotLab/ms_networkary_classifiers@c25f23fc0e4b5d860718769d5da56065cc637a6c 🚀

@@
 
+In all further simulations, the confusion matrix \(\mathbf{C}\) is
+transformed so that it sums to 1.
+
 \hypertarget{what-are-the-baseline-values-of-performance-measures}{%
@@
 In order to examine how these values change w.r.t. the skill and bias,
-we performed a grid exploration of the values of \(s\) (from 0 to 1),
-and of \(\text{logit}(b)\), and visualize the result for a connectance
-of \textbf{TODO}.
+we performed a grid exploration of the values of \(\text{logit}(s)\) and
+\(\text{logit}(b)\) linearly from \(-10\) to \(10\), and visualize the
+result for a connectance of \(\rho = 0.15\), which is within the range
+of usually observed connectance values for empirical food webs.
 
@@
 classifier skill. In other words, at equal skill, maximizing \(\kappa\)
-would lead to select a \emph{more} bias
+would lead to select a \emph{more} biased
 classifier.}\label{fig:analytics}

### 2021-12-10 code/confusionmatrix.jl
update code to put MCC and INF

@@ -35,3 +35,4 @@ function κ(M::ConfusionMatrix)
     return 2.0 * (M.tp * M.tn - M.fn * M.fp) /
            ((M.tp + M.fp) * (M.fp + M.tn) + (M.tp + M.fn) * (M.fn + M.tn))
 end
+mcc(M::ConfusionMatrix) = (M.tp*M.tn-M.fp*M.fn)/sqrt((M.tp+M.fp)*(M.tp+M.fn)*(M.tn+M.fp)*(M.tn+M.fn))

### 2021-12-10 code/main.jl

@@ -25,6 +25,8 @@ for ls in LinRange(-10, 10, 9)
         push!(results, (ρ, logistic(ls), logistic(lb), Symbol("Positive Predictive Value"), ppv(𝐌)))
         push!(results, (ρ, logistic(ls), logistic(lb), :F1, csi(𝐌)))
         push!(results, (ρ, logistic(ls), logistic(lb), Symbol("κ"), κ(𝐌)))
+        push!(results, (ρ, logistic(ls), logistic(lb), :MCC, mcc(𝐌)))
+        push!(results, (ρ, logistic(ls), logistic(lb), :Informedness, informedness(𝐌)))
     end
 end
 
@@ -32,4 +34,4 @@ data(results) *
     mapping(:b => logit => "logit(bias)", :value => "Measure value", layout=:measure => nonnumeric, color=:s => logit => "logit(skill)", group=:s => nonnumeric) *
     (visual(Lines, colormap=:tofino)) |> 
     plt -> draw(plt, facet=(;linkyaxes = :none)) |>
-    plt -> save("analytics.png", plt, px_per_unit = 10)
+    plt -> save("analytics.png", plt, px_per_unit = 3)

### 2021-12-10 draft.tex
Deploying to gh-pages from @ PoisotLab/ms_networkary_classifiers@e77d6c40b71a34d1032e8b719339193704d2bd58 🚀

@@
 confusion matrix, \emph{i.e.} the contingency table reporting true/false
-positive/negative hits. The same approach is used to evaluate
-\emph{e.g.} species occurrence models (Allouche et al., 2006). A
-confusion matrix is laid out as
+positive/negative hits. A confusion matrix is laid out as
 
@@
 
-list of problems to solve - baseline values and response to bias -
-effect of training set bias on performance - which models need the least
-amount of interactions to work
+The same approach is used to evaluate \emph{e.g.} species distribution
+models (SDMs). Indeed, the training and evaluation of SDMs as binary
+classifiers suffers from the same issue of low prevalence. In a previous
+work, Allouche et al. (2006) suggested that \(\kappa\) was a better test
+of model performance than the True Skill Statistic (TSS), which we will
+refer to as Youden's informedness (or \(J\)); these conclusions were
+later criticized by Somodi et al. (2017), who emphasized that
+informedness' relationship to prevalence depends on assumptions about
+bias in the model. Although this work offers recommendations about the
+comparison of models, it doesn't establishes baselines or good practices
+for training on imbalanced ecological data. Within the context of
+networks, there are three specific issues that need to be adressed.
+First, what values of performance measures are we expecting for a
+classifier that has poor performance? This is particularly important as
+it can evaluate whether low prevalence can lull us into a false sense of
+predictive accuracy. Second, independently of the question of model
+evaluation, is low prevalence an issue for \emph{training}, and can we
+remedy it? Finally, because the low amount of data on interaction makes
+a lot of imbalance correction methods (see \emph{e.g.} Branco et al.,
+2015) hard to apply, which indicators can be optimized with the least
+amount of positive interaction data?
 
-summary of the results
+We establish that due to the low prevalence of interactions, even poor
+classifiers applied to food web data will reach a high accuracy; this is
+because the measure is dominated by the accidental correct predictions
+of negatives. The \(F_1\) score and positive predictive values are less
+sensitive to bias, but \textbf{TODO}
+
+Chicco et al. (2021) - MCC maximizes other measures, other measures do
+not maximize MCC, except notably when prevalence is low, or the baseline
+guessing level is uncertain, which applies to networks. In this cases,
+informedness should be used as a maximization criterion.
 
@@
 
+\leavevmode\hypertarget{ref-Branco2015SurPre}{}%
+Branco, P., Torgo, L., \& Ribeiro, R. (2015, May 13). \emph{A Survey of
+Predictive Modelling under Imbalanced Distributions}.
+\url{http://arxiv.org/abs/1505.01658}
+
+\leavevmode\hypertarget{ref-Chicco2021MatCor}{}%
+Chicco, D., Tötsch, N., \& Jurman, G. (2021). The Matthews correlation
+coefficient (MCC) is more reliable than balanced accuracy, bookmaker
+informedness, and markedness in two-class confusion matrix evaluation.
+\emph{BioData Mining}, \emph{14}, 13.
+\url{https://doi.org/10.1186/s13040-021-00244-z}
+
+\leavevmode\hypertarget{ref-Somodi2017PreDep}{}%
+Somodi, I., Lepesi, N., \& Botta‐Dukát, Z. (2017). Prevalence dependence
+in model goodness measures with special emphasis on true skill
+statistics. \emph{Ecology and Evolution}, \emph{7}(3), 863--872.
+\url{https://doi.org/10.1002/ece3.2654}
+
 \leavevmode\hypertarget{ref-Strydom2021RoaPre}{}%

### 2021-12-10 code/main.jl
code update

@@ -17,21 +17,32 @@ end
 
 # Calculations
 results = DataFrame(ρ = Float64[], s = Float64[], b = Float64[], measure = Symbol[], value = Float64[])
-ρ = 0.15
-for ls in LinRange(-10, 10, 9)
-    for lb in LinRange(-10, 10, 150)
-        𝐌 = ConfusionMatrix(ρ=ρ, s=logistic(ls), b=logistic(lb))
-        push!(results, (ρ, logistic(ls), logistic(lb), :Accuracy, accuracy(𝐌)))
-        push!(results, (ρ, logistic(ls), logistic(lb), Symbol("Positive Predictive Value"), ppv(𝐌)))
-        push!(results, (ρ, logistic(ls), logistic(lb), :F1, csi(𝐌)))
-        push!(results, (ρ, logistic(ls), logistic(lb), Symbol("κ"), κ(𝐌)))
-        push!(results, (ρ, logistic(ls), logistic(lb), :MCC, mcc(𝐌)))
-        push!(results, (ρ, logistic(ls), logistic(lb), :Informedness, informedness(𝐌)))
+for ρ in LinRange(0.0, 0.5, 251)
+    if ρ > 0.0
+        for ls in LinRange(-10, 10, 9)
+            for lb in LinRange(-10, 10, 151)
+                𝐌 = ConfusionMatrix(ρ=ρ, s=logistic(ls), b=logistic(lb))
+                push!(results, (ρ, logistic(ls), logistic(lb), :Accuracy, accuracy(𝐌)))
+                push!(results, (ρ, logistic(ls), logistic(lb), :F1, csi(𝐌)))
+                push!(results, (ρ, logistic(ls), logistic(lb), Symbol("Positive Predictive Value"), ppv(𝐌)))
+                push!(results, (ρ, logistic(ls), logistic(lb), :MCC, mcc(𝐌)))
+                push!(results, (ρ, logistic(ls), logistic(lb), Symbol("κ"), κ(𝐌)))
+                push!(results, (ρ, logistic(ls), logistic(lb), :Informedness, informedness(𝐌)))
+            end
+        end
     end
 end
 
-data(results) *
+_co = unique(results.ρ)[last(findmin(abs.(unique(results.ρ) .- 0.1)))]
+
+data(@subset(results, :ρ .== _co)) *
[…]

### 2021-12-10 draft.tex
Deploying to gh-pages from @ PoisotLab/ms_networkary_classifiers@d8734bb468464e922e3370b1e282b19c3a028c70 🚀

@@
 guessing level is uncertain, which applies to networks. In this cases,
-informedness should be used as a maximization criterion.
+informedness should be used as a maximization criterion. Formulating
+guidelines is particularly important, because most of the litterature
+existing on optimizing classifier performance in the life sciences is
+focused on genomics applications (which has very specific challenges,
+see a recent discussion by Whalen et al., 2021), and can give
+contradictory recommendations (Boughorbel et al., 2017; Chicco \&
+Jurman, 2020; Delgado \& Tibau, 2019). This points not to a defficiency
+in the litterature, but rather to a need for domain-specific evaluation
+of how the particular ways in which datasets are biased can affect the
+performance of predictive models.
 
@@
 \begin{figure}
-\hypertarget{fig:analytics}{%
+\hypertarget{fig:bias}{%
 \centering
-\includegraphics{figures/analytics.png}
+\includegraphics{figures/changing-bias.png}
 \caption{Consequences of changing the classifier skills (\(s\)) and bias
@@
 classifier skill. In other words, at equal skill, maximizing \(\kappa\)
-would lead to select a \emph{more} biased
-classifier.}\label{fig:analytics}
+would lead to select a \emph{more} biased classifier.}\label{fig:bias}
+}
+\end{figure}
+
+\hypertarget{are-the-measures-affected-by-connectance}{%
+\subsection{Are the measures affected by
+connectance?}\label{are-the-measures-affected-by-connectance}}
+
+\begin{figure}
+\hypertarget{fig:connectance}{%
+\centering
+\includegraphics{figures/changing-connectance.png}
+\caption{TODO}\label{fig:connectance}
 }
@@
 
+\leavevmode\hypertarget{ref-Boughorbel2017OptCla}{}%
+Boughorbel, S., Jarray, F., \& El-Anbari, M. (2017). Optimal classifier
+for imbalanced data using Matthews Correlation Coefficient metric.
+\emph{PloS One}, \emph{12}(6), e0177678.
+\url{https://doi.org/10.1371/journal.pone.0177678}
+
 \leavevmode\hypertarget{ref-Branco2015SurPre}{}%
@@
 \url{http://arxiv.org/abs/1505.01658}
+
+\leavevmode\hypertarget{ref-Chicco2020AdvMat}{}%
+Chicco, D., \& Jurman, G. (2020). The advantages of the Matthews
+correlation coefficient (MCC) over F1 score and accuracy in binary
+classification evaluation. \emph{BMC Genomics}, \emph{21}(1), 6.
+\url{https://doi.org/10.1186/s12864-019-6413-7}
 
@@
 \url{https://doi.org/10.1186/s13040-021-00244-z}
+
+\leavevmode\hypertarget{ref-Delgado2019WhyCoh}{}%
+Delgado, R., \& Tibau, X.-A. (2019). Why Cohen's Kappa should be avoided
+as performance measure in classification. \emph{PloS One}, \emph{14}(9),
+e0222916. \url{https://doi.org/10.1371/journal.pone.0222916}
 

### 2021-12-11 draft.tex
Deploying to gh-pages from @ PoisotLab/ms_networkary_classifiers@be7007aaeef2c74ad180c22b4d5e984cbccdf2b1 🚀

@@
 
+\begin{itemize}
+\tightlist
+\item
   example on diagnostic test: rare events are hard to detect even with
   really good models
-
-summary of model challenges for networks - Strydom et al. (2021)
-importance of drawing on traits + validation is challenging - Whalen et
-al. (2021) machine learning from genomics
+\item
+  summary of model challenges for networks
+\item
+  Strydom et al. (2021) importance of drawing on traits + validation is
+  challenging + comparing across space
+\end{itemize}
 
@@
 informedness' relationship to prevalence depends on assumptions about
-bias in the model. Although this work offers recommendations about the
-comparison of models, it doesn't establishes baselines or good practices
-for training on imbalanced ecological data. Within the context of
-networks, there are three specific issues that need to be adressed.
-First, what values of performance measures are we expecting for a
-classifier that has poor performance? This is particularly important as
-it can evaluate whether low prevalence can lull us into a false sense of
-predictive accuracy. Second, independently of the question of model
-evaluation, is low prevalence an issue for \emph{training}, and can we
-remedy it? Finally, because the low amount of data on interaction makes
-a lot of imbalance correction methods (see \emph{e.g.} Branco et al.,
-2015) hard to apply, which indicators can be optimized with the least
-amount of positive interaction data?
+bias in the model, and therefore recommend the use of \(\kappa\) as a
+validation of classification performance. Although this work offers
+recommendations about the comparison of models, it doesn't establishes
+baselines or good practices for training on imbalanced ecological data.
+Within the context of networks, there are three specific issues that
+need to be adressed. First, what values of performance measures are we
+expecting for a classifier that has poor performance? This is
+particularly important as it can evaluate whether low prevalence can
+lull us into a false sense of predictive accuracy. Second, independently
+of the question of model evaluation, is low prevalence an issue for
+\emph{training}, and can we remedy it? Finally, because the low amount
+of data on interaction makes a lot of imbalance correction methods (see
+\emph{e.g.} Branco et al., 2015) hard to apply, which indicators can be
+optimized with the least amount of positive interaction data?
+
+In addition to the literature on SDMs, most of the research on machine
+learning application to life sciences is focused on genomics (which has
+very specific challenges, see a recent discussion by Whalen et al.,
+2021); this sub-field has generated largely different recommendations.
+Chicco \& Jurman (2020) suggest using Matthews correlation coefficient
+(MCC) over \(F_1\), as a protection against over-inflation of predicted
+results; Delgado \& Tibau (2019) advocate against the use of Cohen's
+\(\kappa\), again in favor of MCC, as the relative nature of \(\kappa\)
+means that a worse classifier can be picked over a better one;
+similarly, Boughorbel et al. (2017) recommend MCC over other measures of
+performance for imbalanced data, as it has more desirable statistical
+properties. More recently, Chicco et al. (2021) temper the apparent
+supremacy of the MCC, by suggesting it should be replaced by Youden's
+informedness (also known as \(J\), bookmaker's accuracy, and the
+True-Skill Statistic) when the imbalance in the dataset may not be
+representative (Jordano, 2016a, which is the case as networks are
+under-sampled; 2016b; McLeod et al., 2021), when classifiers need to be
+compared across different datasets, and when comparing the results to a
+no-skill (baseline) classifier is important. As these conditions are
+likely to be met with network data, there is a need to evaluate which
+measures of classification accuracy respond in a desirable way.
+
+A lot of binary classifiers are built by using a regressor (whose task
+is to guess the value of the interaction, amd can therefore return
+somethins considered to be a pseudo-probability); in this case, the
+optimal value below which predicitions are assumed to be negative
+(\emph{i.e.} the interaction does not exist) can be determined by
+picking a threshold maximizing some value on the ROC curve or the PR
+curve. The area under these curves (ROC-AUC and PR-AUC henceforth) give
+ideas on the overall goodness of the classifier. Saito \& Rehmsmeier
+(2015) established that the ROC-AUC is biased towards over-estimating
+performance for imbalanced data.
 
@@
 sensitive to bias, but \textbf{TODO}
-
-Chicco et al. (2021) - MCC maximizes other measures, other measures do
-not maximize MCC, except notably when prevalence is low, or the baseline
-guessing level is uncertain, which applies to networks. In this cases,
-informedness should be used as a maximization criterion. Formulating
-guidelines is particularly important, because most of the litterature
-existing on optimizing classifier performance in the life sciences is
-focused on genomics applications (which has very specific challenges,
-see a recent discussion by Whalen et al., 2021), and can give
-contradictory recommendations (Boughorbel et al., 2017; Chicco \&
-Jurman, 2020; Delgado \& Tibau, 2019). This points not to a defficiency
-in the litterature, but rather to a need for domain-specific evaluation
-of how the particular ways in which datasets are biased can affect the
-performance of predictive models.
 
@@
 
+\leavevmode\hypertarget{ref-Jordano2016ChaEco}{}%
+Jordano, P. (2016a). Chasing Ecological Interactions. \emph{PLOS Biol},
+\emph{14}(9), e1002559.
+\url{https://doi.org/10.1371/journal.pbio.1002559}
+
+\leavevmode\hypertarget{ref-Jordano2016SamNet}{}%
+Jordano, P. (2016b). Sampling networks of ecological interactions.
+\emph{Functional Ecology}. \url{https://doi.org/10.1111/1365-2435.12763}
+
+\leavevmode\hypertarget{ref-McLeod2021SamAsy}{}%
+McLeod, A., Leroux, S. J., Gravel, D., Chu, C., Cirtwill, A. R., Fortin,
+M.-J., Galiana, N., Poisot, T., \& Wood, S. A. (2021). Sampling and
+asymptotic network properties of spatial multi-trophic networks.
+\emph{Oikos}, \emph{n/a}(n/a). \url{https://doi.org/10.1111/oik.08650}
+
+\leavevmode\hypertarget{ref-Saito2015PrePlo}{}%
+Saito, T., \& Rehmsmeier, M. (2015). The Precision-Recall Plot Is More
+Informative than the ROC Plot When Evaluating Binary Classifiers on
+Imbalanced Datasets. \emph{PLOS ONE}, \emph{10}(3), e0118432.
+\url{https://doi.org/10.1371/journal.pone.0118432}
+
 \leavevmode\hypertarget{ref-Somodi2017PreDep}{}%

### 2021-12-11 draft.tex
Deploying to gh-pages from @ PoisotLab/ms_networkary_classifiers@9e9861285e5f075d258bb016ab6a58430c5ff49d 🚀

@@
 representative (Jordano, 2016a, which is the case as networks are
-under-sampled; 2016b; McLeod et al., 2021), when classifiers need to be
-compared across different datasets, and when comparing the results to a
-no-skill (baseline) classifier is important. As these conditions are
-likely to be met with network data, there is a need to evaluate which
-measures of classification accuracy respond in a desirable way.
+under-sampled; 2016b), when classifiers need to be compared across
+different datasets (for example when predicting a system in space, where
+undersampling varies locally; McLeod et al., 2021), and when comparing
+the results to a no-skill (baseline) classifier is important. As these
+conditions are likely to be met with network data, there is a need to
+evaluate which measures of classification accuracy respond in a
+desirable way.
 
@@
 (2015) established that the ROC-AUC is biased towards over-estimating
-performance for imbalanced data.
+performance for imbalanced data; on the contrary, the PR-AUC is able to
+identify classifiers that are less able to detect positive interactions
+correctly, with the additional advantage of having a baseline value
+equal to prevalence. Therefore, it is important to assess whether these
+two measures return different results when applied to ecological network
+prediction.
 

### 2021-12-11 code/analytical_results.jl
move code to top-level

@@ -6,7 +6,7 @@ using CSV: CSV
 using StatsFuns
 
 # Confusion matrix utilities
-include("confusionmatrix.jl")
+include(joinpath(@__DIR__, "code", "confusionmatrix.jl"))
 
 function ConfusionMatrix(; s=0.0, b=0.5, ρ=0.05)
     𝐂 = [ρ^2 (1 - ρ)*ρ; (1 - ρ)*ρ (1 - ρ)^2]
@@ -22,9 +22,9 @@ for ρ in LinRange(0.0, 0.5, 251)
         for ls in LinRange(-10, 10, 9)
             for lb in LinRange(-10, 10, 151)
                 𝐌 = ConfusionMatrix(ρ=ρ, s=logistic(ls), b=logistic(lb))
-                push!(results, (ρ, logistic(ls), logistic(lb), :Accuracy, accuracy(𝐌)))
+                #push!(results, (ρ, logistic(ls), logistic(lb), :Accuracy, accuracy(𝐌)))
                 push!(results, (ρ, logistic(ls), logistic(lb), :F1, csi(𝐌)))
-                push!(results, (ρ, logistic(ls), logistic(lb), Symbol("Positive Predictive Value"), ppv(𝐌)))
+                #push!(results, (ρ, logistic(ls), logistic(lb), Symbol("Positive Predictive Value"), ppv(𝐌)))
                 push!(results, (ρ, logistic(ls), logistic(lb), :MCC, mcc(𝐌)))
                 push!(results, (ρ, logistic(ls), logistic(lb), Symbol("κ"), κ(𝐌)))
                 push!(results, (ρ, logistic(ls), logistic(lb), :Informedness, informedness(𝐌)))
@@ -39,10 +39,10 @@ data(@subset(results, :ρ .== _co)) *
     mapping(:b => logit => "logit(bias)", :value => "Measure value", layout=:measure => nonnumeric, color=:s => logit => "logit(skill)", group=:s => nonnumeric) *
     (visual(Lines, colormap=:tofino)) |> 
     plt -> draw(plt, facet=(;linkyaxes = :none)) |>
-    plt -> save("changing-bias.png", plt, px_per_unit = 3)
+    plt -> save(joinpath(@__DIR__, "..", "figures", "changing-bias.png"), plt, px_per_unit = 3)
 
 data(@subset(results, :b .== 0.5)) *
[…]

### 2021-12-11 draft.tex
Deploying to gh-pages from @ PoisotLab/ms_networkary_classifiers@c2429146b928e0e2a7813e8721eac630a5aced5b 🚀

@@
 
-In this section, we will change the values of \(b\) abd \(s\) for a
-given value of \(\rho\), and see how the values of common performance
-measures for binary classification are affected. Specifically, we will
-focus on four quantities: the accuracy
-(\((\text{tp}+\text{tn})/(\text{tp}+\text{tn}+\text{fp}+\text{fn})\)),
-the \emph{balanced} accuracy
-(\(\text{tp}/(2(\text{tp}+\text{fn}))+\text{tn}/(2(\text{tn}+\text{fp}))\)),
-Youden's J
-(\(\text{tp}/(\text{tp}+\text{fn})+\text{tn}/(\text{tn}+\text{fp})-1\)),
-and the \(F_1\) score (\(2\text{tp}/(2\text{tp}+\text{fp}+\text{fn})\)).
+In this section, we will change the values of \(b\), \(s\), and
+\(\rho\), and report how the main measures discussed in the introduction
+(MCC, \(F_1\), \(\kappa\), and informedness) are responding to issues
+with the classifier. Before we do so, it is important to explain why we
+will not focus on accuracy too much. Accuracy is the number of correct
+predictions (\(\text{Tr}(\mathbf{C})\)) divided by the sum of the
+confusion matrix. For a no-skill, no-bias classifier, accuracy is equal
+to \(\rho^2 + (1-\rho)^2\); for \(\rho = 0.05\), this is
+\(\approx 0.90\), and for \(\rho = 0.01\), this is equal to
+\(\approx 0.98\). In other words, the values of accuracy are expected to
+be so high that they are not really informatived (this is simply
+explained by the fact that for \(\rho\) small,
+\(\rho^2 \ll (1-\rho)^2\)). More concerning is the fact that introducing
+bias changes the response of accuracy in unexpected ways. Assuming a
+no-skill classifier, the numerator of accuracy becomes
+\(b\rho^2 + (1-b)(1-\rho)^2\), which increases when \(b\) is low, which
+specifically means that at equal skill, a classifier that under-predicts
+interactions will have higher accuracy than an un-biased classifier.
+These issues are absent from balanced accuracy, but should nevertheless
+lead us to not report accuracy as the primary measure of network
+prediction success; moving forward, we will focus on other measures.
 
-\textbf{Justification} of why these 4
-
-Assuming a no-skill unbiased classifier (\(s=0.5\), \emph{i.e.}
-\(\mathbf{C}=\mathbf{M}\) after normalization), the accuracy is
-\(\rho^2 + (1-\rho)^2\), the balanced accuracy is \(0.5\), Youden's J is
-\(0\), and \(F_1=\rho\). In other words, given a connectance
-\(\rho = 0.05\), we expect that a classifier guessing at random would
-still achieve an accuracy of \(0.905\); for a connectance of
-\(\rho = 0.01\), this accuracy \emph{increases} to over \(0.98\). In
-other words, networks with fewer interactions have inherently higher
-accuracy, because it is easy to predict the overwheling majority of
-non-interactions right.
-
-In order to examine how these values change w.r.t. the skill and bias,
-we performed a grid exploration of the values of \(\text{logit}(s)\) and
-\(\text{logit}(b)\) linearly from \(-10\) to \(10\), and visualize the
-result for a connectance of \(\rho = 0.15\), which is within the range
-of usually observed connectance values for empirical food webs.
+In order to examine how MCC, \(F_1\), \(\kappa\), and informedness
+change w.r.t. the imbalance, skill, and bias, we performed a grid
+exploration of the values of \(\text{logit}(s)\) and \(\text{logit}(b)\)
+linearly from \(-10\) to \(10\), of \(\rho\) linearly in \(]0, 0.5]\),
+which is within the range of usually observed connectance values for
+empirical food webs. Note that at this point, there is no food web model
+to speak of; rather, the confusion matrix we discuss can be obtained for
+any classification task.
 

### 2021-12-11 draft.tex
Deploying to gh-pages from @ PoisotLab/ms_networkary_classifiers@7e00940b0c342f4a700a0ca036ad502368be0f0d 🚀

@@
 
+\hypertarget{definition-of-the-performance-measures}{%
+\subsection{Definition of the performance
+measures}\label{definition-of-the-performance-measures}}
+
+\(\kappa\)
+
+\(F_{\beta}\)
+
+informedness
+
+MCC
+
 \hypertarget{confusion-matrix-with-skill-and-bias}{%
@@
 to speak of; rather, the confusion matrix we discuss can be obtained for
-any classification task.
+any classification task. Based on the previous discussion, the desirable
+properties for a measure of classifier success should be: an increase
+with classifier skill, especially at low bias; a hump-shaped response to
+bias, especially at high skill, and ideally center around
+\(\text{logit}(b)=0\); an increase with prevalence up until
+equiprevalence is reached.
 
@@
 
-\hypertarget{are-the-measures-affected-by-connectance}{%
-\subsection{Are the measures affected by
-connectance?}\label{are-the-measures-affected-by-connectance}}
+In fig.~\ref{fig:bias}, we show that none of the four measures satisfy
+all the considerations at once: \(F_1\) increases with skill, and
+increases monotonously with bias; this is because \(F_1\) does not
+account for true negatives, and the increase in positive detection masks
+the over-prediction of interactions. Informedness varies with skill,
+reaching 0 for a no-skill classifier, but is entirely unsensitive to
+bias. Both MCC and \(\kappa\) have the same behavior, whereby they
+increase with skill. \(\kappa\) peaks at increasing values of biass for
+increasing skill, \emph{i.e.} is likely to lead to the selection of a
+classifier that over-predicts interactions. By contract, MCC peaks at
+the same value, regardless of skill, but this value is not
+\(\text{logit}(b)=0\): unless at very high classifier skill, MCC risks
+leading to a model that over-predicts interactions. In
+fig.~\ref{fig:connectance}, we show that all measures except \(F_1\)
+give a value of 0 for a no-skill classifier, and are forced towars their
+correct maximal value when skill changes (\emph{i.e.} a more connected
+networks will have higher values for a skilled classifierd, and lower
+values for a classifier making mostly mistakes).
 

### 2021-12-11 draft.tex
Deploying to gh-pages from @ PoisotLab/ms_networkary_classifiers@9a12647fe4aedea387c0b211284a79dddb676f28 🚀

@@
 
+These two analyses point to the following recommendations: MCC is indeed
+more appropriate than \(\kappa\), as although sensitive to bias, it is
+sensitive in a consistent way. Informedness is appropriate at
+discriminating between different skills, but confounded by bias. \(F_1\)
+is increasing with bias, and should not be prioritized to evalue the
+performance of the model. The discussion of sensitivity to bias should
+come with a domain-specific caveat: although it is likely that
+interactions documented in ecological networks are correct, a lot of
+non-interactions are simply unobserved; as predictive models are used
+for data-inflation (\emph{i.e.} the prediction of new interactions), it
+is not necessarilly a bad thing to select models that predict more
+interactions than the original dataset, because the original dataset
+misses some interactions.
+
 \hypertarget{numerical-experiments}{%

### 2021-12-11 code/02_simulations.jl
rename code

@@ -18,7 +18,7 @@ function ∫(x::Array{T}, y::Array{T}) where {T<:Number}
 end
 
 # Confusion matrix utilities
-include("confusionmatrix.jl")
+include(joinpath(@__DIR__, "confusionmatrix.jl"))
 
 # Network generation function
 function network(S, breadth)
@@ -70,7 +70,7 @@ candidate_models = [
 
 S = 200
 
-_n_sims = 20000
+_n_sims = 2000
 conditions_breadth = rand(_n_sims) .* 0.1 .+ 0.005
 conditions_bias = rand(_n_sims) .* 0.98 .+ 0.01
 conditions = hcat(conditions_breadth, conditions_bias)
@@ -132,6 +132,7 @@ Threads.@threads for i in 1:size(conditions, 1)
         push!(results[Threads.threadid()], (breadth, bias, mean(𝐲), candidate_model.first, :NPV, npv(𝐌)))
         push!(results[Threads.threadid()], (breadth, bias, mean(𝐲), candidate_model.first, :MKD, markedness(𝐌)))
         push!(results[Threads.threadid()], (breadth, bias, mean(𝐲), candidate_model.first, :F1, f1(𝐌)))
+        push!(results[Threads.threadid()], (breadth, bias, mean(𝐲), candidate_model.first, :MCC, mcc(𝐌)))
     end
 
     # Ensemble model
@@ -172,7 +173,8 @@ Threads.@threads for i in 1:size(conditions, 1)
         push!(results[Threads.threadid()], (breadth, bias, mean(𝐲), :ensemble, :NPV, npv(𝐌)))
         push!(results[Threads.threadid()], (breadth, bias, mean(𝐲), :ensemble, :MKD, markedness(𝐌)))
         push!(results[Threads.threadid()], (breadth, bias, mean(𝐲), :ensemble, :F1, f1(𝐌)))
+        push!(results[Threads.threadid()], (breadth, bias, mean(𝐲), :ensemble, :MCC, mcc(𝐌)))
     end
 end
 
-CSV.write("output.csv", vcat(results...))
+CSV.write(joinpath(@__DIR__, "output.csv"), vcat(results...))

### 2021-12-11 code/03_figures.jl

@@ -5,7 +5,7 @@ using DataFramesMeta
 using StatsBase
 using Statistics
 
-results = DataFrame(CSV.File("output.csv"))
+results = DataFrame(CSV.File(joinpath(@__DIR__, "output.csv")))
 allowmissing!(results, :value)
 replace!(results.value, NaN => missing)
 dropmissing!(results)
@@ -39,10 +39,11 @@ end
 results = leftjoin(results, select(connectance_values, [:runid, :midpoint]), on=:runid)
 
 # Dataviz
-data(@subset(results, :measure .== "F1")) *
+data(@subset(results, :measure .== "MCC")) *
     mapping(:bias, :value, row=:midpoint => nonnumeric, col=:model) *
     (AlgebraOfGraphics.density() * visual(Contour, color=:grey, alpha=0.3) + smooth() * visual(linewidth=2.0)) |>
-    draw
+    plt -> draw(plt, facet=(;linkyaxes = :none)) |>
+    plt -> save(joinpath(@__DIR__, "..", "figures", "changing-bias.png"), plt, px_per_unit = 3)
 
 # Make bins for connectance to get the optimal bias
 connectance_values = unique(select(results, [:runid, :connectance]))

### 2021-12-11 code/connectance.jl

@@ -1,120 +1,0 @@
-using Statistics
-using StatsPlots
-using StatsFuns
-using StatsBase
-using Random
-using Statistics
-using MLJ
-
-# AUC
-function ∫(x::Array{T}, y::Array{T}) where {T<:Number}
-    S = zero(Float64)
-    for i in 2:length(x)
-        S += (x[i] - x[i - 1]) * (y[i] + y[i - 1]) * 0.5
-    end
-    return .-S
-end
-
-# Confusion matrix utilities
-include("confusionmatrix.jl")
-
-# Simple network-generating function with traits
-
-x = sort(rand(200))
-y = sort(rand(200))
-
-# Trait elements
-
-Tree = @load EvoTreeGaussian
-
-breadth = rand(250).*0.15
-
-𝐌 = Vector{ConfusionMatrix}(undef, length(breadth))
-ROCAUC = zeros(Float64, length(breadth))
-AUPRC = zeros(Float64, length(breadth))
-
-training_size = round(Int64, 0.3*length(𝐳))
-
-for (connectance_index, ρ) in enumerate(breadth)
-    
-    𝐱 = repeat(x; inner=length(y))
-    𝐲 = repeat(y; outer=length(x))
-    𝐳 = clamp.(𝐱 .- 0.8.*𝐱 .- ρ, 0, 1) .<= 𝐲 .<= clamp.(𝐱 .- 0.8.*𝐱 .+ ρ, 0, 1)
-    𝐗 = hcat(𝐱, 𝐲)
-    𝐘 = Float64.(𝐳)
-
-    n_positive = round(Int64, training_size*ρ)
-    idx_pos = sample(findall(iszero.(𝐘)), n_positive; replace=true)
-    idx_neg = sample(findall(isone.(𝐘)), training_size-n_positive; replace=true)
-
-    Iₚ = shuffle(vcat(idx_neg, idx_pos))
-    Iₒ = setdiff(eachindex(𝐘), Iₚ)
-    
-    for (midx,m) in enumerate(candidate_models)
-        this_machine = machine(m, table(𝐗), 𝐘)
-        fit!(this_machine, rows=Iₚ) 
-        prediction = mean.(MLJ.predict(this_machine, rows=Iₒ))
-        thresholds = LinRange(minimum(prediction), maximum(prediction), 500)
-        binobs = Bool.(𝐘[Iₒ])
-        M = Vector{ConfusionMatrix}(undef, length(thresholds))
[…]

### 2021-12-11 code/03_figures.jl
results + code update, ready to write last section

@@ -39,11 +39,19 @@ end
 results = leftjoin(results, select(connectance_values, [:runid, :midpoint]), on=:runid)
 
 # Dataviz
-data(@subset(results, :measure .== "MCC")) *
-    mapping(:bias, :value, row=:midpoint => nonnumeric, col=:model) *
+_keepval(f) = f in ["MCC", "INF"]
+data(@subset(results, _keepval.(:measure))) *
+    mapping(:bias => "Training set bias", :value => "Value", row=:midpoint => nonnumeric, col=:model, color=:measure) *
     (AlgebraOfGraphics.density() * visual(Contour, color=:grey, alpha=0.3) + smooth() * visual(linewidth=2.0)) |>
     plt -> draw(plt, facet=(;linkyaxes = :none)) |>
-    plt -> save(joinpath(@__DIR__, "..", "figures", "changing-bias.png"), plt, px_per_unit = 3)
+    plt -> save(joinpath(@__DIR__, "..", "figures", "bias_mcc_inf.png"), plt, px_per_unit = 3)
+
+_keepval(f) = f in ["PRAUC", "ROCAUC"]
+data(@subset(results, _keepval.(:measure))) *
+    mapping(:bias => "Training set bias", :value => "Value", row=:midpoint => nonnumeric, col=:model, color=:measure) *
+    (AlgebraOfGraphics.density() * visual(Contour, color=:grey, alpha=0.3) + smooth() * visual(linewidth=2.0)) |>
+    plt -> draw(plt, facet=(;linkyaxes = :none)) |>
+    plt -> save(joinpath(@__DIR__, "..", "figures", "bias_roc_pr.png"), plt, px_per_unit = 3)
 
 # Make bins for connectance to get the optimal bias
 connectance_values = unique(select(results, [:runid, :connectance]))
@@ -59,10 +67,9 @@ results = leftjoin(results, select(connectance_values, [:runid, :conbin]), on=:r
 bmm = groupby(results, [:model, :measure, :conbin])
 opt = combine(bmm, [:value, :bias] => ((v,b) -> median(b[sortperm(v)[end-min(20, length(v))+1:end]])) => :optimalbias)
 
-_keepval(f) = f in ["PRAUC", "F1", "INF"]
-@subset!(opt, _keepval.(:measure))
-
-data(opt) * 
+_keepval(f) = f in ["MCC", "INF", "PRAUC", "ROCAUC"]
[…]

### 2021-12-11 code/02_simulations.jl
second section

@@ -70,7 +70,7 @@ candidate_models = [
 
 S = 200
 
-_n_sims = 2000
+_n_sims = 500
 conditions_breadth = rand(_n_sims) .* 0.1 .+ 0.005
 conditions_bias = rand(_n_sims) .* 0.98 .+ 0.01
 conditions = hcat(conditions_breadth, conditions_bias)
@@ -133,6 +133,7 @@ Threads.@threads for i in 1:size(conditions, 1)
         push!(results[Threads.threadid()], (breadth, bias, mean(𝐲), candidate_model.first, :MKD, markedness(𝐌)))
         push!(results[Threads.threadid()], (breadth, bias, mean(𝐲), candidate_model.first, :F1, f1(𝐌)))
         push!(results[Threads.threadid()], (breadth, bias, mean(𝐲), candidate_model.first, :MCC, mcc(𝐌)))
+        push!(results[Threads.threadid()], (breadth, bias, mean(𝐲), candidate_model.first, :postbias, (𝐌.tp+𝐌.fp)/(𝐌.tp+𝐌.fn)))
     end
 
     # Ensemble model
@@ -174,6 +175,7 @@ Threads.@threads for i in 1:size(conditions, 1)
         push!(results[Threads.threadid()], (breadth, bias, mean(𝐲), :ensemble, :MKD, markedness(𝐌)))
         push!(results[Threads.threadid()], (breadth, bias, mean(𝐲), :ensemble, :F1, f1(𝐌)))
         push!(results[Threads.threadid()], (breadth, bias, mean(𝐲), :ensemble, :MCC, mcc(𝐌)))
+        push!(results[Threads.threadid()], (breadth, bias, mean(𝐲), :ensemble, :postbias, (𝐌.tp+𝐌.fp)/(𝐌.tp+𝐌.fn)))
     end
 end
 

### 2021-12-11 code/03_figures.jl

@@ -32,7 +32,7 @@ n_connectance_bins = 4
 bins_connectance = LinRange(0.0, 0.2, n_connectance_bins+1)
 for i in 1:n_connectance_bins
     _idx = findall(bins_connectance[i] .<= connectance_values.connectance .< bins_connectance[i+1])
-    connectance_values.midpoint[_idx] .= (bins_connectance[i+1] + bins_connectance[i])/2.0
+    connectance_values.midpoint[_idx] .= round((bins_connectance[i+1] + bins_connectance[i])/2.0; digits=4)
 end
 
 # Join the two dataframes
@@ -52,6 +52,20 @@ data(@subset(results, _keepval.(:measure))) *
     (AlgebraOfGraphics.density() * visual(Contour, color=:grey, alpha=0.3) + smooth() * visual(linewidth=2.0)) |>
     plt -> draw(plt, facet=(;linkyaxes = :none)) |>
     plt -> save(joinpath(@__DIR__, "..", "figures", "bias_roc_pr.png"), plt, px_per_unit = 3)
+
+_keepval(f) = f in ["PT"]
+data(@subset(results, _keepval.(:measure))) *
+    mapping(:bias => "Training set bias", :value => "Prevalence threshold", row=:midpoint => nonnumeric, col=:model) *
+    (AlgebraOfGraphics.density() * visual(Contour, color=:grey, alpha=0.3) + smooth() * visual(linewidth=2.0)) |>
+    plt -> draw(plt, facet=(;linkyaxes = :none)) |>
+    plt -> save(joinpath(@__DIR__, "..", "figures", "bias_pt.png"), plt, px_per_unit = 3)
+
+_keepval(f) = f in ["postbias"]
+data(@subset(results, _keepval.(:measure))) *
+    mapping(:bias => "Training set bias", :value => logistic => "Value", row=:midpoint => nonnumeric, col=:model) *
+    (AlgebraOfGraphics.density() * visual(Contour, color=:grey, alpha=0.3) + smooth() * visual(linewidth=2.0)) |>
+    plt -> draw(plt, facet=(;linkyaxes = :none)) |>
+    plt -> save(joinpath(@__DIR__, "..", "figures", "bias_co.png"), plt, px_per_unit = 3)
 
 # Make bins for connectance to get the optimal bias
 connectance_values = unique(select(results, [:runid, :connectance]))

### 2021-12-11 draft.tex
Deploying to gh-pages from @ PoisotLab/ms_networkary_classifiers@fc87df909ef6e81e24c798d8fb75860abac8e193 🚀
Deploying to gh-pages from @ PoisotLab/ms_networkary_classifiers@a77780db763f722c3c0158980048f93335ff4df9 🚀

@@
 
+In fig.~\ref{fig:biasmccinf}, we present the response of MCC and
+informedness to (i) four levels of network connectance and (ii) a
+gradient of training set bias, for the four component models as well as
+the ensemble. All models reached a higher performance on more connected
+networks, and using more biased training sets (with the exception of
+ridge regression, whose informedness decreased in performance with
+training set bias). In all cases, informedness was extremely high, which
+is an expected consequence of the fact that this is the value we
+optimized to determine the cutoff. MCC increased with training set bias,
+although this increase became less steep with increasing connectance.
+Interestingly, the ensemble almost always outclassed its component
+models.
+
+\begin{figure}
+\hypertarget{fig:biasmccinf}{%
+\centering
+\includegraphics{figures/bias_mcc_inf.png}
+\caption{TODO}\label{fig:biasmccinf}
+}
+\end{figure}
+
+In fig.~\ref{fig:biasrocpr}, we present the same information as
+fig.~\ref{fig:biasmccinf}, this time using ROC-AUC and PR-AUC. ROC-AUC
+is always high, and does not vary with training set bias. On the other
+hand, PR-AUC shows very strong responses, increasing with training set
+bias. It is notable here that two classifiers that seemed to be
+performing well (Decision Tree and Random Forest) based on their MCC are
+not able to reach a high PR-AUC even at higher connectances. As in
+fig.~\ref{fig:biasmccinf}, the ensemble outperforms its component
+models.
+
+\begin{figure}
+\hypertarget{fig:biasrocpr}{%
+\centering
+\includegraphics{figures/bias_roc_pr.png}
+\caption{TODO}\label{fig:biasrocpr}
+}
+\end{figure}
+
+Based on the results presented in fig.~\ref{fig:biasmccinf} and
+fig.~\ref{fig:biasrocpr}, it seems that informedness and ROC-AUC are not
+necessarilly able to discriminate between good and bad classifiers
+(although this result may be an artifact for informedness, as it has
+been optimized when thresholding). On the other hand, MCC and PR-AUC
+show a strong response to training set bias, and may therefore be more
+useful at model comparison.
+
 \hypertarget{required-amount-of-positives-to-get-the-best-performance}{%
@@
 
-\hypertarget{guidelines-for-prediction}{%
-\section{Guidelines for prediction}\label{guidelines-for-prediction}}
+The previous results revealed that the measure of classification
+performance responds both to the bias in the training set \emph{and} to
+the connectance of the network; from a practical point of view,
+assembling a training set requires to withold positive information,
+which in ecological networks are very scarce (and typically more
+valuable than negatives, on which there is a doubt). For this reason,
+across all values of connectance, we measured the training set bias that
+maximized a series of performance measures. When this value is high, the
+training set needs to skew positive in order to get a good model; when
+this value is about 0.5, the training set needs to be artificially
+balanced to optimize the model performance. These results are presented
+in fig.~\ref{fig:optimbias}.
+
+\begin{figure}
+\hypertarget{fig:optimbias}{%
+\centering
+\includegraphics{figures/optim_mcc_inf.png}
+\caption{TODO}\label{fig:optimbias}
+}
+\end{figure}
+
+Interestingly, as long as the connectance of the network was above
+\(\approx 0.1\), the optimal prevalence in the training set is 0.5,
+\emph{i.e.} as many positives as negatives. Low connectance is usually
+achieved for very large networks, due to the scaling relationship
+between richness and links (MacDonald et al., 2020). Therefore, larger
+networks may require \emph{more} biasing of the training set in order to
+be optimally predicted, whereas smaller, more connected networks may
+not. It is worth noting that the optimal bias for the training set
+stabilizes at 0.5 regardless of connectance \emph{and} model \emph{and}
+measure of model evaluation. \# Guidelines for prediction
+
+INF because we trust positives more than negative, but check with MCC
+
+PR-AUC
+
+usually default to training set bias of 0.5
 
@@
 
+\leavevmode\hypertarget{ref-MacDonald2020RevLin}{}%
+MacDonald, A. A. M., Banville, F., \& Poisot, T. (2020). Revisiting the
+Links-Species Scaling Relationship in Food Webs. \emph{Patterns},
+\emph{1}(0). \url{https://doi.org/10.1016/j.patter.2020.100079}
+
 \leavevmode\hypertarget{ref-McLeod2021SamAsy}{}%

### 2021-12-11 code/03_figures.jl
moar figs

@@ -79,11 +79,22 @@ end
 
 results = leftjoin(results, select(connectance_values, [:runid, :conbin]), on=:runid)
 bmm = groupby(results, [:model, :measure, :conbin])
-opt = combine(bmm, [:value, :bias] => ((v,b) -> median(b[sortperm(v)[end-min(20, length(v))+1:end]])) => :optimalbias)
+function optival(v,b)
+    idx = sortperm(v)[end-min(20, length(v))+1:end]
+    return (bias=median(b[idx]), value=median(v[idx]))
+end
+opt = combine(bmm, [:value, :bias] => ((v,b) -> optival(v,b)) => AsTable)
 
 _keepval(f) = f in ["MCC", "INF", "PRAUC", "ROCAUC"]
 data(@subset(opt, _keepval.(:measure))) * 
-    mapping(:conbin, :optimalbias, col=:model, row=:measure) *
+    mapping(:conbin, :bias, col=:model, row=:measure) *
     (AlgebraOfGraphics.density() * visual(Contour, color=:grey, alpha=0.3) + smooth() * visual(linewidth=2.0)) |>
     plt -> draw(plt, facet=(;linkyaxes = :none)) |>
     plt -> save(joinpath(@__DIR__, "..", "figures", "optim_bias.png"), plt, px_per_unit = 3)
+
+_keepval(f) = f in ["MCC", "INF", "PRAUC", "ROCAUC"]
+data(@subset(opt, _keepval.(:measure))) * 
+    mapping(:conbin, :value, col=:model, row=:measure) *
+    (AlgebraOfGraphics.density() * visual(Contour, color=:grey, alpha=0.3) + smooth() * visual(linewidth=2.0)) |>
+    plt -> draw(plt, facet=(;linkyaxes = :none)) |>
+    plt -> save(joinpath(@__DIR__, "..", "figures", "optim_perf.png"), plt, px_per_unit = 3)

### 2021-12-11 code/02_simulations.jl
back to 20k simulations for a final run

@@ -70,7 +70,7 @@ candidate_models = [
 
 S = 200
 
-_n_sims = 500
+_n_sims = 20000
 conditions_breadth = rand(_n_sims) .* 0.1 .+ 0.005
 conditions_bias = rand(_n_sims) .* 0.98 .+ 0.01
 conditions = hcat(conditions_breadth, conditions_bias)

### 2021-12-11 code/03_figures.jl
update y axis linking

@@ -38,33 +38,35 @@ end
 # Join the two dataframes
 results = leftjoin(results, select(connectance_values, [:runid, :midpoint]), on=:runid)
 
+# axis = (xticks = LinearTicks(n),) ??
+
 # Dataviz
 _keepval(f) = f in ["MCC", "INF"]
 data(@subset(results, _keepval.(:measure))) *
     mapping(:bias => "Training set bias", :value => "Value", row=:midpoint => nonnumeric, col=:model, color=:measure) *
     (AlgebraOfGraphics.density() * visual(Contour, color=:grey, alpha=0.3) + smooth() * visual(linewidth=2.0)) |>
-    plt -> draw(plt, facet=(;linkyaxes = :none)) |>
+    plt -> draw(plt, facet=(;linkyaxes = :minimal)) |>
     plt -> save(joinpath(@__DIR__, "..", "figures", "bias_mcc_inf.png"), plt, px_per_unit = 3)
 
 _keepval(f) = f in ["PRAUC", "ROCAUC"]
 data(@subset(results, _keepval.(:measure))) *
     mapping(:bias => "Training set bias", :value => "Value", row=:midpoint => nonnumeric, col=:model, color=:measure) *
     (AlgebraOfGraphics.density() * visual(Contour, color=:grey, alpha=0.3) + smooth() * visual(linewidth=2.0)) |>
-    plt -> draw(plt, facet=(;linkyaxes = :none)) |>
+    plt -> draw(plt, facet=(;linkyaxes = :minimal)) |>
     plt -> save(joinpath(@__DIR__, "..", "figures", "bias_roc_pr.png"), plt, px_per_unit = 3)
 
 _keepval(f) = f in ["PT"]
 data(@subset(results, _keepval.(:measure))) *
     mapping(:bias => "Training set bias", :value => "Prevalence threshold", row=:midpoint => nonnumeric, col=:model) *
     (AlgebraOfGraphics.density() * visual(Contour, color=:grey, alpha=0.3) + smooth() * visual(linewidth=2.0)) |>
-    plt -> draw(plt, facet=(;linkyaxes = :none)) |>
+    plt -> draw(plt, facet=(;linkyaxes = :minimal)) |>
     plt -> save(joinpath(@__DIR__, "..", "figures", "bias_pt.png"), plt, px_per_unit = 3)
 
 _keepval(f) = f in ["postbias"]
[…]

### 2021-12-11 draft.tex
Deploying to gh-pages from @ PoisotLab/ms_networkary_classifiers@a34a2edbabf979682451f47ae08b31595d274fc2 🚀
Deploying to gh-pages from @ PoisotLab/ms_networkary_classifiers@f7197e7d0c68bba2b94787cc9bbb29e02090910c 🚀

@@
 \centering
-\includegraphics{figures/optim_mcc_inf.png}
+\includegraphics{figures/optim_bias.png}
 \caption{TODO}\label{fig:optimbias}
@@
 stabilizes at 0.5 regardless of connectance \emph{and} model \emph{and}
-measure of model evaluation. \# Guidelines for prediction
+measure of model evaluation.
 
-INF because we trust positives more than negative, but check with MCC
+\begin{figure}
+\hypertarget{fig:optimperf}{%
+\centering
+\includegraphics{figures/optim_perf.png}
+\caption{TODO}\label{fig:optimperf}
+}
+\end{figure}
 
-PR-AUC
+\hypertarget{guidelines-for-the-assesment-of-network-predictive-models}{%
+\section{Guidelines for the assesment of network predictive
+models}\label{guidelines-for-the-assesment-of-network-predictive-models}}
 
-usually default to training set bias of 0.5
+The results presented here highlight an interesting paradox: larger
+networks (with lower connectance) require more training set bias in
+order to maximize model performance fig.~\ref{fig:optimbias}, but are
+also more difficult to predict according to MCC and PR-AUC
+fig.~\ref{fig:optimperf}. This suggests that the task of network
+prediction will be difficult regardless of network size: by being
+limited by the \emph{frequency} of interactions when the network is
+large, and by being limited by the \emph{number} of interactions when
+the network is small. Nevertheless, based on the simulations and
+numerical experiments, it is possible to formulate a series of
+recommendations for the evaluation of network prediction models.
+
+First, because we should have more trust in reported interactions than
+in reported absences of interactions, we can draw on previous literature
+to recommend informedness as a measure to decide on a threshold (Chicco
+et al., 2021); this being said, because informedness is insensitive to
+bias, the model performance is better evaluated through the use of MCC
+fig.~\ref{fig:biasmccinf}. Because \(F_1\) is monotonously sensitive to
+classifier bias fig.~\ref{fig:bias} and network connectance
+fig.~\ref{fig:connectance}, MCC should be prefered as a measure of model
+evaluation.
+
+Second, because the PR-AUC responds more to network connectance
+fig.~\ref{fig:optimperf} and training set imbalance
+fig.~\ref{fig:biasrocpr}, it should be used as a measure of model
+performance over the ROC-AUC. This is not to say that ROC-AUC should be
+discarded (in fact, a low ROC-AUC is a sign of an issue with the model),
+but that its interpretation should be guided by the PR-AUC value. This
+again echoes recommendations from other fields (Saito \& Rehmsmeier,
+2015).
+
+Thirdly, regardless of network connectance \emph{or} measure to evaluate
+the model performance, as long as the network connectance is larger than
+\(\approx 0.1\), artificially balancing the training set to have
+equiprevalence will give the best possible results. This was true for
+all models.
+
+Finally, it is noteworthy that the ensemble model was systematically
+better than the component models; even when poor models were included
+(Random Forest and Decision Tree), the ensemble was able to leverage the
+different biases expressed by the models to make an overall more
+accurate prediction. We do not expect that ensembles will \emph{always}
+be better than single models. In a recent multi-model comparison, Becker
+et al. (2021) found that the ensemble was \emph{not} the best model.
+There is no general conclusion to draw from this besides reinforcing the
+need to be pragmatic about which models should be included in the
+ensemble, or whether to use an ensemble at all. In a sense, the
+surprising peformance of the ensemble model should form the basis of the
+last recommendation: optimal training set bias and its interaction with
+connectance and binary classifier is, in a sense, an hyperparameter that
+should be assessed. The distribution of results in
+fig.~\ref{fig:optimbias} and fig.~\ref{fig:optimperf} show that there
+are variations around the trend; furthermore, networks with different
+structures than the one we simulated here may respond in different ways.
 
@@
 1223--1232. \url{https://doi.org/10.1111/j.1365-2664.2006.01214.x}
+
+\leavevmode\hypertarget{ref-Becker2021OptPre}{}%
+Becker, D., Albery, G. F., Sjodin, A. R., Poisot, T., Bergner, L.,
+Dallas, T., Eskew, E. A., Farrell, M. J., Guth, S., Han, B. A., Simmons,
+N. B., Stock, M., Teeling, E. C., \& Carlson, C. J. (2021). Optimizing
+predictive models to prioritize viral discovery in zoonotic reservoirs.
+\emph{bioRxiv}, 2020.05.22.111344.
+\url{https://doi.org/10.1101/2020.05.22.111344}
 

### 2021-12-11 code/03_figures.jl
change to figures / need text update

@@ -72,7 +72,7 @@ data(@subset(results, _keepval.(:measure))) *
 # Make bins for connectance to get the optimal bias
 connectance_values = unique(select(results, [:runid, :connectance]))
 connectance_values.conbin = zeros(Float64, size(connectance_values,1))
-n_connectance_bins = 100
+n_connectance_bins = 200
 bins_connectance = LinRange(0.0, 0.2, n_connectance_bins+1)
 for i in 1:n_connectance_bins
     _idx = findall(bins_connectance[i] .<= connectance_values.connectance .< bins_connectance[i+1])
@@ -82,8 +82,8 @@ end
 results = leftjoin(results, select(connectance_values, [:runid, :conbin]), on=:runid)
 bmm = groupby(results, [:model, :measure, :conbin])
 function optival(v,b)
-    idx = sortperm(v)[end-min(20, length(v))+1:end]
-    return (bias=median(b[idx]), value=median(v[idx]))
+    idx = sortperm(v)[end-min(50, length(v))+1:end]
+    return (bias=mean(b[idx]), value=mean(v[idx]))
 end
 opt = combine(bmm, [:value, :bias] => ((v,b) -> optival(v,b)) => AsTable)
 

### 2021-12-11 code/03_figures.jl
change figures

@@ -72,7 +72,7 @@ data(@subset(results, _keepval.(:measure))) *
 # Make bins for connectance to get the optimal bias
 connectance_values = unique(select(results, [:runid, :connectance]))
 connectance_values.conbin = zeros(Float64, size(connectance_values,1))
-n_connectance_bins = 200
+n_connectance_bins = 150
 bins_connectance = LinRange(0.0, 0.2, n_connectance_bins+1)
 for i in 1:n_connectance_bins
     _idx = findall(bins_connectance[i] .<= connectance_values.connectance .< bins_connectance[i+1])
@@ -91,12 +91,12 @@ _keepval(f) = f in ["MCC", "INF", "PRAUC", "ROCAUC"]
 data(@subset(opt, _keepval.(:measure))) * 
     mapping(:conbin, :bias, col=:model, row=:measure) *
     (AlgebraOfGraphics.density() * visual(Contour, color=:grey, alpha=0.3) + smooth() * visual(linewidth=2.0)) |>
-    plt -> draw(plt, facet=(;linkyaxes = :minimal)) |>
+    plt -> draw(plt, facet=(;linkyaxes = :minimal), axis = (xticks = LinearTicks(3),)) |>
     plt -> save(joinpath(@__DIR__, "..", "figures", "optim_bias.png"), plt, px_per_unit = 3)
 
 _keepval(f) = f in ["MCC", "INF", "PRAUC", "ROCAUC"]
 data(@subset(opt, _keepval.(:measure))) * 
     mapping(:conbin, :value, col=:model, row=:measure) *
     (AlgebraOfGraphics.density() * visual(Contour, color=:grey, alpha=0.3) + smooth() * visual(linewidth=2.0)) |>
-    plt -> draw(plt, facet=(;linkyaxes = :minimal)) |>
+    plt -> draw(plt, facet=(;linkyaxes = :minimal), axis = (xticks = LinearTicks(3),)) |>
     plt -> save(joinpath(@__DIR__, "..", "figures", "optim_perf.png"), plt, px_per_unit = 3)

### 2021-12-11 code/02_simulations.jl
attempt with more traits

@@ -7,6 +7,7 @@ using Statistics
 using MLJ
 using DataFrames
 using CSV: CSV
+using Distributions
 
 # AUC
 function ∫(x::Array{T}, y::Array{T}) where {T<:Number}
@@ -21,24 +22,35 @@ end
 include(joinpath(@__DIR__, "confusionmatrix.jl"))
 
 # Network generation function
-function network(S, breadth)
-    # Get the network
-    x = sort(rand(S))
-    y = sort(rand(S))
-    𝐱 = repeat(x; inner=length(y))
-    𝐲 = repeat(y; outer=length(x))
-    𝐳 =
-        clamp.(𝐱 .- 0.8 .* 𝐱 .- breadth, 0, 1) .<=
-        𝐲 .<=
-        clamp.(𝐱 .- 0.8 .* 𝐱 .+ breadth, 0, 1)
-
-    # Features / labels matrix
-    return (table(hcat(𝐱, 𝐲)), Float64.(𝐳))
+function nichemodel(; C=0.2, S=100)
+    # Beta distribution parameter
+    β = 1.0 / (2.0 * C) - 1.0
+    # Generate body size
+    n = sort(rand(Uniform(0.0, 1.0), S))
+    # Generate random ranges
+    r = n .* rand(Beta(1.0, β), S)
+    # Generate random centroids
+    c = [rand(Uniform(r[i] / 2, n[i])) for i in 1:S]
+    # Smallest species is an obligate producer
+    n[1] = c[1] = 0.0
+    # Fill
+    predictors = zeros(Float64, (S * S, 6))
+    response = zeros(Float64, S * S)
+    for i in 1:S
+        for j in 1:S
+            idx = S * (i - 1) + j
+            predictors[idx, :] = [n[i], r[i], c[i], n[j], r[j], c[j]]
+            response[idx] = c[i] - r[i] / 2 <= n[j] <= c[i] + r[j] / 2
+        end
+    end
+    # Return
+    return table(predictors), response
 end
+
 
 # Prepare the results
 results = [DataFrame(;
-    breadth=Float64[],
+    co=Float64[],
     bias=Float64[],
     connectance=Float64[],
     model=Symbol[],
@@ -47,13 +59,15 @@ results = [DataFrame(;
 ) for _thr in 1:Threads.nthreads()]
 
 # Pick possible models based on a network
-_mock = network(10, 0.2)
+#=
+_mock = nichemodel(S=10, C=0.2)
[…]

### 2021-12-11 code/nichemodel.jl

@@ -1,0 +1,26 @@
+using Distributions
+
+function nichemodel(; C=0.2, S=100)
+    # Beta distribution parameter
+    β = 1.0 / (2.0 * C) - 1.0
+    # Generate body size
+    n = sort(rand(Uniform(0.0, 1.0), S))
+    # Generate random ranges
+    r = n .* rand(Beta(1.0, β), S)
+    # Generate random centroids
+    c = [rand(Uniform(r[i] / 2, n[i])) for i in 1:S]
+    # Smallest species is an obligate producer
+    n[1] = c[1] = 0.0
+    # Fill
+    predictors = zeros(Float64, (S * S, 6))
+    response = zeros(Float64, S * S)
+    for i in 1:S
+        for j in 1:S
+            idx = S * (i - 1) + j
+            predictors[idx, :] = [n[i], r[i], c[i], n[j], r[j], c[j]]
+            response[idx] = c[i] - r[i] / 2 <= n[j] <= c[i] + r[j] / 2
+        end
+    end
+    # Return
+    return table(predictors), response
+end

### 2021-12-11 code/02_simulations.jl
use niche model

@@ -83,8 +83,8 @@ candidate_models = [
 ]
 
 S = 100
-_n_sims = 500
-conditions_co = rand(_n_sims) .* 0.25 .+ 0.01
+_n_sims = 20000
+conditions_co = rand(_n_sims) .* 0.26 .+ 0.001
 conditions_bias = rand(_n_sims) .* 0.98 .+ 0.01
 conditions = hcat(conditions_co, conditions_bias)
 

### 2021-12-11 code/03_figures.jl

@@ -11,7 +11,7 @@ replace!(results.value, NaN => missing)
 dropmissing!(results)
 
 # Drop the models with obviously borked classifiers
-results.runid = hash.(results.connectance .* results.breadth .* results.bias)
+results.runid = hash.(results.connectance .* results.co .* results.bias)
 
 todrop = unique(vcat([
     @subset(results, :measure .== "ACC", :value .== 0.0).runid,
@@ -37,8 +37,6 @@ end
 
 # Join the two dataframes
 results = leftjoin(results, select(connectance_values, [:runid, :midpoint]), on=:runid)
-
-# axis = (xticks = LinearTicks(n),) ??
 
 # Dataviz
 _keepval(f) = f in ["MCC", "INF"]
@@ -72,7 +70,7 @@ data(@subset(results, _keepval.(:measure))) *
 # Make bins for connectance to get the optimal bias
 connectance_values = unique(select(results, [:runid, :connectance]))
 connectance_values.conbin = zeros(Float64, size(connectance_values,1))
-n_connectance_bins = 150
+n_connectance_bins = 100
 bins_connectance = LinRange(0.0, 0.2, n_connectance_bins+1)
 for i in 1:n_connectance_bins
     _idx = findall(bins_connectance[i] .<= connectance_values.connectance .< bins_connectance[i+1])

### 2021-12-11 code/nichemodel.jl

@@ -1,26 +1,0 @@
-using Distributions
-
-function nichemodel(; C=0.2, S=100)
-    # Beta distribution parameter
-    β = 1.0 / (2.0 * C) - 1.0
-    # Generate body size
-    n = sort(rand(Uniform(0.0, 1.0), S))
-    # Generate random ranges
-    r = n .* rand(Beta(1.0, β), S)
-    # Generate random centroids
-    c = [rand(Uniform(r[i] / 2, n[i])) for i in 1:S]
-    # Smallest species is an obligate producer
-    n[1] = c[1] = 0.0
-    # Fill
-    predictors = zeros(Float64, (S * S, 6))
-    response = zeros(Float64, S * S)
-    for i in 1:S
-        for j in 1:S
-            idx = S * (i - 1) + j
-            predictors[idx, :] = [n[i], r[i], c[i], n[j], r[j], c[j]]
-            response[idx] = c[i] - r[i] / 2 <= n[j] <= c[i] + r[j] / 2
-        end
-    end
-    # Return
-    return table(predictors), response
-end

### 2021-12-11 code/03_figures.jl
up

@@ -19,7 +19,7 @@ todrop = unique(vcat([
     @subset(results, :measure .== "TNR", :value .== 0.0).runid,
     @subset(results, :measure .== "FPR", :value .== 1.0).runid,
     @subset(results, :measure .== "FNR", :value .== 1.0).runid,
-    @subset(results, :connectance .>= 0.2).runid
+    @subset(results, :connectance .>= 0.3).runid
 ]...))
 
 tokeep(f) = !(f in todrop)
@@ -28,8 +28,8 @@ tokeep(f) = !(f in todrop)
 # Make bins for connectance to make the plotting more efficient
 connectance_values = unique(select(results, [:runid, :connectance]))
 connectance_values.midpoint = zeros(Float64, size(connectance_values,1))
-n_connectance_bins = 4
-bins_connectance = LinRange(0.0, 0.2, n_connectance_bins+1)
+n_connectance_bins = 6
+bins_connectance = LinRange(0.0, 0.3, n_connectance_bins+1)
 for i in 1:n_connectance_bins
     _idx = findall(bins_connectance[i] .<= connectance_values.connectance .< bins_connectance[i+1])
     connectance_values.midpoint[_idx] .= round((bins_connectance[i+1] + bins_connectance[i])/2.0; digits=4)

### 2021-12-11 code/03_figures.jl
steps

@@ -71,7 +71,7 @@ data(@subset(results, _keepval.(:measure))) *
 connectance_values = unique(select(results, [:runid, :connectance]))
 connectance_values.conbin = zeros(Float64, size(connectance_values,1))
 n_connectance_bins = 100
-bins_connectance = LinRange(0.0, 0.2, n_connectance_bins+1)
+bins_connectance = LinRange(0.0, 0.3, n_connectance_bins+1)
 for i in 1:n_connectance_bins
     _idx = findall(bins_connectance[i] .<= connectance_values.connectance .< bins_connectance[i+1])
     connectance_values.conbin[_idx] .= (bins_connectance[i+1] + bins_connectance[i])/2.0

### 2021-12-12 code/02_simulations.jl
bipartite model with traits

@@ -22,52 +22,30 @@ end
 include(joinpath(@__DIR__, "confusionmatrix.jl"))
 
 # Network generation function
-function nichemodel(; C=0.2, S=100)
-    # Beta distribution parameter
-    β = 1.0 / (2.0 * C) - 1.0
-    # Generate body size
-    n = sort(rand(Uniform(0.0, 1.0), S))
-    # Generate random ranges
-    r = n .* rand(Beta(1.0, β), S)
-    # Generate random centroids
-    c = [rand(Uniform(r[i] / 2, n[i])) for i in 1:S]
-    # Smallest species is an obligate producer
-    n[1] = c[1] = 0.0
-    # Fill
-    predictors = zeros(Float64, (S * S, 6))
-    response = zeros(Float64, S * S)
-    for i in 1:S
-        for j in 1:S
-            idx = S * (i - 1) + j
-            predictors[idx, :] = [n[i], r[i], c[i], n[j], r[j], c[j]]
-            response[idx] = c[i] - r[i] / 2 <= n[j] <= c[i] + r[j] / 2
-        end
-    end
-    # Return
-    return table(predictors), response
+L(x::T, y::T; r::T=0.1) where {T<:Number} = (x-r/2.0) ≤ y ≤ (x+r/2.0)  ? one(T) : zero(T)
+function network(S, ξ)
+    infectivity = Beta(6.0, 8.0)
+    resistance = Beta(2.0, 8.0)
+    𝐱ᵥ = sort(rand(infectivity, S))
+    𝐱ₕ = sort(rand(resistance, S))
+    𝐱₁ = repeat(𝐱ᵥ; outer=length(𝐱ₕ))
+    𝐱₂ = repeat(𝐱ₕ; inner=length(𝐱ᵥ))
+    𝐱₃ = abs.(𝐱₁ .- 𝐱₂)
+    𝐲 = [L(𝐱₁[i], 𝐱₂[i]; r=ξ) for i in 1:(S*S)]
+    𝐱 = table(hcat(𝐱₁, 𝐱₂, 𝐱₃))
+    return (𝐱, 𝐲)
 end
 
 
 # Prepare the results
 results = [DataFrame(;
-    co=Float64[],
+    breadth=Float64[],
     bias=Float64[],
     connectance=Float64[],
     model=Symbol[],
     measure=Symbol[],
     value=Float64[],
 ) for _thr in 1:Threads.nthreads()]
-
-# Pick possible models based on a network
[…]

### 2021-12-12 code/02_simulations.jl
run on narval, draft figures for writing

@@ -32,7 +32,8 @@ function network(S, ξ)
     𝐱₂ = repeat(𝐱ₕ; inner=length(𝐱ᵥ))
     𝐱₃ = abs.(𝐱₁ .- 𝐱₂)
     𝐲 = [L(𝐱₁[i], 𝐱₂[i]; r=ξ) for i in 1:(S*S)]
-    𝐱 = table(hcat(𝐱₁, 𝐱₂, 𝐱₃))
+    #𝐱 = table(hcat(𝐱₁, 𝐱₂, 𝐱₃))
+    𝐱 = table(hcat(𝐱₁, 𝐱₂))
     return (𝐱, 𝐲)
 end
 
@@ -61,8 +62,8 @@ candidate_models = [
 ]
 
 S = 100
-_n_sims = 300
-conditions_breadth = rand(_n_sims) .* 0.22 .+ 0.01
+_n_sims = 2000
+conditions_breadth = rand(_n_sims) .* 0.4 .+ 0.05
 conditions_bias = rand(_n_sims) .* 0.98 .+ 0.01
 conditions = hcat(conditions_breadth, conditions_bias)
 

### 2021-12-12 code/03_figures.jl

@@ -11,7 +11,7 @@ replace!(results.value, NaN => missing)
 dropmissing!(results)
 
 # Drop the models with obviously borked classifiers
-results.runid = hash.(results.connectance .* results.co .* results.bias)
+results.runid = hash.(results.connectance .* results.breadth .* results.bias)
 
 todrop = unique(vcat([
     @subset(results, :measure .== "ACC", :value .== 0.0).runid,
@@ -19,7 +19,9 @@ todrop = unique(vcat([
     @subset(results, :measure .== "TNR", :value .== 0.0).runid,
     @subset(results, :measure .== "FPR", :value .== 1.0).runid,
     @subset(results, :measure .== "FNR", :value .== 1.0).runid,
-    @subset(results, :connectance .>= 0.3).runid
+    @subset(results, :measure .== "PRAUC", :value .<= 0.0).runid,
+    @subset(results, :connectance .>= 0.45).runid,
+    @subset(results, :connectance .<= 0.05).runid
 ]...))
 
 tokeep(f) = !(f in todrop)
@@ -28,8 +30,8 @@ tokeep(f) = !(f in todrop)
 # Make bins for connectance to make the plotting more efficient
 connectance_values = unique(select(results, [:runid, :connectance]))
 connectance_values.midpoint = zeros(Float64, size(connectance_values,1))
-n_connectance_bins = 6
-bins_connectance = LinRange(0.0, 0.3, n_connectance_bins+1)
+n_connectance_bins = 4
+bins_connectance = LinRange(0.05, 0.45, n_connectance_bins+1)
 for i in 1:n_connectance_bins
     _idx = findall(bins_connectance[i] .<= connectance_values.connectance .< bins_connectance[i+1])
     connectance_values.midpoint[_idx] .= round((bins_connectance[i+1] + bins_connectance[i])/2.0; digits=4)
@@ -62,7 +64,7 @@ data(@subset(results, _keepval.(:measure))) *
 
 _keepval(f) = f in ["postbias"]
 data(@subset(results, _keepval.(:measure))) *
-    mapping(:bias => "Training set bias", :value => logistic => "Value", row=:midpoint => nonnumeric, col=:model) *
+    mapping(:bias => "Training set bias", :value => (v -> logistic(v-1)) => "Value", row=:midpoint => nonnumeric, col=:model) *
[…]

### 2021-12-12 draft.tex
Deploying to gh-pages from @ PoisotLab/ms_networkary_classifiers@9c57188839a677c080e4176c35e8c6b2dc8c20eb 🚀
Deploying to gh-pages from @ PoisotLab/ms_networkary_classifiers@9193efcaedf173b7998a227d8497953cfac19103 🚀

@@
 \section{Baseline values}\label{baseline-values}}
+
+Intro
 
@@
 
+\textbf{Acknowledgements:} We acknowledge that this study was conducted
+on land within the traditional unceded territory of the Saint Lawrence
+Iroquoian, Anishinabewaki, Mohawk, Huron-Wendat, and Omàmiwininiwak
+nations. This research was enabled in part by support provided by Calcul
+Québec (www.calculquebec.ca) and Compute Canada (www.computecanada.ca).
+
 \hypertarget{references}{%

### 2021-12-12 draft.tex
Deploying to gh-pages from @ PoisotLab/ms_networkary_classifiers@00313593cd843082ea8043d1bff842745e049e2d 🚀

@@
 
-The dataset used for numerical experiments is composed of 20000 such
+The dataset used for numerical experiments is composed of 64000 such
 \((\xi, \nu)\) pairs, on which four learners are trained: a decision
@@
 \([0,1]\)), then put through the same thresholding process. The complete
-code to run the simulations is given as an appendix.
+code to run the simulations is given as an appendix; running the final
+simulation required 4.8 core days (approx. 117 hours).
 
@@
 nations. This research was enabled in part by support provided by Calcul
-Québec (www.calculquebec.ca) and Compute Canada (www.computecanada.ca).
+Québec (www.calculquebec.ca) and Compute Canada (www.computecanada.ca)
+through the Narval general purpose cluster. TP is supported by a NSERC
+Discovery Grant and Discovery Acceleration Supplement, and by a grant
+from the Institut de Valorisation des Données (IVADO).
 

### 2021-12-12 draft.tex
Deploying to gh-pages from @ PoisotLab/ms_networkary_classifiers@7cce168731bec98eb0897cbd064f425372221f1e 🚀

@@
 
+\hypertarget{illustration-on-empirical-food-webs}{%
+\section{Illustration on empirical food
+webs}\label{illustration-on-empirical-food-webs}}
+
+In this section, we use a collection of XX empirical networks
+
+\hypertarget{using-t-sbn-and-rdpg-to-simulate-arbitrarily-imperfect-classifier}{%
+\subsection{Using t-SBN and RDPG to simulate arbitrarily imperfect
+classifier}\label{using-t-sbn-and-rdpg-to-simulate-arbitrarily-imperfect-classifier}}
+
+use t-SVD to only use some information
+
+cumulative eigenvalues up to rank n give amount of information recovered
+
+can be thresholded
+
+how do values react to using fewer ranks
+
+\hypertarget{effect-of-classifier-performance-on-metrics}{%
+\subsection{Effect of classifier performance on
+metrics}\label{effect-of-classifier-performance-on-metrics}}
+
+\ldots{}
+
 \hypertarget{numerical-experiments}{%

### 2021-12-12 code/03_figures.jl
update the figures with 64k runs

@@ -4,6 +4,7 @@ using CSV: CSV
 using DataFramesMeta
 using StatsBase
 using Statistics
+using StatsFuns
 
 results = DataFrame(CSV.File(joinpath(@__DIR__, "output.csv")))
 allowmissing!(results, :value)
@@ -20,7 +21,7 @@ todrop = unique(vcat([
     @subset(results, :measure .== "FPR", :value .== 1.0).runid,
     @subset(results, :measure .== "FNR", :value .== 1.0).runid,
     @subset(results, :measure .== "PRAUC", :value .<= 0.0).runid,
-    @subset(results, :connectance .>= 0.45).runid,
+    @subset(results, :connectance .>= 0.35).runid,
     @subset(results, :connectance .<= 0.05).runid
 ]...))
 
@@ -30,8 +31,8 @@ tokeep(f) = !(f in todrop)
 # Make bins for connectance to make the plotting more efficient
 connectance_values = unique(select(results, [:runid, :connectance]))
 connectance_values.midpoint = zeros(Float64, size(connectance_values,1))
-n_connectance_bins = 4
-bins_connectance = LinRange(0.05, 0.45, n_connectance_bins+1)
+n_connectance_bins = 5
+bins_connectance = LinRange(0.05, 0.35, n_connectance_bins+1)
 for i in 1:n_connectance_bins
     _idx = findall(bins_connectance[i] .<= connectance_values.connectance .< bins_connectance[i+1])
     connectance_values.midpoint[_idx] .= round((bins_connectance[i+1] + bins_connectance[i])/2.0; digits=4)
@@ -73,7 +74,7 @@ data(@subset(results, _keepval.(:measure))) *
 connectance_values = unique(select(results, [:runid, :connectance]))
 connectance_values.conbin = zeros(Float64, size(connectance_values,1))
 n_connectance_bins = 100
-bins_connectance = LinRange(0.05, 0.45, n_connectance_bins+1)
+bins_connectance = LinRange(0.05, 0.35, n_connectance_bins+1)
 for i in 1:n_connectance_bins
     _idx = findall(bins_connectance[i] .<= connectance_values.connectance .< bins_connectance[i+1])
     connectance_values.conbin[_idx] .= (bins_connectance[i+1] + bins_connectance[i])/2.0

### 2021-12-12 code/03_figures.jl
fig update

@@ -21,7 +21,7 @@ todrop = unique(vcat([
     @subset(results, :measure .== "FPR", :value .== 1.0).runid,
     @subset(results, :measure .== "FNR", :value .== 1.0).runid,
     @subset(results, :measure .== "PRAUC", :value .<= 0.0).runid,
-    @subset(results, :connectance .>= 0.35).runid,
+    @subset(results, :connectance .>= 0.25).runid,
     @subset(results, :connectance .<= 0.05).runid
 ]...))
 
@@ -32,7 +32,7 @@ tokeep(f) = !(f in todrop)
 connectance_values = unique(select(results, [:runid, :connectance]))
 connectance_values.midpoint = zeros(Float64, size(connectance_values,1))
 n_connectance_bins = 5
-bins_connectance = LinRange(0.05, 0.35, n_connectance_bins+1)
+bins_connectance = LinRange(0.05, 0.25, n_connectance_bins+1)
 for i in 1:n_connectance_bins
     _idx = findall(bins_connectance[i] .<= connectance_values.connectance .< bins_connectance[i+1])
     connectance_values.midpoint[_idx] .= round((bins_connectance[i+1] + bins_connectance[i])/2.0; digits=4)
@@ -74,7 +74,7 @@ data(@subset(results, _keepval.(:measure))) *
 connectance_values = unique(select(results, [:runid, :connectance]))
 connectance_values.conbin = zeros(Float64, size(connectance_values,1))
 n_connectance_bins = 100
-bins_connectance = LinRange(0.05, 0.35, n_connectance_bins+1)
+bins_connectance = LinRange(0.05, 0.4, n_connectance_bins+1)
 for i in 1:n_connectance_bins
     _idx = findall(bins_connectance[i] .<= connectance_values.connectance .< bins_connectance[i+1])
     connectance_values.conbin[_idx] .= (bins_connectance[i+1] + bins_connectance[i])/2.0

### 2021-12-12 code/04_rdpg.jl
SVD experiment

@@ -1,0 +1,68 @@
+using EcologicalNetworks
+using AlgebraOfGraphics, CairoMakie
+using DataFrames
+using CSV: CSV
+using DataFramesMeta
+using LinearAlgebra
+using Statistics
+
+N = simplify.(nz_stream_foodweb())
+
+# AUC
+function ∫(x::Array{T}, y::Array{T}) where {T<:Number}
+    S = zero(Float64)
+    for i in 2:length(x)
+        S += (x[i] - x[i - 1]) * (y[i] + y[i - 1]) * 0.5
+    end
+    return .-S
+end
+
+# Confusion matrix utilities
+include(joinpath(@__DIR__, "confusionmatrix.jl"))
+
+
+results = DataFrame(network=Int64[], connectance=Float64[], rank=Int64[], varexpl=Float64[], measure=Symbol[], value=Float64[])
+
+for n in 1:length(N)
+    for r in 1:45
+        _svd = svd(N[n])
+        varexpl = sum(_svd.S[1:r])/sum(_svd.S)
+        prediction = prod(rdpg(N[n], r))
+        thresholds = LinRange(minimum(vec(prediction)), maximum(vec(prediction)), 200)
+        M = Vector{ConfusionMatrix}(undef, length(thresholds))
+        binobs = adjacency(N[n])
+        for (i, τ) in enumerate(thresholds)
+            binpred = prediction .>= τ
+            tp = sum(binobs .& binpred)
+            tn = sum(.!binobs .& .!binpred)
+            fp = sum(.!binobs .& binpred)
+            fn = sum(binobs .& .!binpred)
+            M[i] = ConfusionMatrix(tp, tn, fp, fn)
+        end
+        X = UnipartiteNetwork(prediction .>= thresholds[last(findmax(informedness.(M)))])
+        ROCAUC = ∫(fpr.(M), tpr.(M))
+        AUPRC = ∫(tpr.(M), ppv.(M))
+        𝐌 = M[last(findmax(informedness.(M)))]
+        push!(results, (n, connectance(N[n]), r, varexpl, :ROCAUC, ROCAUC))
+        push!(results, (n, connectance(N[n]), r, varexpl, :PRAUC, AUPRC))
+        push!(results, (n, connectance(N[n]), r, varexpl, :INF, informedness(𝐌)))
+        push!(results, (n, connectance(N[n]), r, varexpl, :MCC, mcc(𝐌)))
[…]

### 2021-12-12 code/03_figures.jl
add a figure

@@ -44,14 +44,14 @@ results = leftjoin(results, select(connectance_values, [:runid, :midpoint]), on=
 # Dataviz
 _keepval(f) = f in ["MCC", "INF"]
 data(@subset(results, _keepval.(:measure))) *
-    mapping(:bias => "Training set bias", :value => "Value", row=:midpoint => nonnumeric, col=:model, color=:measure) *
+    mapping(:bias => "Training set bias", :value => "Value", row=:midpoint => nonnumeric, col=:model, color=:measure => "Measure") *
     (AlgebraOfGraphics.density() * visual(Contour, color=:grey, alpha=0.3) + smooth() * visual(linewidth=2.0)) |>
     plt -> draw(plt, facet=(;linkyaxes = :minimal)) |>
     plt -> save(joinpath(@__DIR__, "..", "figures", "bias_mcc_inf.png"), plt, px_per_unit = 3)
 
 _keepval(f) = f in ["PRAUC", "ROCAUC"]
 data(@subset(results, _keepval.(:measure))) *
-    mapping(:bias => "Training set bias", :value => "Value", row=:midpoint => nonnumeric, col=:model, color=:measure) *
+    mapping(:bias => "Training set bias", :value => "Value", row=:midpoint => nonnumeric, col=:model, color=:measure => "Measure") *
     (AlgebraOfGraphics.density() * visual(Contour, color=:grey, alpha=0.3) + smooth() * visual(linewidth=2.0)) |>
     plt -> draw(plt, facet=(;linkyaxes = :minimal)) |>
     plt -> save(joinpath(@__DIR__, "..", "figures", "bias_roc_pr.png"), plt, px_per_unit = 3)
@@ -65,7 +65,7 @@ data(@subset(results, _keepval.(:measure))) *
 
 _keepval(f) = f in ["postbias"]
 data(@subset(results, _keepval.(:measure))) *
-    mapping(:bias => "Training set bias", :value => (v -> logistic(v-1)) => "Value", row=:midpoint => nonnumeric, col=:model) *
+    mapping(:bias => "Training set bias", :value => (v -> logistic(v-1)) => "Relative connectance", row=:midpoint => nonnumeric, col=:model) *
     (AlgebraOfGraphics.density() * visual(Contour, color=:grey, alpha=0.3) + smooth() * visual(linewidth=2.0)) |>
[…]

### 2021-12-12 draft.tex
Deploying to gh-pages from @ PoisotLab/ms_networkary_classifiers@7cc292d1076ac679f470425645eb6e4e1eaaac37 🚀

@@
 \includegraphics{figures/changing-connectance.png}
-\caption{TODO}\label{fig:connectance}
+\caption{As in fig.~\ref{fig:bias}, consequences of changing connectance
+for different levels of classifier skill, assuming no classifier bias.
+Informedness, \(\kappa\), and MCC do increase with connectance, but only
+when the classifier is not no-skill; by way of contrast, a more
+connected network will give a higher \(F_1\) value even with a no-skill
+classifier.}\label{fig:connectance}
 }
@@
 for data-inflation (\emph{i.e.} the prediction of new interactions), it
-is not necessarilly a bad thing to select models that predict more
-interactions than the original dataset, because the original dataset
-misses some interactions.
+is not necessarilly a bad thing in practice to select models that
+predict more interactions than the original dataset, because the
+original dataset misses some interactions. Furthermore, the weight of
+positive interactions could be adjusted if some information about the
+extent of undersampling exists (\emph{e.g.} Branco et al., 2015).
 

### 2021-12-12 draft.tex
Deploying to gh-pages from @ PoisotLab/ms_networkary_classifiers@5e6aba817564f072386d39b379e62faa00c327eb 🚀

@@
 \item
-  Strydom et al. (2021) importance of drawing on traits + validation is
-  challenging + comparing across space
+  Strydom, Catchen, et al. (2021) importance of drawing on traits +
+  validation is challenging + comparing across space
 \end{itemize}
@@
 MCC
+
+ROC-AUC
+
+PR-AUC
 
@@
 
-In this section, we use a collection of XX empirical networks
-
-\hypertarget{using-t-sbn-and-rdpg-to-simulate-arbitrarily-imperfect-classifier}{%
-\subsection{Using t-SBN and RDPG to simulate arbitrarily imperfect
-classifier}\label{using-t-sbn-and-rdpg-to-simulate-arbitrarily-imperfect-classifier}}
-
-use t-SVD to only use some information
-
-cumulative eigenvalues up to rank n give amount of information recovered
-
-can be thresholded
-
-how do values react to using fewer ranks
+In this section, we use a collection of 22 empirical networks (Thompson
+\& Townsend, 2004; Thompson \& Townsend, 1999) from streams in NEw
+Zealand; these networks were previously analyzed by Pomeranz et al.
+(2018), who found that a variety of rule-based models tended to increase
+the number of interactions. Our goal here is not to infer which
+interactions are missing, but rather to estimate (i) how much precise
+knowledge of the network influences the measures of classification
+performance, and (ii) whether an increase in performance also results in
+more ecologically realistic networks. In order to do so, we use a
+trucated Singular Value Decomposition (t-SVD) coupled to a Random Dot
+Product Graph (RDPG; \emph{e.g.} Dalla Riva \& Stouffer, 2015) at
+increasingly high dimensions to reconstruct an approximation of the
+network. Recent work by Poisot et al. (2021) and Strydom, Bouskila, et
+al. (2021) show that t-SVD and RDPG are able to infer biotic
+interactions in, respectively, host-virus and predator-prey systems. The
+t-SVD/RDPG approach is covered in detail by Strydom, Bouskila, et al.
+(2021); by approximating the network at different ranks, where rank 1 is
+the coarsest, we can generate a matrix that can be thresholded and
+compared to the original network. The variance explained at each rank is
+given by the relative sum of the eigenvalues of the SVD up to the chosen
+rank, and measures how much information about the network we can capture
+(Strydom, Dalla Riva, et al., 2021).
 
@@
 
-\ldots{}
+By changing the rank of approximation from 1 (variance explained is
+minimized) to 45 (variance explained is close to 1), we show in
+fig.~\ref{fig:svdperf} that although all measures of classification
+performance increase the more we know about the network, MCC is
+initially lower compared to informedness; similarly, ROC-AUC is
+consistently higher compared to PR-AUC.
+
+\begin{figure}
+\hypertarget{fig:svdperf}{%
+\centering
+\includegraphics{figures/svd_perf.png}
+\caption{TODO}\label{fig:svdperf}
+}
+\end{figure}
+
+\hypertarget{effect-of-classifier-performance-on-network-properties}{%
+\subsection{Effect of classifier performance on network
+properties}\label{effect-of-classifier-performance-on-network-properties}}
+
+\begin{figure}
+\centering
+\includegraphics{figures/svd_ecol.png}
+\caption{TODO}
+\end{figure}
+
+connectance, spectral radius (Staniczenko et al., 2013), entropy (Stock
+et al., 2021)
 
@@
 
+\leavevmode\hypertarget{ref-DallaRiva2015ExpEvo}{}%
+Dalla Riva, G. V., \& Stouffer, D. B. (2015). Exploring the evolutionary
+signature of food webs' backbones using functional traits. \emph{Oikos},
+\emph{125}(4), 446--456. \url{https://doi.org/10.1111/oik.02305}
+
 \leavevmode\hypertarget{ref-Delgado2019WhyCoh}{}%
@@
 
+\leavevmode\hypertarget{ref-Poisot2021ImpMam}{}%
+Poisot, T., Ouellet, M.-A., Mollentze, N., Farrell, M. J., Becker, D.
+J., Albery, G. F., Gibb, R. J., Seifert, S. N., \& Carlson, C. J. (2021,
+May 31). \emph{Imputing the mammalian virome with linear filtering and
+singular value decomposition}. \url{http://arxiv.org/abs/2105.14973}
+
+\leavevmode\hypertarget{ref-Pomeranz2018InfPre}{}%
+Pomeranz, J. P., Thompson, R. M., Poisot, T., \& Harding, J. S. (2018).
+Inferring predator-prey interactions in food webs. \emph{Methods in
+Ecology and Evolution}, \emph{0}.
+\url{https://doi.org/10.1111/2041-210X.13125}
+
 \leavevmode\hypertarget{ref-Saito2015PrePlo}{}%
@@
 
+\leavevmode\hypertarget{ref-Staniczenko2013GhoNes}{}%
+Staniczenko, P. P. A., Kopp, J. C., \& Allesina, S. (2013). The ghost of
+nestedness in ecological networks. \emph{Nature Communications},
+\emph{4}(1), 1391. \url{https://doi.org/10.1038/ncomms2422}
+
+\leavevmode\hypertarget{ref-Stock2021DisInf}{}%
+Stock, M., Hoebeke, L., \& De Baets, B. (2021). Disentangling the
+Information in Species Interaction Networks. \emph{Entropy},
+\emph{23}(6, 6), 703. \url{https://doi.org/10.3390/e23060703}
+
+\leavevmode\hypertarget{ref-Strydom2021FooWeb}{}%
+Strydom, T., Bouskila, S., Banville, F., Barros, C., Caron, D., Farrell,
+M. J., Fortin, M.-J., Hemming, V., Mercier, B., Pollock, L., Runghen,
+R., Riva, G. V. D., \& Poisot, T. (2021). \emph{Food web reconstruction
+through phylogenetic transfer of low-rank network representation}.
+\url{https://doi.org/10.32942/osf.io/y7sdz}
+
 \leavevmode\hypertarget{ref-Strydom2021RoaPre}{}%
@@
 \url{https://doi.org/10.1098/rstb.2021.0063}
+
+\leavevmode\hypertarget{ref-Strydom2021SvdEnt}{}%
+Strydom, T., Dalla Riva, G. V., \& Poisot, T. (2021). SVD Entropy
+Reveals the High Complexity of Ecological Networks. \emph{Frontiers in
+Ecology and Evolution}, \emph{9}.
+\url{https://doi.org/10.3389/fevo.2021.623141}
+
+\leavevmode\hypertarget{ref-Thompson2004LanUse}{}%
+Thompson, R. M., \& Townsend, C. R. (2004). Land‐use influences on New
+Zealand stream communities: Effects on species composition, functional
+organisation, and food‐web structure. \emph{New Zealand Journal of
+Marine and Freshwater Research}, \emph{38}(4), 595--608.
+\url{https://doi.org/10.1080/00288330.2004.9517265}
+
+\leavevmode\hypertarget{ref-Thompson1999EffSea}{}%
+Thompson, R. M., \& Townsend, C. R. (1999). The Effect of Seasonal
+Variation on the Community Structure and Food-Web Attributes of Two
+Streams: Implications for Food-Web Science. \emph{Oikos}, \emph{87}(1),
+75--88. \url{https://doi.org/10.2307/3546998}
 

### 2021-12-13 draft.tex
Deploying to gh-pages from @ PoisotLab/ms_networkary_classifiers@0fce0931a5ebc09dbe834c2b1cd1595c42b2a141 🚀

--- 2021-12-12 draft.tex
@@
 
+ROC-AUC
+
+PR-AUC
+
 \hypertarget{confusion-matrix-with-skill-and-bias}{%
@@
 
-\hypertarget{illustration-on-empirical-food-webs}{%
-\section{Illustration on empirical food
-webs}\label{illustration-on-empirical-food-webs}}
+\hypertarget{numerical-experiments-on-training-strategy}{%
+\section{Numerical experiments on training
+strategy}\label{numerical-experiments-on-training-strategy}}
 
-In this section, we use a collection of XX empirical networks
+In the following section, we will generate random bipartite networks
+(this works without loss of generality on unipartite networks), and
+train four binary classifiers (as well as an ensemble model using the
+sum of ranged outputs from the component models) on 30\% of the
+interaction data. Networks are generated by picking a random
+infectiousness trait \(v_i\) for 100 species (from a \(B(6,8)\)
+distribution), and a resistance trait \(h_j\) for 100 species (from a
+\(B(2,8)\) distribution). There is an interaction between \(i\) and
+\(j\) when \(v_i-\xi/2 \le h_j \le v_i+\xi/2\), where \(\xi\) is a
+constant regulating the connectance of the network (there is an almost
+1:1 relationship between \(\xi\) and connectance), and varies uniformly
+in \([0.05, 0.35]\). This model gives fully interval networks that are
+close analogues to the bacteria--phage model of Weitz et al. (2005),
+with both a modular structure and a non-uniform degree distribution.
+This model is easy to learn: when trained with features
+\([v_i, h_j, \text{abs}(v_i, h_j)] ^T\) to predict the interactions
+between \(i\) and \(j\), all four models presented below were able to
+reach almost perfect predictions all the time (data not presented here)
+-- this is in part because the rule is fixed for all interactions. In
+order to make the problem more difficult to solve, we use \([v_i, h_j]\)
+as a feature vector, and therefore the models will have to uncover that
+the rule for interaction is \(\text{abs}(v_i, h_j) \le \xi\).
 
-\hypertarget{using-t-sbn-and-rdpg-to-simulate-arbitrarily-imperfect-classifier}{%
-\subsection{Using t-SBN and RDPG to simulate arbitrarily imperfect
-classifier}\label{using-t-sbn-and-rdpg-to-simulate-arbitrarily-imperfect-classifier}}
-
-use t-SVD to only use some information
-
-cumulative eigenvalues up to rank n give amount of information recovered
-
-can be thresholded
-
-how do values react to using fewer ranks
-
-\hypertarget{effect-of-classifier-performance-on-metrics}{%
-\subsection{Effect of classifier performance on
-metrics}\label{effect-of-classifier-performance-on-metrics}}
-
-\ldots{}
-
-\hypertarget{numerical-experiments}{%
-\section{Numerical experiments}\label{numerical-experiments}}
-
-In the following section, we will generate random networks, and train
-four binary classifiers (as well as an ensemble model using the sum of
-the outputs) on 30\% of the interaction data. Networks are generated by
-picking random generality \(g\) and vulnerability \(v\) traits for
-\(S = 200\) species uniformly on the unit interval, and assigning an
-interaction from species \(i\) to species \(j\) if
-\(0.2g_i-\xi \le v_j \le 0.2g_i+\xi\), where \(\xi\) is a constant
-regulating the connectance of the networks, and varies uniformly in
-\([5\times 10^{-3}, 10^{-1}]\). This model gives fully interval networks
-that are close analogues to the niche model (Williams \& Martinez,
-2000), but has the benefit of only relying on two features
-(\(g_i, v_j\)), and having the exact same rule for all interactions. It
-is, therefore, a simple case which most classifiers should be able to
-learn.
-
-The training sample is composed of 30\% of the \(4\times 10^4\) possible
-entries in the network, \emph{i.e.} \(n=12000\). Out of these
-interactions, we pick a proportion \(\nu\) (the training set bias) to be
-positive, so that the training set has \(\nu n\) interactions, and
-\((1-\nu) n\) non-interactions. We vary \(\nu\) uniformly in \(]0,1[\).
-This allows to evaluate how the measures of binary classification
-performance respond to artificially rebalanced dataset for a given
-network connectance. Note that both \(\xi\) and \(\nu\) are sampled from
-a distribution rather than being picked on a grid; this is because there
-is no direct relationship between the value of \(\xi\) and the
-connectance of the simulated network, and therefore the precise value of
-\(\xi\) is not relevant for the analysis of the results.
+The training sample is composed of 30\% of the \(10^4\) possible entries
+in the network, \emph{i.e.} \(n=3000\). Out of these interactions, we
+pick a proportion \(\nu\) (the training set bias) to be positive, so
+that the training set has \(\nu n\) interactions, and \((1-\nu) n\)
+non-interactions. We vary \(\nu\) uniformly in \(]0,1[\). This allows to
+evaluate how the measures of binary classification performance respond
+to artificially rebalanced dataset for a given network connectance. The
+rest of the dataset (\(n=7000\) pairs of species) is used as a testing
+set, on which all furher measures are calculated. Note that although the
+training set is balanced, the testing set is not, and retains (part of)
+the imbalance of the original data.
 
 The dataset used for numerical experiments is composed of 64000 such
-\((\xi, \nu)\) pairs, on which four learners are trained: a decision
+\((\xi, \nu)\) pairs, on which four machines are trained: a decision
 tree regressor, a boosted regression tree, a ridge regressor, and a
@@
 package (Blaom et al., 2020; Blaom \& Vollmer, 2020) in Julia 1.7
-(Bezanson et al., 2017). In order to pick the best adjacency matrix for
-a given learner, we performed a thresholding approach using 500 steps on
-predictions from the testing set, and picking the threshold that
-maximized Youden's informedness, which is usually the optimized target
-for imbalanced classification. During the thresholding step, we measured
-the area under the receiving-operator characteristic (ROC-AUC) and
-precision-recall (PR-AUC) curves, as measures of overall performance
-over the range of returned values. We report the ROC-AUC and PR-AUC, as
-well as a suite of other measures as introduced in the next section, for
-the best threshold. The ensemble model was generated by summing the
-predictions of all component models on the testing set (ranged in
-\([0,1]\)), then put through the same thresholding process. The complete
-code to run the simulations is given as an appendix; running the final
-simulation required 4.8 core days (approx. 117 hours).
+(Bezanson et al., 2017). All machines use the default parameterization;
+this is an obvious deviation from best practices, as the hyperparameters
+of any machine require training before its application on a real
+dataset. As we use 64000 such datasets, this would require 256000 unique
+instances of tweaking thehyperparameters, which is not realistic.
+Therefore, we assume that the default parameterizations are comparable
+across networks. All machines return a quantitative prediction, usually
+(but not necessarilly) in \([0,1]\), which is proportional (but not
+necessarilly linearly) to the probability of an interaction between
+\(i\) and \(j\).
+
+In order to pick the best adjacency matrix for a given trained machine,
+we performed a thresholding approach using 500 steps on predictions from
+the testing set, and picking the threshold that maximized Youden's
+informedness, which is usually the optimized target for imbalanced
+classification. During the thresholding step, we measured the area under
+the receiving-operator characteristic (ROC-AUC) and precision-recall
+(PR-AUC) curves, as measures of overall performance over the range of
+returned values. We report the ROC-AUC and PR-AUC, as well as a suite of
+other measures as introduced in the next section, for the best
+threshold. The ensemble model was generated by summing the predictions
+of all component models on the testing set (ranged in \([0,1]\)), then
+put through the same thresholding process. The complete code to run the
+simulations is given as an appendix; running the final simulation
+required 4.8 core days (approx. 117 hours).
 
@@
 \end{figure}
+
+\hypertarget{do-better-classification-accuracy-result-in-more-realistic-networks}{%
+\section{Do better classification accuracy result in more realistic
+networks?}\label{do-better-classification-accuracy-result-in-more-realistic-networks}}
 
@@
 
+\leavevmode\hypertarget{ref-Weitz2005CoeArm}{}%
+Weitz, J. S., Hartman, H., \& Levin, S. A. (2005). Coevolutionary arms
+races between bacteria and bacteriophage. \emph{Proceedings of the
+National Academy of Sciences of the United States of America},
+\emph{102}(27), 9535--9540.
+\url{https://doi.org/10.1073/pnas.0504062102}
+
 \leavevmode\hypertarget{ref-Whalen2021NavPit}{}%
@@
 
-\leavevmode\hypertarget{ref-Williams2000SimRul}{}%
-Williams, R., \& Martinez, N. (2000). Simple rules yield complex food
-webs. \emph{Nature}, \emph{404}, 180--183.
-\href{http://userwww.sfsu.edu/\%20}{http://userwww.sfsu.edu/}
-
 \end{CSLReferences}

### 2021-12-13 draft.tex
Deploying to gh-pages from @ PoisotLab/ms_networkary_classifiers@e91d67f78b3d2ceb0c2fdfcc532e299741595d66 🚀

@@
 conditions was met: the accuracy was 0, the true positive or true
-negative rates were 0, the connectance was larger than 0.2. This removes
-both the obviously failed model runs, and the networks that are more
-densely connected compared to the connectance of empirical food webs
-(and are therefore less difficult to predict, being less imbalanced).
+negative rates were 0, the connectance was larger than 0.25. This
+removes both the obviously failed model runs, and the networks that are
+more densely connected compared to the connectance of empirical food
+webs (and are therefore less difficult to predict, being less
+imbalanced; preliminary analyses of data with a connectance larger than
+3 revealed that all machines reached consistently high performance).
 
@@
 In fig.~\ref{fig:biasmccinf}, we present the response of MCC and
-informedness to (i) four levels of network connectance and (ii) a
+informedness to (i) five levels of network connectance and (ii) a
 gradient of training set bias, for the four component models as well as
@@
 Interestingly, the ensemble almost always outclassed its component
-models.
+models. In a few cases, both MCC and informedness stared decreasing when
+the training set bias got too close to one, which suggests that it is
+possible to over-correct the imbalance.
 
@@
 maximized a series of performance measures. When this value is high, the
-training set needs to skew positive in order to get a good model; when
-this value is about 0.5, the training set needs to be artificially
-balanced to optimize the model performance. These results are presented
-in fig.~\ref{fig:optimbias}.
+training set needs to skew more positive in order to get a performant
+model; when this value is about 0.5, the training set needs to be
+artificially balanced to optimize the model performance. These results
+are presented in fig.~\ref{fig:optimbias}.
 
@@
 
-Interestingly, as long as the connectance of the network was above
-\(\approx 0.1\), the optimal prevalence in the training set is 0.5,
-\emph{i.e.} as many positives as negatives. Low connectance is usually
-achieved for very large networks, due to the scaling relationship
-between richness and links (MacDonald et al., 2020). Therefore, larger
-networks may require \emph{more} biasing of the training set in order to
-be optimally predicted, whereas smaller, more connected networks may
-not. It is worth noting that the optimal bias for the training set
-stabilizes at 0.5 regardless of connectance \emph{and} model \emph{and}
-measure of model evaluation.
+The more ``optimistic'' measures (ROC-AUC and informedness) required a
+biasing of the dataset from about 0.4 to 0.75 to be maximized, with the
+amount of bias required decreasing only slightly with the connectance of
+the original network. MCC and PR-AUC required values of training set
+bias from 0.75 to almost 1 to be optimized, which is in line with the
+results of the previous section, \emph{i.e.} they are more stringent
+tests of model performance.
 
@@
 \end{figure}
+
+When trained at their optimal training set bias, performance still had a
+significant impact on the performance of some machines
+fig.~\ref{fig:optimperf}. Notably, Decision Tree, Random Forest, and
+Ridge Regression had low values of PR-AUC. In all cases, the Boosted
+Regression Tree was reaching very good predictions (esepcially for
+connectances larger than 0.1), and the ensemble was almost always
+scoring perfectly. This suggests that all the models are biased in
+different ways, and that the averaging in the ensemble is able to
+correct these biases. We do not expect this last result to have any
+generality, and provide a discussion of a recent exemple in which the
+ensemble was performing worse than its components models.
 
@@
 
-\leavevmode\hypertarget{ref-MacDonald2020RevLin}{}%
-MacDonald, A. A. M., Banville, F., \& Poisot, T. (2020). Revisiting the
-Links-Species Scaling Relationship in Food Webs. \emph{Patterns},
-\emph{1}(0). \url{https://doi.org/10.1016/j.patter.2020.100079}
-
 \leavevmode\hypertarget{ref-McLeod2021SamAsy}{}%

### 2021-12-13 code/04_rdpg.jl
table

@@ -1,68 +1,0 @@
-using EcologicalNetworks
-using AlgebraOfGraphics, CairoMakie
-using DataFrames
-using CSV: CSV
-using DataFramesMeta
-using LinearAlgebra
-using Statistics
-
-N = simplify.(nz_stream_foodweb())
-
-# AUC
-function ∫(x::Array{T}, y::Array{T}) where {T<:Number}
-    S = zero(Float64)
-    for i in 2:length(x)
-        S += (x[i] - x[i - 1]) * (y[i] + y[i - 1]) * 0.5
-    end
-    return .-S
-end
-
-# Confusion matrix utilities
-include(joinpath(@__DIR__, "confusionmatrix.jl"))
-
-
-results = DataFrame(network=Int64[], connectance=Float64[], rank=Int64[], varexpl=Float64[], measure=Symbol[], value=Float64[])
-
-for n in 1:length(N)
-    for r in 1:45
-        _svd = svd(N[n])
-        varexpl = sum(_svd.S[1:r])/sum(_svd.S)
-        prediction = prod(rdpg(N[n], r))
-        thresholds = LinRange(minimum(vec(prediction)), maximum(vec(prediction)), 200)
-        M = Vector{ConfusionMatrix}(undef, length(thresholds))
-        binobs = adjacency(N[n])
-        for (i, τ) in enumerate(thresholds)
-            binpred = prediction .>= τ
-            tp = sum(binobs .& binpred)
-            tn = sum(.!binobs .& .!binpred)
-            fp = sum(.!binobs .& binpred)
-            fn = sum(binobs .& .!binpred)
-            M[i] = ConfusionMatrix(tp, tn, fp, fn)
-        end
-        X = UnipartiteNetwork(prediction .>= thresholds[last(findmax(informedness.(M)))])
-        ROCAUC = ∫(fpr.(M), tpr.(M))
-        AUPRC = ∫(tpr.(M), ppv.(M))
-        𝐌 = M[last(findmax(informedness.(M)))]
-        push!(results, (n, connectance(N[n]), r, varexpl, :ROCAUC, ROCAUC))
-        push!(results, (n, connectance(N[n]), r, varexpl, :PRAUC, AUPRC))
-        push!(results, (n, connectance(N[n]), r, varexpl, :INF, informedness(𝐌)))
-        push!(results, (n, connectance(N[n]), r, varexpl, :MCC, mcc(𝐌)))
[…]

### 2021-12-13 code/04_sim2.jl

@@ -1,0 +1,166 @@
+using Statistics
+using StatsFuns
+using StatsBase
+using DataFramesMeta
+using Random
+using Statistics
+using MLJ
+using DataFrames
+using CSV: CSV
+using Distributions
+using AlgebraOfGraphics, CairoMakie
+using EcologicalNetworks
+using Random
+
+Random.seed!(1)
+
+# AUC
+function ∫(x::Array{T}, y::Array{T}) where {T<:Number}
+    S = zero(Float64)
+    for i in 2:length(x)
+        S += (x[i] - x[i - 1]) * (y[i] + y[i - 1]) * 0.5
+    end
+    return .-S
+end
+
+# Confusion matrix utilities
+include(joinpath(@__DIR__, "confusionmatrix.jl"))
+
+# Network generation function
+function L(x::T, y::T; r::T=0.1) where {T<:Number}
+    return (x - r / 2.0) ≤ y ≤ (x + r / 2.0) ? one(T) : zero(T)
+end
+function network(S, ξ)
+    infectivity = Beta(6.0, 8.0)
+    resistance = Beta(2.0, 8.0)
+    𝐱ᵥ = sort(rand(infectivity, S))
+    𝐱ₕ = sort(rand(resistance, S))
+    𝐱₁ = repeat(𝐱ᵥ; outer=length(𝐱ₕ))
+    𝐱₂ = repeat(𝐱ₕ; inner=length(𝐱ᵥ))
+    𝐱₃ = abs.(𝐱₁ .- 𝐱₂)
+    𝐲 = [L(𝐱₁[i], 𝐱₂[i]; r=ξ) for i in 1:(S * S)]
+    #𝐱 = table(hcat(𝐱₁, 𝐱₂, 𝐱₃))
+    𝐱 = table(hcat(𝐱₁, 𝐱₂))
+    return (𝐱, 𝐲)
+end
+
+DecisionTree = @load DecisionTreeRegressor pkg = DecisionTree
+RandomForest = @load RandomForestRegressor pkg = DecisionTree
+BoostedRegressor = @load EvoTreeRegressor pkg = EvoTrees
+RidgeRegressor = @load RidgeRegressor pkg = MLJLinearModels
+
+candidate_models = [
+    :DecTree => DecisionTree(),
+    :BRT => BoostedRegressor(),
+    :RF => RandomForest(),
+    :RR => RidgeRegressor(),
+]
+
+S = 50
+
[…]

### 2021-12-13 draft.tex
Deploying to gh-pages from @ PoisotLab/ms_networkary_classifiers@86ebd98bc34ce6f8d63cf608c9fdb2c6906dd657 🚀

@@
 
+In this last section, we generate a network using the same model as
+before, with \(S = 50\) species, a connectance of \(\approx 0.16\)
+(\(\xi = 0.19\)), and a training set bias of \(0.7\). The prediction
+made on the complete dataset is presented in fig.~\ref{fig:ecovalid}.
+Visualizing the results this way highlights the importance of
+exploratory data analysis: whereas all models return a network with
+interactions laying mostly on the diagonal (as expected), the Ridge
+Regression is quite obviously biased. Despite this, we can see that the
+ensemble is close to the initial dataset.
+
+\includegraphics{figures/valid_ensemble.png}
+
+The trained models were then thresholded (again by optimising
+informedness), and their predictions transformed back into networks for
+analysis; specifically, we measured the connectance, nestedness (REF),
+modularity (REF), and entropy (REF). The results are presented in the
+following table:
+
+\begin{longtable}[]{@{}lllllllll@{}}
+\toprule
+Model & MCC & Inf. & ROC-AUC & PR-AUC & Conn. & \(\eta\) & \(Q\) &
+Entropy\tabularnewline
+\midrule
+\endhead
+Decision tree & 0.83 & 0.68 & 0.95 & 0.15 & 0.18 & 0.53 & 0.49 &
+8.86\tabularnewline
+BRT & 0.76 & 0.89 & 0.95 & \textbf{0.65} & 0.22 & 0.63 & 0.43 &
+9.14\tabularnewline
+Random Forest & \textbf{0.89} & \textbf{0.94} & \textbf{0.99} & 0.41 &
+\textbf{0.17} & \textbf{0.48} & \textbf{0.49} &
+\textbf{8.80}\tabularnewline
+Ridge Regression & 0.67 & 0.85 & 0.89 & 0.38 & 0.27 & 1.0 & 0.26 &
+9.40\tabularnewline
+Ensemble & 0.84 & 0.91 & \textbf{0.99} & \textbf{0.94} & 0.19 & 0.54 &
+0.48 & 8.92\tabularnewline
+Data & & & & & 0.16 & 0.45 & 0.49 & 8.71\tabularnewline
+\bottomrule
+\end{longtable}
+
 \hypertarget{guidelines-for-the-assesment-of-network-predictive-models}{%

### 2021-12-13 draft.tex
Deploying to gh-pages from @ PoisotLab/ms_networkary_classifiers@925c31f38baa91496e12e982e1cc699c0cb6afe0 🚀

@@
 analysis; specifically, we measured the connectance, nestedness (REF),
-modularity (REF), and entropy (REF). The results are presented in the
-following table:
+modularity (REF), and entropy (REF). The results are presented in
+tbl.~\ref{tbl:comparison}. The random forest model is an interesting
+instance here: it produces the network that looks the most like the
+original dataset, despite having a very low PR-AUC, suggesting it hits
+high recall at the cost of low precision. Although the ensemble was
+about to reach a very high PR-AUC (and a very high ROC-AUC), this did
+not necessarilly translate into more accurate reconstructions of the
+structure of the network. This result bears elaborating. Measures of
+model performance capture how much of the interactions and
+non-interactions are correctly identified. As long as these predictions
+are not perfect, some interactions will be predicted at the ``wrong''
+position in the network; these measures cannot describe the structural
+effect of these mistakes. On the other hand, measures of network
+structure can have the same value with interactions that fall at
+drastically different positions; this is in part because a lot of these
+measures covary with connectance, and in part because as long as these
+values are not 0 or their respective maximum, there is a large number of
+network configurations that can have the same value.
 
+\hypertarget{tbl:comparison}{}
 \begin{longtable}[]{@{}lllllllll@{}}
+\caption{\label{tbl:comparison}Values of four performance metrics, and
+four network structure metrics, for the predictions presented in
+fig.~\ref{fig:ecovalid}. The values in \textbf{bold} indicate the best
+value for each column (including ties).}\tabularnewline
+\toprule
+Model & MCC & Inf. & ROC-AUC & PR-AUC & Conn. & \(\eta\) & \(Q\) &
+Entropy\tabularnewline
+\midrule
+\endfirsthead
 \toprule
@@
 \endhead
-Decision tree & 0.83 & 0.68 & 0.95 & 0.15 & 0.18 & 0.53 & 0.49 &
-8.86\tabularnewline
-BRT & 0.76 & 0.89 & 0.95 & \textbf{0.65} & 0.22 & 0.63 & 0.43 &
+Decision tree & 0.83 & 0.68 & 0.95 & 0.15 & 0.18 & 0.53 & \textbf{0.49}
+& 8.86\tabularnewline
+BRT & 0.76 & 0.89 & 0.95 & 0.65 & 0.22 & 0.63 & 0.43 &
 9.14\tabularnewline

### 2021-12-13 preprint.tex
Deploying to gh-pages from @ PoisotLab/ms_networkary_classifiers@925c31f38baa91496e12e982e1cc699c0cb6afe0 🚀

@@
 analysis; specifically, we measured the connectance, nestedness (REF),
-modularity (REF), and entropy (REF). The results are presented in the
-following table:
-
+modularity (REF), and entropy (REF). The results are presented in
+tbl.~\ref{tbl:comparison}. The random forest model is an interesting
+instance here: it produces the network that looks the most like the
+original dataset, despite having a very low PR-AUC, suggesting it hits
+high recall at the cost of low precision. Although the ensemble was
+about to reach a very high PR-AUC (and a very high ROC-AUC), this did
+not necessarilly translate into more accurate reconstructions of the
+structure of the network. This result bears elaborating. Measures of
+model performance capture how much of the interactions and
+non-interactions are correctly identified. As long as these predictions
+are not perfect, some interactions will be predicted at the ``wrong''
+position in the network; these measures cannot describe the structural
+effect of these mistakes. On the other hand, measures of network
+structure can have the same value with interactions that fall at
+drastically different positions; this is in part because a lot of these
+measures covary with connectance, and in part because as long as these
+values are not 0 or their respective maximum, there is a large number of
+network configurations that can have the same value.
+
+\hypertarget{tbl:comparison}{}
 \begin{longtable}[]{@{}lllllllll@{}}
+\caption{\label{tbl:comparison}Values of four performance metrics, and
+four network structure metrics, for the predictions presented in
+fig.~\ref{fig:ecovalid}. The values in \textbf{bold} indicate the best
+value for each column (including ties).}\tabularnewline
+\toprule
+Model & MCC & Inf. & ROC-AUC & PR-AUC & Conn. & \(\eta\) & \(Q\) &
+Entropy\tabularnewline
+\midrule
+\endfirsthead
 \toprule

### 2021-12-13 draft.tex
Deploying to gh-pages from @ PoisotLab/ms_networkary_classifiers@0acf3eab6214f5fef9a85ed8bd54e8e397320da2 🚀
Deploying to gh-pages from @ PoisotLab/ms_networkary_classifiers@e6b6520e90d1098c2015de1f62973bcd08345f29 🚀
Deploying to gh-pages from @ PoisotLab/ms_networkary_classifiers@529ac2db9673226635fdaebe9f4decf3458f1e77 🚀

@@
 
+\begin{figure}
+\hypertarget{fig:ecovalid}{%
+\centering
 \includegraphics{figures/valid_ensemble.png}
+\caption{TODO}\label{fig:ecovalid}
+}
+\end{figure}
 
@@
 analysis; specifically, we measured the connectance, nestedness (REF),
-modularity (REF), and entropy (REF). The results are presented in
+and modularity (REF). The results are presented in
 tbl.~\ref{tbl:comparison}. The random forest model is an interesting
@@
 \hypertarget{tbl:comparison}{}
-\begin{longtable}[]{@{}lllllllll@{}}
+\begin{longtable}[]{@{}rccccccc@{}}
 \caption{\label{tbl:comparison}Values of four performance metrics, and
-four network structure metrics, for the predictions presented in
+three network structure metrics, for the predictions presented in
 fig.~\ref{fig:ecovalid}. The values in \textbf{bold} indicate the best
@@
 \toprule
-Model & MCC & Inf. & ROC-AUC & PR-AUC & Conn. & \(\eta\) & \(Q\) &
-Entropy\tabularnewline
+Model & MCC & Inf. & ROC-AUC & PR-AUC & Conn. & \(\eta\) &
+\(Q\)\tabularnewline
 \midrule
@@
 \toprule
-Model & MCC & Inf. & ROC-AUC & PR-AUC & Conn. & \(\eta\) & \(Q\) &
-Entropy\tabularnewline
+Model & MCC & Inf. & ROC-AUC & PR-AUC & Conn. & \(\eta\) &
+\(Q\)\tabularnewline
 \midrule
 \endhead
-Decision tree & 0.83 & 0.68 & 0.95 & 0.15 & 0.18 & 0.53 & \textbf{0.49}
-& 8.86\tabularnewline
-BRT & 0.76 & 0.89 & 0.95 & 0.65 & 0.22 & 0.63 & 0.43 &
-9.14\tabularnewline
+Decision tree & 0.83 & 0.68 & 0.95 & 0.15 & 0.18 & 0.53 &
+\textbf{0.49}\tabularnewline
+BRT & 0.76 & 0.89 & 0.95 & 0.65 & 0.22 & 0.63 & 0.43\tabularnewline
 Random Forest & \textbf{0.89} & \textbf{0.94} & \textbf{0.99} & 0.41 &
-\textbf{0.17} & \textbf{0.48} & \textbf{0.49} &
-\textbf{8.80}\tabularnewline
-Ridge Regression & 0.67 & 0.85 & 0.89 & 0.38 & 0.27 & 1.0 & 0.26 &
-9.40\tabularnewline
+\textbf{0.17} & \textbf{0.48} & \textbf{0.49}\tabularnewline
+Ridge Regression & 0.67 & 0.85 & 0.89 & 0.38 & 0.27 & 1.0 &
+0.26\tabularnewline
 Ensemble & 0.84 & 0.91 & \textbf{0.99} & \textbf{0.94} & 0.19 & 0.54 &
-0.48 & 8.92\tabularnewline
-Data & & & & & 0.16 & 0.45 & 0.49 & 8.71\tabularnewline
+0.48\tabularnewline
+Data & & & & & 0.16 & 0.45 & 0.49\tabularnewline
 \bottomrule

### 2021-12-13 code/04_sim2.jl
last section done, ccl and intro need updating

@@ -33,12 +33,12 @@ end
 function network(S, ξ)
     infectivity = Beta(6.0, 8.0)
     resistance = Beta(2.0, 8.0)
-    𝐱ᵥ = sort(rand(infectivity, S))
-    𝐱ₕ = sort(rand(resistance, S))
+    𝐱ᵥ = sort(rand(infectivity, S[1]))
+    𝐱ₕ = sort(rand(resistance, S[2]))
     𝐱₁ = repeat(𝐱ᵥ; outer=length(𝐱ₕ))
     𝐱₂ = repeat(𝐱ₕ; inner=length(𝐱ᵥ))
     𝐱₃ = abs.(𝐱₁ .- 𝐱₂)
-    𝐲 = [L(𝐱₁[i], 𝐱₂[i]; r=ξ) for i in 1:(S * S)]
+    𝐲 = [L(𝐱₁[i], 𝐱₂[i]; r=ξ) for i in 1:length(𝐱₁)]
     #𝐱 = table(hcat(𝐱₁, 𝐱₂, 𝐱₃))
     𝐱 = table(hcat(𝐱₁, 𝐱₂))
     return (𝐱, 𝐲)
@@ -56,7 +56,8 @@ candidate_models = [
     :RR => RidgeRegressor(),
 ]
 
-S = 50
+S = (50,80)
+
 
 𝐗, 𝐲 = network(S, 0.19)
 bias = 0.7
@@ -154,13 +155,3 @@ data(results) *
     visual(Heatmap, colormap=Reverse(:deep)) |>
     plt -> draw(plt, facet=(;linkyaxes = :minimal), axis = (xticks = LinearTicks(3),)) |>
     plt -> save(joinpath(@__DIR__, "..", "figures", "valid_ensemble.png"), plt, px_per_unit = 3)
-
-for v in ("DecTree", "BRT", "RF", "RR", "Ensemble", "Dataset")
-    @info v
-    @info (
-        reshape(@subset(results, :model .== v).guess, (S, S)) .|>
-        Bool
-    ) |>
-        BipartiteNetwork |>
-        n -> Q(brim(lp(n)...)...)
-end

### 2021-12-13 code/05_sim3.jl

@@ -1,0 +1,176 @@
+using Statistics
+using StatsFuns
+using StatsBase
+using DataFramesMeta
+using Random
+using Statistics
+using MLJ
+using DataFrames
+using CSV: CSV
+using Distributions
+using AlgebraOfGraphics, CairoMakie
+using EcologicalNetworks
+using Random
+
+Random.seed!(1)
+
+# AUC
+function ∫(x::Array{T}, y::Array{T}) where {T<:Number}
+    S = zero(Float64)
+    for i in 2:length(x)
+        S += (x[i] - x[i - 1]) * (y[i] + y[i - 1]) * 0.5
+    end
+    return .-S
+end
+
+# Confusion matrix utilities
+include(joinpath(@__DIR__, "confusionmatrix.jl"))
+
+# Network generation function
+function L(x::T, y::T; r::T=0.1) where {T<:Number}
+    return (x - r / 2.0) ≤ y ≤ (x + r / 2.0) ? one(T) : zero(T)
+end
+function network(S, ξ)
+    infectivity = Beta(6.0, 8.0)
+    resistance = Beta(2.0, 8.0)
+    𝐱ᵥ = sort(rand(infectivity, S[1]))
+    𝐱ₕ = sort(rand(resistance, S[2]))
+    𝐱₁ = repeat(𝐱ᵥ; outer=length(𝐱ₕ))
+    𝐱₂ = repeat(𝐱ₕ; inner=length(𝐱ᵥ))
+    𝐱₃ = abs.(𝐱₁ .- 𝐱₂)
+    𝐲 = [L(𝐱₁[i], 𝐱₂[i]; r=ξ) for i in 1:length(𝐱₁)]
+    #𝐱 = table(hcat(𝐱₁, 𝐱₂, 𝐱₃))
+    𝐱 = table(hcat(𝐱₁, 𝐱₂))
+    return (𝐱, 𝐲)
+end
+
+function R(x, m, M)
+    x = x .- m
+    x = x ./ (M - m)
+    return x
+end
+
+R(x) = R(x, extrema(x)...)
+
+
+DecisionTree = @load DecisionTreeRegressor pkg = DecisionTree
+RandomForest = @load RandomForestRegressor pkg = DecisionTree
+BoostedRegressor = @load EvoTreeRegressor pkg = EvoTrees
+RidgeRegressor = @load RidgeRegressor pkg = MLJLinearModels
+
+candidate_models = [
+    :DecTree => DecisionTree(),
[…]

### 2021-12-13 draft.tex
Deploying to gh-pages from @ PoisotLab/ms_networkary_classifiers@d25924fed7ca63135b35efc4c9eb3fc04798bde0 🚀

@@
 
-\(F_{\beta}\)
-
 informedness
@@
 
-PR-AUC
+\(F_{\beta}\)
+
+PR-AUC - baseline is prevalence, wtf are recall and precision, you may
+ask?
 
@@
 In this last section, we generate a network using the same model as
-before, with \(S = 50\) species, a connectance of \(\approx 0.16\)
-(\(\xi = 0.19\)), and a training set bias of \(0.7\). The prediction
-made on the complete dataset is presented in fig.~\ref{fig:ecovalid}.
-Visualizing the results this way highlights the importance of
-exploratory data analysis: whereas all models return a network with
-interactions laying mostly on the diagonal (as expected), the Ridge
-Regression is quite obviously biased. Despite this, we can see that the
-ensemble is close to the initial dataset.
+before, with \(S_1, S_2 = 50, 80\) species, a connectance of
+\(\approx 0.16\) (\(\xi = 0.19\)), and a training set bias of \(0.7\).
+The prediction made on the complete dataset is presented in
+fig.~\ref{fig:ecovalid}. Visualizing the results this way highlights the
+importance of exploratory data analysis: whereas all models return a
+network with interactions laying mostly on the diagonal (as expected),
+the Ridge Regression is quite obviously biased. Despite this, we can see
+that the ensemble is close to the initial dataset.
 
@@
 analysis; specifically, we measured the connectance, nestedness (REF),
-and modularity (REF). The results are presented in
-tbl.~\ref{tbl:comparison}. The random forest model is an interesting
-instance here: it produces the network that looks the most like the
-original dataset, despite having a very low PR-AUC, suggesting it hits
-high recall at the cost of low precision. Although the ensemble was
-about to reach a very high PR-AUC (and a very high ROC-AUC), this did
-not necessarilly translate into more accurate reconstructions of the
-structure of the network. This result bears elaborating. Measures of
-model performance capture how much of the interactions and
-non-interactions are correctly identified. As long as these predictions
-are not perfect, some interactions will be predicted at the ``wrong''
-position in the network; these measures cannot describe the structural
-effect of these mistakes. On the other hand, measures of network
-structure can have the same value with interactions that fall at
+and modularity (REF). This process was repeated 250 times, and the
+results are presented in tbl.~\ref{tbl:comparison}. The random forest
+model is an interesting instance here: it produces the network that
+looks the most like the original dataset, despite having a very low
+PR-AUC, suggesting it hits high recall at the cost of low precision.
+Although the ensemble was about to reach a very high PR-AUC (and a very
+high ROC-AUC), this did not necessarilly translate into more accurate
+reconstructions of the structure of the network. This result bears
+elaborating. Measures of model performance capture how much of the
+interactions and non-interactions are correctly identified. As long as
+these predictions are not perfect, some interactions will be predicted
+at the ``wrong'' position in the network; these measures cannot describe
+the structural effect of these mistakes. On the other hand, measures of
+network structure can have the same value with interactions that fall at
 drastically different positions; this is in part because a lot of these
@@
 values are not 0 or their respective maximum, there is a large number of
-network configurations that can have the same value.
+network configurations that can have the same value. That ROC-AUC is
+consistently larger than PR-AUC may be a case of this measure masking
+models that are not, individually, strong predictors (Jeni et al.,
+2013).
 
@@
 \caption{\label{tbl:comparison}Values of four performance metrics, and
-three network structure metrics, for the predictions presented in
-fig.~\ref{fig:ecovalid}. The values in \textbf{bold} indicate the best
-value for each column (including ties).}\tabularnewline
+three network structure metrics, for 250 independent predictions similar
+to the ones presented in fig.~\ref{fig:ecovalid}. The values in
+\textbf{bold} indicate the best value for each column (including ties).
+Because the values have been rounded, values of 1.0 for the ROC-AUC
+column indicate an average \(\ge 0.99\).}\tabularnewline
 \toprule
@@
 \endhead
-Decision tree & 0.83 & 0.68 & 0.95 & 0.15 & 0.18 & 0.53 &
-\textbf{0.49}\tabularnewline
-BRT & 0.76 & 0.89 & 0.95 & 0.65 & 0.22 & 0.63 & 0.43\tabularnewline
-Random Forest & \textbf{0.89} & \textbf{0.94} & \textbf{0.99} & 0.41 &
-\textbf{0.17} & \textbf{0.48} & \textbf{0.49}\tabularnewline
-Ridge Regression & 0.67 & 0.85 & 0.89 & 0.38 & 0.27 & 1.0 &
-0.26\tabularnewline
-Ensemble & 0.84 & 0.91 & \textbf{0.99} & \textbf{0.94} & 0.19 & 0.54 &
-0.48\tabularnewline
-Data & & & & & 0.16 & 0.45 & 0.49\tabularnewline
+Decision tree & 0.85 & 0.92 & 0.97 & 0.12 & 0.21 & 0.76 &
+0.31\tabularnewline
+BRT & 0.90 & 0.90 & 0.98 & 0.86 & 0.23 & 0.82 & 0.27\tabularnewline
+Random Forest & \textbf{0.90} & \textbf{0.96} & \textbf{1.00} & 0.27 &
+\textbf{0.20} & \textbf{0.72} & \textbf{0.32}\tabularnewline
+Ridge Regression & 0.80 & 0.91 & 0.95 & 0.58 & 0.24 & 1.0 &
+0.18\tabularnewline
+Ensemble & 0.88 & 0.94 & \textbf{1.00} & \textbf{0.96} & \textbf{0.20} &
+0.75 & 0.31\tabularnewline
+Data & & & & & 0.18 & 0.66 & 0.34\tabularnewline
 \bottomrule
@@
 
+\leavevmode\hypertarget{ref-Jeni2013FacImb}{}%
+Jeni, L. A., Cohn, J. F., \& De La Torre, F. (2013). Facing Imbalanced
+Data--Recommendations for the Use of Performance Metrics. \emph{2013
+Humaine Association Conference on Affective Computing and Intelligent
+Interaction}, 245--251. \url{https://doi.org/10.1109/ACII.2013.47}
+
 \leavevmode\hypertarget{ref-Jordano2016ChaEco}{}%

### 2021-12-13 preprint.tex
Deploying to gh-pages from @ PoisotLab/ms_networkary_classifiers@d25924fed7ca63135b35efc4c9eb3fc04798bde0 🚀

@@
 
+informedness
+
+MCC
+
+ROC-AUC
+
 \(F_{\beta}\)
 
-informedness
-
-MCC
-
-ROC-AUC
-
-PR-AUC
+PR-AUC - baseline is prevalence, wtf are recall and precision, you may
+ask?
 

### 2021-12-13 draft.tex
Deploying to gh-pages from @ PoisotLab/ms_networkary_classifiers@cae9ebd294a5cdb86e6726c5c2b15722c29476a8 🚀

@@
 
+There is an immense amount of measures to evaluate the performance of
+classification tasks (Ferri et al., 2009). Here we will focus on five of
+them with high relevance for imbalanced learning (He \& Ma, 2013); three
+threshold metrics (\(\kappa\), informedness, and the Matthews
+Correlation Coefficient), and two ranking metrics (the ROC-AUC and
+PR-AUC). The \(\kappa\) measure of agreement (Landis \& Koch, 1977)
+establishes the extent to which two observers (here the data and the
+prediction) agree, and is measured as
+
+\[
+2\frac{tp\times tn - fn\times fp}{(tp+fp)\times (fp+tn)+(tn+fp)\times (tn+fn)} \,.
+\]
+
+Informedness (\textbf{Youden1950IndRat?}) (also known as bookmaker
+informedness or the True Skill Statistic) is
+\(\text{TPR}+\text{TNR}-1\), where \(\text{TPR}= tp/(tp+fn)\) and
+\(\text{TNR} = tn/(tn+fp)\); informedness can be used to find the
+optimal cutpoint in thresholding analyses (Schisterman et al., 2005).
+The MCC is defined as
+
+\[
+\frac{tp\times tn - fn\times fp}{\srt{(tp+fp)\times (tp+fn)\times (tn+fp)\times (tn+fn)}} \,.
+\]
+
+Finally, \(F_1\) is the harmonic mean of precision (the chance that a
+positive even was correctly classified) and sensitivity (the ability to
+correctly classify positive events), and is defined as
+
+\[
+2\frac{tp}{2\times tp + fp + fn}\,.
+\]
+
+A lot of binary classifiers are built by using a regressor (whose task
+is to guess the value of the interaction, amd can therefore return
+somethins considered to be a pseudo-probability); in this case, the
+optimal value below which predicitions are assumed to be negative
+(\emph{i.e.} the interaction does not exist) can be determined by
+picking a threshold maximizing some value on the ROC curve or the PR
+curve. The area under these curves (ROC-AUC and PR-AUC henceforth) give
+ideas on the overall goodness of the classifier. Saito \& Rehmsmeier
+(2015) established that the ROC-AUC is biased towards over-estimating
+performance for imbalanced data; on the contrary, the PR-AUC is able to
+identify classifiers that are less able to detect positive interactions
+correctly, with the additional advantage of having a baseline value
+equal to prevalence. Therefore, it is important to assess whether these
+two measures return different results when applied to ecological network
+prediction. The ROC curve is defined by the false positive rate on the
+\(x\) axis, and the true positive rate on the \(y\) axis, and the PR
+curve is defined by the true positive rate on the \(x\) axis, and the
+positive predictive value on the \(y\) axis. By comparison with the
+previous paragraph, it is obvious that \(F_1\) has ties to the PR curve
+(being close to the expected PR-AUC), and that informedness has ties to
+the ROC curve (whereby the threshold maximizing informedness is also the
+point of maximal inflection on the ROC curve). One important difference
+between ROC and PR is that the later does not prominently account for
+the size of the true negative compartments: in short, it is more
+sensitive to the correct positive predictions. In a context of strong
+imbalance, PR-AUC is therefore a more stringent test of model
+performance.
+
 The same approach is used to evaluate \emph{e.g.} species distribution
 models (SDMs). Indeed, the training and evaluation of SDMs as binary
-classifiers suffers from the same issue of low prevalence. In a previous
+classifiers suffers from the same issue of low prevalence. In previous
 work, Allouche et al. (2006) suggested that \(\kappa\) was a better test
-of model performance than the True Skill Statistic (TSS), which we will
-refer to as Youden's informedness (or \(J\)); these conclusions were
-later criticized by Somodi et al. (2017), who emphasized that
-informedness' relationship to prevalence depends on assumptions about
-bias in the model, and therefore recommend the use of \(\kappa\) as a
-validation of classification performance. Although this work offers
-recommendations about the comparison of models, it doesn't establishes
-baselines or good practices for training on imbalanced ecological data.
-Within the context of networks, there are three specific issues that
-need to be adressed. First, what values of performance measures are we
-expecting for a classifier that has poor performance? This is
-particularly important as it can evaluate whether low prevalence can
-lull us into a false sense of predictive accuracy. Second, independently
-of the question of model evaluation, is low prevalence an issue for
-\emph{training}, and can we remedy it? Finally, because the low amount
-of data on interaction makes a lot of imbalance correction methods (see
-\emph{e.g.} Branco et al., 2015) hard to apply, which indicators can be
-optimized with the least amount of positive interaction data?
+of model performance than the True Skill Statistic (TSS; which we refer
+to as Youden's informedness); these conclusions were later criticized by
+Somodi et al. (2017), who emphasized that informedness' relationship to
+prevalence depends on assumptions about bias in the model, and therefore
+recommend the use of \(\kappa\) as a validation of classification
+performance. Although this work offers recommendations about the
+comparison of models, it doesn't establishes baselines or good practices
+for training on imbalanced ecological data. Within the context of
+networks, there are three specific issues that need to be adressed.
+First, what values of performance measures are we expecting for a
+classifier that has poor performance? This is particularly important as
+it can evaluate whether low prevalence can lull us into a false sense of
+predictive accuracy. Second, independently of the question of model
+evaluation, is low prevalence an issue for \emph{training}, and can we
+remedy it? Finally, because the low amount of data on interaction makes
+a lot of imbalance correction methods (see \emph{e.g.} Branco et al.,
+2015) hard to apply, which indicators can be optimized with the least
+amount of positive interaction data?
 
@@
 
-A lot of binary classifiers are built by using a regressor (whose task
-is to guess the value of the interaction, amd can therefore return
-somethins considered to be a pseudo-probability); in this case, the
-optimal value below which predicitions are assumed to be negative
-(\emph{i.e.} the interaction does not exist) can be determined by
-picking a threshold maximizing some value on the ROC curve or the PR
-curve. The area under these curves (ROC-AUC and PR-AUC henceforth) give
-ideas on the overall goodness of the classifier. Saito \& Rehmsmeier
-(2015) established that the ROC-AUC is biased towards over-estimating
-performance for imbalanced data; on the contrary, the PR-AUC is able to
-identify classifiers that are less able to detect positive interactions
-correctly, with the additional advantage of having a baseline value
-equal to prevalence. Therefore, it is important to assess whether these
-two measures return different results when applied to ecological network
-prediction.
-
 We establish that due to the low prevalence of interactions, even poor
@@
 because the measure is dominated by the accidental correct predictions
-of negatives. The \(F_1\) score and positive predictive values are less
-sensitive to bias, but \textbf{TODO}
+of negatives. On simulated confusion matrices with ranges of imbalance
+that are credible for ecological networks, MCC had the most desirable
+behavior, and informedness is a linear measure of classifier skill. By
+performing simulations with four models and an ensemble,we show that
+informedness and ROC-AUC are consistently high on network data, and that
+MCC and PR-AUC are more accurate measures of the effective performance
+of the classifier. Finally, by measuring the structure of predicted
+networks, we highlight an interesting paradox: the models with the best
+performance measures are not the models with the closest reconstructed
+network structure. We discuss these results in the context of
+establishing guidelines for the prediction of ecological interactions.
 
@@
 Intro
-
-\hypertarget{definition-of-the-performance-measures}{%
-\subsection{Definition of the performance
-measures}\label{definition-of-the-performance-measures}}
-
-\(\kappa\)
-
-informedness
-
-MCC
-
-ROC-AUC
-
-\(F_{\beta}\)
-
-PR-AUC - baseline is prevalence, wtf are recall and precision, you may
-ask?
 
@@
 0.31\tabularnewline
-BRT & 0.90 & 0.90 & 0.98 & 0.86 & 0.23 & 0.82 & 0.27\tabularnewline
+BRT & \textbf{0.90} & 0.90 & 0.98 & 0.86 & 0.23 & 0.82 &
+0.27\tabularnewline
 Random Forest & \textbf{0.90} & \textbf{0.96} & \textbf{1.00} & 0.27 &
@@
 
-The results presented here highlight an interesting paradox: larger
-networks (with lower connectance) require more training set bias in
-order to maximize model performance fig.~\ref{fig:optimbias}, but are
-also more difficult to predict according to MCC and PR-AUC
-fig.~\ref{fig:optimperf}. This suggests that the task of network
-prediction will be difficult regardless of network size: by being
-limited by the \emph{frequency} of interactions when the network is
-large, and by being limited by the \emph{number} of interactions when
-the network is small. Nevertheless, based on the simulations and
-numerical experiments, it is possible to formulate a series of
-recommendations for the evaluation of network prediction models.
+The results presented here highlight an interesting paradox: although
+the Random Forest was ultimately able to get a correct estimate of
+network structure tbl.~\ref{tbl:comaprison}, it ultimately remains a
+poor classifier, as evidenced by its low PR-AUC. This suggests that the
+goal of predicting \emph{interactions} and predicting \emph{networks}
+may not be solvable in the same way -- of course a perfect classifier of
+interactions would make a perfect network prediction; but even the best
+scoring predictor of interactions (the ensemble model) had not
+necessarilly the best prediction of network structure. The tasks of
+predicting networks structure and of predicting interactions within
+networks are essentially two different ones. For some applications
+(\emph{.e.g.} comparison of network structure across gradients), one may
+care more about a robust estimate of the structure, at the cost at
+putting some interactions at the wrong place. For other applications
+(\emph{e.g.} identifying pairs of interacting species), one may
+conversely care more about getting as many pairs right, even though the
+mistakes accumulate in the form of a slightly worse estimate of network
+structure. How these two approaches can be reconciled is undoubtedly a
+task for further research. Despite this apprent tension at the heart of
+the predictive exercise, we can use the results presented here to
+suggest a number of guidelines.
 
@@
 discarded (in fact, a low ROC-AUC is a sign of an issue with the model),
-but that its interpretation should be guided by the PR-AUC value. This
-again echoes recommendations from other fields (Saito \& Rehmsmeier,
-2015).
-
-Thirdly, regardless of network connectance \emph{or} measure to evaluate
-the model performance, as long as the network connectance is larger than
-\(\approx 0.1\), artificially balancing the training set to have
-equiprevalence will give the best possible results. This was true for
-all models.
+but that its interpretation should be guided by the PR-AUC value.
+Specifically, a high ROC-AUC is not informative, as it can be associated
+to a low PR-AUC (see \emph{e.g.} Random Forest in
+tbl.~\ref{tbl:comparison}) This again echoes recommendations from other
+fields (Jeni et al., 2013; Saito \& Rehmsmeier, 2015).
+
+Thirdly, regardless of network connectance, maximizing informedness
+required a training set bias of about 0.5, and maximizing the MCC
+required a training set bias of 0.7 and more. This has an important
+consequence in ecological networks, for which the pool of positive cases
+(interactions) to draw from is typically small: the most parsimonious
+measure (\emph{i.e.} the one requiring to discard the least amount of
+information to train the model) will give the best validation potential,
+and is probably the informedness (maximizing informedness is the
+generally accepted default for imbalanced classification; Schisterman et
+al., 2005).
 
 Finally, it is noteworthy that the ensemble model was systematically
-better than the component models; even when poor models were included
-(Random Forest and Decision Tree), the ensemble was able to leverage the
-different biases expressed by the models to make an overall more
-accurate prediction. We do not expect that ensembles will \emph{always}
-be better than single models. In a recent multi-model comparison, Becker
-et al. (2021) found that the ensemble was \emph{not} the best model.
-There is no general conclusion to draw from this besides reinforcing the
-need to be pragmatic about which models should be included in the
-ensemble, or whether to use an ensemble at all. In a sense, the
-surprising peformance of the ensemble model should form the basis of the
-last recommendation: optimal training set bias and its interaction with
-connectance and binary classifier is, in a sense, an hyperparameter that
-should be assessed. The distribution of results in
-fig.~\ref{fig:optimbias} and fig.~\ref{fig:optimperf} show that there
-are variations around the trend; furthermore, networks with different
-structures than the one we simulated here may respond in different ways.
+better than the component models; even when the models were individually
+far form perfect, the ensemble was able to leverage the different biases
+expressed by the models to make an overall more accurate prediction. We
+do not expect that ensembles will \emph{always} be better than single
+models. In a recent multi-model comparison, Becker et al. (2021) found
+that the ensemble was \emph{not} the best model. There is no general
+conclusion to draw from this besides reinforcing the need to be
+pragmatic about which models should be included in the ensemble, or
+whether to use an ensemble at all. In a sense, the surprising peformance
+of the ensemble model should form the basis of the last recommendation:
+optimal training set bias and its interaction with connectance and
+binary classifier is, in a sense, an hyperparameter that should be
+assessed. The distribution of results in fig.~\ref{fig:optimbias} and
+fig.~\ref{fig:optimperf} show that there are variations around the
+trend; furthermore, networks with different structures than the one we
+simulated here may respond in different ways.
 
@@
 
+\leavevmode\hypertarget{ref-Ferri2009ExpCom}{}%
+Ferri, C., Hernández-Orallo, J., \& Modroiu, R. (2009). An experimental
+comparison of performance measures for classification. \emph{Pattern
+Recognition Letters}, \emph{30}(1), 27--38.
+\url{https://doi.org/10.1016/j.patrec.2008.08.010}
+
+\leavevmode\hypertarget{ref-He2013ImbLea}{}%
+He, H., \& Ma, Y. (Eds.). (2013). \emph{Imbalanced Learning:
+Foundations, Algorithms, and Applications} (1st edition). Wiley-IEEE
+Press.
+
 \leavevmode\hypertarget{ref-Jeni2013FacImb}{}%
@@
 
+\leavevmode\hypertarget{ref-Landis1977MeaObs}{}%
+Landis, J. R., \& Koch, G. G. (1977). The Measurement of Observer
+Agreement for Categorical Data. \emph{Biometrics}, \emph{33}(1),
+159--174. \url{https://doi.org/10.2307/2529310}
+
 \leavevmode\hypertarget{ref-McLeod2021SamAsy}{}%
@@
 \url{https://doi.org/10.1371/journal.pone.0118432}
+
+\leavevmode\hypertarget{ref-Schisterman2005OptCut}{}%
+Schisterman, E. F., Perkins, N. J., Liu, A., \& Bondell, H. (2005).
+Optimal Cut-point and Its Corresponding Youden Index to Discriminate
+Individuals Using Pooled Blood Samples. \emph{Epidemiology},
+\emph{16}(1), 73--81.
+\url{https://doi.org/10.1097/01.ede.0000147512.81966.ba}
 

### 2021-12-13 draft.tex
Deploying to gh-pages from @ PoisotLab/ms_networkary_classifiers@62b5758e463482662104e6b7cbba81cb30dafd48 🚀

@@
 \thispagestyle{empty}
-{\Large\bfseries\sffamily Guidelines for the supervised learning of
+{\Large\bfseries\sffamily Validating machine learning predictions of
 species interactions}
@@
 \begin{enumerate}
-    \item The prediction of species interaction networks is gaining
-momentum as a way to circumvent limitations in data volume. Yet,
-ecological networks are challenging to predict because they are
-typically small and sparse. Dealing with extreme class imbalance is a
-challenge for most binary classifiers, and there are currently no
-guidelines as to how predictive models can be trained.%
+    \item The prediction of species interactions is gaining momentum as
+a way to circumvent limitations in data volume. Yet, ecological networks
+are challenging to predict because they are typically small and sparse.
+Dealing with extreme class imbalance is a challenge for most binary
+classifiers, and there are currently no guidelines as to how predictive
+models can be trained for this specific problem.%
     \item Using simple mathematical arguments and numerical experiments
@@
 connectance is strong. The amount of correction to apply to the training
-dataset depends as a function of the classifier and the network
-connectance.%
+dataset depends on network connectance, on the measure to be optimized,
+and only weakly on the classifier.%
     \item These results reveal that training machines to predict
@@
 
-\begin{itemize}
-\tightlist
-\item
-  example on diagnostic test: rare events are hard to detect even with
-  really good models
-\item
-  summary of model challenges for networks
-\item
-  Strydom et al. (2021) importance of drawing on traits + validation is
-  challenging + comparing across space
-\end{itemize}
+The accuracy paradox is the basis of a number of problems in statistical
+education, and lies in the fact that, when the desired class is rare, a
+model that gets less and less performant will become more and more
+accurate and useful, simply by (i) underpredicting true positive cases
+and (ii) over-predicting false negatives. In other words, accuracy,
+defined as the proportion of predictions that are correct, is often
+useless as a measure of how predictive a model is. This is particularly
+true in ecological networks; the desired class (presence of an
+interaction between two species) is the one we care most about, and by
+far the least commmon. Herein lies the core challenge of predicting
+species interactions: the extreme imbalance between classes makes the
+training of predictive models difficult. In a recent contribution,
+Strydom et al. (2021) highlight that predictive models of interactions
+can likely be improved by adding information (in the form of,
+\emph{e.g.} traits), but that we do not have robust guidelines as to how
+the predictive ability of these models should be evaluated, nor about
+how the models should be trained. Here, by relying on simple derivations
+and a series of simulations, we formulate a number of such guidelines,
+specifically for the case of binary classifiers derived from thresholded
+values.
 
@@
 
-Intro
-
-\hypertarget{confusion-matrix-with-skill-and-bias}{%
-\subsection{Confusion matrix with skill and
-bias}\label{confusion-matrix-with-skill-and-bias}}
-
 In this section, we will assume a network of connectance \(\rho\),
@@
 terms, as we will work on the confusion matrix, which ends up expressing
-\emph{relative} values).
+\emph{relative} values). We will apply skill and bias to this matrix,
+and measure how a selection of performance metrics respond to changes in
+these values, in order to assess their suitability for model evaluation.
+
+\hypertarget{confusion-matrix-with-skill-and-bias}{%
+\subsection{Confusion matrix with skill and
+bias}\label{confusion-matrix-with-skill-and-bias}}
 

### 2021-12-13 preprint.tex
Deploying to gh-pages from @ PoisotLab/ms_networkary_classifiers@62b5758e463482662104e6b7cbba81cb30dafd48 🚀

@@
 
-\noindent {\LARGE \textbf{\textsf{Guidelines for the supervised learning
-of species interactions}}}
+\noindent {\LARGE \textbf{\textsf{Validating machine learning
+predictions of species interactions}}}
 

### 2021-12-13 draft.tex
Deploying to gh-pages from @ PoisotLab/ms_networkary_classifiers@2305e0d8834b4252a839573bb30ef9fe5525ff75 🚀

@@
 
-Informedness (\textbf{Youden1950IndRat?}) (also known as bookmaker
-informedness or the True Skill Statistic) is
-\(\text{TPR}+\text{TNR}-1\), where \(\text{TPR}= tp/(tp+fn)\) and
-\(\text{TNR} = tn/(tn+fp)\); informedness can be used to find the
-optimal cutpoint in thresholding analyses (Schisterman et al., 2005).
-The MCC is defined as
+Informedness (Youden, 1950) (also known as bookmaker informedness or the
+True Skill Statistic) is \(\text{TPR}+\text{TNR}-1\), where
+\(\text{TPR}= tp/(tp+fn)\) and \(\text{TNR} = tn/(tn+fp)\); informedness
+can be used to find the optimal cutpoint in thresholding analyses
+(Schisterman et al., 2005). The MCC is defined as
 
@@
 
+\leavevmode\hypertarget{ref-Youden1950IndRat}{}%
+Youden, W. J. (1950). Index for rating diagnostic tests. \emph{Cancer},
+\emph{3}(1), 32--35.
+\url{https://doi.org/10.1002/1097-0142(1950)3:1\%3C32::AID-CNCR2820030106\%3E3.0.CO;2-3}
+
 \end{CSLReferences}

### 2021-12-13 draft.tex
Deploying to gh-pages from @ PoisotLab/ms_networkary_classifiers@2308ff5b6b0725a98a3dfe4d0d5fd9f0f24228b9 🚀

@@
 \[
-\frac{tp\times tn - fn\times fp}{\srt{(tp+fp)\times (tp+fn)\times (tn+fp)\times (tn+fn)}} \,.
+\frac{tp\times tn - fn\times fp}{\sqrt{(tp+fp)\times (tp+fn)\times (tn+fp)\times (tn+fn)}} \,.
 \]
@@
 the Random Forest was ultimately able to get a correct estimate of
-network structure tbl.~\ref{tbl:comaprison}, it ultimately remains a
+network structure tbl.~\ref{tbl:comparison}, it ultimately remains a
 poor classifier, as evidenced by its low PR-AUC. This suggests that the

### 2022-01-10 draft.tex
Deploying to gh-pages from @ PoisotLab/ms_networkary_classifiers@461578cc1bc819deaaa27cf8d194d3fe39ed926d 🚀

@@
 \thispagestyle{empty}
-{\Large\bfseries\sffamily Validating machine learning predictions of
-species interactions}
+{\Large\bfseries\sffamily Guidelines for the validation of machine
+learning predictions of species interactions}
 \vskip 5em
@@
 unbiasing to apply to the training dataset.%
-    \item Classifier accuracy and the ROC-AUC are not informative
+    \item Neither classifier accuracy nor the ROC-AUC are informative
 measures for the performance of interaction prediction. PR-AUC is a
-fairer assesment of performance. In some cases, even standard measures
+fairer assessment of performance. In some cases, even standard measures
 can lead to selecting a more biased classifier because the effect of
@@
 species interactions: the extreme imbalance between classes makes the
-training of predictive models difficult. In a recent contribution,
-Strydom et al. (2021) highlight that predictive models of interactions
-can likely be improved by adding information (in the form of,
-\emph{e.g.} traits), but that we do not have robust guidelines as to how
-the predictive ability of these models should be evaluated, nor about
-how the models should be trained. Here, by relying on simple derivations
-and a series of simulations, we formulate a number of such guidelines,
-specifically for the case of binary classifiers derived from thresholded
-values.
+training of predictive models difficult. The connectance of empirical
+networks is usually well under 20\%, with larger networks having a lower
+connectance (MacDonald et al., 2020). Recent contributions (Becker et
+al., 2021; Strydom et al., 2021) highlight that predictive models of
+interactions can likely be improved by adding information (in the form
+of, \emph{e.g.} traits), but that we do not have robust guidelines as to
+how the predictive ability of these models should be evaluated, nor
+about how the models should be trained. Here, by relying on simple
+derivations and a series of simulations, we formulate a number of such
+guidelines, specifically for the case of binary classifiers derived from
+thresholded values.
 
@@
 
-There is an immense amount of measures to evaluate the performance of
+There is an immense diversity of measures to evaluate the performance of
 classification tasks (Ferri et al., 2009). Here we will focus on five of
 them with high relevance for imbalanced learning (He \& Ma, 2013); three
-threshold metrics (\(\kappa\), informedness, and the Matthews
-Correlation Coefficient), and two ranking metrics (the ROC-AUC and
-PR-AUC). The \(\kappa\) measure of agreement (Landis \& Koch, 1977)
-establishes the extent to which two observers (here the data and the
-prediction) agree, and is measured as
+threshold metrics (\(\kappa\), informedness, and MMC, the Matthews
+Correlation Coefficient), and two ranking metrics (the areas under the
+Receiving Operator Characteristic and the Precision-Recall curves; resp.
+ROC-ACU and PR-AUC). The \(\kappa\) measure of agreement (Landis \&
+Koch, 1977) establishes the extent to which two observers (here the data
+and the prediction) agree, and is measured as
 
@@
 A lot of binary classifiers are built by using a regressor (whose task
-is to guess the value of the interaction, amd can therefore return
-somethins considered to be a pseudo-probability); in this case, the
-optimal value below which predicitions are assumed to be negative
-(\emph{i.e.} the interaction does not exist) can be determined by
-picking a threshold maximizing some value on the ROC curve or the PR
-curve. The area under these curves (ROC-AUC and PR-AUC henceforth) give
-ideas on the overall goodness of the classifier. Saito \& Rehmsmeier
-(2015) established that the ROC-AUC is biased towards over-estimating
-performance for imbalanced data; on the contrary, the PR-AUC is able to
-identify classifiers that are less able to detect positive interactions
-correctly, with the additional advantage of having a baseline value
-equal to prevalence. Therefore, it is important to assess whether these
-two measures return different results when applied to ecological network
-prediction. The ROC curve is defined by the false positive rate on the
-\(x\) axis, and the true positive rate on the \(y\) axis, and the PR
-curve is defined by the true positive rate on the \(x\) axis, and the
-positive predictive value on the \(y\) axis. By comparison with the
-previous paragraph, it is obvious that \(F_1\) has ties to the PR curve
-(being close to the expected PR-AUC), and that informedness has ties to
-the ROC curve (whereby the threshold maximizing informedness is also the
-point of maximal inflection on the ROC curve). One important difference
-between ROC and PR is that the later does not prominently account for
-the size of the true negative compartments: in short, it is more
-sensitive to the correct positive predictions. In a context of strong
-imbalance, PR-AUC is therefore a more stringent test of model
-performance.
+is to guess the value of the interaction, and can therefore return a
+value considered to be a pseudo-probability); in this case, the optimal
+value below which predictions are assumed to be negative (\emph{i.e.}
+the interaction does not exist) can be determined by picking a threshold
+maximizing some value on the ROC curve or the PR curve. The area under
+these curves (ROC-AUC and PR-AUC henceforth) give ideas on the overall
+goodness of the classifier. Saito \& Rehmsmeier (2015) established that
+the ROC-AUC is biased towards over-estimating performance for imbalanced
+data; on the contrary, the PR-AUC is able to identify classifiers that
+are less able to detect positive interactions correctly, with the
+additional advantage of having a baseline value equal to prevalence.
+Therefore, it is important to assess whether these two measures return
+different results when applied to ecological network prediction. The ROC
+curve is defined by the false positive rate on the \(x\) axis, and the
+true positive rate on the \(y\) axis, and the PR curve is defined by the
+true positive rate on the \(x\) axis, and the positive predictive value
+on the \(y\) axis. By comparison with the previous paragraph, it is
+obvious that \(F_1\) has ties to the PR curve (being close to the
+expected PR-AUC), and that informedness has ties to the ROC curve
+(whereby the threshold maximizing informedness is also the point of
+maximal inflection on the ROC curve). One important difference between
+ROC and PR is that the later does not prominently account for the size
+of the true negative compartments: in short, it is more sensitive to the
+correct positive predictions. In a context of strong imbalance, PR-AUC
+is therefore a more stringent test of model performance.
 
@@
 models (SDMs). Indeed, the training and evaluation of SDMs as binary
-classifiers suffers from the same issue of low prevalence. In previous
-work, Allouche et al. (2006) suggested that \(\kappa\) was a better test
-of model performance than the True Skill Statistic (TSS; which we refer
-to as Youden's informedness); these conclusions were later criticized by
+classifiers suffers from the same issue of low prevalence; this is not
+surprising that the two fields (SDMs and network predictions) would
+share methods and their attached conceptual issues, as they suffer from
+data limitations, class imbalance, and the conversion of quantitative
+prediction into a binary classification. In previous work, Allouche et
+al. (2006) suggested that \(\kappa\) was a better test of model
+performance than the True Skill Statistic (TSS; which we refer to as
+Youden's informedness); these conclusions were later criticized by
 Somodi et al. (2017), who emphasized that informedness' relationship to
@@
 classifiers applied to food web data will reach a high accuracy; this is
-because the measure is dominated by the accidental correct predictions
+because the measure is dominated by the accidentally correct predictions
 of negatives. On simulated confusion matrices with ranges of imbalance
@@
 behavior, and informedness is a linear measure of classifier skill. By
-performing simulations with four models and an ensemble,we show that
+performing simulations with four models and an ensemble, we show that
 informedness and ROC-AUC are consistently high on network data, and that
@@
 richness), and \((1-\rho) S^2\) non-interactions. Therefore, the vector
-describing the \emph{true} state of the network is a column vector
+describing the \emph{true} state of the network (assumed to be an
+unweighted, directed network) is a column vector
 \(\mathbf{o}^T = [\rho (1-\rho)]\) (we can safely drop the \(S^2\)
@@
 for data-inflation (\emph{i.e.} the prediction of new interactions), it
-is not necessarilly a bad thing in practice to select models that
-predict more interactions than the original dataset, because the
-original dataset misses some interactions. Furthermore, the weight of
-positive interactions could be adjusted if some information about the
-extent of undersampling exists (\emph{e.g.} Branco et al., 2015).
+is not necessarily a bad thing in practice to select models that predict
+more interactions than the original dataset, because the original
+dataset misses some interactions. Furthermore, the weight of positive
+interactions could be adjusted if some information about the extent of
+undersampling exists (\emph{e.g.} Branco et al., 2015). In a recent
+large-scale imputation of interactions in the mammal-virus networks,
+Poisot et al. (2021) for example estimated that 93\% of interactions are
+yet to be documented.
 
@@
 order to make the problem more difficult to solve, we use \([v_i, h_j]\)
-as a feature vector, and therefore the models will have to uncover that
-the rule for interaction is \(\text{abs}(v_i, h_j) \le \xi\).
+as a feature vector (\emph{i.e.} the traits on which the models are
+trained), and therefore the models will have to uncover that the rule
+for interaction is \(\text{abs}(v_i, h_j) \le \xi\).
 
@@
 dataset. As we use 64000 such datasets, this would require 256000 unique
-instances of tweaking thehyperparameters, which is not realistic.
+instances of tweaking the hyperparameters, which is not realistic.
 Therefore, we assume that the default parameterizations are comparable
 across networks. All machines return a quantitative prediction, usually
-(but not necessarilly) in \([0,1]\), which is proportional (but not
-necessarilly linearly) to the probability of an interaction between
-\(i\) and \(j\).
+(but not necessarily) in \([0,1]\), which is proportional (but not
+necessarily linearly) to the probability of an interaction between \(i\)
+and \(j\).
 
@@
 fig.~\ref{fig:biasrocpr}, it seems that informedness and ROC-AUC are not
-necessarilly able to discriminate between good and bad classifiers
+necessarily able to discriminate between good and bad classifiers
 (although this result may be an artifact for informedness, as it has
@@
 PR-AUC, suggesting it hits high recall at the cost of low precision.
-Although the ensemble was about to reach a very high PR-AUC (and a very
-high ROC-AUC), this did not necessarilly translate into more accurate
+Although the ensemble was able to reach a very high PR-AUC (and a very
+high ROC-AUC), this did not necessarily translate into more accurate
 reconstructions of the structure of the network. This result bears
@@
 scoring predictor of interactions (the ensemble model) had not
-necessarilly the best prediction of network structure. The tasks of
+necessarily the best prediction of network structure. The tasks of
 predicting networks structure and of predicting interactions within
@@
 structure. How these two approaches can be reconciled is undoubtedly a
-task for further research. Despite this apprent tension at the heart of
+task for further research. Despite this apparent tension at the heart of
 the predictive exercise, we can use the results presented here to
@@
 through the Narval general purpose cluster. TP is supported by a NSERC
-Discovery Grant and Discovery Acceleration Supplement, and by a grant
-from the Institut de Valorisation des Données (IVADO).
+Discovery Grant and Discovery Acceleration Supplement, by funding to the
+Viral Emergence Research Initiative (VERENA) consortium including NSF
+BII 2021909, and by a grant from the Institut de Valorisation des
+Données (IVADO).
 
@@
 \emph{Oikos}, \emph{n/a}(n/a). \url{https://doi.org/10.1111/oik.08650}
+
+\leavevmode\hypertarget{ref-Poisot2021ImpMam}{}%
+Poisot, T., Ouellet, M.-A., Mollentze, N., Farrell, M. J., Becker, D.
+J., Albery, G. F., Gibb, R. J., Seifert, S. N., \& Carlson, C. J. (2021,
+May 31). \emph{Imputing the mammalian virome with linear filtering and
+singular value decomposition}. \url{http://arxiv.org/abs/2105.14973}
 

### 2022-01-10 preprint.tex
Deploying to gh-pages from @ PoisotLab/ms_networkary_classifiers@461578cc1bc819deaaa27cf8d194d3fe39ed926d 🚀

@@
 
-\noindent {\LARGE \textbf{\textsf{Validating machine learning
-predictions of species interactions}}}
+\noindent {\LARGE \textbf{\textsf{Guidelines for the validation of
+machine learning predictions of species interactions}}}
 

### 2022-01-10 draft.tex
Deploying to gh-pages from @ PoisotLab/ms_networkary_classifiers@ce958bac6b767933b6a0914f98b693d70979a3d6 🚀

@@
 species interactions: the extreme imbalance between classes makes the
-training of predictive models difficult. The connectance of empirical
-networks is usually well under 20\%, with larger networks having a lower
-connectance (MacDonald et al., 2020). Recent contributions (Becker et
-al., 2021; Strydom et al., 2021) highlight that predictive models of
-interactions can likely be improved by adding information (in the form
-of, \emph{e.g.} traits), but that we do not have robust guidelines as to
-how the predictive ability of these models should be evaluated, nor
-about how the models should be trained. Here, by relying on simple
-derivations and a series of simulations, we formulate a number of such
-guidelines, specifically for the case of binary classifiers derived from
-thresholded values.
+training of predictive models difficult, and their validation even more
+so as we do not reliably know which negatives are true. The connectance
+of empirical networks is usually well under 20\%, with larger networks
+having a lower connectance (MacDonald et al., 2020). Recent
+contributions (Becker et al., 2021; Strydom et al., 2021) highlight that
+predictive models of interactions can likely be improved by adding
+information (in the form of, \emph{e.g.} traits), but that we do not
+have robust guidelines as to how the predictive ability of these models
+should be evaluated, nor about how the models should be trained. Here,
+by relying on simple derivations and a series of simulations, we
+formulate a number of such guidelines, specifically for the case of
+binary classifiers derived from thresholded values.
 
@@
 sensitive in a consistent way. Informedness is appropriate at
-discriminating between different skills, but confounded by bias. \(F_1\)
-is increasing with bias, and should not be prioritized to evalue the
-performance of the model. The discussion of sensitivity to bias should
-come with a domain-specific caveat: although it is likely that
-interactions documented in ecological networks are correct, a lot of
-non-interactions are simply unobserved; as predictive models are used
-for data-inflation (\emph{i.e.} the prediction of new interactions), it
-is not necessarily a bad thing in practice to select models that predict
-more interactions than the original dataset, because the original
-dataset misses some interactions. Furthermore, the weight of positive
-interactions could be adjusted if some information about the extent of
-undersampling exists (\emph{e.g.} Branco et al., 2015). In a recent
-large-scale imputation of interactions in the mammal-virus networks,
-Poisot et al. (2021) for example estimated that 93\% of interactions are
-yet to be documented.
+discriminating between different skills, but confounded by bias. As both
+of these measures bring valuable information on the model behavior, we
+will retain them for future analyses. \(F_1\) is increasing with bias,
+and should not be prioritized to evalue the performance of the model.
+The discussion of sensitivity to bias should come with a domain-specific
+caveat: although it is likely that interactions documented in ecological
+networks are correct, a lot of non-interactions are simply unobserved;
+as predictive models are used for data-inflation (\emph{i.e.} the
+prediction of new interactions), it is not necessarily a bad thing in
+practice to select models that predict more interactions than the
+original dataset, because the original dataset misses some interactions.
+Furthermore, the weight of positive interactions could be adjusted if
+some information about the extent of undersampling exists (\emph{e.g.}
+Branco et al., 2015). In a recent large-scale imputation of interactions
+in the mammal-virus networks, Poisot et al. (2021) for example estimated
+that 93\% of interactions are yet to be documented.
 
@@
 results of the previous section, \emph{i.e.} they are more stringent
-tests of model performance.
+tests of model performance. These results suggest that learning from a
+dataset with very low connectance can be a different task than for more
+connected networks: it becomes increasingly important to caputre the
+mechanisms that make an interaction \emph{exist}, and therefore having a
+slightly more biased training dataset might be beneficial. As
+connectance increases, the need for biased training sets is less
+prominent, as learning the rules for which interactions \emph{do not}
+exist starts gaining importance.
 
@@
 
-When trained at their optimal training set bias, performance still had a
+When trained at their optimal training set bias, connectance still had a
 significant impact on the performance of some machines
@@
 Iroquoian, Anishinabewaki, Mohawk, Huron-Wendat, and Omàmiwininiwak
-nations. This research was enabled in part by support provided by Calcul
-Québec (www.calculquebec.ca) and Compute Canada (www.computecanada.ca)
-through the Narval general purpose cluster. TP is supported by a NSERC
-Discovery Grant and Discovery Acceleration Supplement, by funding to the
-Viral Emergence Research Initiative (VERENA) consortium including NSF
-BII 2021909, and by a grant from the Institut de Valorisation des
-Données (IVADO).
+nations. We thank Colin J. Carlson, Michael D. Catchen, Giulio Valentino
+Dalla Riva, and Tanya Strydom for inputs on earlier versions of this
+manuscript. This research was enabled in part by support provided by
+Calcul Québec (www.calculquebec.ca) and Compute Canada
+(www.computecanada.ca) through the Narval general purpose cluster. TP is
+supported by a NSERC Discovery Grant and Discovery Acceleration
+Supplement, by funding to the Viral Emergence Research Initiative
+(VERENA) consortium including NSF BII 2021909, and by a grant from the
+Institut de Valorisation des Données (IVADO).
 

### 2022-01-10 draft.tex
Deploying to gh-pages from @ PoisotLab/ms_networkary_classifiers@af265955124089f8d193e322f084cf61a3e86e08 🚀

@@
 \includegraphics{figures/bias_mcc_inf.png}
-\caption{TODO}\label{fig:biasmccinf}
+\caption{Response of MCC and Informedness to changes in the training set
+bias for a fixed connectance (rows). Both of these values approach 1 for
+a good model. Informedness is consistently high, and by contrast, MCC
+increases with additional training set bias. Across all models, training
+on a more connected network is easier.}\label{fig:biasmccinf}
 }
@@
 \includegraphics{figures/bias_roc_pr.png}
-\caption{TODO}\label{fig:biasrocpr}
+\caption{Response of ROC-AUC and PR-AUC to changes in the training set
+bias for a fixed connectance (rows). ROC-AUC is consistently high, and
+therefore not properly able to separate good from poor classifiers. On
+the other hand, PR-AUC responds to changes in the training set. As in
+fig.~\ref{fig:biasmccinf}, training on more connected networks is
+easier.}\label{fig:biasrocpr}
 }
@@
 \includegraphics{figures/optim_bias.png}
-\caption{TODO}\label{fig:optimbias}
+\caption{Value of the optimal training set bias for the different models
+and measures evaluated here, over a range of connectances. Informedness
+was reliably maximized for balanced training sets, and kept this
+behavior across models. For other measures, larger connectances in the
+true network allowed lower biases in the training set. In a large number
+of cases, ``over-correcting'' by having training sets with more than
+half instances representing interactions would maximize the values of
+the model performance measures.}\label{fig:optimbias}
 }
@@
 \includegraphics{figures/optim_perf.png}
-\caption{TODO}\label{fig:optimperf}
+\caption{When trained on their optimally biased training set, most
+models were able to maximize their performance; this is not true for
+decision tree, which had a very low PR-AUC, and to some extent for ridge
+regression who had a slow increase with network connectance. The
+ensemble had a consistently high performance despite incorporating poor
+models.}\label{fig:optimperf}
 }
@@
 \includegraphics{figures/valid_ensemble.png}
-\caption{TODO}\label{fig:ecovalid}
+\caption{Visualisation of the models predictions for one instance of a
+network prediction problem (shown in the ``Dataset'' panel). This figure
+reveals how inspecting the details of the prediction is important:
+indeed, although the performance measures hint at the fact that ridge
+regression is mediocre, this figure reveals that it is making
+predictions that correspond to a network with an entirely different
+topology (namely, nested as opposed to diagonal).}\label{fig:ecovalid}
 }

### 2022-01-10 draft.tex
Deploying to gh-pages from @ PoisotLab/ms_networkary_classifiers@026deab5c8025bdff8bbbfe770cdde8d08049f2d 🚀

@@
 unweighted, directed network) is a column vector
-\(\mathbf{o}^T = [\rho (1-\rho)]\) (we can safely drop the \(S^2\)
+\(\mathbf{o}^T = [\rho, (1-\rho)]\) (we can safely drop the \(S^2\)
 terms, as we will work on the confusion matrix, which ends up expressing
@@
 \(\mathbf{M}\) for a no-skill classifier is given by the element-wise
-product of these vectors \(\mathbf{o} \odot \mathbf{p}\), \emph{i.e.}
+(Hadamard, outer) product of these vectors
+\(\mathbf{o} \odot \mathbf{p}\), \emph{i.e.}
 

### 2022-02-25 draft.tex
Deploying to gh-pages from @ PoisotLab/ms_networkary_classifiers@503c15f72088d6e5443c3b54bbd5062fe51628d3 🚀

@@
 so as we do not reliably know which negatives are true. The connectance
-of empirical networks is usually well under 20\%, with larger networks
-having a lower connectance (MacDonald et al., 2020). Recent
-contributions (Becker et al., 2021; Strydom et al., 2021) highlight that
-predictive models of interactions can likely be improved by adding
-information (in the form of, \emph{e.g.} traits), but that we do not
-have robust guidelines as to how the predictive ability of these models
-should be evaluated, nor about how the models should be trained. Here,
-by relying on simple derivations and a series of simulations, we
-formulate a number of such guidelines, specifically for the case of
-binary classifiers derived from thresholded values.
+(the proportion of realized interactions, usually the number of
+interactions divided by the number of species pairs) of empirical
+networks is usually well under 20\%, with larger networks having a lower
+connectance (MacDonald et al., 2020). Recent contributions (Strydom et
+al., 2021; \textbf{Becker2021OptPre?}) highlight that predictive models
+of interactions can likely be improved by adding information (in the
+form of, \emph{e.g.} traits), but that we do not have robust guidelines
+as to how the predictive ability of these models should be evaluated,
+nor about how the models should be trained. Here, by relying on simple
+derivations and a series of simulations, we formulate a number of such
+guidelines, specifically for the case of binary classifiers derived from
+thresholded values.
 
@@
 can be used to find the optimal cutpoint in thresholding analyses
-(Schisterman et al., 2005). The MCC is defined as
+(Schisterman et al., 2005). The formula for informedness is
+
+\[\frac{tp}{tp+fn}+\frac{tn}{tn+fp}-1\,.\]
+
+The MCC is defined as
 
@@
 comparison of models, it doesn't establishes baselines or good practices
-for training on imbalanced ecological data. Within the context of
-networks, there are three specific issues that need to be adressed.
-First, what values of performance measures are we expecting for a
-classifier that has poor performance? This is particularly important as
-it can evaluate whether low prevalence can lull us into a false sense of
-predictive accuracy. Second, independently of the question of model
-evaluation, is low prevalence an issue for \emph{training}, and can we
-remedy it? Finally, because the low amount of data on interaction makes
-a lot of imbalance correction methods (see \emph{e.g.} Branco et al.,
-2015) hard to apply, which indicators can be optimized with the least
-amount of positive interaction data?
+for training on imbalanced ecological data. Steen et al. (2021) show
+that, when applying spatial thinning (a process that has no analogue in
+networks), the best approach to train ML-based SDMs varies according to
+the balancing of the dataset, and the evaluation measures used. This
+suggests that there is no single ``recipe'' that is guaranteed to give
+the best model. Within the context of networks, there are three specific
+issues that need to be adressed. First, what values of performance
+measures are we expecting for a classifier that has poor performance?
+This is particularly important as it can evaluate whether low prevalence
+can lull us into a false sense of predictive accuracy. Second,
+independently of the question of model evaluation, is low prevalence an
+issue for \emph{training}, and can we remedy it? Finally, because the
+low amount of data on interaction makes a lot of imbalance correction
+methods (see \emph{e.g.} Branco et al., 2015) hard to apply, which
+indicators can be optimized with the least amount of positive
+interaction data?
 
@@
 True-Skill Statistic) when the imbalance in the dataset may not be
-representative (Jordano, 2016a, which is the case as networks are
-under-sampled; 2016b), when classifiers need to be compared across
-different datasets (for example when predicting a system in space, where
-undersampling varies locally; McLeod et al., 2021), and when comparing
-the results to a no-skill (baseline) classifier is important. As these
-conditions are likely to be met with network data, there is a need to
-evaluate which measures of classification accuracy respond in a
-desirable way.
+representative, which is the case as species interaction networks are
+often under-sampled (Jordano, 2016b, 2016a), when classifiers need to be
+compared across different datasets (for example when predicting a system
+in space, where undersampling varies locally; McLeod et al., 2021), and
+when comparing the results to a no-skill (baseline) classifier is
+important. As these conditions are likely to be met with network data,
+there is a need to evaluate which measures of classification accuracy
+respond in a desirable way.
 
@@
 exploration of the values of \(\text{logit}(s)\) and \(\text{logit}(b)\)
-linearly from \(-10\) to \(10\), of \(\rho\) linearly in \(]0, 0.5]\),
-which is within the range of usually observed connectance values for
-empirical food webs. Note that at this point, there is no food web model
-to speak of; rather, the confusion matrix we discuss can be obtained for
-any classification task. Based on the previous discussion, the desirable
+linearly from \(-10\) to \(10\); \(\text{logit}(x) = -10\) means that
+\(x\) is essentially 0, and \(\text{logit}(x) = 10\) means it is
+essentially 1 -- this choice was motivated by the fact that most
+responses are non-linear with regards to bias and skill. The values or
+\(\rho\) were taken linearly in \(]0, 0.5]\), which is within the range
+of usually observed connectance values for empirical species interaction
+networks. Note that at this point, there is no food web model to speak
+of; rather, the confusion matrix we discuss can be obtained for any
+classification task. Based on the previous discussion, the desirable
 properties for a measure of classifier success should be: an increase
@@
 sum of ranged outputs from the component models) on 30\% of the
-interaction data. Networks are generated by picking a random
-infectiousness trait \(v_i\) for 100 species (from a \(B(6,8)\)
-distribution), and a resistance trait \(h_j\) for 100 species (from a
-\(B(2,8)\) distribution). There is an interaction between \(i\) and
-\(j\) when \(v_i-\xi/2 \le h_j \le v_i+\xi/2\), where \(\xi\) is a
-constant regulating the connectance of the network (there is an almost
-1:1 relationship between \(\xi\) and connectance), and varies uniformly
-in \([0.05, 0.35]\). This model gives fully interval networks that are
-close analogues to the bacteria--phage model of Weitz et al. (2005),
-with both a modular structure and a non-uniform degree distribution.
-This model is easy to learn: when trained with features
-\([v_i, h_j, \text{abs}(v_i, h_j)] ^T\) to predict the interactions
-between \(i\) and \(j\), all four models presented below were able to
-reach almost perfect predictions all the time (data not presented here)
--- this is in part because the rule is fixed for all interactions. In
-order to make the problem more difficult to solve, we use \([v_i, h_j]\)
-as a feature vector (\emph{i.e.} the traits on which the models are
-trained), and therefore the models will have to uncover that the rule
-for interaction is \(\text{abs}(v_i, h_j) \le \xi\).
+interaction data. In practice, testing usually uses 70\% of the total
+data; for ecological networks, where interactions are sparse \emph{and}
+the number of species is low, this may not be the best solution, as the
+testing set becomes constrained not by the \emph{proportion} of
+interactions, but by their \emph{number}. Preliminary experiments using
+different splits revealed no qualitative change in the results. Networks
+are generated by picking a random infectiousness trait \(v_i\) for 100
+species (from a beta distribution \(B(\alpha=6,\beta=8)\) distribution),
+and a resistance trait \(h_j\) for 100 species (from
+\(B(\alpha=2,\beta=8)\) distribution). There is an interaction between
+\(i\) and \(j\) when \(v_i-\xi/2 \le h_j \le v_i+\xi/2\), where \(\xi\)
+is a constant regulating the connectance of the network (there is an
+almost 1:1 relationship between \(\xi\) and connectance), and varies
+uniformly in \([0.05, 0.35]\). This model gives fully interval networks
+that are close analogues to the bacteria--phage model of Weitz et al.
+(2005), with both a modular structure and a non-uniform degree
+distribution. This dataset is easy for almost any algorithm to learn:
+when trained with features \([v_i, h_j, \text{abs}(v_i, h_j)] ^T\) to
+predict the interactions between \(i\) and \(j\), all four models
+presented below were able to reach almost perfect predictions all the
+time (data not presented here) -- this is in part because the rule
+(there is maximum value of the distance between traits for which there
+is an interaction) is fixed for all interactions. In order to make the
+problem more difficult to solve, we use \([v_i, h_j]\) as a feature
+vector (\emph{i.e.} the traits on which the models are trained), and
+therefore the models will have to uncover that the rule for interaction
+is \(\text{abs}(v_i, h_j) \le \xi\). The models therefore all have the
+form
+
+\[
+\begin{bmatrix}
+           x_{1} \\
+           x_{2} \\
+           \vdots \\
+           x_{m}
+         \end{bmatrix}
+\]
 
@@
 do not expect that ensembles will \emph{always} be better than single
-models. In a recent multi-model comparison, Becker et al. (2021) found
-that the ensemble was \emph{not} the best model. There is no general
-conclusion to draw from this besides reinforcing the need to be
+models. In a recent multi-model comparison, (\textbf{Becker2021OptPre?})
+found that the ensemble was \emph{not} the best model. There is no
+general conclusion to draw from this besides reinforcing the need to be
 pragmatic about which models should be included in the ensemble, or
@@
 
-\leavevmode\hypertarget{ref-Becker2021OptPre}{}%
-Becker, D., Albery, G. F., Sjodin, A. R., Poisot, T., Bergner, L.,
-Dallas, T., Eskew, E. A., Farrell, M. J., Guth, S., Han, B. A., Simmons,
-N. B., Stock, M., Teeling, E. C., \& Carlson, C. J. (2021). Optimizing
-predictive models to prioritize viral discovery in zoonotic reservoirs.
-\emph{bioRxiv}, 2020.05.22.111344.
-\url{https://doi.org/10.1101/2020.05.22.111344}
-
 \leavevmode\hypertarget{ref-Bezanson2017JulFre}{}%
@@
 \leavevmode\hypertarget{ref-Blaom2020FleMod}{}%
-Blaom, A. D., \& Vollmer, S. J. (2020, December 31). \emph{Flexible
-model composition in machine learning and its implementation in MLJ}.
-\url{http://arxiv.org/abs/2012.15505}
+Blaom, A. D., \& Vollmer, S. J. (2020). Flexible model composition in
+machine learning and its implementation in MLJ. \emph{arXiv:2012.15505
+{[}cs{]}}. \url{http://arxiv.org/abs/2012.15505}
 
@@
 \leavevmode\hypertarget{ref-Branco2015SurPre}{}%
-Branco, P., Torgo, L., \& Ribeiro, R. (2015, May 13). \emph{A Survey of
-Predictive Modelling under Imbalanced Distributions}.
-\url{http://arxiv.org/abs/1505.01658}
+Branco, P., Torgo, L., \& Ribeiro, R. (2015). A Survey of Predictive
+Modelling under Imbalanced Distributions. \emph{arXiv:1505.01658
+{[}cs{]}}. \url{http://arxiv.org/abs/1505.01658}
 
@@
 Jeni, L. A., Cohn, J. F., \& De La Torre, F. (2013). Facing Imbalanced
-Data--Recommendations for the Use of Performance Metrics. \emph{2013
+DataRecommendations for the Use of Performance Metrics. \emph{2013
 Humaine Association Conference on Affective Computing and Intelligent
@@
 Poisot, T., Ouellet, M.-A., Mollentze, N., Farrell, M. J., Becker, D.
-J., Albery, G. F., Gibb, R. J., Seifert, S. N., \& Carlson, C. J. (2021,
-May 31). \emph{Imputing the mammalian virome with linear filtering and
-singular value decomposition}. \url{http://arxiv.org/abs/2105.14973}
+J., Albery, G. F., Gibb, R. J., Seifert, S. N., \& Carlson, C. J.
+(2021). Imputing the mammalian virome with linear filtering and singular
+value decomposition. \emph{arXiv:2105.14973 {[}q-Bio{]}}.
+\url{http://arxiv.org/abs/2105.14973}
 
@@
 \leavevmode\hypertarget{ref-Somodi2017PreDep}{}%
-Somodi, I., Lepesi, N., \& Botta‐Dukát, Z. (2017). Prevalence dependence
+Somodi, I., Lepesi, N., \& Botta-Dukát, Z. (2017). Prevalence dependence
 in model goodness measures with special emphasis on true skill
@@
 \url{https://doi.org/10.1002/ece3.2654}
+
+\leavevmode\hypertarget{ref-Steen2021SpaThi}{}%
+Steen, V. A., Tingley, M. W., Paton, P. W. C., \& Elphick, C. S. (2021).
+Spatial thinning and class balancing: Key choices lead to variation in
+the performance of species distribution models with citizen science
+data. \emph{Methods in Ecology and Evolution}, \emph{12}(2), 216--226.
+\url{https://doi.org/10.1111/2041-210X.13525}
 

### 2022-02-26 draft.tex
Deploying to gh-pages from @ PoisotLab/ms_networkary_classifiers@bd34c7c24c585e30759086973bd04b941ba011bd 🚀

@@
 
-The accuracy paradox is the basis of a number of problems in statistical
-education, and lies in the fact that, when the desired class is rare, a
-model that gets less and less performant will become more and more
-accurate and useful, simply by (i) underpredicting true positive cases
-and (ii) over-predicting false negatives. In other words, accuracy,
-defined as the proportion of predictions that are correct, is often
-useless as a measure of how predictive a model is. This is particularly
-true in ecological networks; the desired class (presence of an
-interaction between two species) is the one we care most about, and by
-far the least commmon. Herein lies the core challenge of predicting
-species interactions: the extreme imbalance between classes makes the
-training of predictive models difficult, and their validation even more
-so as we do not reliably know which negatives are true. The connectance
-(the proportion of realized interactions, usually the number of
-interactions divided by the number of species pairs) of empirical
-networks is usually well under 20\%, with larger networks having a lower
-connectance (MacDonald et al., 2020). Recent contributions (Strydom et
-al., 2021; \textbf{Becker2021OptPre?}) highlight that predictive models
-of interactions can likely be improved by adding information (in the
-form of, \emph{e.g.} traits), but that we do not have robust guidelines
-as to how the predictive ability of these models should be evaluated,
-nor about how the models should be trained. Here, by relying on simple
-derivations and a series of simulations, we formulate a number of such
-guidelines, specifically for the case of binary classifiers derived from
-thresholded values.
-
-Binary classifiers are usually assessed by measuring properties of their
-confusion matrix, \emph{i.e.} the contingency table reporting true/false
-positive/negative hits. A confusion matrix is laid out as
+Ecological networks are a backbone for key ecological and evolutionary
+processes; yet enumerating all of the interactions they contain is a
+daunting task, as it scales with \(S^2\), \emph{i.e.} the squared
+species richness TODO. Recent contributions to the field of ecological
+network prediction (Strydom et al., 2021; \textbf{Becker2021OptPre?})
+highlight that although interactions can be predicted by adding
+ecologically relevant information (in the form of, \emph{e.g.} traits),
+we do not have robust guidelines as to how the predictive ability of
+these models should be evaluated, nor about how the models should be
+trained. Here, by relying on simple derivations and a series of
+simulations, we formulate a number of such guidelines, specifically for
+the case of binary classifiers derived from thresholded values.
+Specifically, we conduct an investigation of the models in terms of
+their skill (ability to make the right prediction), bias (trends towards
+systematically over-predicting one class), class imbalance (the relative
+number of cases representing interactions), and show how these effects
+interact. We conclude on the fact that models with the best
+interaction-scale predictive score do not necessarily result in the most
+accurate representation of the network.
+
+The prediction of ecological interactions shares conceptual and
+methodological issues with two fields in biology: species distribution
+models (SDMs), and genomics. SDMs suffers from issues affecting
+interactions prediction, namely low prevalence (due to sparsity of
+observations/interactions) and data aggregation (due to bias in sampling
+some locations/species). In previous work, Allouche et al. (2006)
+suggested that \(\kappa\) was a better test of model performance than
+the True Skill Statistic (TSS; which we refer to as Youden's
+informedness); these conclusions were later criticized by Somodi et al.
+(2017), who emphasized that informedness' is affected both by prevalence
+and bias. Although this work offers recommendations about the comparison
+of models, it doesn't establishes baselines or good practices for
+training on imbalanced ecological data, or ways to remedy the imbalance.
+Steen et al. (2021) show that, when applying spatial thinning (a process
+that has no analogue in networks), the best approach to train ML-based
+SDMs varies according to the balancing of the dataset, and the
+evaluation measures used. This suggests that there is no single
+``recipe'' that is guaranteed to give the best model. By contrast to
+networks, SDMs have the advantage of being able to both thin datasets to
+remove some of the sampling bias (\emph{e.g.} Inman et al., 2021), but
+also to create pseudo-absences to inflate the number of supposed
+negatives in the dataset (\emph{e.g.} Iturbide et al., 2015).
+
+An immense body of research on machine learning application to life
+sciences is focused on genomics (which has very specific challenges, see
+a recent discussion by Whalen et al., 2021); this sub-field has
+generated recommendations that do not necessarily match the current
+best-practices for SDMs, and therefore hint at the importance of
+domain-specific guidelines. Chicco \& Jurman (2020) suggest using
+Matthews correlation coefficient (MCC) over \(F_1\), as a protection
+against over-inflation of predicted results; Delgado \& Tibau (2019)
+advocate against the use of Cohen's \(\kappa\), again in favor of MCC,
+as the relative nature of \(\kappa\) means that a worse classifier can
+be picked over a better one; similarly, Boughorbel et al. (2017)
+recommend MCC over other measures of performance for imbalanced data, as
+it has more desirable statistical properties. More recently, Chicco et
+al. (2021) temper the apparent supremacy of the MCC, by suggesting it
+should be replaced by Youden's informedness (also known as \(J\),
+bookmaker's accuracy, and the True-Skill Statistic) when the imbalance
+in the dataset may not be representative of the actual imbalance.
+
+Species interaction networks are often under-sampled (Jordano, 2016b,
+2016a), and this under-sampling is structured taxonomically (Beauchesne
+et al., 2016), structurally (de Aguiar et al., 2019) and spatially
+(Poisot, Bergeron, et al., 2021; Wood et al., 2015). As a consequence,
+networks suffer from data deficiencies both within and between datasets.
+This implies that the comparison of classifiers across space, when
+undersampling varies locally (see \emph{e.g.} McLeod et al., 2021) is
+non-trivial. Furthermore, the baseline value of classifiers performance
+measures under various conditions of skill, bias, and prevalence, has to
+be identified to allow researchers to evaluate whether their interaction
+prediction model is indeed learning. Taken together, these
+considerations highlight three specific issues for ecological networks.
+First, what values of performance measures are indicative of a
+classifier with no skill? This is particularly important as it can
+evaluate whether low prevalence can lull us into a false sense of
+predictive accuracy. Second, independently of the question of model
+evaluation, is low prevalence an issue for \emph{training} or
+\emph{testing}, and can we remedy it? Finally, because the low amount of
+data on interaction makes a lot of imbalance correction methods (see
+\emph{e.g.} Branco et al., 2015) hard to apply, which indicators can be
+optimized by sacrificing least amount of positive interaction data?
+
+It may sound counter-intuitive to care so deeply about how good a
+classifier with no-skill is, as by definition, is has no skill. The
+necessity of this exercise has its roots in the paradox of accuracy:
+when the desired class (``two species interact'') is rare, a model that
+gets less ecologically performant by only predicting the opposite class
+(``these two species do not interact'') sees its accuracy increase;
+because most of the guesses have ``these two species do not interact''
+as a correct answer, a model that never predicts interactions would be
+right an overwhelming majority of the time; it would also be utterly
+useless. Herein lies the core challenge of predicting species
+interactions: the extreme imbalance between classes makes the training
+of predictive models difficult, and their validation even more so as we
+do not reliably know which negatives are true. The connectance (the
+proportion of realized interactions, usually the number of interactions
+divided by the number of species pairs) of empirical networks is usually
+well under 20\%, with larger networks having a lower connectance
+(MacDonald et al., 2020), and therefore being increasingly difficult to
+predict.
+
+\hypertarget{a-primer-on-binary-classifier-evaluation}{%
+\section{A primer on binary classifier
+evaluation}\label{a-primer-on-binary-classifier-evaluation}}
+
+Binary classifiers, which it to say, machine learning algorithms whose
+answer is a categorical value, are usually assessed by measuring
+properties of their confusion matrix, \emph{i.e.} the contingency table
+reporting true/false positive/negative hits. A confusion matrix is laid
+out as
 
@@
     \text{fn} & \text{tn}
-\end{pmatrix} \,,\]
-
-wherein \(\text{tp}\) is the number of interactions predicted as
-positive, \(\text{tn}\) is the number of non-interactions predicted as
-negative, \(\text{fp}\) is the number of non-interactions predicted as
-positive, and \(\text{fn}\) is the number of interactions predicted as
-negative. Almost all measures based on the confusion matrix express
-rates of error or success as proportions, and therefore the values of
-these components matter in a \emph{relative} way. At a coarse scale, a
-classifier is \emph{accurate} when the trace of the matrix divided by
-the sum of the matrix is close to 1, with other measures focusing on
-different ways in which the classifer is wrong.
+\end{pmatrix} \,.\]
+
+In this matrix, tp is the number of times the model predicts an
+interaction that exists in the network (true positive), fp is the number
+of times the model predicts an interaction that does not exist in the
+network (false positive), fn is the number of times the model fails to
+predict an interaction that actually exists in the network (false
+negatives), and tn is the number of times the model correctly predicts
+that an interaction does not exist (true negatives). From these values,
+we can derive a number of measures of model performance (see Strydom et
+al., 2021 for a review of their interpretation in the context of
+networks). At a coarse scale, a classifier is \emph{accurate} when the
+trace of the matrix divided by the sum of the matrix is close to 1, with
+other measures informing us on how the predictions fail.
 
@@
 classification tasks (Ferri et al., 2009). Here we will focus on five of
-them with high relevance for imbalanced learning (He \& Ma, 2013); three
-threshold metrics (\(\kappa\), informedness, and MMC, the Matthews
-Correlation Coefficient), and two ranking metrics (the areas under the
-Receiving Operator Characteristic and the Precision-Recall curves; resp.
-ROC-ACU and PR-AUC). The \(\kappa\) measure of agreement (Landis \&
-Koch, 1977) establishes the extent to which two observers (here the data
-and the prediction) agree, and is measured as
+them with high relevance for imbalanced learning (He \& Ma, 2013). The
+choice of metrics with relevance to class-imbalanced problems is
+fundamental, because as Japkowicz (2013) unambiguously concluded,
+``relatively robust procedures used for unskewed data can break down
+miserably when the data is skewed.'' Following Japkowicz (2013), we
+focus on two ranking metrics (the areas under the Receiving Operator
+Characteristic and Precision Recall curves), and three threshold metrics
+(\(\kappa\), informedness, and MCC; we will briefly discuss \(F_1\) but
+show early on that it has undesirable properties).
+
+The \(\kappa\) measure (Landis \& Koch, 1977) establishes the extent to
+which two observers (the network and the prediction) agree, and is
+measured as
 
@@
 the interaction does not exist) can be determined by picking a threshold
-maximizing some value on the ROC curve or the PR curve. The area under
-these curves (ROC-AUC and PR-AUC henceforth) give ideas on the overall
-goodness of the classifier. Saito \& Rehmsmeier (2015) established that
-the ROC-AUC is biased towards over-estimating performance for imbalanced
-data; on the contrary, the PR-AUC is able to identify classifiers that
-are less able to detect positive interactions correctly, with the
-additional advantage of having a baseline value equal to prevalence.
-Therefore, it is important to assess whether these two measures return
-different results when applied to ecological network prediction. The ROC
-curve is defined by the false positive rate on the \(x\) axis, and the
-true positive rate on the \(y\) axis, and the PR curve is defined by the
-true positive rate on the \(x\) axis, and the positive predictive value
-on the \(y\) axis. By comparison with the previous paragraph, it is
-obvious that \(F_1\) has ties to the PR curve (being close to the
-expected PR-AUC), and that informedness has ties to the ROC curve
-(whereby the threshold maximizing informedness is also the point of
-maximal inflection on the ROC curve). One important difference between
-ROC and PR is that the later does not prominently account for the size
-of the true negative compartments: in short, it is more sensitive to the
-correct positive predictions. In a context of strong imbalance, PR-AUC
-is therefore a more stringent test of model performance.
-
-The same approach is used to evaluate \emph{e.g.} species distribution
-models (SDMs). Indeed, the training and evaluation of SDMs as binary
-classifiers suffers from the same issue of low prevalence; this is not
-surprising that the two fields (SDMs and network predictions) would
-share methods and their attached conceptual issues, as they suffer from
-data limitations, class imbalance, and the conversion of quantitative
-prediction into a binary classification. In previous work, Allouche et
-al. (2006) suggested that \(\kappa\) was a better test of model
-performance than the True Skill Statistic (TSS; which we refer to as
-Youden's informedness); these conclusions were later criticized by
-Somodi et al. (2017), who emphasized that informedness' relationship to
-prevalence depends on assumptions about bias in the model, and therefore
-recommend the use of \(\kappa\) as a validation of classification
-performance. Although this work offers recommendations about the
-comparison of models, it doesn't establishes baselines or good practices
-for training on imbalanced ecological data. Steen et al. (2021) show
-that, when applying spatial thinning (a process that has no analogue in
-networks), the best approach to train ML-based SDMs varies according to
-the balancing of the dataset, and the evaluation measures used. This
-suggests that there is no single ``recipe'' that is guaranteed to give
-the best model. Within the context of networks, there are three specific
-issues that need to be adressed. First, what values of performance
-measures are we expecting for a classifier that has poor performance?
-This is particularly important as it can evaluate whether low prevalence
-can lull us into a false sense of predictive accuracy. Second,
-independently of the question of model evaluation, is low prevalence an
-issue for \emph{training}, and can we remedy it? Finally, because the
-low amount of data on interaction makes a lot of imbalance correction
-methods (see \emph{e.g.} Branco et al., 2015) hard to apply, which
-indicators can be optimized with the least amount of positive
-interaction data?
-
-In addition to the literature on SDMs, most of the research on machine
-learning application to life sciences is focused on genomics (which has
-very specific challenges, see a recent discussion by Whalen et al.,
-2021); this sub-field has generated largely different recommendations.
-Chicco \& Jurman (2020) suggest using Matthews correlation coefficient
-(MCC) over \(F_1\), as a protection against over-inflation of predicted
-results; Delgado \& Tibau (2019) advocate against the use of Cohen's
-\(\kappa\), again in favor of MCC, as the relative nature of \(\kappa\)
-means that a worse classifier can be picked over a better one;
-similarly, Boughorbel et al. (2017) recommend MCC over other measures of
-performance for imbalanced data, as it has more desirable statistical
-properties. More recently, Chicco et al. (2021) temper the apparent
-supremacy of the MCC, by suggesting it should be replaced by Youden's
-informedness (also known as \(J\), bookmaker's accuracy, and the
-True-Skill Statistic) when the imbalance in the dataset may not be
-representative, which is the case as species interaction networks are
-often under-sampled (Jordano, 2016b, 2016a), when classifiers need to be
-compared across different datasets (for example when predicting a system
-in space, where undersampling varies locally; McLeod et al., 2021), and
-when comparing the results to a no-skill (baseline) classifier is
-important. As these conditions are likely to be met with network data,
-there is a need to evaluate which measures of classification accuracy
-respond in a desirable way.
-
-We establish that due to the low prevalence of interactions, even poor
-classifiers applied to food web data will reach a high accuracy; this is
-because the measure is dominated by the accidentally correct predictions
-of negatives. On simulated confusion matrices with ranges of imbalance
-that are credible for ecological networks, MCC had the most desirable
-behavior, and informedness is a linear measure of classifier skill. By
-performing simulations with four models and an ensemble, we show that
-informedness and ROC-AUC are consistently high on network data, and that
-MCC and PR-AUC are more accurate measures of the effective performance
-of the classifier. Finally, by measuring the structure of predicted
-networks, we highlight an interesting paradox: the models with the best
-performance measures are not the models with the closest reconstructed
-network structure. We discuss these results in the context of
-establishing guidelines for the prediction of ecological interactions.
-
-\hypertarget{baseline-values}{%
-\section{Baseline values}\label{baseline-values}}
+maximizing some value on the ROC or the PR curve. The area under these
+curves (ROC-AUC and PR-AUC henceforth) give ideas on the overall
+goodness of the classifier, and the ideal threshold is the point on
+these curves that minimizes the tradeoff represented in these curves.
+Saito \& Rehmsmeier (2015) established that the ROC-AUC is biased
+towards over-estimating performance for imbalanced data; on the
+contrary, the PR-AUC is able to identify classifiers that are less able
+to detect positive interactions correctly, with the additional advantage
+of having a baseline value equal to prevalence. Therefore, it is
+important to assess whether these two measures return different results
+when applied to ecological network prediction. The ROC curve is defined
+by the false positive rate on the \(x\) axis, and the true positive rate
+on the \(y\) axis, and the PR curve is defined by the true positive rate
+on the \(x\) axis, and the positive predictive value on the \(y\) axis.
+By comparison with the previous paragraph, it is obvious that \(F_1\)
+and MCC have ties to the PR curve (being close to the expected PR-AUC),
+and that informedness has ties to the ROC curve (whereby the threshold
+maximizing informedness is also the point of maximal inflection on the
+ROC curve). One important difference between ROC and PR is that the
+later does not prominently account for the size of the true negative
+compartments: in short, it is more sensitive to the correct positive
+predictions. In a context of strong imbalance, PR-AUC is therefore a
+more stringent test of model performance.
+
+\hypertarget{baseline-values-for-the-threshold-metrics}{%
+\section{Baseline values for the threshold
+metrics}\label{baseline-values-for-the-threshold-metrics}}
 
@@
 Branco et al., 2015). In a recent large-scale imputation of interactions
-in the mammal-virus networks, Poisot et al. (2021) for example estimated
-that 93\% of interactions are yet to be documented.
+in the mammal-virus networks, Poisot, Ouellet, et al. (2021) for example
+estimated that 93\% of interactions are yet to be documented.
 
@@
 
+We establish that due to the low prevalence of interactions, even poor
+classifiers applied to food web data will reach a high accuracy; this is
+because the measure is dominated by the accidentally correct predictions
+of negatives. On simulated confusion matrices with ranges of imbalance
+that are credible for ecological networks, MCC had the most desirable
+behavior, and informedness is a linear measure of classifier skill. By
+performing simulations with four models and an ensemble, we show that
+informedness and ROC-AUC are consistently high on network data, and that
+MCC and PR-AUC are more accurate measures of the effective performance
+of the classifier. Finally, by measuring the structure of predicted
+networks, we highlight an interesting paradox: the models with the best
+performance measures are not the models with the closest reconstructed
+network structure. We discuss these results in the context of
+establishing guidelines for the prediction of ecological interactions.
+
 The results presented here highlight an interesting paradox: although
@@
 
+\leavevmode\hypertarget{ref-Beauchesne2016ThiOut}{}%
+Beauchesne, D., Desjardins-Proulx, Archambault, P., \& Gravel, D.
+(2016). Thinking Outside the Boxpredicting Biotic Interactions in
+Data-poor Environments. \emph{Vie Et Milieu-Life and enVironment},
+\emph{66}(3-4), 333--342.
+
 \leavevmode\hypertarget{ref-Bezanson2017JulFre}{}%
@@
 
+\leavevmode\hypertarget{ref-deAguiar2019RevBia}{}%
+de Aguiar, M. A. M., Newman, E. A., Pires, M. M., Yeakel, J. D.,
+Boettiger, C., Burkle, L. A., Gravel, D., Guimarães, P. R., O'Donnell,
+J. L., Poisot, T., Fortin, M.-J., \& Hembry, D. H. (2019). Revealing
+biases in the sampling of ecological interaction networks. \emph{PeerJ},
+\emph{7}, e7566. \url{https://doi.org/10.7717/peerj.7566}
+
 \leavevmode\hypertarget{ref-Delgado2019WhyCoh}{}%
@@
 
+\leavevmode\hypertarget{ref-Inman2021ComSam}{}%
+Inman, R., Franklin, J., Esque, T., \& Nussear, K. (2021). Comparing
+sample bias correction methods for species distribution modeling using
+virtual species. \emph{Ecosphere}, \emph{12}(3), e03422.
+\url{https://doi.org/10.1002/ecs2.3422}
+
+\leavevmode\hypertarget{ref-Iturbide2015FraSpe}{}%
+Iturbide, M., Bedia, J., Herrera, S., del Hierro, O., Pinto, M., \&
+Gutiérrez, J. M. (2015). A framework for species distribution modelling
+with improved pseudo-absence generation. \emph{Ecological Modelling},
+\emph{312}, 166--174.
+\url{https://doi.org/10.1016/j.ecolmodel.2015.05.018}
+
+\leavevmode\hypertarget{ref-Japkowicz2013AssMet}{}%
+Japkowicz, N. (2013). Assessment Metrics for Imbalanced Learning. In
+\emph{Imbalanced Learning} (pp. 187--206). John Wiley \& Sons, Ltd.
+\url{https://doi.org/10.1002/9781118646106.ch8}
+
 \leavevmode\hypertarget{ref-Jeni2013FacImb}{}%
@@
 \emph{Oikos}, \emph{n/a}(n/a). \url{https://doi.org/10.1111/oik.08650}
+
+\leavevmode\hypertarget{ref-Poisot2021GloKno}{}%
+Poisot, T., Bergeron, G., Cazelles, K., Dallas, T., Gravel, D.,
+MacDonald, A., Mercier, B., Violet, C., \& Vissault, S. (2021). Global
+knowledge gaps in species interaction networks data. \emph{Journal of
+Biogeography}, jbi.14127. \url{https://doi.org/10.1111/jbi.14127}
 
@@
 
+\leavevmode\hypertarget{ref-Wood2015EffSpa}{}%
+Wood, S. A., Russell, R., Hanson, D., Williams, R. J., \& Dunne, J. A.
+(2015). Effects of spatial scale of sampling on food web structure.
+\emph{Ecology and Evolution}, \emph{5}(17), 3769--3782.
+\url{https://doi.org/10.1002/ece3.1640}
+
 \leavevmode\hypertarget{ref-Youden1950IndRat}{}%

### 2022-02-26 code/01_analytical_results.jl
updates

@@ -39,10 +39,10 @@ data(@subset(results, :ρ .== _co)) *
     mapping(:b => logit => "logit(bias)", :value => "Measure value", layout=:measure => nonnumeric, color=:s => logit => "logit(skill)", group=:s => nonnumeric) *
     (visual(Lines, colormap=:tofino)) |> 
     plt -> draw(plt, facet=(;linkyaxes = :none)) |>
-    plt -> save(joinpath(@__DIR__, "..", "figures", "changing-bias.png"), plt, px_per_unit = 3)
+    plt -> save(joinpath(@__DIR__, "figures", "changing-bias.png"), plt, px_per_unit = 3)
 
 data(@subset(results, :b .== 0.5)) *
     mapping(:ρ => "Connectance", :value => "Measure value", layout=:measure => nonnumeric, color=:s => logit => "logit(skill)", group=:s => nonnumeric) *
     (visual(Lines, colormap=:tofino)) |> 
     plt -> draw(plt, facet=(;linkyaxes = :none)) |>
-    plt -> save(joinpath(@__DIR__, "..", "figures", "changing-connectance.png"), plt, px_per_unit = 3)
+    plt -> save(joinpath(@__DIR__, figures", "changing-connectance.png"), plt, px_per_unit = 3)

### 2022-02-26 code/confusionmatrix.jl

@@ -1,8 +1,8 @@
 struct ConfusionMatrix
-    tp::Float64
-    tn::Float64
-    fp::Float64
-    fn::Float64
+    tp
+    tn
+    fp
+    fn
     function ConfusionMatrix(a,b,c,d)
         s = a+b+c+d
         return new(a/s, b/s, c/s, d/s)
@@ -10,6 +10,7 @@ struct ConfusionMatrix
 end
 
 ConfusionMatrix(A::Matrix{Float64}) = ConfusionMatrix(A[1,1], A[2,2], A[2,1], A[1,2])
+ConfusionMatrix(A::Matrix) = ConfusionMatrix(A[1,1], A[2,2], A[2,1], A[1,2])
 
 tpr(M::ConfusionMatrix) = M.tp / (M.tp + M.fn)
 tnr(M::ConfusionMatrix) = M.tn / (M.tn + M.fp)

### 2022-02-26 draft.tex
Deploying to gh-pages from @ PoisotLab/ms_networkary_classifiers@ff3a0ed160a7f03f2cd8bc675822c2f3b0bb1bad 🚀

@@
 True Skill Statistic) is \(\text{TPR}+\text{TNR}-1\), where
-\(\text{TPR}= tp/(tp+fn)\) and \(\text{TNR} = tn/(tn+fp)\); informedness
+\(\text{TPR}= tp/(tp+fn)\) and \(\text{TNR} = tn/(tn+fp)\). Informedness
 can be used to find the optimal cutpoint in thresholding analyses
-(Schisterman et al., 2005). The formula for informedness is
+(Schisterman et al., 2005); indeed, the maximal informedness corresponds
+to the point on the ROC curve that is closest to the perfect classifier
+point. The formula for informedness is
 
@@
 
-Finally, \(F_1\) is the harmonic mean of precision (the chance that a
-positive even was correctly classified) and sensitivity (the ability to
-correctly classify positive events), and is defined as
+Finally, \(F_1\) is the harmonic mean of precision (the chance that
+interaction was correctly detected as such) and sensitivity (the ability
+to correctly classify interactions), and is defined as
 
@@
 unweighted, directed network) is a column vector
-\(\mathbf{o}^T = [\rho, (1-\rho)]\) (we can safely drop the \(S^2\)
-terms, as we will work on the confusion matrix, which ends up expressing
-\emph{relative} values). We will apply skill and bias to this matrix,
-and measure how a selection of performance metrics respond to changes in
-these values, in order to assess their suitability for model evaluation.
+\(\mathbf{o}^T = [\rho, \qquad (1-\rho)]\) (we can safely drop the
+\(S^2\) terms, as we will work on the confusion matrix, which ends up
+expressing \emph{relative} values). We will apply skill and bias to this
+matrix, and measure how a selection of performance metrics respond to
+changes in these values, in order to assess their suitability for model
+evaluation.
 
@@
 
-Note that when \(s=0\), \(\text{Tr}(\mathbf{M}) = 0\) (the classifier is
+When \(s=0\), \(\text{Tr}(\mathbf{M}) = 0\) (the classifier is
 \emph{always} wrong), when \(s=0.5\), the classifier is no-skill and
@@
 In all further simulations, the confusion matrix \(\mathbf{C}\) is
-transformed so that it sums to 1.
+transformed so that it sums to unity, \emph{i.e.} the entries are the
+\emph{proportions} of guesses.
 
@@
 \(\rho\), and report how the main measures discussed in the introduction
-(MCC, \(F_1\), \(\kappa\), and informedness) are responding to issues
-with the classifier. Before we do so, it is important to explain why we
-will not focus on accuracy too much. Accuracy is the number of correct
-predictions (\(\text{Tr}(\mathbf{C})\)) divided by the sum of the
-confusion matrix. For a no-skill, no-bias classifier, accuracy is equal
-to \(\rho^2 + (1-\rho)^2\); for \(\rho = 0.05\), this is
-\(\approx 0.90\), and for \(\rho = 0.01\), this is equal to
-\(\approx 0.98\). In other words, the values of accuracy are expected to
-be so high that they are not really informatived (this is simply
-explained by the fact that for \(\rho\) small,
-\(\rho^2 \ll (1-\rho)^2\)). More concerning is the fact that introducing
-bias changes the response of accuracy in unexpected ways. Assuming a
-no-skill classifier, the numerator of accuracy becomes
+(MCC, \(F_1\), \(\kappa\), and informedness) respond. Before we do so,
+it is important to explain why we will not focus on accuracy too much.
+Accuracy is the number of correct predictions
+(\(\text{Tr}(\mathbf{C})\)) divided by the sum of the confusion matrix.
+For a no-skill, no-bias classifier, accuracy is equal to
+\(\rho^2 + (1-\rho)^2\); for \(\rho = 0.05\), this is \(\approx 0.90\),
+and for \(\rho = 0.01\), this is equal to \(\approx 0.98\). In other
+words, the values of accuracy are high enough to be uninformative (for
+\(\rho\) small, \(\rho^2 \ll (1-\rho)^2\)). More concerning is the fact
+that introducing bias changes the response of accuracy in unexpected
+ways. Assuming a no-skill classifier, the numerator of accuracy becomes
 \(b\rho^2 + (1-b)(1-\rho)^2\), which increases when \(b\) is low, which
 specifically means that at equal skill, a classifier that under-predicts
-interactions will have higher accuracy than an un-biased classifier.
-These issues are absent from balanced accuracy, but should nevertheless
-lead us to not report accuracy as the primary measure of network
-prediction success; moving forward, we will focus on other measures.
+interactions will have higher accuracy than an un-biased classifier
+(because the value of accuracy is dominated by the size of tn, which
+will increase). These issues are absent from balanced accuracy, but
+should nevertheless lead us to not report accuracy as the primary
+measure of network prediction success; moving forward, we will focus on
+other measures.
 
@@
 \(\rho\) were taken linearly in \(]0, 0.5]\), which is within the range
-of usually observed connectance values for empirical species interaction
-networks. Note that at this point, there is no food web model to speak
-of; rather, the confusion matrix we discuss can be obtained for any
-classification task. Based on the previous discussion, the desirable
-properties for a measure of classifier success should be: an increase
-with classifier skill, especially at low bias; a hump-shaped response to
-bias, especially at high skill, and ideally center around
-\(\text{logit}(b)=0\); an increase with prevalence up until
-equiprevalence is reached.
+of connectance for species interaction networks. Note that at this
+point, there is no network model to speak of; the confusion matrix we
+discuss can be obtained for any classification task. Based on the
+previous discussion, the desirable properties for a measure of
+classifier success should be: an increase with classifier skill,
+especially at low bias; a hump-shaped response to bias, especially at
+high skill, and ideally centered around \(\text{logit}(b)=0\); an
+increase with prevalence up until equiprevalence is reached.
 
@@
 is \(\text{abs}(v_i, h_j) \le \xi\). The models therefore all have the
-form
+following form, where \(i_{i,j}\) is an interaction from species \(i\)
+to species \(j\):
 
@@
 \begin{bmatrix}
-           x_{1} \\
-           x_{2} \\
+    i_{1,1} \\
+    i_{1,2} \\
     \vdots \\
-           x_{m}
+    i_{m,n-1} \\
+    i_{m,n}
+\end{bmatrix}
+\propto
+\begin{bmatrix}
+           v_{1} & h_{1} \\
+           v_{1} & h_{2} \\
+           \vdots & \vdots \\
+           v_{m} & h_{n-1} \\
+           v_{m} & h_{n}
          \end{bmatrix}
@@
 
+TODO informedness and accuracy should be easy to beat, make sure the
+model is better than them!
+
 The results presented here highlight an interesting paradox: although

### 2022-02-26 draft.tex
Deploying to gh-pages from @ PoisotLab/ms_networkary_classifiers@59ea7282a67f66615b11db9c9a0529c061308c24 🚀

@@
 species richness TODO. Recent contributions to the field of ecological
-network prediction (Strydom et al., 2021; \textbf{Becker2021OptPre?})
-highlight that although interactions can be predicted by adding
-ecologically relevant information (in the form of, \emph{e.g.} traits),
-we do not have robust guidelines as to how the predictive ability of
-these models should be evaluated, nor about how the models should be
-trained. Here, by relying on simple derivations and a series of
-simulations, we formulate a number of such guidelines, specifically for
-the case of binary classifiers derived from thresholded values.
-Specifically, we conduct an investigation of the models in terms of
-their skill (ability to make the right prediction), bias (trends towards
-systematically over-predicting one class), class imbalance (the relative
-number of cases representing interactions), and show how these effects
-interact. We conclude on the fact that models with the best
+network prediction (Pichler et al., 2020; Strydom et al., 2021;
+\textbf{Becker2021OptPre?}) highlight that although interactions can be
+predicted by adding ecologically relevant information (in the form of,
+\emph{e.g.} traits), we do not have robust guidelines as to how the
+predictive ability of these models should be evaluated, nor about how
+the models should be trained. Here, by relying on simple derivations and
+a series of simulations, we formulate a number of such guidelines,
+specifically for the case of binary classifiers derived from thresholded
+values. Specifically, we conduct an investigation of the models in terms
+of their skill (ability to make the right prediction), bias (trends
+towards systematically over-predicting one class), class imbalance (the
+relative number of cases representing interactions), and show how these
+effects interact. We conclude on the fact that models with the best
 interaction-scale predictive score do not necessarily result in the most
@@
 miserably when the data is skewed.'' Following Japkowicz (2013), we
-focus on two ranking metrics (the areas under the Receiving Operator
+focus on two ranking metrics (the areas under the Receiver Operating
 Characteristic and Precision Recall curves), and three threshold metrics
@@
 necessarily linearly) to the probability of an interaction between \(i\)
-and \(j\).
+and \(j\). Nevertheless, the ROC-AUC and PR-AUC (and therefore the
+thresholds) can be measured by integrating over the domain of the values
+return by each machine.
 
@@
 the testing set, and picking the threshold that maximized Youden's
-informedness, which is usually the optimized target for imbalanced
-classification. During the thresholding step, we measured the area under
-the receiving-operator characteristic (ROC-AUC) and precision-recall
+informedness. During the thresholding step, we measured the area under
+the receiver operating characteristic (ROC-AUC) and precision-recall
 (PR-AUC) curves, as measures of overall performance over the range of
@@
 put through the same thresholding process. The complete code to run the
-simulations is given as an appendix; running the final simulation
-required 4.8 core days (approx. 117 hours).
+simulations is available at \texttt{10.17605/OSF.IO/JKEWD}; the
+simulations require approx. 5 core days.
 
@@
 imbalanced; preliminary analyses of data with a connectance larger than
-3 revealed that all machines reached consistently high performance).
+0.3 revealed that all machines reached consistently high performance).
 
@@
 
+\leavevmode\hypertarget{ref-Pichler2020MacLea}{}%
+Pichler, M., Boreux, V., Klein, A.-M., Schleuning, M., \& Hartig, F.
+(2020). Machine learning algorithms to infer trait-matching and predict
+species interactions in ecological networks. \emph{Methods in Ecology
+and Evolution}, \emph{11}(2), 281--293.
+\url{https://doi.org/10.1111/2041-210X.13329}
+
 \leavevmode\hypertarget{ref-Poisot2021GloKno}{}%

### 2022-02-26 code/02_simulations.jl
update the code

@@ -62,7 +62,7 @@ candidate_models = [
 ]
 
 S = 100
-_n_sims = 2000
+_n_sims = 200
 conditions_breadth = rand(_n_sims) .* 0.4 .+ 0.05
 conditions_bias = rand(_n_sims) .* 0.98 .+ 0.01
 conditions = hcat(conditions_breadth, conditions_bias)
@@ -72,14 +72,23 @@ Threads.@threads for i in 1:size(conditions, 1)
     @info i, Threads.threadid(), breadth, bias
 
     𝐗, 𝐲 = network(S, breadth)
+    _real_co = mean(𝐲)
 
-    # Training and testing sets
-    training_size = round(Int64, 0.3 * length(𝐲))
+    training_size = round(Int64, 0.5 * length(𝐲))
     n_positive = round(Int64, training_size * bias)
     idx_pos = sample(findall(iszero.(𝐲)), n_positive; replace=true)
     idx_neg = sample(findall(isone.(𝐲)), training_size - n_positive; replace=true)
     Iₚ = shuffle(vcat(idx_neg, idx_pos))
     Iₒ = setdiff(eachindex(𝐲), Iₚ)
+
+    # Tweak the testing set to have the correct connectance - this results in a SMALLER testing set
+    _test_pos = sum(𝐲[Iₒ])
+    _expected_neg = round(Int64, _test_pos/_real_co - _test_pos)
+    _observed_neg = round(Int64, length(Iₒ)-sum(𝐲[Iₒ]))
+    tst_neg = sample(findall(iszero.(𝐲[Iₒ])), max(_observed_neg - _expected_neg, 1); replace=true)
+    deleteat!(Iₒ, sort(unique(tst_neg)))
+    #
+
     m = []
     for candidate_model in candidate_models
         this_machine = machine(candidate_model.second, 𝐗, 𝐲)

### 2022-02-27 code/02_simulations.jl
some code tweaks while the simulations run on narval

@@ -30,9 +30,7 @@ function network(S, ξ)
     𝐱ₕ = sort(rand(resistance, S))
     𝐱₁ = repeat(𝐱ᵥ; outer=length(𝐱ₕ))
     𝐱₂ = repeat(𝐱ₕ; inner=length(𝐱ᵥ))
-    𝐱₃ = abs.(𝐱₁ .- 𝐱₂)
     𝐲 = [L(𝐱₁[i], 𝐱₂[i]; r=ξ) for i in 1:(S*S)]
-    #𝐱 = table(hcat(𝐱₁, 𝐱₂, 𝐱₃))
     𝐱 = table(hcat(𝐱₁, 𝐱₂))
     return (𝐱, 𝐲)
 end
@@ -48,24 +46,42 @@ results = [DataFrame(;
     value=Float64[],
 ) for _thr in 1:Threads.nthreads()]
 
-# these regression machines go brrr as f u c k
-DecisionTree = @load DecisionTreeRegressor pkg = DecisionTree
-RandomForest = @load RandomForestRegressor pkg = DecisionTree
-BoostedRegressor = @load EvoTreeRegressor pkg = EvoTrees
-RidgeRegressor = @load RidgeRegressor pkg = MLJLinearModels
+models() do model
+    matching(model, network(10, 0.2)...) &&
+    model.package_name != "ScikitLearn" # For a strange reason, ScikitLearn machines drop the core
+end
+
+# These regression machines go brrr as f u c k
+DecisionTree = @load DecisionTreeRegressor pkg = DecisionTree verbosity=0
+RandomForest = @load RandomForestRegressor pkg = DecisionTree verbosity=0
+BoostedRegressor = @load EvoTreeRegressor pkg = EvoTrees verbosity=0
+RidgeRegressor = @load RidgeRegressor pkg = MLJLinearModels verbosity=0
+LinearRegressor = @load LinearRegressor pkg = MLJLinearModels verbosity=0
 
 candidate_models = [
     :DecTree => DecisionTree(),
     :BRT => BoostedRegressor(),
     :RF => RandomForest(),
-    :RR => RidgeRegressor()
+    :RR => RidgeRegressor(),
+    :LR => LinearRegressor()
 ]
 
+function rescaler(m, M)
+    return function f(x)
+        y = copy(x)
+        y = (y .- minimum(y))./(maximum(y)-minimum(y))
+        y = y .* (M-m) .+ m
+        return y
+    end
+end
+
 S = 100
[…]

### 2022-02-27 code/03_figures.jl

@@ -56,6 +56,34 @@ data(@subset(results, _keepval.(:measure))) *
     plt -> draw(plt, facet=(;linkyaxes = :minimal)) |>
     plt -> save(joinpath(@__DIR__, "..", "figures", "bias_roc_pr.png"), plt, px_per_unit = 3)
 
+_keepval(f) = f in ["PPV", "NPV"]
+data(@subset(results, _keepval.(:measure))) *
+    mapping(:bias => "Training set bias", :value => "Value", row=:midpoint => nonnumeric, col=:model, color=:measure => "Measure") *
+    (AlgebraOfGraphics.density() * visual(Contour, color=:grey, alpha=0.3) + smooth() * visual(linewidth=2.0)) |>
+    plt -> draw(plt, facet=(;linkyaxes = :minimal)) |>
+    plt -> save(joinpath(@__DIR__, "..", "figures", "bias_predvalues.png"), plt, px_per_unit = 3)
+
+_keepval(f) = f in ["ACC", "MKD"]
+data(@subset(results, _keepval.(:measure))) *
+    mapping(:bias => "Training set bias", :value => "Value", row=:midpoint => nonnumeric, col=:model, color=:measure => "Measure") *
+    (AlgebraOfGraphics.density() * visual(Contour, color=:grey, alpha=0.3) + smooth() * visual(linewidth=2.0)) |>
+    plt -> draw(plt, facet=(;linkyaxes = :minimal)) |>
+    plt -> save(joinpath(@__DIR__, "..", "figures", "bias_acc.png"), plt, px_per_unit = 3)
+
+_keepval(f) = f in ["FOR", "FDR"]
+data(@subset(results, _keepval.(:measure))) *
+    mapping(:bias => "Training set bias", :value => "Value", row=:midpoint => nonnumeric, col=:model, color=:measure => "Measure") *
+    (AlgebraOfGraphics.density() * visual(Contour, color=:grey, alpha=0.3) + smooth() * visual(linewidth=2.0)) |>
+    plt -> draw(plt, facet=(;linkyaxes = :minimal)) |>
+    plt -> save(joinpath(@__DIR__, "..", "figures", "bias_falserates.png"), plt, px_per_unit = 3)
+
+_keepval(f) = f in ["FPR", "FNR", "TPR", "TNR"]
+data(@subset(results, _keepval.(:measure))) *
[…]

### 2022-02-27 code/04_sim2.jl

@@ -58,7 +58,6 @@ candidate_models = [
 
 S = (50,80)
 
-
 𝐗, 𝐲 = network(S, 0.19)
 bias = 0.7
 

### 2022-02-27 code/02_simulations.jl
some code

@@ -9,6 +9,9 @@ using DataFrames
 using CSV: CSV
 using Distributions
 
+# Simulation suffix
+_suffix = get(ENV, "SLURM_ARRAY_TASK_ID", "global")
+
 # AUC
 function ∫(x::Array{T}, y::Array{T}) where {T<:Number}
     S = zero(Float64)
@@ -30,7 +33,9 @@ function network(S, ξ)
     𝐱ₕ = sort(rand(resistance, S))
     𝐱₁ = repeat(𝐱ᵥ; outer=length(𝐱ₕ))
     𝐱₂ = repeat(𝐱ₕ; inner=length(𝐱ᵥ))
+    𝐱₃ = abs.(𝐱₁ .- 𝐱₂)
     𝐲 = [L(𝐱₁[i], 𝐱₂[i]; r=ξ) for i in 1:(S*S)]
+    #𝐱 = table(hcat(𝐱₁, 𝐱₂, 𝐱₃))
     𝐱 = table(hcat(𝐱₁, 𝐱₂))
     return (𝐱, 𝐲)
 end
@@ -46,42 +51,24 @@ results = [DataFrame(;
     value=Float64[],
 ) for _thr in 1:Threads.nthreads()]
 
-models() do model
-    matching(model, network(10, 0.2)...) &&
-    model.package_name != "ScikitLearn" # For a strange reason, ScikitLearn machines drop the core
-end
-
-# These regression machines go brrr as f u c k
-DecisionTree = @load DecisionTreeRegressor pkg = DecisionTree verbosity=0
-RandomForest = @load RandomForestRegressor pkg = DecisionTree verbosity=0
-BoostedRegressor = @load EvoTreeRegressor pkg = EvoTrees verbosity=0
-RidgeRegressor = @load RidgeRegressor pkg = MLJLinearModels verbosity=0
-LinearRegressor = @load LinearRegressor pkg = MLJLinearModels verbosity=0
+# these regression machines go brrr as f u c k
+DecisionTree = @load DecisionTreeRegressor pkg = DecisionTree
+RandomForest = @load RandomForestRegressor pkg = DecisionTree
+BoostedRegressor = @load EvoTreeRegressor pkg = EvoTrees
+RidgeRegressor = @load RidgeRegressor pkg = MLJLinearModels
 
 candidate_models = [
     :DecTree => DecisionTree(),
     :BRT => BoostedRegressor(),
     :RF => RandomForest(),
-    :RR => RidgeRegressor(),
-    :LR => LinearRegressor()
[…]

### 2022-02-27 code/XX_aggregate_simulations.jl
add code to merge the multicore simulations

@@ -1,0 +1,20 @@
+using DataFrames
+using CSV: CSV
+
+function main()::Int64
+    try
+        simulations = readdir(joinpath(@__DIR__, "simulations"))
+        all_data = [
+            DataFrame(CSV.File(joinpath(@__DIR__, "simulations", file))) for
+            file in simulations
+        ]
+        CSV.write(joinpath(@__DIR__, "output.csv"), reduce(vcat, all_data))
+    catch err
+        @info err
+        return 1
+    end
+    return 0
+end
+
+main()
+

### 2022-02-27 code/02_simulations.jl
update the simulations with LR as well

@@ -40,7 +40,6 @@ function network(S, ξ)
     return (𝐱, 𝐲)
 end
 
-
 # Prepare the results
 results = [DataFrame(;
     breadth=Float64[],
@@ -56,12 +55,14 @@ DecisionTree = @load DecisionTreeRegressor pkg = DecisionTree
 RandomForest = @load RandomForestRegressor pkg = DecisionTree
 BoostedRegressor = @load EvoTreeRegressor pkg = EvoTrees
 RidgeRegressor = @load RidgeRegressor pkg = MLJLinearModels
+LinearRegressor = @load LinearRegressor pkg = GLM
 
 candidate_models = [
-    :DecTree => DecisionTree(),
+    Symbol("Decision tree") => DecisionTree(),
     :BRT => BoostedRegressor(),
-    :RF => RandomForest(),
-    :RR => RidgeRegressor()
+    Symbol("Random Forest") => RandomForest(),
+    Symbol("Ridge regression") => RidgeRegressor(),
+    Symbol("Linear regression") => LinearRegressor()
 ]
 
 S = 100
@@ -160,26 +161,26 @@ Threads.@threads for i in 1:size(conditions, 1)
         ROCAUC = ∫(fpr.(M), tpr.(M))
         AUPRC = ∫(tpr.(M), ppv.(M))
         𝐌 = M[last(findmax(informedness.(M)))]
-        push!(results[Threads.threadid()], (breadth, bias, mean(𝐲), :ensemble, :ROCAUC, ROCAUC))
-        push!(results[Threads.threadid()], (breadth, bias, mean(𝐲), :ensemble, :PRAUC, AUPRC))
-        push!(results[Threads.threadid()], (breadth, bias, mean(𝐲), :ensemble, :CSI, csi(𝐌)))
-        push!(results[Threads.threadid()], (breadth, bias, mean(𝐲), :ensemble, :BA, balanced(𝐌)))
-        push!(results[Threads.threadid()], (breadth, bias, mean(𝐲), :ensemble, :ACC, accuracy(𝐌)))
-        push!(results[Threads.threadid()], (breadth, bias, mean(𝐲), :ensemble, :INF, informedness(𝐌)))
-        push!(results[Threads.threadid()], (breadth, bias, mean(𝐲), :ensemble, :PT, pt(𝐌)))
[…]

### 2022-02-27 code/05_networkproperties.jl
rename sim file

@@ -57,12 +57,14 @@ DecisionTree = @load DecisionTreeRegressor pkg = DecisionTree
 RandomForest = @load RandomForestRegressor pkg = DecisionTree
 BoostedRegressor = @load EvoTreeRegressor pkg = EvoTrees
 RidgeRegressor = @load RidgeRegressor pkg = MLJLinearModels
+LinearRegressor = @load LinearRegressor pkg = GLM
 
 candidate_models = [
-    :DecTree => DecisionTree(),
+    Symbol("Decision tree") => DecisionTree(),
     :BRT => BoostedRegressor(),
-    :RF => RandomForest(),
-    :RR => RidgeRegressor(),
+    Symbol("Random Forest") => RandomForest(),
+    Symbol("Ridge regression") => RidgeRegressor(),
+    Symbol("Linear regression") => LinearRegressor()
 ]
 
 S = (50,80)

### 2022-02-27 code/02_simulations.jl
fix filenames + network comparison code

@@ -55,7 +55,7 @@ DecisionTree = @load DecisionTreeRegressor pkg = DecisionTree
 RandomForest = @load RandomForestRegressor pkg = DecisionTree
 BoostedRegressor = @load EvoTreeRegressor pkg = EvoTrees
 RidgeRegressor = @load RidgeRegressor pkg = MLJLinearModels
-LinearRegressor = @load LinearRegressor pkg = GLM
+LinearRegressor = @load LinearRegressor pkg = MLJLinearModels
 
 candidate_models = [
     Symbol("Decision tree") => DecisionTree(),

### 2022-02-27 code/04_networkviz.jl

@@ -48,12 +48,14 @@ DecisionTree = @load DecisionTreeRegressor pkg = DecisionTree
 RandomForest = @load RandomForestRegressor pkg = DecisionTree
 BoostedRegressor = @load EvoTreeRegressor pkg = EvoTrees
 RidgeRegressor = @load RidgeRegressor pkg = MLJLinearModels
+LinearRegressor = @load LinearRegressor pkg = MLJLinearModels
 
 candidate_models = [
-    :DecTree => DecisionTree(),
+    Symbol("Decision tree") => DecisionTree(),
     :BRT => BoostedRegressor(),
-    :RF => RandomForest(),
-    :RR => RidgeRegressor(),
+    Symbol("Random Forest") => RandomForest(),
+    Symbol("Ridge regression") => RidgeRegressor(),
+    Symbol("Linear regression") => LinearRegressor()
 ]
 
 S = (50,80)

### 2022-02-27 code/05_networkproperties.jl

@@ -57,7 +57,7 @@ DecisionTree = @load DecisionTreeRegressor pkg = DecisionTree
 RandomForest = @load RandomForestRegressor pkg = DecisionTree
 BoostedRegressor = @load EvoTreeRegressor pkg = EvoTrees
 RidgeRegressor = @load RidgeRegressor pkg = MLJLinearModels
-LinearRegressor = @load LinearRegressor pkg = GLM
+LinearRegressor = @load LinearRegressor pkg = MLJLinearModels
 
 candidate_models = [
     Symbol("Decision tree") => DecisionTree(),

### 2022-02-27 code/04_networkviz.jl
update the code for network viz

@@ -129,6 +129,7 @@ push!(Ms, 𝐌)
 ens_thres = thresholds[last(findmax(informedness.(M)))]
 
 mnames = ["DecTree", "BRT", "RF", "RR"]
+mnames = [String(p.first) for p in candidate_models]
 results = DataFrame(;
     infectivity=Float64[],
     resistance=Float64[],
@@ -152,7 +153,7 @@ for i in 1:length(𝐲)
 end
 
 data(results) *
-    mapping(:infectivity => "Infectivity trait", :resistance => "Resistance trait", :prediction => "Interaction probability"; layout=:model => sorter("BRT", "RF", "DecTree", "RR", "Ensemble", "Dataset")) *
+    mapping(:infectivity => "Infectivity trait", :resistance => "Resistance trait", :prediction => "Interaction probability"; layout=:model => sorter("BRT", "Random Forest", "Decision tree", "Ridge regression", "Linear regression", "Ensemble", "Dataset")) *
     visual(Heatmap, colormap=Reverse(:deep)) |>
     plt -> draw(plt, facet=(;linkyaxes = :minimal), axis = (xticks = LinearTicks(3),)) |>
     plt -> save(joinpath(@__DIR__, "..", "figures", "valid_ensemble.png"), plt, px_per_unit = 3)

### 2022-02-27 code/04_networkviz.jl
update dataviz plot

@@ -12,7 +12,7 @@ using AlgebraOfGraphics, CairoMakie
 using EcologicalNetworks
 using Random
 
-Random.seed!(1)
+Random.seed!(5)
 
 # AUC
 function ∫(x::Array{T}, y::Array{T}) where {T<:Number}
@@ -39,7 +39,7 @@ function network(S, ξ)
     𝐱₂ = repeat(𝐱ₕ; inner=length(𝐱ᵥ))
     𝐱₃ = abs.(𝐱₁ .- 𝐱₂)
     𝐲 = [L(𝐱₁[i], 𝐱₂[i]; r=ξ) for i in 1:length(𝐱₁)]
-    #𝐱 = table(hcat(𝐱₁, 𝐱₂, 𝐱₃))
+    # 𝐱 = table(hcat(𝐱₁, 𝐱₂, 𝐱₃))
     𝐱 = table(hcat(𝐱₁, 𝐱₂))
     return (𝐱, 𝐲)
 end
@@ -58,10 +58,10 @@ candidate_models = [
     Symbol("Linear regression") => LinearRegressor()
 ]
 
-S = (50,80)
+S = (50,50)
 
-𝐗, 𝐲 = network(S, 0.19)
-bias = 0.7
+𝐗, 𝐲 = network(S, 0.20)
+bias = 0.5
 
 training_size = round(Int64, 0.3 * length(𝐲))
 n_positive = round(Int64, training_size * bias)
@@ -136,24 +136,31 @@ results = DataFrame(;
     model=String[],
     prediction=Float64[],
     guess=Float64[],
+    truth = Float64[]
 )
 for j in 1:length(m)
     pr = MLJ.predict(m[j].first)
     prr = R(pr)
     thr = R(m[j].second, extrema(pr)...)
     for i in 1:length(𝐲)
-        push!(results, (𝐗.x1[i], 𝐗.x2[i], mnames[j], prr[i], prr[i] >= thr))
+        push!(results, (𝐗.x1[i], 𝐗.x2[i], mnames[j], prr[i], prr[i] >= thr, 𝐲[i]))
     end
 end
 for i in 1:length(𝐲)
-    push!(results, (𝐗.x1[i], 𝐗.x2[i], "Ensemble", ens_pred[i], ens_pred[i] >= ens_thres))
[…]

### 2022-02-27 code/05_networkproperties.jl
update for network properties code

@@ -13,7 +13,7 @@ using EcologicalNetworks
 using Random
 using Combinatorics
 
-Random.seed!(1)
+Random.seed!(3)
 
 # AUC
 function ∫(x::Array{T}, y::Array{T}) where {T<:Number}
@@ -67,7 +67,7 @@ candidate_models = [
     Symbol("Linear regression") => LinearRegressor()
 ]
 
-S = (50,80)
+S = (50,60)
 
 results = DataFrame(
     run = Int64[],
@@ -80,9 +80,9 @@ _m1 = BipartiteNetwork([true true; true true])
 _m2 = BipartiteNetwork([true false; true true])
 _m3 = BipartiteNetwork([true false; false true])
 
-for i in 1:250
+for i in 1:500
 
-    𝐗, 𝐲 = network(S, 0.19)
+    𝐗, 𝐲 = network(S, 0.20)
 
     net = BipartiteNetwork(reshape(Bool.(𝐲), S))
     push!(results, (i, :data, :Connectance, connectance(net)))

### 2022-02-27 code/05_networkproperties.jl
fig updates

@@ -67,7 +67,7 @@ candidate_models = [
     Symbol("Linear regression") => LinearRegressor()
 ]
 
-S = (50,60)
+S = (50,40)
 
 results = DataFrame(
     run = Int64[],
@@ -76,26 +76,27 @@ results = DataFrame(
     value = Float64[]
 )
 
-_m1 = BipartiteNetwork([true true; true true])
-_m2 = BipartiteNetwork([true false; true true])
-_m3 = BipartiteNetwork([true false; false true])
+function asymmetry(B::BipartiteNetwork)
+    s1 = species(B; dims=1)
+    s2 = species(B; dims=2)
+    k1 = [degree(B)[s] for s in s1]
+    k2 = [degree(B)[s] for s in s2]'
+    C = abs.(k1 .- k2) ./ (k1 .+ k2) .* adjacency(B)
+    return sum(C)/links(B)
+end
 
 for i in 1:500
 
-    𝐗, 𝐲 = network(S, 0.20)
+    𝐗, 𝐲 = network(S, 0.15)
 
     net = BipartiteNetwork(reshape(Bool.(𝐲), S))
     push!(results, (i, :data, :Connectance, connectance(net)))
     push!(results, (i, :data, :Nestedness, η(net)))
     push!(results, (i, :data, :Modularity, Q(brim(lp(net)...)...)))
-    #push!(results, (i, :data, :MotifOne, length(find_motif(net, _m1))))
-    #push!(results, (i, :data, :MotifTwo, length(find_motif(net, _m2))))
-    #push!(results, (i, :data, :MotifThree, length(find_motif(net, _m3))))
-    
+    push!(results, (i, :data, :Asymmetry, asymmetry(net)))
+    bias = 0.5
 
-    bias = 0.7
-
-    training_size = round(Int64, 0.3 * length(𝐲))
+    training_size = round(Int64, 0.7 * length(𝐲))
     n_positive = round(Int64, training_size * bias)
     idx_pos = sample(findall(iszero.(𝐲)), n_positive; replace=true)
     idx_neg = sample(findall(isone.(𝐲)), training_size - n_positive; replace=true)
@@ -143,9 +144,7 @@ for i in 1:500
         push!(results, (i, candidate_model.first, :Connectance, connectance(net)))
         push!(results, (i, candidate_model.first, :Nestedness, η(net)))
         push!(results, (i, candidate_model.first, :Modularity, Q(brim(lp(net)...)...)))
[…]

### 2022-02-27 code/03_figures.jl
change layout for main figures

@@ -18,8 +18,6 @@ todrop = unique(vcat([
     @subset(results, :measure .== "ACC", :value .== 0.0).runid,
     @subset(results, :measure .== "TPR", :value .== 0.0).runid,
     @subset(results, :measure .== "TNR", :value .== 0.0).runid,
-    @subset(results, :measure .== "FPR", :value .== 1.0).runid,
-    @subset(results, :measure .== "FNR", :value .== 1.0).runid,
     @subset(results, :measure .== "PRAUC", :value .<= 0.0).runid,
     @subset(results, :connectance .>= 0.25).runid,
     @subset(results, :connectance .<= 0.05).runid
@@ -44,8 +42,8 @@ results = leftjoin(results, select(connectance_values, [:runid, :midpoint]), on=
 # Dataviz
 _keepval(f) = f in ["MCC", "INF"]
 data(@subset(results, _keepval.(:measure))) *
-    mapping(:bias => "Training set bias", :value => "Value", row=:midpoint => nonnumeric, col=:model, color=:measure => "Measure") *
-    (AlgebraOfGraphics.density() * visual(Contour, color=:grey, alpha=0.3) + smooth() * visual(linewidth=2.0)) |>
+    mapping(:bias => "Training set bias", :value => "Value", row=:midpoint => nonnumeric, col=:measure => "Measure", color=:model => "Model") *
+    (smooth() * visual(linewidth=2.0)) |>
     plt -> draw(plt, facet=(;linkyaxes = :minimal)) |>
     plt -> save(joinpath(@__DIR__, "..", "figures", "bias_mcc_inf.png"), plt, px_per_unit = 3)
 

### 2022-02-27 code/03_figures.jl
remove figures that are not in the paper

@@ -54,48 +54,6 @@ data(@subset(results, _keepval.(:measure))) *
     plt -> draw(plt, facet=(;linkyaxes = :minimal)) |>
     plt -> save(joinpath(@__DIR__, "..", "figures", "bias_roc_pr.png"), plt, px_per_unit = 3)
 
-_keepval(f) = f in ["PPV", "NPV"]
-data(@subset(results, _keepval.(:measure))) *
-    mapping(:bias => "Training set bias", :value => "Value", row=:midpoint => nonnumeric, col=:model, color=:measure => "Measure") *
-    (AlgebraOfGraphics.density() * visual(Contour, color=:grey, alpha=0.3) + smooth() * visual(linewidth=2.0)) |>
-    plt -> draw(plt, facet=(;linkyaxes = :minimal)) |>
-    plt -> save(joinpath(@__DIR__, "..", "figures", "bias_predvalues.png"), plt, px_per_unit = 3)
-
-_keepval(f) = f in ["ACC", "MKD"]
-data(@subset(results, _keepval.(:measure))) *
-    mapping(:bias => "Training set bias", :value => "Value", row=:midpoint => nonnumeric, col=:model, color=:measure => "Measure") *
-    (AlgebraOfGraphics.density() * visual(Contour, color=:grey, alpha=0.3) + smooth() * visual(linewidth=2.0)) |>
-    plt -> draw(plt, facet=(;linkyaxes = :minimal)) |>
-    plt -> save(joinpath(@__DIR__, "..", "figures", "bias_acc.png"), plt, px_per_unit = 3)
-
-_keepval(f) = f in ["FOR", "FDR"]
-data(@subset(results, _keepval.(:measure))) *
-    mapping(:bias => "Training set bias", :value => "Value", row=:midpoint => nonnumeric, col=:model, color=:measure => "Measure") *
-    (AlgebraOfGraphics.density() * visual(Contour, color=:grey, alpha=0.3) + smooth() * visual(linewidth=2.0)) |>
-    plt -> draw(plt, facet=(;linkyaxes = :minimal)) |>
-    plt -> save(joinpath(@__DIR__, "..", "figures", "bias_falserates.png"), plt, px_per_unit = 3)
-
-_keepval(f) = f in ["FPR", "FNR", "TPR", "TNR"]
-data(@subset(results, _keepval.(:measure))) *
[…]

### 2022-02-27 code/05_networkproperties.jl
update with betadiv

@@ -85,11 +85,16 @@ function asymmetry(B::BipartiteNetwork)
     return sum(C)/links(B)
 end
 
+function betadiv(n1, n2)
+    return KGL08(βos(n1, n2))
+end
+
 for i in 1:500
 
     𝐗, 𝐲 = network(S, 0.15)
 
     net = BipartiteNetwork(reshape(Bool.(𝐲), S))
+    realnet = copy(net)
     push!(results, (i, :data, :Connectance, connectance(net)))
     push!(results, (i, :data, :Nestedness, η(net)))
     push!(results, (i, :data, :Modularity, Q(brim(lp(net)...)...)))
@@ -145,6 +150,7 @@ for i in 1:500
         push!(results, (i, candidate_model.first, :Nestedness, η(net)))
         push!(results, (i, candidate_model.first, :Modularity, Q(brim(lp(net)...)...)))
         push!(results, (i, candidate_model.first, :Asymmetry, asymmetry(net)))
+        push!(results, (i, candidate_model.first, :Betadiv, betadiv(realnet, net)))
 
     end
 
@@ -178,6 +184,7 @@ for i in 1:500
     push!(results, (i, :ensemble, :Nestedness, η(net)))
     push!(results, (i, :ensemble, :Modularity, Q(brim(lp(net)...)...)))
     push!(results, (i, :ensemble, :Asymmetry, asymmetry(net)))
+    push!(results, (i, :ensemble, :Betadiv, betadiv(realnet, net)))
 
 end
 

### 2022-02-27 code/02_simulations.jl
update final model selection

@@ -51,22 +51,20 @@ results = [DataFrame(;
 ) for _thr in 1:Threads.nthreads()]
 
 # these regression machines go brrr as f u c k
-DecisionTree = @load DecisionTreeRegressor pkg = DecisionTree
-RandomForest = @load RandomForestRegressor pkg = DecisionTree
-BoostedRegressor = @load EvoTreeRegressor pkg = EvoTrees
-RidgeRegressor = @load RidgeRegressor pkg = MLJLinearModels
-LinearRegressor = @load LinearRegressor pkg = MLJLinearModels
+DecisionTree = @load DecisionTreeRegressor pkg = DecisionTree verbosity=0
+RandomForest = @load RandomForestRegressor pkg = DecisionTree verbosity=0
+BoostedRegressor = @load EvoTreeRegressor pkg = EvoTrees verbosity=0
+KNNRegressor = @load KNNRegressor pkg = NearestNeighborModels verbosity=0
 
 candidate_models = [
     Symbol("Decision tree") => DecisionTree(),
     :BRT => BoostedRegressor(),
     Symbol("Random Forest") => RandomForest(),
-    Symbol("Ridge regression") => RidgeRegressor(),
-    Symbol("Linear regression") => LinearRegressor()
+    :kNN => KNNRegressor(),
 ]
 
 S = 100
-_n_sims = 640
+_n_sims = 600
 conditions_breadth = rand(_n_sims) .* 0.4 .+ 0.05
 conditions_bias = rand(_n_sims) .* 0.98 .+ 0.01
 conditions = hcat(conditions_breadth, conditions_bias)

### 2022-02-27 code/04_networkviz.jl

@@ -44,18 +44,16 @@ function network(S, ξ)
     return (𝐱, 𝐲)
 end
 
-DecisionTree = @load DecisionTreeRegressor pkg = DecisionTree
-RandomForest = @load RandomForestRegressor pkg = DecisionTree
-BoostedRegressor = @load EvoTreeRegressor pkg = EvoTrees
-RidgeRegressor = @load RidgeRegressor pkg = MLJLinearModels
-LinearRegressor = @load LinearRegressor pkg = MLJLinearModels
+DecisionTree = @load DecisionTreeRegressor pkg = DecisionTree verbosity=0
+RandomForest = @load RandomForestRegressor pkg = DecisionTree verbosity=0
+BoostedRegressor = @load EvoTreeRegressor pkg = EvoTrees verbosity=0
+KNNRegressor = @load KNNRegressor pkg = NearestNeighborModels verbosity=0
 
 candidate_models = [
     Symbol("Decision tree") => DecisionTree(),
     :BRT => BoostedRegressor(),
     Symbol("Random Forest") => RandomForest(),
-    Symbol("Ridge regression") => RidgeRegressor(),
-    Symbol("Linear regression") => LinearRegressor()
+    :kNN => KNNRegressor(),
 ]
 
 S = (50,50)
@@ -128,7 +126,6 @@ AUPRC = ∫(tpr.(M), ppv.(M))
 push!(Ms, 𝐌)
 ens_thres = thresholds[last(findmax(informedness.(M)))]
 
-mnames = ["DecTree", "BRT", "RF", "RR"]
 mnames = [String(p.first) for p in candidate_models]
 results = DataFrame(;
     infectivity=Float64[],
@@ -154,13 +151,7 @@ for i in 1:length(𝐲)
 end
 
 data(results) *
-    mapping(:infectivity => "Infectivity trait", :resistance => "Resistance trait", :prediction => "Prediction score"; layout=:model => sorter("BRT", "Random Forest", "Decision tree", "Ridge regression", "Linear regression", "Ensemble", "Dataset")) *
-    visual(Heatmap, colormap=:Greys) |>
+    mapping(:infectivity => "Infectivity trait", :resistance => "Resistance trait", :prediction => "Prediction score"; layout=:model => sorter("BRT", "Random Forest", "Decision tree", "kNN", "Ensemble", "Dataset")) *
+    visual(Heatmap, colormap=Reverse(:deep)) |>
[…]

### 2022-02-27 code/05_networkproperties.jl

@@ -53,18 +53,16 @@ end
 
 R(x) = R(x, extrema(x)...)
 
-DecisionTree = @load DecisionTreeRegressor pkg = DecisionTree
-RandomForest = @load RandomForestRegressor pkg = DecisionTree
-BoostedRegressor = @load EvoTreeRegressor pkg = EvoTrees
-RidgeRegressor = @load RidgeRegressor pkg = MLJLinearModels
-LinearRegressor = @load LinearRegressor pkg = MLJLinearModels
+DecisionTree = @load DecisionTreeRegressor pkg = DecisionTree verbosity=0
+RandomForest = @load RandomForestRegressor pkg = DecisionTree verbosity=0
+BoostedRegressor = @load EvoTreeRegressor pkg = EvoTrees verbosity=0
+KNNRegressor = @load KNNRegressor pkg = NearestNeighborModels verbosity=0
 
 candidate_models = [
     Symbol("Decision tree") => DecisionTree(),
     :BRT => BoostedRegressor(),
     Symbol("Random Forest") => RandomForest(),
-    Symbol("Ridge regression") => RidgeRegressor(),
-    Symbol("Linear regression") => LinearRegressor()
+    :kNN => KNNRegressor(),
 ]
 
 S = (50,40)
@@ -89,16 +87,16 @@ function betadiv(n1, n2)
     return KGL08(βos(n1, n2))
 end
 
-for i in 1:500
+for i in 1:100
 
     𝐗, 𝐲 = network(S, 0.15)
 
     net = BipartiteNetwork(reshape(Bool.(𝐲), S))
     realnet = copy(net)
-    push!(results, (i, :data, :Connectance, connectance(net)))
-    push!(results, (i, :data, :Nestedness, η(net)))
-    push!(results, (i, :data, :Modularity, Q(brim(lp(net)...)...)))
-    push!(results, (i, :data, :Asymmetry, asymmetry(net)))
+    push!(results, (i, :data, :Co, connectance(net)))
+    push!(results, (i, :data, :Nest, η(net)))
+    push!(results, (i, :data, :Mod, Q(brim(lp(net)...)...)))
+    push!(results, (i, :data, :Asymm, asymmetry(net)))
     bias = 0.5
 
     training_size = round(Int64, 0.7 * length(𝐲))
@@ -138,6 +136,9 @@ for i in 1:500
         push!(results, (i, candidate_model.first, :PRAUC, AUPRC))
         push!(results, (i, candidate_model.first, :MCC, mcc(𝐌)))
[…]

### 2022-02-28 code/05_networkproperties.jl
update code for the summary table

@@ -31,6 +31,7 @@ include(joinpath(@__DIR__, "confusionmatrix.jl"))
 function L(x::T, y::T; r::T=0.1) where {T<:Number}
     return (x - r / 2.0) ≤ y ≤ (x + r / 2.0) ? one(T) : zero(T)
 end
+
 function network(S, ξ)
     infectivity = Beta(6.0, 8.0)
     resistance = Beta(2.0, 8.0)
@@ -38,11 +39,17 @@ function network(S, ξ)
     𝐱ₕ = sort(rand(resistance, S[2]))
     𝐱₁ = repeat(𝐱ᵥ; outer=length(𝐱ₕ))
     𝐱₂ = repeat(𝐱ₕ; inner=length(𝐱ᵥ))
-    𝐱₃ = abs.(𝐱₁ .- 𝐱₂)
     𝐲 = [L(𝐱₁[i], 𝐱₂[i]; r=ξ) for i in 1:length(𝐱₁)]
-    #𝐱 = table(hcat(𝐱₁, 𝐱₂, 𝐱₃))
     𝐱 = table(hcat(𝐱₁, 𝐱₂))
     return (𝐱, 𝐲)
+end
+
+function swaps!(𝐲, s)
+    p = findall(isone, 𝐲)
+    n = findall(iszero, 𝐲)
+    𝐲[rand(p, s)] .= 0.0
+    𝐲[rand(n, s)] .= 1.0
+    return 𝐲
 end
 
 function R(x, m, M)
@@ -87,9 +94,10 @@ function betadiv(n1, n2)
     return KGL08(βos(n1, n2))
 end
 
-for i in 1:100
+for i in 1:300
 
     𝐗, 𝐲 = network(S, 0.15)
+    # swaps!(𝐲, 15)
 
     net = BipartiteNetwork(reshape(Bool.(𝐲), S))
     realnet = copy(net)

### 2022-02-28 code/03_figures.jl
updates

@@ -19,7 +19,7 @@ todrop = unique(vcat([
     @subset(results, :measure .== "TPR", :value .== 0.0).runid,
     @subset(results, :measure .== "TNR", :value .== 0.0).runid,
     @subset(results, :measure .== "PRAUC", :value .<= 0.0).runid,
-    @subset(results, :connectance .>= 0.25).runid,
+    @subset(results, :connectance .>= 0.50).runid,
     @subset(results, :connectance .<= 0.05).runid
 ]...))
 
@@ -30,7 +30,7 @@ tokeep(f) = !(f in todrop)
 connectance_values = unique(select(results, [:runid, :connectance]))
 connectance_values.midpoint = zeros(Float64, size(connectance_values,1))
 n_connectance_bins = 5
-bins_connectance = LinRange(0.05, 0.25, n_connectance_bins+1)
+bins_connectance = LinRange(0.05, 0.50, n_connectance_bins+1)
 for i in 1:n_connectance_bins
     _idx = findall(bins_connectance[i] .<= connectance_values.connectance .< bins_connectance[i+1])
     connectance_values.midpoint[_idx] .= round((bins_connectance[i+1] + bins_connectance[i])/2.0; digits=4)
@@ -39,10 +39,27 @@ end
 # Join the two dataframes
 results = leftjoin(results, select(connectance_values, [:runid, :midpoint]), on=:runid)
 
+# With connectance first
+_keepval(f) = f in ["INF", "ROCAUC", "MCC", "PRAUC"]
+
+dt = data(@subset(results, _keepval.(:measure)))
+mp = mapping(
+    :bias => "Training set bias",
+    :value => "Value";
+    row=:measure => "Measure",
+    col=:midpoint => nonnumeric => "Connectance",
+    color=:model => "Model",
+)
+ly = smooth() * visual(; linewidth=1.0)
+
+dt * mp * ly |>
+    plt -> draw(plt, facet=(;linkyaxes = :minimal)) |>
+    plt -> save(joinpath(@__DIR__, "..", "figures", "bias_combined.png"), plt, px_per_unit = 3)
+
 # Dataviz
 _keepval(f) = f in ["MCC", "INF"]
 data(@subset(results, _keepval.(:measure))) *
-    mapping(:bias => "Training set bias", :value => "Value", row=:midpoint => nonnumeric, col=:measure => "Measure", color=:model => "Model") *
[…]

### 2022-02-28 code/02_simulations.jl
super broken version

@@ -10,35 +10,17 @@ using CSV: CSV
 using Distributions
 
 # Simulation suffix
-_suffix = get(ENV, "SLURM_ARRAY_TASK_ID", "global")
-
-# AUC
-function ∫(x::Array{T}, y::Array{T}) where {T<:Number}
-    S = zero(Float64)
-    for i in 2:length(x)
-        S += (x[i] - x[i - 1]) * (y[i] + y[i - 1]) * 0.5
-    end
-    return .-S
-end
+_jobid = get(ENV, "SLURM_ARRAY_TASK_ID", 1)
+_jobcount = get(ENV, "SLURM_ARRAY_TASK_COUNT", 1)
 
 # Confusion matrix utilities
 include(joinpath(@__DIR__, "confusionmatrix.jl"))
 
-# Network generation function
-L(x::T, y::T; r::T=0.1) where {T<:Number} = (x-r/2.0) ≤ y ≤ (x+r/2.0)  ? one(T) : zero(T)
-function network(S, ξ)
-    infectivity = Beta(6.0, 8.0)
-    resistance = Beta(2.0, 8.0)
-    𝐱ᵥ = sort(rand(infectivity, S))
-    𝐱ₕ = sort(rand(resistance, S))
-    𝐱₁ = repeat(𝐱ᵥ; outer=length(𝐱ₕ))
-    𝐱₂ = repeat(𝐱ₕ; inner=length(𝐱ᵥ))
-    𝐱₃ = abs.(𝐱₁ .- 𝐱₂)
-    𝐲 = [L(𝐱₁[i], 𝐱₂[i]; r=ξ) for i in 1:(S*S)]
-    #𝐱 = table(hcat(𝐱₁, 𝐱₂, 𝐱₃))
-    𝐱 = table(hcat(𝐱₁, 𝐱₂))
-    return (𝐱, 𝐲)
-end
+# Network simulation utilities
+include(joinpath(@__DIR__, "networksimulation.jl"))
+
+# Choice of models
+include(joinpath(@__DIR__, "modelchoice.jl"))
 
 # Prepare the results
 results = [DataFrame(;
@@ -50,26 +32,20 @@ results = [DataFrame(;
     value=Float64[],
 ) for _thr in 1:Threads.nthreads()]
 
-# these regression machines go brrr as f u c k
-DecisionTree = @load DecisionTreeRegressor pkg = DecisionTree verbosity=0
[…]

### 2022-02-28 code/confusionmatrix.jl

@@ -37,3 +37,12 @@ function κ(M::ConfusionMatrix)
            ((M.tp + M.fp) * (M.fp + M.tn) + (M.tp + M.fn) * (M.fn + M.tn))
 end
 mcc(M::ConfusionMatrix) = (M.tp*M.tn-M.fp*M.fn)/sqrt((M.tp+M.fp)*(M.tp+M.fn)*(M.tn+M.fp)*(M.tn+M.fn))
+
+function ∫(x::Array{T}, y::Array{T}) where {T<:Number}
+    S = zero(Float64)
+    for i in 2:length(x)
+        S += (x[i] - x[i - 1]) * (y[i] + y[i - 1]) * 0.5
+    end
+    return .-S
+end
+

### 2022-02-28 code/modelchoice.jl

@@ -1,0 +1,13 @@
+# these regression machines go brrr as f u c k
+DecisionTree = @load DecisionTreeRegressor pkg = DecisionTree verbosity=0
+RandomForest = @load RandomForestRegressor pkg = DecisionTree verbosity=0
+BoostedRegressor = @load EvoTreeRegressor pkg = EvoTrees verbosity=0
+KNNRegressor = @load KNNRegressor pkg = NearestNeighborModels verbosity=0
+
+# Object for candidate models
+candidate_models = [
+    Symbol("Decision tree") => DecisionTree(),
+    :BRT => BoostedRegressor(),
+    Symbol("Random Forest") => RandomForest(),
+    :kNN => KNNRegressor(),
+]

### 2022-02-28 code/networksimulation.jl

@@ -1,0 +1,16 @@
+L(x::T, y::T; r::T=0.1) where {T<:Number} = (x-r/2.0) ≤ y ≤ (x+r/2.0)  ? one(T) : zero(T)
+
+function network(S, ξ)
+    infectivity = Beta(6.0, 8.0)
+    resistance = Beta(2.0, 8.0)
+    𝐱ᵥ = sort(rand(infectivity, S))
+    𝐱ₕ = sort(rand(resistance, S))
+    𝐱₁ = repeat(𝐱ᵥ; outer=length(𝐱ₕ))
+    𝐱₂ = repeat(𝐱ₕ; inner=length(𝐱ᵥ))
+    𝐱₃ = abs.(𝐱₁ .- 𝐱₂)
+    𝐲 = [L(𝐱₁[i], 𝐱₂[i]; r=ξ) for i in 1:(S*S)]
+    #𝐱 = table(hcat(𝐱₁, 𝐱₂, 𝐱₃))
+    𝐱 = table(hcat(𝐱₁, 𝐱₂))
+    return (𝐱, 𝐲)
+end
+

### 2022-02-28 code/02_simulations.jl
massive simplifcation to the simulation code

@@ -22,140 +22,108 @@ include(joinpath(@__DIR__, "networksimulation.jl"))
 # Choice of models
 include(joinpath(@__DIR__, "modelchoice.jl"))
 
-# Prepare the results
-results = [DataFrame(;
-    breadth=Float64[],
-    bias=Float64[],
-    connectance=Float64[],
-    model=Symbol[],
-    measure=Symbol[],
-    value=Float64[],
-) for _thr in 1:Threads.nthreads()]
+# Function to split the sets
+include(joinpath(@__DIR__, "setsplitter.jl"))
 
-# Block for the gridded simulations
+# Prepare the results
+results = [
+    DataFrame(;
+        links=Int64[], bias=Float64[], model=Symbol[], measure=Symbol[], value=Float64[]
+    ) for _thr in 1:Threads.nthreads()
+]
+
+# Network size
 S = (100, 100)
 
-grid_size = 80
-links = LinRange(minimum(S)+10, round(Int64, 0.5*prod(S)), grid_size)
+# Number of simulation blocks
+grid_size = 6
+
+# List of conditions
+conditions = [(i, j) for i in 1:grid_size for j in 1:grid_size]
+
+links = round.(Int64, LinRange(minimum(S) + 10, round(Int64, 0.5 * prod(S)), grid_size))
 biases = LinRange(0.02, 0.98, grid_size)
 
-conditions = hcat(conditions_breadth, conditions_bias)
-
-Threads.@threads for connectance in connectances
-    # Get the network for this connectance run
-    for bias in biases
-    end
+# Prepare networks for the gridded simulations
+nets = []
+for link in links
+    approx_conn = link / prod(S)
+    candidates = [network(S, approx_conn) for i in 1:10_000]
+    idx = last(findmin(n -> abs(approx_conn - mean(last(n))), candidates))
+    push!(nets, candidates[idx])
 end
-    breadth, bias = conditions[i, :]
-    @info i, Threads.threadid(), breadth, bias
 
-    𝐗, 𝐲 = network(S, breadth)
-    _real_co = mean(𝐲)
+# Function to range the predictions
+R = (v) -> (v .- minimum(v)) ./ (maximum(v) - minimum(v))
 
-    training_size = round(Int64, 0.5 * length(𝐲))
[…]

### 2022-02-28 code/confusionmatrix.jl

@@ -46,3 +46,19 @@ function ∫(x::Array{T}, y::Array{T}) where {T<:Number}
     return .-S
 end
 
+function threshold(obs, pred; levels=500)
+    thresholds = LinRange(minimum(pred), maximum(pred), levels)
+    M = Vector{ConfusionMatrix}(undef, length(thresholds))
+    for (i, τ) in enumerate(thresholds)
+        binpred = pred .>= τ
+        tp = sum(obs .& binpred)
+        tn = sum(.!obs .& .!binpred)
+        fp = sum(.!obs .& binpred)
+        fn = sum(obs .& .!binpred)
+        M[i] = ConfusionMatrix(tp, tn, fp, fn)
+    end
+    ROCAUC = ∫(fpr.(M), tpr.(M))
+    AUPRC = ∫(tpr.(M), ppv.(M))
+    𝐌 = M[last(findmax(informedness.(M)))]
+    return 𝐌, ROCAUC, AUPRC
+end

### 2022-02-28 code/networksimulation.jl

@@ -3,13 +3,11 @@ L(x::T, y::T; r::T=0.1) where {T<:Number} = (x-r/2.0) ≤ y ≤ (x+r/2.0)  ? one
 function network(S, ξ)
     infectivity = Beta(6.0, 8.0)
     resistance = Beta(2.0, 8.0)
-    𝐱ᵥ = sort(rand(infectivity, S))
-    𝐱ₕ = sort(rand(resistance, S))
+    𝐱ᵥ = sort(rand(infectivity, S[1]))
+    𝐱ₕ = sort(rand(resistance, S[2]))
     𝐱₁ = repeat(𝐱ᵥ; outer=length(𝐱ₕ))
     𝐱₂ = repeat(𝐱ₕ; inner=length(𝐱ᵥ))
-    𝐱₃ = abs.(𝐱₁ .- 𝐱₂)
-    𝐲 = [L(𝐱₁[i], 𝐱₂[i]; r=ξ) for i in 1:(S*S)]
-    #𝐱 = table(hcat(𝐱₁, 𝐱₂, 𝐱₃))
+    𝐲 = [L(𝐱₁[i], 𝐱₂[i]; r=ξ) for i in 1:prod(S)]
     𝐱 = table(hcat(𝐱₁, 𝐱₂))
     return (𝐱, 𝐲)
 end

### 2022-02-28 code/setsplitter.jl

@@ -1,0 +1,25 @@
+function split_indices(𝐲, bias; split=0.5)
+    # Get the training size
+    training_size = round(Int64, split * length(𝐲))
+
+    # Number of positive samples in the training set
+    n_positive = round(Int64, training_size * bias)
+
+    # Locate positive and negative cases
+    idx_pos = sample(findall(iszero.(𝐲)), n_positive; replace=true)
+    idx_neg = sample(findall(isone.(𝐲)), training_size - n_positive; replace=true)
+
+    # Get the training set and the initial testing set
+    Iₚ = shuffle(vcat(idx_neg, idx_pos))
+    Iₒ = setdiff(eachindex(𝐲), Iₚ)
+
+    # Correct the testing set to have it keep the connectance of the network
+    _test_pos = sum(𝐲[Iₒ])
+    _expected_neg = round(Int64, _test_pos/mean(𝐲) - _test_pos)
+    _observed_neg = round(Int64, length(Iₒ)-sum(𝐲[Iₒ]))
+    tst_neg = sample(findall(iszero.(𝐲[Iₒ])), max(_observed_neg - _expected_neg, 1); replace=true)
+    deleteat!(Iₒ, sort(unique(tst_neg)))
+    
+    # Return the INDICES of the training and testing set, in that order
+    return Iₚ, Iₒ
+end

### 2022-02-28 code/02_simulations.jl
update the grid size for narval simuls

@@ -36,7 +36,7 @@ results = [
 S = (100, 100)
 
 # Number of simulation blocks
-grid_size = 6
+grid_size = 25
 
 # List of conditions
 conditions = [(i, j) for i in 1:grid_size for j in 1:grid_size]

### 2022-02-28 code/03_figures.jl
update fig code

@@ -11,47 +11,39 @@ allowmissing!(results, :value)
 replace!(results.value, NaN => missing)
 dropmissing!(results)
 
-# Drop the models with obviously borked classifiers
-results.runid = hash.(results.connectance .* results.breadth .* results.bias)
+bmm = groupby(results, [:model, :measure, :links, :bias])
+raw = combine(bmm, :value => mean => :value)
 
-todrop = unique(vcat([
-    @subset(results, :measure .== "ACC", :value .== 0.0).runid,
-    @subset(results, :measure .== "TPR", :value .== 0.0).runid,
-    @subset(results, :measure .== "TNR", :value .== 0.0).runid,
-    @subset(results, :measure .== "PRAUC", :value .<= 0.0).runid,
-    @subset(results, :connectance .>= 0.50).runid,
-    @subset(results, :connectance .<= 0.05).runid
-]...))
+raw.connectance = raw.links ./ (100*100)
 
-tokeep(f) = !(f in todrop)
-@subset!(results, tokeep.(:runid))
+# Give each run a UUID
+# results.runid = hash.(results.link .* results.bias .* results.bias)
 
 # Make bins for connectance to make the plotting more efficient
-connectance_values = unique(select(results, [:runid, :connectance]))
-connectance_values.midpoint = zeros(Float64, size(connectance_values,1))
-n_connectance_bins = 5
-bins_connectance = LinRange(0.05, 0.50, n_connectance_bins+1)
-for i in 1:n_connectance_bins
-    _idx = findall(bins_connectance[i] .<= connectance_values.connectance .< bins_connectance[i+1])
-    connectance_values.midpoint[_idx] .= round((bins_connectance[i+1] + bins_connectance[i])/2.0; digits=4)
-end
+# connectance_values = unique(select(results, [:runid, :connectance]))
+# connectance_values.midpoint = zeros(Float64, size(connectance_values,1))
+# n_connectance_bins = 5
+# bins_connectance = LinRange(0.01, 0.50, n_connectance_bins+1)
+# for i in 1:n_connectance_bins
+#     _idx = findall(bins_connectance[i] .<= connectance_values.connectance .< bins_connectance[i+1])
+#     connectance_values.midpoint[_idx] .= round((bins_connectance[i+1] + bins_connectance[i])/2.0; digits=4)
+# end
 
-# Join the two dataframes
[…]

### 2022-02-28 code/02_simulations.jl
new run with some more grid points

@@ -36,7 +36,7 @@ results = [
 S = (100, 100)
 
 # Number of simulation blocks
-grid_size = 25
+grid_size = 35
 
 # List of conditions
 conditions = [(i, j) for i in 1:grid_size for j in 1:grid_size]

### 2022-02-28 code/03_figures.jl

@@ -14,83 +14,54 @@ dropmissing!(results)
 bmm = groupby(results, [:model, :measure, :links, :bias])
 raw = combine(bmm, :value => mean => :value)
 
+# Derived measure of connectance
 raw.connectance = raw.links ./ (100*100)
 
-# Give each run a UUID
-# results.runid = hash.(results.link .* results.bias .* results.bias)
-
-# Make bins for connectance to make the plotting more efficient
-# connectance_values = unique(select(results, [:runid, :connectance]))
-# connectance_values.midpoint = zeros(Float64, size(connectance_values,1))
-# n_connectance_bins = 5
-# bins_connectance = LinRange(0.01, 0.50, n_connectance_bins+1)
-# for i in 1:n_connectance_bins
-#     _idx = findall(bins_connectance[i] .<= connectance_values.connectance .< bins_connectance[i+1])
-#     connectance_values.midpoint[_idx] .= round((bins_connectance[i+1] + bins_connectance[i])/2.0; digits=4)
-# end
-
-# # Join the two dataframes
-# results = leftjoin(results, select(connectance_values, [:runid, :midpoint]), on=:runid)
-
-# With connectance first
+# Show the core measures by connectance
 _keepval(f) = f in ["ROC", "Y", "PR", "mcc"]
-
 dt = data(@subset(raw, _keepval.(:measure)))
 mp = mapping(
     :bias => "Training set bias",
     :value => "Value";
     color=:connectance => "Network connectance",
     row=:measure => "Measure",
-    col=:model => "Model",
+    col=:model => "Model"
 )
-ly = visual(Scatter, colormap = :cividis, markersize = 4)
+ly = visual(Scatter, colormap = :deep, markersize = 4)
 dt * mp * ly |>
     plt -> draw(plt, facet=(;linkyaxes = :minimal)) |>
-    plt -> save(joinpath(@__DIR__, "..", "figures", "bias_combined.png"), plt, px_per_unit = 3)
+    plt -> save(joinpath(@__DIR__, "..", "figures", "bias_by_connectance.png"), plt, px_per_unit = 3)
 
-# Dataviz
-_keepval(f) = f in ["MCC", "INF"]
-data(@subset(results, _keepval.(:measure))) *
[…]

### 2022-03-01 code/02_simulations.jl
wrap 02

@@ -109,7 +109,9 @@ Threads.@threads for i in 1:length(conditions)
             push!(results[Threads.threadid()], (link, bias, model.first, :ROC, ROC))
             push!(results[Threads.threadid()], (link, bias, model.first, :PR, PR))
             for (mname, mfunc) in measures
-                push!(results[Threads.threadid()], (link, bias, model.first, mname, mfunc(𝐌)))
+                push!(
+                    results[Threads.threadid()], (link, bias, model.first, mname, mfunc(𝐌))
+                )
             end
         end
     end

### 2022-03-01 code/03_figures.jl

@@ -27,7 +27,7 @@ mp = mapping(
     row=:measure => "Measure",
     col=:model => "Model"
 )
-ly = visual(Scatter, colormap = :deep, markersize = 4)
+ly = visual(Scatter, colormap = :bamako, markersize = 4)
 dt * mp * ly |>
     plt -> draw(plt, facet=(;linkyaxes = :minimal)) |>
     plt -> save(joinpath(@__DIR__, "..", "figures", "bias_by_connectance.png"), plt, px_per_unit = 3)

### 2022-03-01 code/03_figures.jl
new figures ready for writing

@@ -23,9 +23,9 @@ dt = data(@subset(raw, _keepval.(:measure)))
 mp = mapping(
     :bias => "Training set bias",
     :value => "Value";
-    color=:connectance => "Network connectance",
-    row=:measure => "Measure",
-    col=:model => "Model"
+    color = :connectance => "Network connectance",
+    row = :measure => "Measure",
+    col = :model => "Model",
 )
 ly = visual(Scatter, colormap = :bamako, markersize = 4)
 dt * mp * ly |>
@@ -41,27 +41,27 @@ end
 opt = combine(bmm, [:value, :bias] => ((v,b) -> optival(v,b)) => AsTable)
 
 _keepval(f) = f in ["ROC", "Y", "PR", "mcc"]
+
 dt = data(@subset(opt, _keepval.(:measure)))
 mp = mapping(
     :connectance => "Connectance",
-    :bias => "Optimal bias";
-    row=:measure => "Measure",
-    col=:model => "Model"
+    :bias => "Optimal training balance";
+    layout = :measure => "Measure",
+    color = :model => sorter("Decision tree", "kNN", "Random Forest", "BRT", "Ensemble") => "Model"
 )
-ly = visual(Scatter, colormap = :deep, markersize = 4)
+ly = visual(ScatterLines, colormap = :deep, markersize = 4)
 dt * mp * ly |>
     plt -> draw(plt, facet=(;linkyaxes = :minimal)) |>
     plt -> save(joinpath(@__DIR__, "..", "figures", "optimal_bias.png"), plt, px_per_unit = 3)
 
-_keepval(f) = f in ["ROC", "Y", "PR", "mcc"]
 dt = data(@subset(opt, _keepval.(:measure)))
 mp = mapping(
     :connectance => "Connectance",
-    :value => "Optimal value";
-    row=:measure => "Measure",
-    col=:model => "Model"
+    :value => "Value at optimal training balance";
+    layout = :measure => "Measure",
+    color = :model => sorter("Decision tree", "kNN", "Random Forest", "BRT", "Ensemble") => "Model"
 )
-ly = visual(Scatter, colormap = :deep, markersize = 4)
+ly = visual(ScatterLines, colormap = :deep, markersize = 4)
 dt * mp * ly |>
     plt -> draw(plt, facet=(;linkyaxes = :minimal)) |>
     plt -> save(joinpath(@__DIR__, "..", "figures", "optimal_value.png"), plt, px_per_unit = 3)

### 2022-03-01 code/02_simulations.jl
update the figures

@@ -98,16 +98,17 @@ Threads.@threads for i in 1:length(conditions)
         fit!(this_machine; rows=Iₚ)
 
         # Predict on the validation set
-        prediction = MLJ.predict(this_machine; rows=Iₒ)
+        prediction = R(MLJ.predict(this_machine; rows=Iₒ))
 
         # Thresholding analysis
-        𝐌, ROC, PR = threshold(𝐲[Iₒ] .> 0.0, prediction)
+        𝐌, ROC, PR, τ = threshold(𝐲[Iₒ] .> 0.0, prediction)
 
         # Write the outputs
         if !isnan(ROC)
             predictions[model.first] = R(prediction)
             push!(results[Threads.threadid()], (link, bias, model.first, :ROC, ROC))
             push!(results[Threads.threadid()], (link, bias, model.first, :PR, PR))
+            push!(results[Threads.threadid()], (link, bias, model.first, :threshold, τ))
             for (mname, mfunc) in measures
                 push!(
                     results[Threads.threadid()], (link, bias, model.first, mname, mfunc(𝐌))
@@ -119,9 +120,10 @@ Threads.@threads for i in 1:length(conditions)
     # Ensemble model
     if !isempty(predictions)
         predictions[:Ensemble] = R(mean(hcat(collect(values(predictions))...); dims=2))
-        𝐌, ROC, PR = threshold(𝐲[Iₒ] .> 0.0, predictions[:Ensemble])
+        𝐌, ROC, PR, τ = threshold(𝐲[Iₒ] .> 0.0, predictions[:Ensemble])
         push!(results[Threads.threadid()], (link, bias, :Ensemble, :ROC, ROC))
         push!(results[Threads.threadid()], (link, bias, :Ensemble, :PR, PR))
+        push!(results[Threads.threadid()], (link, bias, :Ensemble, :threshold, τ))
         for (mname, mfunc) in measures
             push!(results[Threads.threadid()], (link, bias, :Ensemble, mname, mfunc(𝐌)))
         end

### 2022-03-01 code/04_networkviz.jl

@@ -12,146 +12,110 @@ using AlgebraOfGraphics, CairoMakie
 using EcologicalNetworks
 using Random
 
-Random.seed!(5)
-
-# AUC
-function ∫(x::Array{T}, y::Array{T}) where {T<:Number}
-    S = zero(Float64)
-    for i in 2:length(x)
-        S += (x[i] - x[i - 1]) * (y[i] + y[i - 1]) * 0.5
-    end
-    return .-S
-end
+Random.seed!(1)
 
 # Confusion matrix utilities
 include(joinpath(@__DIR__, "confusionmatrix.jl"))
 
-# Network generation function
-function L(x::T, y::T; r::T=0.1) where {T<:Number}
-    return (x - r / 2.0) ≤ y ≤ (x + r / 2.0) ? one(T) : zero(T)
-end
-function network(S, ξ)
-    infectivity = Beta(6.0, 8.0)
-    resistance = Beta(2.0, 8.0)
-    𝐱ᵥ = sort(rand(infectivity, S[1]))
-    𝐱ₕ = sort(rand(resistance, S[2]))
-    𝐱₁ = repeat(𝐱ᵥ; outer=length(𝐱ₕ))
-    𝐱₂ = repeat(𝐱ₕ; inner=length(𝐱ᵥ))
-    𝐱₃ = abs.(𝐱₁ .- 𝐱₂)
-    𝐲 = [L(𝐱₁[i], 𝐱₂[i]; r=ξ) for i in 1:length(𝐱₁)]
-    # 𝐱 = table(hcat(𝐱₁, 𝐱₂, 𝐱₃))
-    𝐱 = table(hcat(𝐱₁, 𝐱₂))
-    return (𝐱, 𝐲)
-end
+# Network simulation utilities
+include(joinpath(@__DIR__, "networksimulation.jl"))
 
-DecisionTree = @load DecisionTreeRegressor pkg = DecisionTree verbosity=0
-RandomForest = @load RandomForestRegressor pkg = DecisionTree verbosity=0
-BoostedRegressor = @load EvoTreeRegressor pkg = EvoTrees verbosity=0
-KNNRegressor = @load KNNRegressor pkg = NearestNeighborModels verbosity=0
+# Choice of models
+include(joinpath(@__DIR__, "modelchoice.jl"))
 
-candidate_models = [
-    Symbol("Decision tree") => DecisionTree(),
-    :BRT => BoostedRegressor(),
[…]

### 2022-03-01 code/confusionmatrix.jl

@@ -60,5 +60,6 @@ function threshold(obs, pred; levels=500)
     ROCAUC = ∫(fpr.(M), tpr.(M))
     AUPRC = ∫(tpr.(M), ppv.(M))
     𝐌 = M[last(findmax(informedness.(M)))]
-    return 𝐌, ROCAUC, AUPRC
+    τ = thresholds[last(findmax(informedness.(M)))]
+    return 𝐌, ROCAUC, AUPRC, τ 
 end

### 2022-03-01 code/03_figures.jl
update

@@ -21,11 +21,11 @@ raw.connectance = raw.links ./ (100*100)
 _keepval(f) = f in ["ROC", "Y", "PR", "mcc"]
 dt = data(@subset(raw, _keepval.(:measure)))
 mp = mapping(
-    :bias => "Training set bias",
+    :bias => "Training balance",
     :value => "Value";
     color = :connectance => "Network connectance",
     row = :measure => "Measure",
-    col = :model => "Model",
+    col = :model => sorter("Decision tree", "kNN", "Random Forest", "BRT", "Ensemble") => "Model"
 )
 ly = visual(Scatter, colormap = :bamako, markersize = 4)
 dt * mp * ly |>

### 2022-03-01 draft.tex
Deploying to gh-pages from @ PoisotLab/ms_networkary_classifiers@1d32b44983ade554d3e68ee64c93f37464e0de05 🚀

@@
 in the network, \emph{i.e.} \(n=3000\). Out of these interactions, we
-pick a proportion \(\nu\) (the training set bias) to be positive, so
+pick a proportion \(\nu\) (the training set balance) to be positive, so
 that the training set has \(\nu n\) interactions, and \((1-\nu) n\)
@@
 put through the same thresholding process. The complete code to run the
-simulations is available at \texttt{10.17605/OSF.IO/JKEWD}; the
-simulations require approx. 5 core days.
+simulations is available at \texttt{10.17605/OSF.IO/JKEWD}.
 
@@
 
-\hypertarget{effect-of-training-set-bias-on-performance}{%
-\subsection{Effect of training set bias on
-performance}\label{effect-of-training-set-bias-on-performance}}
-
-In fig.~\ref{fig:biasmccinf}, we present the response of MCC and
-informedness to (i) five levels of network connectance and (ii) a
-gradient of training set bias, for the four component models as well as
-the ensemble. All models reached a higher performance on more connected
-networks, and using more biased training sets (with the exception of
-ridge regression, whose informedness decreased in performance with
-training set bias). In all cases, informedness was extremely high, which
-is an expected consequence of the fact that this is the value we
-optimized to determine the cutoff. MCC increased with training set bias,
-although this increase became less steep with increasing connectance.
+\hypertarget{effect-of-training-set-balance-on-performance}{%
+\subsection{Effect of training set balance on
+performance}\label{effect-of-training-set-balance-on-performance}}
+
+In fig.~\ref{fig:biasco}, we present the response of two thresholding
+measures (PR-AUC and ROC-AUC) and two ranking measures (Informedness and
+MCC) to a grid of 35 values of training set balance, and 35 values of
+connectance, for the four component models as well as the ensemble.
+ROC-AUC is always high, and does not vary with training set balance. On
+the other hand, PR-AUC shows very strong responses, increasing with
+training set balance. It is notable here that two classifiers that
+seemed to be performing well (Decision Tree and Random Forest) based on
+their MCC are not able to reach a high PR-AUC even at higher
+connectances.All models reached a higher performance on more connected
+networks, and using more balanced training sets. In all cases,
+informedness was extremely high, which is an expected consequence of the
+fact that this is the value we optimized to determine the cutoff. MCC
+increased with training set balance, although this increase became less
+steep with increasing connectance. Three of the models (kNN, decision
+tree, and random forest) only increased their PR-AUC sharply when the
+training set was heavily imbalanced towards more interactions.
 Interestingly, the ensemble almost always outclassed its component
-models. In a few cases, both MCC and informedness stared decreasing when
-the training set bias got too close to one, which suggests that it is
-possible to over-correct the imbalance.
+models. For larger connectances (less difficult networks to predict, as
+they are more balanced), MCC and informedness stared decreasing when the
+training set bias got too close to one, suggesting that a training set
+balance of 0.5 may often be appropriate if these measures are the one to
+optimize.
 
 \begin{figure}
-\hypertarget{fig:biasmccinf}{%
+\hypertarget{fig:biasco}{%
 \centering
-\includegraphics{figures/bias_mcc_inf.png}
-\caption{Response of MCC and Informedness to changes in the training set
-bias for a fixed connectance (rows). Both of these values approach 1 for
-a good model. Informedness is consistently high, and by contrast, MCC
-increases with additional training set bias. Across all models, training
-on a more connected network is easier.}\label{fig:biasmccinf}
+\includegraphics{figures/bias_by_connectance.png}
+\caption{Response of MCC, Informedness, ROC-AUC, and PR-AUC to changes
+in the training set balance (on the \(x\) axis) for a series of
+increasing connectances (color). All of these values approach 1 for a
+good model, but should be lower when the prediction is more difficult.
+Informedness is consistently high, and by contrast, MCC increases with
+additional training set balance. Across all models, training on a more
+connected network is easier. ROC-AUC is consistently high, and therefore
+not properly able to separate good from poor classifiers. On the other
+hand, PR-AUC responds to changes in the training set.}\label{fig:biasco}
 }
@@
 
-In fig.~\ref{fig:biasrocpr}, we present the same information as
-fig.~\ref{fig:biasmccinf}, this time using ROC-AUC and PR-AUC. ROC-AUC
-is always high, and does not vary with training set bias. On the other
-hand, PR-AUC shows very strong responses, increasing with training set
-bias. It is notable here that two classifiers that seemed to be
-performing well (Decision Tree and Random Forest) based on their MCC are
-not able to reach a high PR-AUC even at higher connectances. As in
-fig.~\ref{fig:biasmccinf}, the ensemble outperforms its component
-models.
-
-\begin{figure}
-\hypertarget{fig:biasrocpr}{%
-\centering
-\includegraphics{figures/bias_roc_pr.png}
-\caption{Response of ROC-AUC and PR-AUC to changes in the training set
-bias for a fixed connectance (rows). ROC-AUC is consistently high, and
-therefore not properly able to separate good from poor classifiers. On
-the other hand, PR-AUC responds to changes in the training set. As in
-fig.~\ref{fig:biasmccinf}, training on more connected networks is
-easier.}\label{fig:biasrocpr}
-}
-\end{figure}
-
-Based on the results presented in fig.~\ref{fig:biasmccinf} and
-fig.~\ref{fig:biasrocpr}, it seems that informedness and ROC-AUC are not
-necessarily able to discriminate between good and bad classifiers
-(although this result may be an artifact for informedness, as it has
-been optimized when thresholding). On the other hand, MCC and PR-AUC
-show a strong response to training set bias, and may therefore be more
-useful at model comparison.
+Based on the results presented in fig.~\ref{fig:biasco}, it seems that
+informedness and ROC-AUC are not necessarily able to discriminate
+between good and bad classifiers (although this result may be an
+artifact for informedness, as it has been optimized when thresholding).
+On the other hand, MCC and PR-AUC show a strong response to training set
+balance, and may therefore be more useful at model comparison.
 
@@
 valuable than negatives, on which there is a doubt). For this reason,
-across all values of connectance, we measured the training set bias that
-maximized a series of performance measures. When this value is high, the
-training set needs to skew more positive in order to get a performant
-model; when this value is about 0.5, the training set needs to be
-artificially balanced to optimize the model performance. These results
-are presented in fig.~\ref{fig:optimbias}.
+across all values of connectance, we measured the training set balance
+that maximized a series of performance measures. When this value is
+high, the training set needs to skew more positive in order to get a
+performant model; when this value is about 0.5, the training set needs
+to be artificially balanced to optimize the model performance. These
+results are presented in fig.~\ref{fig:optimbias}.
 
@@
 \includegraphics{figures/optim_bias.png}
-\caption{Value of the optimal training set bias for the different models
-and measures evaluated here, over a range of connectances. Informedness
-was reliably maximized for balanced training sets, and kept this
-behavior across models. For other measures, larger connectances in the
-true network allowed lower biases in the training set. In a large number
-of cases, ``over-correcting'' by having training sets with more than
-half instances representing interactions would maximize the values of
-the model performance measures.}\label{fig:optimbias}
+\caption{Value of the optimal training set balance for the different
+models and measures evaluated here, over a range of connectances.
+Informedness was reliably maximized for balanced training sets, and kept
+this behavior across models. For other measures, larger connectances in
+the true network allowed lower biases in the training set. In a large
+number of cases, ``over-correcting'' by having training sets with more
+than half instances representing interactions would maximize the values
+of the model performance measures.}\label{fig:optimbias}
 }
@@
 the original network. MCC and PR-AUC required values of training set
-bias from 0.75 to almost 1 to be optimized, which is in line with the
+balance from 0.75 to almost 1 to be optimized, which is in line with the
 results of the previous section, \emph{i.e.} they are more stringent
@@
 dataset with very low connectance can be a different task than for more
-connected networks: it becomes increasingly important to caputre the
+connected networks: it becomes increasingly important to capture the
 mechanisms that make an interaction \emph{exist}, and therefore having a
@@
 \begin{figure}
-\hypertarget{fig:optimperf}{%
+\hypertarget{fig:optimvalue}{%
 \centering
-\includegraphics{figures/optim_perf.png}
+\includegraphics{figures/optim_value.png}
 \caption{When trained on their optimally biased training set, most
@@
 ensemble had a consistently high performance despite incorporating poor
-models.}\label{fig:optimperf}
+models.}\label{fig:optimvalue}
 }
@@
 
-When trained at their optimal training set bias, connectance still had a
-significant impact on the performance of some machines
-fig.~\ref{fig:optimperf}. Notably, Decision Tree, Random Forest, and
+When trained at their optimal training set balance, connectance still
+had a significant impact on the performance of some machines
+fig.~\ref{fig:optimvalue}. Notably, Decision Tree, Random Forest, and
 Ridge Regression had low values of PR-AUC. In all cases, the Boosted
-Regression Tree was reaching very good predictions (esepcially for
+Regression Tree was reaching very good predictions (especially for
 connectances larger than 0.1), and the ensemble was almost always
@@
 correct these biases. We do not expect this last result to have any
-generality, and provide a discussion of a recent exemple in which the
+generality, and provide a discussion of a recent example in which the
 ensemble was performing worse than its components models.
@@
 before, with \(S_1, S_2 = 50, 80\) species, a connectance of
-\(\approx 0.16\) (\(\xi = 0.19\)), and a training set bias of \(0.7\).
-The prediction made on the complete dataset is presented in
-fig.~\ref{fig:ecovalid}. Visualizing the results this way highlights the
-importance of exploratory data analysis: whereas all models return a
-network with interactions laying mostly on the diagonal (as expected),
-the Ridge Regression is quite obviously biased. Despite this, we can see
-that the ensemble is close to the initial dataset.
+\(\approx 0.16\) (\(\xi = 0.19\)), and a training set balance of
+\(0.5\), as fig.~\ref{fig:optimbias} suggests this is the optimal
+training set balance for this range of connectance. The prediction made
+on the complete dataset is presented in fig.~\ref{fig:ecovalid}.
+Visualizing the results this way highlights the importance of
+exploratory data analysis: whereas all models return a network with
+interactions laying mostly on the diagonal (as expected), the Ridge
+Regression is quite obviously biased. Despite this, we can see that the
+ensemble is close to the initial dataset.
 
@@
 analysis; specifically, we measured the connectance, nestedness (REF),
-and modularity (REF). This process was repeated 250 times, and the
-results are presented in tbl.~\ref{tbl:comparison}. The random forest
-model is an interesting instance here: it produces the network that
-looks the most like the original dataset, despite having a very low
-PR-AUC, suggesting it hits high recall at the cost of low precision.
-Although the ensemble was able to reach a very high PR-AUC (and a very
-high ROC-AUC), this did not necessarily translate into more accurate
-reconstructions of the structure of the network. This result bears
-elaborating. Measures of model performance capture how much of the
-interactions and non-interactions are correctly identified. As long as
-these predictions are not perfect, some interactions will be predicted
-at the ``wrong'' position in the network; these measures cannot describe
-the structural effect of these mistakes. On the other hand, measures of
-network structure can have the same value with interactions that fall at
+modularity (REF), asymmetry, and network dissimilarity (REF). This
+process was repeated 250 times, and the results are presented in
+tbl.~\ref{tbl:comparison}. The random forest model is an interesting
+instance here: it produces the network that looks the most like the
+original dataset, despite having a very low PR-AUC, suggesting it hits
+high recall at the cost of low precision. Although the ensemble was able
+to reach a very high PR-AUC (and a very high ROC-AUC), this did not
+necessarily translate into more accurate reconstructions of the
+structure of the network. This result bears elaborating. Measures of
+model performance capture how much of the interactions and
+non-interactions are correctly identified. As long as these predictions
+are not perfect, some interactions will be predicted at the ``wrong''
+position in the network; these measures cannot describe the structural
+effect of these mistakes. On the other hand, measures of network
+structure can have the same value with interactions that fall at
 drastically different positions; this is in part because a lot of these
@@
 \caption{\label{tbl:comparison}Values of four performance metrics, and
-three network structure metrics, for 250 independent predictions similar
+five network structure metrics, for 500 independent predictions similar
 to the ones presented in fig.~\ref{fig:ecovalid}. The values in
@@
 Second, because the PR-AUC responds more to network connectance
-fig.~\ref{fig:optimperf} and training set imbalance
+fig.~\ref{fig:optimvalue} and training set imbalance
 fig.~\ref{fig:biasrocpr}, it should be used as a measure of model
@@
 Thirdly, regardless of network connectance, maximizing informedness
-required a training set bias of about 0.5, and maximizing the MCC
+required a training set balance of about 0.5, and maximizing the MCC
 required a training set bias of 0.7 and more. This has an important
@@
 assessed. The distribution of results in fig.~\ref{fig:optimbias} and
-fig.~\ref{fig:optimperf} show that there are variations around the
+fig.~\ref{fig:optimvalue} show that there are variations around the
 trend; furthermore, networks with different structures than the one we

### 2022-03-02 code/06_overfit.jl
some updates

@@ -1,0 +1,141 @@
+using Statistics
+using StatsFuns
+using StatsBase
+using DataFramesMeta
+using Random
+using Statistics
+using MLJ
+using DataFrames
+using CSV: CSV
+using Distributions
+using UUIDs
+using ProgressMeter
+
+# Simulation suffix
+_jobid = get(ENV, "SLURM_ARRAY_TASK_ID", 1)
+_jobcount = get(ENV, "SLURM_ARRAY_TASK_COUNT", 1)
+
+# Confusion matrix utilities
+include(joinpath(@__DIR__, "confusionmatrix.jl"))
+
+# Network simulation utilities
+include(joinpath(@__DIR__, "networksimulation.jl"))
+
+# Choice of models
+include(joinpath(@__DIR__, "modelchoice.jl"))
+
+# Function to split the sets
+include(joinpath(@__DIR__, "setsplitter.jl"))
+
+# Prepare the results
+results = [
+    DataFrame(;
+        id=UUID[],
+        connectance=Float64[],
+        model=Symbol[],
+        set=Symbol[],
+        measure=Symbol[],
+        value=Float64[],
+    ) for _thr in 1:Threads.nthreads()
+]
+
+# Network size
+S = (40, 50)
+
+# List of conditions
+links = sort(floor.(Int, repeat(LinRange(45, 0.5*(40*50), 10), 60)))
+
+# Prepare networks for the gridded simulations
+nets = []
+@showprogress for i in 1:length(links)
+    link = links[i]
+    approx_conn = link / prod(S)
+    candidates = [network(S, approx_conn) for i in 1:1_000]
+    idx = last(findmin(n -> abs(approx_conn - mean(last(n))), candidates))
+    push!(nets, candidates[idx])
+end
+
+# Function to range the predictions
+R = (v) -> (v .- minimum(v)) ./ (maximum(v) - minimum(v))
+
+# Dictionary of measures
+measures = Dict(:acc => accuracy, :bac => balanced, :mcc => mcc, :Y => informedness)
+
+# Simulations
+Threads.@threads for i in 1:length(links)
+    bias = 0.5
+
+    runid = uuid4()
+
+    𝐗, 𝐲 = nets[i]
+
+    # Indices to split the dataset
+    Iₚ, Iₒ = split_indices(𝐲, bias)
+
+    # Vector to store the predictions
+
+    for current_set in [(:training, Iₚ), (:testing, Iₒ), (:full, 1:prod(S))]
+        setname, setrow = current_set
[…]

### 2022-03-02 code/07_fitviz.jl

@@ -1,0 +1,29 @@
+using AlgebraOfGraphics, CairoMakie
+using DataFrames
+using CSV: CSV
+using DataFramesMeta
+using StatsBase
+using Statistics
+using StatsFuns
+
+results = DataFrame(CSV.File(joinpath(@__DIR__, "overfit.csv")))
+allowmissing!(results, :value)
+replace!(results.value, NaN => missing)
+dropmissing!(results)
+
+bmm = groupby(results, [:model, :measure, :connectance, :set])
+raw = combine(bmm, :value => mean => :value)
+
+# Show the overfit
+_keepval(f) = f in ["acc"]
+dt = data(@subset(raw, _keepval.(:measure)))
+mp = mapping(
+    :connectance => "Connectance",
+    :value => "Value";
+    color = :set => nonnumeric => "Dataset",
+    layout = :model => sorter("Decision tree", "kNN", "Random Forest", "BRT", "Ensemble") => "Model"
+)
+ly = visual(Lines)
+dt * mp * ly |>
+    plt -> draw(plt, facet=(;linkyaxes = :all)) |>
+    plt -> save(joinpath(@__DIR__, "..", "figures", "overfit.png"), plt, px_per_unit = 3)

### 2022-03-02 draft.tex
Deploying to gh-pages from @ PoisotLab/ms_networkary_classifiers@975948874c7ec87793df7c94132500727cfc4872 🚀
Deploying to gh-pages from @ PoisotLab/ms_networkary_classifiers@97286fcf1eae26f80ccdff7614dc8e698c1a6f82 🚀

@@
 Ecological networks are a backbone for key ecological and evolutionary
-processes; yet enumerating all of the interactions they contain is a
-daunting task, as it scales with \(S^2\), \emph{i.e.} the squared
-species richness TODO. Recent contributions to the field of ecological
-network prediction (Pichler et al., 2020; Strydom et al., 2021;
-\textbf{Becker2021OptPre?}) highlight that although interactions can be
-predicted by adding ecologically relevant information (in the form of,
-\emph{e.g.} traits), we do not have robust guidelines as to how the
-predictive ability of these models should be evaluated, nor about how
-the models should be trained. Here, by relying on simple derivations and
-a series of simulations, we formulate a number of such guidelines,
-specifically for the case of binary classifiers derived from thresholded
-values. Specifically, we conduct an investigation of the models in terms
-of their skill (ability to make the right prediction), bias (trends
-towards systematically over-predicting one class), class imbalance (the
-relative number of cases representing interactions), and show how these
-effects interact. We conclude on the fact that models with the best
-interaction-scale predictive score do not necessarily result in the most
-accurate representation of the network.
+processes; yet enumerating all of the interactions between \(S\) species
+is a daunting task, as it scales with \(S^2\), \emph{i.e.} the squared
+species richness (\textbf{Martinez1992ConCon?}). Recent contributions to
+the field of ecological network prediction (Pichler et al., 2020;
+Strydom et al., 2021; \textbf{Becker2022OptPre?}) highlight that
+although interactions can be predicted by adding ecologically relevant
+information (in the form of, \emph{e.g.} traits), we do not have robust
+guidelines as to how the predictive ability of these models should be
+evaluated, nor about how the models should be trained. Here, by relying
+on simple derivations and a series of simulations, we formulate a number
+of such guidelines, specifically for the case of binary classifiers
+derived from thresholded values. Specifically, we conduct an
+investigation of the models in terms of their skill (ability to make the
+right prediction), bias (trends towards systematically over-predicting
+one class), class imbalance (the relative number of cases representing
+interactions), and show how these effects interact. We conclude on the
+fact that models with the best interaction-scale predictive score do not
+necessarily result in the most accurate representation of the network.
 
@@
 
-In the following section, we will generate random bipartite networks
-(this works without loss of generality on unipartite networks), and
-train four binary classifiers (as well as an ensemble model using the
-sum of ranged outputs from the component models) on 30\% of the
+In the following section, we will generate random bipartite networks,
+and train four binary classifiers (as well as an ensemble model using
+the sum of ranged outputs from the component models) on 50\% of the
 interaction data. In practice, testing usually uses 70\% of the total
@@
 \(i\) and \(j\) when \(v_i-\xi/2 \le h_j \le v_i+\xi/2\), where \(\xi\)
-is a constant regulating the connectance of the network (there is an
-almost 1:1 relationship between \(\xi\) and connectance), and varies
-uniformly in \([0.05, 0.35]\). This model gives fully interval networks
-that are close analogues to the bacteria--phage model of Weitz et al.
-(2005), with both a modular structure and a non-uniform degree
-distribution. This dataset is easy for almost any algorithm to learn:
-when trained with features \([v_i, h_j, \text{abs}(v_i, h_j)] ^T\) to
-predict the interactions between \(i\) and \(j\), all four models
-presented below were able to reach almost perfect predictions all the
-time (data not presented here) -- this is in part because the rule
-(there is maximum value of the distance between traits for which there
-is an interaction) is fixed for all interactions. In order to make the
+is a constant regulating the connectance of the network (visual
+exploration of the parameters show that there is an almost 1:1
+relationship between \(\xi\) and connectance), and varies uniformly in
+\([0.05, 0.35]\). This model gives fully interval networks that are
+close analogues to the bacteria--phage model of Weitz et al. (2005),
+with both a modular structure and a non-uniform degree distribution.
+This dataset is easy for almost any algorithm to learn: when trained
+with features \([v_i, h_j, \text{abs}(v_i, h_j)] ^T\) to predict the
+interactions between \(i\) and \(j\), all four models presented below
+were able to reach almost perfect predictions all the time (data not
+presented here) -- this is in part because the rule (there is maximum
+value of the distance between traits for which there is an interaction)
+is fixed for all interactions, and any method able to learn non-linear
+relationships should infer it without issues. In order to make the
 problem more difficult to solve, we use \([v_i, h_j]\) as a feature
@@
 
-The training sample is composed of 30\% of the \(10^4\) possible entries
-in the network, \emph{i.e.} \(n=3000\). Out of these interactions, we
+The training sample is composed of 50\% of the \(10^4\) possible entries
+in the network, \emph{i.e.} \(n=5000\). Out of these interactions, we
 pick a proportion \(\nu\) (the training set balance) to be positive, so
@@
 to artificially rebalanced dataset for a given network connectance. The
-rest of the dataset (\(n=7000\) pairs of species) is used as a testing
-set, on which all furher measures are calculated. Note that although the
-training set is balanced, the testing set is not, and retains (part of)
-the imbalance of the original data.
+rest of the dataset is used as a testing set, on which all further
+measures are calculated. Note that although the training set is balanced
+arbitrarily, the testing set is assembled so that it has the exact
+connectance of the entire network; this ensures that the model is
+evaluated under the class imbalance where the predictions will be made,
+which represents a more meaningful evaluation. Note also that although
+the simulated networks are bipartite, the algorithms have no
+``knowledge'' of the network structure, and simply look at pairs of
+species; therefore, the approach outlined here would also work for
+unipartite networks.
 
@@
 \centering
-\includegraphics{figures/optim_bias.png}
+\includegraphics{figures/optimal_bias.png}
 \caption{Value of the optimal training set balance for the different
@@
 \centering
-\includegraphics{figures/optim_value.png}
+\includegraphics{figures/optimal_value.png}
 \caption{When trained on their optimally biased training set, most
@@
 do not expect that ensembles will \emph{always} be better than single
-models. In a recent multi-model comparison, (\textbf{Becker2021OptPre?})
+models. In a recent multi-model comparison, (\textbf{Becker2022OptPre?})
 found that the ensemble was \emph{not} the best model. There is no

### 2022-03-02 draft.tex
Deploying to gh-pages from @ PoisotLab/ms_networkary_classifiers@943bf0a37a1706bd64fc5717e74f7377744217d6 🚀

@@
 \includegraphics{figures/valid_ensemble.png}
-\caption{Visualisation of the models predictions for one instance of a
-network prediction problem (shown in the ``Dataset'' panel). This figure
-reveals how inspecting the details of the prediction is important:
-indeed, although the performance measures hint at the fact that ridge
-regression is mediocre, this figure reveals that it is making
-predictions that correspond to a network with an entirely different
-topology (namely, nested as opposed to diagonal).}\label{fig:ecovalid}
+\caption{Visualisation of the raw (un-thresholded) models predictions
+for one instance of a network prediction problem (shown in the
+``Dataset'' panel). Increasing the value of the \(\xi\) parameter would
+make the diagonal structure ``broader,'' leading to more interactions. A
+visual inspection of the results is important, as it highlights how some
+models can ``miss'' parts of the network; by combining them in an
+ensemble, these gaps compensate one another, and lead (in this case) to
+a better prediction.}\label{fig:ecovalid}
 }
@@
 
-\hypertarget{guidelines-for-the-assesment-of-network-predictive-models}{%
-\section{Guidelines for the assesment of network predictive
-models}\label{guidelines-for-the-assesment-of-network-predictive-models}}
+\hypertarget{guidelines-for-the-assessment-of-network-predictive-models}{%
+\section{Guidelines for the assessment of network predictive
+models}\label{guidelines-for-the-assessment-of-network-predictive-models}}
+
+It is noteworthy that the ensemble model was systematically better than
+the component models. We do not expect that ensembles will \emph{always}
+be better than single models. Networks with different structures than
+the one we simulated here may respond in different ways, especially if
+the rules are fuzzier than the simple rule we used here. In a recent
+multi-model comparison involving supervised and unsupervised learning,
+(\textbf{Becker2022OptPre?}) found that the ensemble was \emph{not} the
+best model. There is no general conclusion to draw from this besides
+reinforcing the need to be pragmatic about which models should be
+included in the ensemble, or whether to use an ensemble at all. In a
+sense, the surprising performance of the ensemble model should form the
+basis of the last recommendation: optimal training set balance and its
+interaction with connectance and the specific binary classifier used is,
+in a sense, an hyperparameter that should be assessed. The distribution
+of results in fig.~\ref{fig:optimbias} and fig.~\ref{fig:optimvalue}
+show that there are variations around the trend. TODO fix this sentence
+
+(\textbf{Xie2017ComCom?}) - most networks have SBM structure ---
+(\textbf{Valdovinos2019MutNet?}) matching + forbidden
+(\textbf{Olesen2011MisFor?}), coming up with matching rules that
+allow/forbid links can make predictions easier
+(\textbf{Strona2017ForPer?}); but valid for most type of interactions as
+long as we can match yes/no to numeric predictors
 
@@
 
-Finally, it is noteworthy that the ensemble model was systematically
-better than the component models; even when the models were individually
-far form perfect, the ensemble was able to leverage the different biases
-expressed by the models to make an overall more accurate prediction. We
-do not expect that ensembles will \emph{always} be better than single
-models. In a recent multi-model comparison, (\textbf{Becker2022OptPre?})
-found that the ensemble was \emph{not} the best model. There is no
-general conclusion to draw from this besides reinforcing the need to be
-pragmatic about which models should be included in the ensemble, or
-whether to use an ensemble at all. In a sense, the surprising peformance
-of the ensemble model should form the basis of the last recommendation:
-optimal training set bias and its interaction with connectance and
-binary classifier is, in a sense, an hyperparameter that should be
-assessed. The distribution of results in fig.~\ref{fig:optimbias} and
-fig.~\ref{fig:optimvalue} show that there are variations around the
-trend; furthermore, networks with different structures than the one we
-simulated here may respond in different ways.
-
 \textbf{Acknowledgements:} We acknowledge that this study was conducted

### 2022-03-02 draft.tex
Deploying to gh-pages from @ PoisotLab/ms_networkary_classifiers@5ab839c230e8422cceddc0ce741a933b95bbf9fd 🚀

@@
 
-The dataset used for numerical experiments is composed of 64000 such
-\((\xi, \nu)\) pairs, on which four machines are trained: a decision
-tree regressor, a boosted regression tree, a ridge regressor, and a
-random forest regressor. All models were taken from the \texttt{MLJ.jl}
-package (Blaom et al., 2020; Blaom \& Vollmer, 2020) in Julia 1.7
-(Bezanson et al., 2017). All machines use the default parameterization;
-this is an obvious deviation from best practices, as the hyperparameters
-of any machine require training before its application on a real
-dataset. As we use 64000 such datasets, this would require 256000 unique
-instances of tweaking the hyperparameters, which is not realistic.
-Therefore, we assume that the default parameterizations are comparable
-across networks. All machines return a quantitative prediction, usually
-(but not necessarily) in \([0,1]\), which is proportional (but not
-necessarily linearly) to the probability of an interaction between \(i\)
-and \(j\). Nevertheless, the ROC-AUC and PR-AUC (and therefore the
-thresholds) can be measured by integrating over the domain of the values
-return by each machine.
+The dataset used for numerical experiments is composed of a grid of 35
+values of connectance (from 0.011 to 0.5) and 35 values of \(\nu\) (from
+0.02 to 0.98); for each pair of values, 500 networks are generated and
+predicted. For each network, we train four machines: a trait-based k-NN
+(\emph{e.g.} \textbf{Desjardins-Proulx2017EcoInt?}), a regression tree,
+a regression random forest, and a boosted regression tree. Following
+results from Pichler et al. (2020), linear models have not been
+considered (in any way, the relationship in the simulated networks is
+non-linear). The point of these numerical experiments is \emph{not} to
+recommend the best model (this is likely problem-specific), but to
+highlight a series of recommendations that would work for supervised
+learning tasks. All models were taken from the \texttt{MLJ.jl} package
+(Blaom et al., 2020; Blaom \& Vollmer, 2020) in Julia 1.7 (Bezanson et
+al., 2017). All machines use the default parameterization; this is an
+obvious deviation from best practices, as the hyperparameters of any
+machine require training before its application on a real dataset. As we
+use 612500 such datasets, this would require over 2 millions unique
+instances of tweaking the hyperparameters, which is prohibitive from a
+computing time point of view. An important thing to keep in mind is that
+the problem we simulate has been designed to be simple to solve: we
+expect all machines with sensible default parameters to fare well ---
+the results presented in the later sections show that this assumption is
+warranted, and we further checked that the models do not overfit by
+ensuring that there is never more than 5\% of difference between the
+accuracy on the training and testing sets. All machines return a
+quantitative prediction, usually (but not necessarily) in \([0,1]\),
+which is proportional (but not necessarily linearly) to the probability
+of an interaction between \(i\) and \(j\). The ROC-AUC and PR-AUC (and
+therefore the thresholds) can be measured by integrating over the domain
+of the values return by each machine, but in order to make the
+average-based ensemble model more meaningful, all predictions are
+expressed in \([0,1]\).
 
@@
 After the simulations were completed, we removed all runs (\emph{i.e.}
-pairs of \(\xi\) and \(\nu\)) for which at least one of the following
-conditions was met: the accuracy was 0, the true positive or true
-negative rates were 0, the connectance was larger than 0.25. This
+triples of model, \(\xi\), and \(\nu\)) for which at least one of the
+following conditions was met: the accuracy was 0, the true positive or
+true negative rates were 0, the connectance was larger than 0.25. This
 removes both the obviously failed model runs, and the networks that are
@@
 the connectance of the network; from a practical point of view,
-assembling a training set requires to withold positive information,
+assembling a training set requires one to withhold positive information,
 which in ecological networks are very scarce (and typically more
@@
 
-(\textbf{Xie2017ComCom?}) - most networks have SBM structure ---
-(\textbf{Valdovinos2019MutNet?}) matching + forbidden
-(\textbf{Olesen2011MisFor?}), coming up with matching rules that
-allow/forbid links can make predictions easier
-(\textbf{Strona2017ForPer?}); but valid for most type of interactions as
-long as we can match yes/no to numeric predictors
-
 We establish that due to the low prevalence of interactions, even poor
@@
 
+One key element for real-life data that can make the prediction exercise
+more tractable is that some interactions can safely be assumed to be
+impossible; indeed, a lot of networks admit a stochastic block model as
+a good approximation (\emph{e.g.} \textbf{Xie2017ComCom?}). In
+ecological networks, this can be due to spatial constrains
+(\textbf{Valdovinos2019MutNet?}), or to the long-standing knowledge that
+some links are ``forbidden'' due to traits (\textbf{Olesen2011MisFor?})
+or abundances (\textbf{Canard2014EmpEva?}). The matching rules
+(\textbf{Strona2017ForPer?}) can be incorporated in the model either by
+adding compatibility traits, or by \emph{only} training the model on
+pairs of species that are not likely to be forbidden links. Besides
+forbidden links, a real-life case that may arise is multi-interaction or
+multi-layer networks (\textbf{Pilosof2017MulNat?}). These can be studied
+using the same general approach outlined here, either by assuming that
+pairs of species can interact in more than one way (wherein one would
+train a model for each type of interaction, based on the relevant
+predictors), or by assuming that pairs of species can only have one type
+of interaction (wherein this becomes a multi-label classification
+problem).
+
 \textbf{Acknowledgements:} We acknowledge that this study was conducted

### 2022-05-20 draft.tex
Deploying to gh-pages from @ PoisotLab/ms_networkary_classifiers@fb8b223211b9147a644676f6ddaec4102610bd72 🚀

@@
 is a daunting task, as it scales with \(S^2\), \emph{i.e.} the squared
-species richness (\textbf{Martinez1992ConCon?}). Recent contributions to
-the field of ecological network prediction (Pichler et al., 2020;
-Strydom et al., 2021; \textbf{Becker2022OptPre?}) highlight that
-although interactions can be predicted by adding ecologically relevant
-information (in the form of, \emph{e.g.} traits), we do not have robust
-guidelines as to how the predictive ability of these models should be
-evaluated, nor about how the models should be trained. Here, by relying
-on simple derivations and a series of simulations, we formulate a number
-of such guidelines, specifically for the case of binary classifiers
-derived from thresholded values. Specifically, we conduct an
-investigation of the models in terms of their skill (ability to make the
-right prediction), bias (trends towards systematically over-predicting
-one class), class imbalance (the relative number of cases representing
-interactions), and show how these effects interact. We conclude on the
-fact that models with the best interaction-scale predictive score do not
-necessarily result in the most accurate representation of the network.
+species richness (Martinez, 1992). Recent contributions to the field of
+ecological network prediction (Becker et al., 2022; Pichler et al.,
+2020; Strydom et al., 2021) highlight that although interactions can be
+predicted by adding ecologically relevant information (in the form of,
+\emph{e.g.} traits), we do not have robust guidelines as to how the
+predictive ability of these models should be evaluated, nor about how
+the models should be trained. Here, by relying on simple derivations and
+a series of simulations, we formulate a number of such guidelines,
+specifically for the case of binary classifiers derived from thresholded
+values. Specifically, we conduct an investigation of the models in terms
+of their skill (ability to make the right prediction), bias (trends
+towards systematically over-predicting one class), class imbalance (the
+relative number of cases representing interactions), and show how these
+effects interact. We conclude on the fact that models with the best
+interaction-scale predictive score do not necessarily result in the most
+accurate representation of the network.
 
@@
 
-Species interaction networks are often under-sampled (Jordano, 2016b,
-2016a), and this under-sampling is structured taxonomically (Beauchesne
+Species interaction networks are often under-sampled (Jordano, 2016a,
+2016b), and this under-sampling is structured taxonomically (Beauchesne
 et al., 2016), structurally (de Aguiar et al., 2019) and spatially
@@
 \caption{Consequences of changing the classifier skills (\(s\)) and bias
-(\(s\)) for a connectance \(\rho=0.15\), on accuracy, \(F_1\), postive
-predictive value, and \(\kappa\). Accuracy increases with skill, but
-also increases when the bias tends towards estimating \emph{fewer}
-interactions. The \(F_1\) score increases with skill but also increases
-when the bias tends towards estimating \emph{more} interactions; PPV
-behaves in the same way. Interestingly, \(\kappa\) responds as expected
-to skill (being negative whenever \(s < 0.5\)), and peaks for values of
-\(b \approx 0.5\); nevertheless, the value of bias for which \(\kappa\)
-is maximized in \emph{not} \(b=0.5\), but instead increases with
-classifier skill. In other words, at equal skill, maximizing \(\kappa\)
-would lead to select a \emph{more} biased classifier.}\label{fig:bias}
+(\(s\)) for a connectance \(\rho=0.15\), on \(F_1\), informedness, MCC,
+and \(\kappa\). Accuracy increases with skill, but also increases when
+the bias tends towards estimating \emph{fewer} interactions. The \(F_1\)
+score increases with skill but also increases when the bias tends
+towards estimating \emph{more} interactions. Interestingly, \(\kappa\)
+responds as expected to skill (being negative whenever \(s < 0.5\)), and
+peaks for values of \(b \approx 0.5\); nevertheless, the value of bias
+for which \(\kappa\) is maximized in \emph{not} \(b=0.5\), but instead
+increases with classifier skill. In other words, at equal skill,
+maximizing \(\kappa\) would lead to select a \emph{more} biased
+classifier.}\label{fig:bias}
 }
@@
 bias. Both MCC and \(\kappa\) have the same behavior, whereby they
-increase with skill. \(\kappa\) peaks at increasing values of biass for
+increase with skill. \(\kappa\) peaks at increasing values of bias for
 increasing skill, \emph{i.e.} is likely to lead to the selection of a
@@
 predicted. For each network, we train four machines: a trait-based k-NN
-(\emph{e.g.} \textbf{Desjardins-Proulx2017EcoInt?}), a regression tree,
-a regression random forest, and a boosted regression tree. Following
+(\emph{e.g.} Desjardins-Proulx et al., 2017), a regression tree, a
+regression random forest, and a boosted regression tree. Following
 results from Pichler et al. (2020), linear models have not been
@@
 models were able to maximize their performance; this is not true for
-decision tree, which had a very low PR-AUC, and to some extent for ridge
-regression who had a slow increase with network connectance. The
-ensemble had a consistently high performance despite incorporating poor
+decision tree, k-NN, and to a lower extent RF. The ensemble had a
+consistently high performance despite incorporating poor
 models.}\label{fig:optimvalue}
@@
 had a significant impact on the performance of some machines
-fig.~\ref{fig:optimvalue}. Notably, Decision Tree, Random Forest, and
-Ridge Regression had low values of PR-AUC. In all cases, the Boosted
-Regression Tree was reaching very good predictions (especially for
-connectances larger than 0.1), and the ensemble was almost always
-scoring perfectly. This suggests that all the models are biased in
-different ways, and that the averaging in the ensemble is able to
-correct these biases. We do not expect this last result to have any
+fig.~\ref{fig:optimvalue}. Notably, Decision Tree, and k-NN, as well as
+Random forest to a lower extent, had low values of PR-AUC. In all cases,
+the Boosted Regression Tree was reaching very good predictions
+(especially for connectances larger than 0.1), and the ensemble was
+almost always scoring perfectly. This suggests that all the models are
+biased in different ways, and that the averaging in the ensemble is able
+to correct these biases. We do not expect this last result to have any
 generality, and provide a discussion of a recent example in which the
@@
 on the complete dataset is presented in fig.~\ref{fig:ecovalid}.
-Visualizing the results this way highlights the importance of
-exploratory data analysis: whereas all models return a network with
-interactions laying mostly on the diagonal (as expected), the Ridge
-Regression is quite obviously biased. Despite this, we can see that the
-ensemble is close to the initial dataset.
 
@@
 \textbf{0.20} & \textbf{0.72} & \textbf{0.32}\tabularnewline
-Ridge Regression & 0.80 & 0.91 & 0.95 & 0.58 & 0.24 & 1.0 &
-0.18\tabularnewline
+k-NN & 0.80 & 0.91 & 0.95 & 0.58 & 0.24 & 1.0 & 0.18\tabularnewline
 Ensemble & 0.88 & 0.94 & \textbf{1.00} & \textbf{0.96} & \textbf{0.20} &
@@
 multi-model comparison involving supervised and unsupervised learning,
-(\textbf{Becker2022OptPre?}) found that the ensemble was \emph{not} the
-best model. There is no general conclusion to draw from this besides
+Becker et al. (2022) found that the ensemble was \emph{not} the best
+model. There is no general conclusion to draw from this besides
 reinforcing the need to be pragmatic about which models should be
@@
 generally accepted default for imbalanced classification; Schisterman et
-al., 2005).
+al., 2005). This last result suggests that the amount of bias \emph{is}
+an hyper-parameter that must be fine-tuned, as using the wrong bias can
+lead to under-performing models.
 
@@
 impossible; indeed, a lot of networks admit a stochastic block model as
-a good approximation (\emph{e.g.} \textbf{Xie2017ComCom?}). In
-ecological networks, this can be due to spatial constrains
-(\textbf{Valdovinos2019MutNet?}), or to the long-standing knowledge that
-some links are ``forbidden'' due to traits (\textbf{Olesen2011MisFor?})
-or abundances (\textbf{Canard2014EmpEva?}). The matching rules
-(\textbf{Strona2017ForPer?}) can be incorporated in the model either by
-adding compatibility traits, or by \emph{only} training the model on
-pairs of species that are not likely to be forbidden links. Besides
-forbidden links, a real-life case that may arise is multi-interaction or
-multi-layer networks (\textbf{Pilosof2017MulNat?}). These can be studied
-using the same general approach outlined here, either by assuming that
-pairs of species can interact in more than one way (wherein one would
-train a model for each type of interaction, based on the relevant
-predictors), or by assuming that pairs of species can only have one type
-of interaction (wherein this becomes a multi-label classification
-problem).
+a good approximation (\emph{e.g.} Xie et al., 2017). In ecological
+networks, this can be due to spatial constrains (Valdovinos, 2019), or
+to the long-standing knowledge that some links are ``forbidden'' due to
+traits (Olesen et al., 2011) or abundances (Canard et al., 2014). The
+matching rules (Olito \& Fox, 2015; Strona \& Veech, 2017) can be
+incorporated in the model either by adding compatibility traits, or by
+\emph{only} training the model on pairs of species that are not likely
+to be forbidden links. Besides forbidden links, a real-life case that
+may arise is multi-interaction or multi-layer networks (Pilosof et al.,
+2017). These can be studied using the same general approach outlined
+here, either by assuming that pairs of species can interact in more than
+one way (wherein one would train a model for each type of interaction,
+based on the relevant predictors), or by assuming that pairs of species
+can only have one type of interaction (wherein this becomes a
+multi-label classification problem).
 
@@
 Beauchesne, D., Desjardins-Proulx, Archambault, P., \& Gravel, D.
-(2016). Thinking Outside the Boxpredicting Biotic Interactions in
+(2016). Thinking Outside the Box--predicting Biotic Interactions in
 Data-poor Environments. \emph{Vie Et Milieu-Life and enVironment},
 \emph{66}(3-4), 333--342.
+
+\leavevmode\hypertarget{ref-Becker2022OptPre}{}%
+Becker, D. J., Albery, G. F., Sjodin, A. R., Poisot, T., Bergner, L. M.,
+Chen, B., Cohen, L. E., Dallas, T. A., Eskew, E. A., Fagre, A. C.,
+Farrell, M. J., Guth, S., Han, B. A., Simmons, N. B., Stock, M.,
+Teeling, E. C., \& Carlson, C. J. (2022). Optimising predictive models
+to prioritise viral discovery in zoonotic reservoirs. \emph{The Lancet
+Microbe}. \url{https://doi.org/10.1016/S2666-5247(21)00245-7}
 
@@
 \leavevmode\hypertarget{ref-Blaom2020FleMod}{}%
-Blaom, A. D., \& Vollmer, S. J. (2020). Flexible model composition in
-machine learning and its implementation in MLJ. \emph{arXiv:2012.15505
-{[}cs{]}}. \url{http://arxiv.org/abs/2012.15505}
+Blaom, A. D., \& Vollmer, S. J. (2020). \emph{Flexible model composition
+in machine learning and its implementation in MLJ}.
+\url{http://arxiv.org/abs/2012.15505}
 
@@
 \leavevmode\hypertarget{ref-Branco2015SurPre}{}%
-Branco, P., Torgo, L., \& Ribeiro, R. (2015). A Survey of Predictive
-Modelling under Imbalanced Distributions. \emph{arXiv:1505.01658
-{[}cs{]}}. \url{http://arxiv.org/abs/1505.01658}
+Branco, P., Torgo, L., \& Ribeiro, R. (2015). \emph{A Survey of
+Predictive Modelling under Imbalanced Distributions}.
+\url{http://arxiv.org/abs/1505.01658}
+
+\leavevmode\hypertarget{ref-Canard2014EmpEva}{}%
+Canard, E. F., Mouquet, N., Mouillot, D., Stanko, M., Miklisova, D., \&
+Gravel, D. (2014). Empirical evaluation of neutral interactions in
+host-parasite networks. \emph{The American Naturalist}, \emph{183}(4),
+468--479. \url{https://doi.org/10.1086/675363}
 
@@
 
+\leavevmode\hypertarget{ref-Desjardins-Proulx2017EcoInt}{}%
+Desjardins-Proulx, P., Laigle, I., Poisot, T., \& Gravel, D. (2017).
+Ecological interactions and the Netflix problem. \emph{PeerJ},
+\emph{5}(e3644). \url{https://doi.org/10.7717/peerj.3644}
+
 \leavevmode\hypertarget{ref-Ferri2009ExpCom}{}%
@@
 Jeni, L. A., Cohn, J. F., \& De La Torre, F. (2013). Facing Imbalanced
-DataRecommendations for the Use of Performance Metrics. \emph{2013
+Data--Recommendations for the Use of Performance Metrics. \emph{2013
 Humaine Association Conference on Affective Computing and Intelligent
@@
 \emph{1}(0). \url{https://doi.org/10.1016/j.patter.2020.100079}
+
+\leavevmode\hypertarget{ref-Martinez1992ConCon}{}%
+Martinez, N. D. (1992). Constant Connectance in Community Food Webs.
+\emph{The American Naturalist}, \emph{139}(6), 1208--1218.
+\url{http://www.jstor.org/stable/2462337}
 
@@
 
+\leavevmode\hypertarget{ref-Olesen2011MisFor}{}%
+Olesen, J. M., Bascompte, J., Dupont, Y. L., Elberling, H., Rasmussen,
+C., \& Jordano, P. (2011). Missing and forbidden links in mutualistic
+networks. \emph{Proc. R. Soc. B}, \emph{278}(1706), 725--732.
+\url{https://doi.org/10.1098/rspb.2010.1371}
+
+\leavevmode\hypertarget{ref-Olito2015SpeTra}{}%
+Olito, C., \& Fox, J. W. (2015). Species traits and abundances predict
+metrics of plant--pollinator network structure, but not pairwise
+interactions. \emph{Oikos}, \emph{124}, 428--436.
+
 \leavevmode\hypertarget{ref-Pichler2020MacLea}{}%
-Pichler, M., Boreux, V., Klein, A.-M., Schleuning, M., \& Hartig, F.
-(2020). Machine learning algorithms to infer trait-matching and predict
+Pichler, M., Boreux, V., Klein, A., Schleuning, M., \& Hartig, F.
+(2020). Machine learning algorithms to infer trait‐matching and predict
 species interactions in ecological networks. \emph{Methods in Ecology
@@
 \url{https://doi.org/10.1111/2041-210X.13329}
+
+\leavevmode\hypertarget{ref-Pilosof2017MulNat}{}%
+Pilosof, S., Porter, M. A., Pascual, M., \& Kéfi, S. (2017). The
+multilayer nature of ecological networks. \emph{Nature Ecology \&
+Evolution}, \emph{1}, 0101.
+\url{https://doi.org/10.1038/s41559-017-0101}
 
@@
 J., Albery, G. F., Gibb, R. J., Seifert, S. N., \& Carlson, C. J.
-(2021). Imputing the mammalian virome with linear filtering and singular
-value decomposition. \emph{arXiv:2105.14973 {[}q-Bio{]}}.
-\url{http://arxiv.org/abs/2105.14973}
+(2021). \emph{Imputing the mammalian virome with linear filtering and
+singular value decomposition}. \url{http://arxiv.org/abs/2105.14973}
 
@@
 \leavevmode\hypertarget{ref-Somodi2017PreDep}{}%
-Somodi, I., Lepesi, N., \& Botta-Dukát, Z. (2017). Prevalence dependence
+Somodi, I., Lepesi, N., \& Botta‐Dukát, Z. (2017). Prevalence dependence
 in model goodness measures with special emphasis on true skill
@@
 \url{https://doi.org/10.1111/2041-210X.13525}
+
+\leavevmode\hypertarget{ref-Strona2017ForPer}{}%
+Strona, G., \& Veech, J. A. (2017). Forbidden versus permitted
+interactions: Disentangling processes from patterns in ecological
+network analysis. \emph{Ecology and Evolution}, \emph{7}(14),
+5476--5481. \url{https://doi.org/10.1002/ece3.3102}
 
@@
 
+\leavevmode\hypertarget{ref-Valdovinos2019MutNet}{}%
+Valdovinos, F. S. (2019). Mutualistic networks: Moving closer to a
+predictive theory. \emph{Ecology Letters}, \emph{0}(0).
+\url{https://doi.org/10.1111/ele.13279}
+
 \leavevmode\hypertarget{ref-Weitz2005CoeArm}{}%
@@
 
+\leavevmode\hypertarget{ref-Xie2017ComCom}{}%
+Xie, J.-R., Zhang, P., Zhang, H.-F., \& Wang, B.-H. (2017). Completeness
+of Community Structure in Networks. \emph{Scientific Reports},
+\emph{7}(1), 5269. \url{https://doi.org/10.1038/s41598-017-05585-6}
+
 \leavevmode\hypertarget{ref-Youden1950IndRat}{}%

### 2022-06-08 draft.tex
Deploying to gh-pages from @ PoisotLab/ms_networkary_classifiers@66d41abcb568ec783472153faab2d2933e7f2bcd 🚀

@@
 \hypertarget{tbl:comparison}{}
-\begin{longtable}[]{@{}rccccccc@{}}
+\begin{longtable}[]{@{}rccccccccc@{}}
 \caption{\label{tbl:comparison}Values of four performance metrics, and
@@
 \toprule
-Model & MCC & Inf. & ROC-AUC & PR-AUC & Conn. & \(\eta\) &
-\(Q\)\tabularnewline
+Model & MCC & Inf. & ROC-AUC & PR-AUC & Conn. & \(\eta\) & \(Q\) & \(A\)
+& Jaccard\tabularnewline
 \midrule
@@
 \toprule
-Model & MCC & Inf. & ROC-AUC & PR-AUC & Conn. & \(\eta\) &
-\(Q\)\tabularnewline
+Model & MCC & Inf. & ROC-AUC & PR-AUC & Conn. & \(\eta\) & \(Q\) & \(A\)
+& Jaccard\tabularnewline
 \midrule
 \endhead
-Decision tree & 0.85 & 0.92 & 0.97 & 0.12 & 0.21 & 0.76 &
-0.31\tabularnewline
-BRT & \textbf{0.90} & 0.90 & 0.98 & 0.86 & 0.23 & 0.82 &
-0.27\tabularnewline
-Random Forest & \textbf{0.90} & \textbf{0.96} & \textbf{1.00} & 0.27 &
-\textbf{0.20} & \textbf{0.72} & \textbf{0.32}\tabularnewline
-k-NN & 0.80 & 0.91 & 0.95 & 0.58 & 0.24 & 1.0 & 0.18\tabularnewline
-Ensemble & 0.88 & 0.94 & \textbf{1.00} & \textbf{0.96} & \textbf{0.20} &
-0.75 & 0.31\tabularnewline
-Data & & & & & 0.18 & 0.66 & 0.34\tabularnewline
+Decision tree & 0.59 & 0.94 & 0.97 & 0.04 & 0.17 & 0.64 & 0.37 &
+\textbf{0.42} & 0.1\tabularnewline
+BRT & 0.46 & 0.91 & 0.97 & 0.36 & 0.2 & 0.78 & 0.29 & 0.41 &
+0.19\tabularnewline
+Random Forest & 0.72 & \textbf{0.98} & 0.99 & 0.1 & \textbf{0.16} &
+\textbf{0.61} & 0.38 & \textbf{0.42} & \textbf{0.06}\tabularnewline
+k-NN & 0.71 & \textbf{0.98} & 0.99 & 0.02 & \textbf{0.16} &
+\textbf{0.61} & \textbf{0.39} & \textbf{0.42} &
+\textbf{0.06}\tabularnewline
+\emph{Ensemble} & \textbf{0.74} & \textbf{0.98} & \textbf{1.0} &
+\textbf{0.79} & \textbf{0.16} & \textbf{0.61} & 0.38 & \textbf{0.42} &
+\textbf{0.06}\tabularnewline
+\emph{Data} & & & & & 0.16 & 0.56 & 0.41 & 0.42 & 0.0\tabularnewline
 \bottomrule
@@
 networks, we highlight an interesting paradox: the models with the best
-performance measures are not the models with the closest reconstructed
-network structure. We discuss these results in the context of
-establishing guidelines for the prediction of ecological interactions.
-
-TODO informedness and accuracy should be easy to beat, make sure the
-model is better than them!
+performance measures are not necessarilly the models with the closest
+reconstructed network structure. We discuss these results in the context
+of establishing guidelines for the prediction of ecological
+interactions.
 
@@
 fig.~\ref{fig:connectance}, MCC should be prefered as a measure of model
-evaluation.
+evaluation and comparison.
 
@@
 
-Thirdly, regardless of network connectance, maximizing informedness
-required a training set balance of about 0.5, and maximizing the MCC
-required a training set bias of 0.7 and more. This has an important
-consequence in ecological networks, for which the pool of positive cases
-(interactions) to draw from is typically small: the most parsimonious
-measure (\emph{i.e.} the one requiring to discard the least amount of
+Third, for the same reasons that should give more weight to PR-AUC than
+ROC-AUC, accuracy alone should not be used as a measure of model
+performance, but rather as an expectation of how well the model should
+behave given the balance in the set on which predictions are made; this
+is because, as derived earlier, the expected accuracy for a no-skill
+no-bias classifier is \(\rho^2 + (1-\rho)^2\) (where \(\rho\) is the
+class balance), which will very easily by large. This pitfall is notably
+illustrated in a recent model (Caron et al., 2022) wherein the authors
+claim that using a training set of \(n = 10^4\) with only 100 positive
+interactions (representing 0.1\% of the total interactions) reached a
+good accuracy. Reporting a good accuracy isn't informative is this
+accuracy isn't (i) compared to the baseline expected value under the
+class balance, and (ii) interpreted in the context of a measure that
+isn't sensitive to true negatives (like MCC).
+
+Finally, network connectance (\emph{i.e.} the empirical class imbalance)
+should inform the composition of the training and testing set. In the
+approach outlined here, we treat the class imbalance as an
+hyper-parameter, but \emph{test} the model on a set that has the same
+class imbalance as the actual dataset. This is an important change to
+the overall methodology so far, as it ensure that the prediction
+environment matches the testing environment, and so the values on the
+testing set can be directly compared to the values for the actual
+prediction. A striking result fig.~\ref{fig:optimbias} is that
+Informedness was almost always maximal at 50/50 balance (regardless of
+connectance), whereas MCC required \emph{more} positives to be maximized
+when connectance \emph{increases}. This has an important consequence in
+ecological networks, for which the pool of positive cases (interactions)
+to draw from is typically small: the most parsimonious measure
+(\emph{i.e.} the one requiring to discard the least amount of
 information to train the model) will give the best validation potential,
@@
 generally accepted default for imbalanced classification; Schisterman et
-al., 2005). This last result suggests that the amount of bias \emph{is}
-an hyper-parameter that must be fine-tuned, as using the wrong bias can
-lead to under-performing models.
+al., 2005). This last result further strengthens the assumption that the
+amount of bias \emph{is} an hyper-parameter that must be fine-tuned, as
+using the wrong bias can lead to models with excess biass or lower
+skill.
 
@@
 \emph{only} training the model on pairs of species that are not likely
-to be forbidden links. Besides forbidden links, a real-life case that
-may arise is multi-interaction or multi-layer networks (Pilosof et al.,
-2017). These can be studied using the same general approach outlined
-here, either by assuming that pairs of species can interact in more than
-one way (wherein one would train a model for each type of interaction,
-based on the relevant predictors), or by assuming that pairs of species
-can only have one type of interaction (wherein this becomes a
-multi-label classification problem).
+to be forbidden links; having this information would allow to assemble
+training/testing sets that have true negatives, and in this situation,
+it may be possible to use the more usual 70/30 split. Besides forbidden
+links, a real-life case that may arise is multi-interaction or
+multi-layer networks (Pilosof et al., 2017). These can be studied using
+the same general approach outlined here, either by assuming that pairs
+of species can interact in more than one way (wherein one would train a
+model for each type of interaction, based on the relevant predictors),
+or by assuming that pairs of species can only have one type of
+interaction (wherein this becomes a multi-label classification problem).
 
@@
 
+\leavevmode\hypertarget{ref-Caron2022AddElt}{}%
+Caron, D., Maiorano, L., Thuiller, W., \& Pollock, L. J. (2022).
+Addressing the Eltonian shortfall with trait-based interaction models.
+\emph{Ecology Letters}, \emph{25}(4), 889--899.
+\url{https://doi.org/10.1111/ele.13966}
+
 \leavevmode\hypertarget{ref-Chicco2020AdvMat}{}%

### 2022-06-09 draft.tex
Deploying to gh-pages from @ PoisotLab/ms_networkary_classifiers@e4608d1df93d7206dc91c3c754a12d9a884e5ac0 🚀

@@
 \thispagestyle{empty}
-{\Large\bfseries\sffamily Guidelines for the validation of machine
-learning predictions of species interactions}
+{\Large\bfseries\sffamily Guidelines for the prediction of species
+interactions through binary classification}
 \vskip 5em
@@
 networks is a challenging task, and that in virtually all cases, the
-composition of the training set needs to be experimented on before
-performing the actual training. We discuss these consequences in the
-context of the low volume of data.%
+composition of the training set needs to be fine-tuned before performing
+the actual training. We discuss these consequences in the context of the
+low volume of data.%
 \end{enumerate}
@@
 \caption{When trained on their optimally biased training set, most
-models were able to maximize their performance; this is not true for
-decision tree, k-NN, and to a lower extent RF. The ensemble had a
-consistently high performance despite incorporating poor
-models.}\label{fig:optimvalue}
+models were able to maximize their performance; this is not true when
+measuring PR-AUC for decision tree, k-NN, and to a lower extent RF. The
+ensemble had a consistently high performance despite incorporating
+low-performing models.}\label{fig:optimvalue}
 }
@@
 informedness), and their predictions transformed back into networks for
-analysis; specifically, we measured the connectance, nestedness (REF),
-modularity (REF), asymmetry, and network dissimilarity (REF). This
-process was repeated 250 times, and the results are presented in
-tbl.~\ref{tbl:comparison}. The random forest model is an interesting
+analysis; specifically, we measured the connectance, nestedness
+(\(\eta\); \textbf{Bastolla2009ArcMut?}), modularity (\(Q\);
+\textbf{Barber2007ModCom?}), asymmetry (\(A\);
+\textbf{Delmas2018AnaEco?}), and Jaccard network dissimilarity (Canard
+et al., 2014). This process was repeated 250 times, and the results are
+presented in tbl.~\ref{tbl:comparison}. The k-NN model is an interesting
 instance here: it produces the network that looks the most like the
-original dataset, despite having a very low PR-AUC, suggesting it hits
-high recall at the cost of low precision. Although the ensemble was able
-to reach a very high PR-AUC (and a very high ROC-AUC), this did not
-necessarily translate into more accurate reconstructions of the
-structure of the network. This result bears elaborating. Measures of
-model performance capture how much of the interactions and
-non-interactions are correctly identified. As long as these predictions
-are not perfect, some interactions will be predicted at the ``wrong''
-position in the network; these measures cannot describe the structural
-effect of these mistakes. On the other hand, measures of network
-structure can have the same value with interactions that fall at
-drastically different positions; this is in part because a lot of these
-measures covary with connectance, and in part because as long as these
-values are not 0 or their respective maximum, there is a large number of
-network configurations that can have the same value. That ROC-AUC is
-consistently larger than PR-AUC may be a case of this measure masking
-models that are not, individually, strong predictors (Jeni et al.,
-2013).
+original dataset, despite having the lowest PR-AUC, suggesting it hits
+high recall at the cost of low precision. The ensemble was able to reach
+a very high PR-AUC (and a very high ROC-AUC), which translated into more
+accurate reconstructions of the structure of the network (with the
+exception of modulairty, which is underestimated by \(0.03\)). This
+result bears elaborating. Measures of model performance capture how much
+of the interactions and non-interactions are correctly identified. As
+long as these predictions are not perfect, some interactions will be
+predicted at the ``wrong'' position in the network; these measures
+cannot describe the structural effect of these mistakes. On the other
+hand, measures of network structure can have the same value with
+interactions that fall at drastically different positions; this is in
+part because a lot of these measures covary with connectance, and in
+part because as long as these values are not 0 or their respective
+maximum, there is a large number of network configurations that can have
+the same value. That ROC-AUC is consistently larger than PR-AUC may be a
+case of this measure masking models that are not, individually, strong
+predictors (Jeni et al., 2013). In this specif example, the combination
+of individually ``adequate'' models resulted in an extremely strong
+ensemble, suggesting that the correct prediction of interactions (as
+measured by MCC, Inf., ROC-AUC, and PR-AUC) \emph{and} network
+properties is indeed a feasible task under appropriately
+hyper-parameterized models.
 
@@
 
+We establish that due to the low prevalence of interactions, even poor
+classifiers applied to food web data will reach a high accuracy; this is
+because the measure is dominated by the accidentally correct predictions
+of negatives. On simulated confusion matrices with ranges of imbalance
+that are credible for ecological networks, MCC had the most desirable
+behavior, and informedness is a linear measure of classifier skill. By
+performing simulations with four models and an ensemble, we show that
+informedness and ROC-AUC are consistently high on network data, whereas
+MCC and PR-AUC are more accurate measures of the effective performance
+of the classifier. Finally, by measuring the structure of predicted
+networks, we highlight an interesting paradox: the models with the best
+performance measures are not necessarilly the models with the closest
+reconstructed network structure. We discuss these results in the context
+of establishing guidelines for the prediction of ecological
+interactions.
+
 It is noteworthy that the ensemble model was systematically better than
@@
 Becker et al. (2022) found that the ensemble was \emph{not} the best
-model. There is no general conclusion to draw from this besides
+model, and was specifically under-performing compared to models using
+biological traits. This may be because the dataset of Becker et al.
+(2022) was known to be under-sampled, and so the network itself
+contained less information than the network and species traits. There is
+no general conclusion to draw from either these results or ours, besides
 reinforcing the need to be pragmatic about which models should be
-included in the ensemble, or whether to use an ensemble at all. In a
+included in the ensemble, and whether to use an ensemble at all. In a
 sense, the surprising performance of the ensemble model should form the
-basis of the last recommendation: optimal training set balance and its
-interaction with connectance and the specific binary classifier used is,
-in a sense, an hyperparameter that should be assessed. The distribution
-of results in fig.~\ref{fig:optimbias} and fig.~\ref{fig:optimvalue}
-show that there are variations around the trend. TODO fix this sentence
-
-We establish that due to the low prevalence of interactions, even poor
-classifiers applied to food web data will reach a high accuracy; this is
-because the measure is dominated by the accidentally correct predictions
-of negatives. On simulated confusion matrices with ranges of imbalance
-that are credible for ecological networks, MCC had the most desirable
-behavior, and informedness is a linear measure of classifier skill. By
-performing simulations with four models and an ensemble, we show that
-informedness and ROC-AUC are consistently high on network data, and that
-MCC and PR-AUC are more accurate measures of the effective performance
-of the classifier. Finally, by measuring the structure of predicted
-networks, we highlight an interesting paradox: the models with the best
-performance measures are not necessarilly the models with the closest
-reconstructed network structure. We discuss these results in the context
-of establishing guidelines for the prediction of ecological
-interactions.
+basis of the first broad recommendation: optimal training set balance
+and its interaction with connectance and the specific binary classifier
+used is, in a sense, an hyperparameter that should be assessed. The
+distribution of results in fig.~\ref{fig:optimbias} and
+fig.~\ref{fig:optimvalue} show that there are variations around the
+trend, and multiple models should probably be trained on their
+``optimal'' training/testing set, as opposed to the same ones.
 
 The results presented here highlight an interesting paradox: although
-the Random Forest was ultimately able to get a correct estimate of
-network structure tbl.~\ref{tbl:comparison}, it ultimately remains a
-poor classifier, as evidenced by its low PR-AUC. This suggests that the
-goal of predicting \emph{interactions} and predicting \emph{networks}
-may not be solvable in the same way -- of course a perfect classifier of
-interactions would make a perfect network prediction; but even the best
-scoring predictor of interactions (the ensemble model) had not
-necessarily the best prediction of network structure. The tasks of
-predicting networks structure and of predicting interactions within
-networks are essentially two different ones. For some applications
-(\emph{.e.g.} comparison of network structure across gradients), one may
-care more about a robust estimate of the structure, at the cost at
-putting some interactions at the wrong place. For other applications
-(\emph{e.g.} identifying pairs of interacting species), one may
-conversely care more about getting as many pairs right, even though the
-mistakes accumulate in the form of a slightly worse estimate of network
-structure. How these two approaches can be reconciled is undoubtedly a
-task for further research. Despite this apparent tension at the heart of
+the k-NN model was ultimately able to get a correct estimate of network
+structure tbl.~\ref{tbl:comparison}, it ultimately remains a poor
+classifier, as evidenced by its low PR-AUC. This suggests that the goal
+of predicting \emph{interactions} and predicting \emph{networks} may not
+always be solvable in the same way -- of course a perfect classifier of
+interactions would make a perfect network prediction; indeed, the best
+scoring predictor of interactions (the ensemble model) had the best
+prediction of network structure. The tasks of predicting networks
+structure and of predicting interactions within networks are essentially
+two different ones. For some applications (\emph{.e.g.} comparison of
+network structure across gradients), one may care more about a robust
+estimate of the structure, at the cost at putting some interactions at
+the wrong place. For other applications (\emph{e.g.} identifying pairs
+of interacting species), one may conversely care more about getting as
+many pairs right, even though the mistakes accumulate in the form of a
+slightly worse estimate of network structure. How these two approaches
+can be reconciled is something to evaluate on a case-by-case basis,
+especially since there is no guarantee that an esemble model will always
+be the most precise one. Despite this apparent tension at the heart of
 the predictive exercise, we can use the results presented here to
@@
 
-First, because we should have more trust in reported interactions than
-in reported absences of interactions, we can draw on previous literature
-to recommend informedness as a measure to decide on a threshold (Chicco
-et al., 2021); this being said, because informedness is insensitive to
-bias, the model performance is better evaluated through the use of MCC
-fig.~\ref{fig:biasmccinf}. Because \(F_1\) is monotonously sensitive to
-classifier bias fig.~\ref{fig:bias} and network connectance
+First, because we have more trust in reported interactions than in
+reported absences of interactions (which are overwhelmingly
+\emph{pseudo}-absences), we can draw on previous literature to recommend
+informedness as a measure to decide on a threshold for binary
+classification (Chicco et al., 2021); this being said, because
+informedness is insensitive to bias (although it is a linear measure of
+skill), the overall model performance is better evaluated through the
+use of MCC fig.~\ref{fig:biasmccinf}. Because \(F_1\) is monotonously
+sensitive to classifier bias fig.~\ref{fig:bias} and network connectance
 fig.~\ref{fig:connectance}, MCC should be prefered as a measure of model
-evaluation and comparison.
-
-Second, because the PR-AUC responds more to network connectance
+evaluation and comparison. When dealing with multiple models, we
+therefore suggest to find the optimal threshold using informedness, and
+to pick the best model using MCC (assuming one does not want to use an
+ensemble model).
+
+Second, accuracy alone should not be the main measure of model
+performance, but rather an expectation of how well the model should
+behave given the class balance in the set on which predictions are made;
+this is because, as derived earlier, the expected accuracy for a
+no-skill no-bias classifier is \(\rho^2 + (1-\rho)^2\) (where \(\rho\)
+is the class balance), which will most often be large. This pitfall is
+notably illustrated in a recent food-web model (Caron et al., 2022)
+wherein the authors, using a training set of \(n = 10^4\) with only 100
+positive interactions (representing 0.1\% of the total interactions),
+reached a good accuracy. Reporting a good accuracy is not informative,
+especially when accuracy isn't (i) compared to the baseline expected
+value under the given class balance, and (ii) interpreted in the context
+of a measure that is not sensitive to the chance prediction of many
+negatives (like MCC).
+
+Third, because the PR-AUC responds more to network connectance
 fig.~\ref{fig:optimvalue} and training set imbalance
-fig.~\ref{fig:biasrocpr}, it should be used as a measure of model
-performance over the ROC-AUC. This is not to say that ROC-AUC should be
-discarded (in fact, a low ROC-AUC is a sign of an issue with the model),
-but that its interpretation should be guided by the PR-AUC value.
-Specifically, a high ROC-AUC is not informative, as it can be associated
-to a low PR-AUC (see \emph{e.g.} Random Forest in
+fig.~\ref{fig:biasrocpr} than ROC-AUC, it should be used as a measure of
+model performance over the ROC-AUC. This is not to say that ROC-AUC
+should be discarded (in fact, a low ROC-AUC is undoubtedly a sign of an
+issue with the model), but that its interpretation should be guided by
+the PR-AUC value. Specifically, a high ROC-AUC is not informative, as it
+can be associated to a low PR-AUC (see \emph{e.g.} Random Forest in
 tbl.~\ref{tbl:comparison}) This again echoes recommendations from other
-fields (Jeni et al., 2013; Saito \& Rehmsmeier, 2015).
-
-Third, for the same reasons that should give more weight to PR-AUC than
-ROC-AUC, accuracy alone should not be used as a measure of model
-performance, but rather as an expectation of how well the model should
-behave given the balance in the set on which predictions are made; this
-is because, as derived earlier, the expected accuracy for a no-skill
-no-bias classifier is \(\rho^2 + (1-\rho)^2\) (where \(\rho\) is the
-class balance), which will very easily by large. This pitfall is notably
-illustrated in a recent model (Caron et al., 2022) wherein the authors
-claim that using a training set of \(n = 10^4\) with only 100 positive
-interactions (representing 0.1\% of the total interactions) reached a
-good accuracy. Reporting a good accuracy isn't informative is this
-accuracy isn't (i) compared to the baseline expected value under the
-class balance, and (ii) interpreted in the context of a measure that
-isn't sensitive to true negatives (like MCC).
+fields (Jeni et al., 2013; Saito \& Rehmsmeier, 2015). We therefore
+expect to see high ROC-AUC values, and then to pick the model that
+maximizes the PR-AUC value. Taken together with the previous two
+guidelines, we strongly encourage to (i) ensure that accuracy and
+ROC-AUC are high (in the case of accuracy, higher than expected under
+no-skill no-bias situation), and (ii) to discuss the performance of the
+model in terms of the most discriminant measures, \emph{i.e.} PR-AUC and
+MCC.
 
 Finally, network connectance (\emph{i.e.} the empirical class imbalance)
-should inform the composition of the training and testing set. In the
-approach outlined here, we treat the class imbalance as an
-hyper-parameter, but \emph{test} the model on a set that has the same
-class imbalance as the actual dataset. This is an important change to
-the overall methodology so far, as it ensure that the prediction
-environment matches the testing environment, and so the values on the
-testing set can be directly compared to the values for the actual
-prediction. A striking result fig.~\ref{fig:optimbias} is that
-Informedness was almost always maximal at 50/50 balance (regardless of
-connectance), whereas MCC required \emph{more} positives to be maximized
-when connectance \emph{increases}. This has an important consequence in
-ecological networks, for which the pool of positive cases (interactions)
-to draw from is typically small: the most parsimonious measure
-(\emph{i.e.} the one requiring to discard the least amount of
-information to train the model) will give the best validation potential,
-and is probably the informedness (maximizing informedness is the
-generally accepted default for imbalanced classification; Schisterman et
-al., 2005). This last result further strengthens the assumption that the
-amount of bias \emph{is} an hyper-parameter that must be fine-tuned, as
-using the wrong bias can lead to models with excess biass or lower
-skill.
+should inform the composition of the training and testing set, because
+it is an ecologically relevant value. In the approach outlined here, we
+treat the class imbalance of the training set as an hyper-parameter, but
+\emph{test} the model on a set that has the same class imbalance as the
+actual dataset. This is an important distinction, as it ensure that the
+prediction environment matches the testing environment (as we cannot
+manipulate the connectance of the empirical dataset on which the
+predictions will be made), and so the values measured on the testing set
+(or validation set if the data volume allows one to exists) can be
+directly compared to the values for the actual prediction. A striking
+result fig.~\ref{fig:optimbias} is that Informedness was almost always
+maximal at 50/50 balance (regardless of connectance), whereas MCC
+required \emph{more} positives to be maximized when connectance
+\emph{increases}, matching the idea that it is a more stringent measure
+of performance. This has an important consequence in ecological
+networks, for which the pool of positive cases (interactions) to draw
+from is typically small: the most parsimonious measure (\emph{i.e.} the
+one requiring to discard the least amount of interactions to train the
+model) will give the best validation potential, and in this light is
+very likely informedness (maximizing informedness is, in fact, the
+generally accepted default for imbalanced classification regardless of
+the problem domain; Schisterman et al., 2005). This last result further
+strengthens the assumption that the amount of bias \emph{is} an
+hyper-parameter that must be fine-tuned, as using the wrong bias can
+lead to models with lower performance; for this reason, it makes sense
+to not train all models on the same training/testing set, but rather to
+optimize the set composition for each of them.
 
@@
 manuscript. This research was enabled in part by support provided by
-Calcul Québec (www.calculquebec.ca) and Compute Canada
-(www.computecanada.ca) through the Narval general purpose cluster. TP is
-supported by a NSERC Discovery Grant and Discovery Acceleration
-Supplement, by funding to the Viral Emergence Research Initiative
-(VERENA) consortium including NSF BII 2021909, and by a grant from the
-Institut de Valorisation des Données (IVADO).
+Calcul Québec (www.calculquebec.ca) through the Narval general purpose
+cluster. TP is supported by the Fondation Courtois, a NSERC Discovery
+Grant and Discovery Acceleration Supplement, by funding to the Viral
+Emergence Research Initiative (VERENA) consortium including NSF BII
+2021909, and by a grant from the Institut de Valorisation des Données
+(IVADO).
 

### 2022-06-09 preprint.tex
Deploying to gh-pages from @ PoisotLab/ms_networkary_classifiers@e4608d1df93d7206dc91c3c754a12d9a884e5ac0 🚀

@@
 
-\noindent {\LARGE \textbf{\textsf{Guidelines for the validation of
-machine learning predictions of species interactions}}}
+\noindent {\LARGE \textbf{\textsf{Guidelines for the prediction of
+species interactions through binary classification}}}
 

### 2022-06-09 draft.tex
Deploying to gh-pages from @ PoisotLab/ms_networkary_classifiers@b8ef26e57717e6fa9014a0df28be203a8e88aca9 🚀

@@
 their MCC are not able to reach a high PR-AUC even at higher
-connectances.All models reached a higher performance on more connected
+connectances. All models reached a higher performance on more connected
 networks, and using more balanced training sets. In all cases,
@@
 had a significant impact on the performance of some machines
-fig.~\ref{fig:optimvalue}. Notably, Decision Tree, and k-NN, as well as
-Random forest to a lower extent, had low values of PR-AUC. In all cases,
-the Boosted Regression Tree was reaching very good predictions
+(fig.~\ref{fig:optimvalue}). Notably, Decision Tree, and k-NN, as well
+as Random forest to a lower extent, had low values of PR-AUC. In all
+cases, the Boosted Regression Tree was reaching very good predictions
 (especially for connectances larger than 0.1), and the ensemble was
@@
 analysis; specifically, we measured the connectance, nestedness
-(\(\eta\); \textbf{Bastolla2009ArcMut?}), modularity (\(Q\);
-\textbf{Barber2007ModCom?}), asymmetry (\(A\);
-\textbf{Delmas2018AnaEco?}), and Jaccard network dissimilarity (Canard
-et al., 2014). This process was repeated 250 times, and the results are
-presented in tbl.~\ref{tbl:comparison}. The k-NN model is an interesting
-instance here: it produces the network that looks the most like the
-original dataset, despite having the lowest PR-AUC, suggesting it hits
-high recall at the cost of low precision. The ensemble was able to reach
-a very high PR-AUC (and a very high ROC-AUC), which translated into more
-accurate reconstructions of the structure of the network (with the
-exception of modulairty, which is underestimated by \(0.03\)). This
-result bears elaborating. Measures of model performance capture how much
-of the interactions and non-interactions are correctly identified. As
-long as these predictions are not perfect, some interactions will be
-predicted at the ``wrong'' position in the network; these measures
-cannot describe the structural effect of these mistakes. On the other
-hand, measures of network structure can have the same value with
-interactions that fall at drastically different positions; this is in
-part because a lot of these measures covary with connectance, and in
-part because as long as these values are not 0 or their respective
-maximum, there is a large number of network configurations that can have
-the same value. That ROC-AUC is consistently larger than PR-AUC may be a
-case of this measure masking models that are not, individually, strong
-predictors (Jeni et al., 2013). In this specif example, the combination
-of individually ``adequate'' models resulted in an extremely strong
-ensemble, suggesting that the correct prediction of interactions (as
-measured by MCC, Inf., ROC-AUC, and PR-AUC) \emph{and} network
-properties is indeed a feasible task under appropriately
-hyper-parameterized models.
+(\(\eta\); Bastolla et al., 2009), modularity (\(Q\); Barber, 2007),
+asymmetry (\(A\); Delmas et al., 2018), and Jaccard network
+dissimilarity (Canard et al., 2014). This process was repeated 250
+times, and the results are presented in tbl.~\ref{tbl:comparison}. The
+k-NN model is an interesting instance here: it produces the network that
+looks the most like the original dataset, despite having the lowest
+PR-AUC, suggesting it hits high recall at the cost of low precision. The
+ensemble was able to reach a very high PR-AUC (and a very high ROC-AUC),
+which translated into more accurate reconstructions of the structure of
+the network (with the exception of modulairty, which is underestimated
+by \(0.03\)). This result bears elaborating. Measures of model
+performance capture how much of the interactions and non-interactions
+are correctly identified. As long as these predictions are not perfect,
+some interactions will be predicted at the ``wrong'' position in the
+network; these measures cannot describe the structural effect of these
+mistakes. On the other hand, measures of network structure can have the
+same value with interactions that fall at drastically different
+positions; this is in part because a lot of these measures covary with
+connectance, and in part because as long as these values are not 0 or
+their respective maximum, there is a large number of network
+configurations that can have the same value. That ROC-AUC is
+consistently larger than PR-AUC may be a case of this measure masking
+models that are not, individually, strong predictors (Jeni et al.,
+2013). In this specif example, the combination of individually
+``adequate'' models resulted in an extremely strong ensemble, suggesting
+that the correct prediction of interactions (as measured by MCC, Inf.,
+ROC-AUC, and PR-AUC) \emph{and} network properties is indeed a feasible
+task under appropriately hyper-parameterized models.
 
@@
 the k-NN model was ultimately able to get a correct estimate of network
-structure tbl.~\ref{tbl:comparison}, it ultimately remains a poor
-classifier, as evidenced by its low PR-AUC. This suggests that the goal
-of predicting \emph{interactions} and predicting \emph{networks} may not
-always be solvable in the same way -- of course a perfect classifier of
-interactions would make a perfect network prediction; indeed, the best
-scoring predictor of interactions (the ensemble model) had the best
-prediction of network structure. The tasks of predicting networks
-structure and of predicting interactions within networks are essentially
-two different ones. For some applications (\emph{.e.g.} comparison of
-network structure across gradients), one may care more about a robust
-estimate of the structure, at the cost at putting some interactions at
-the wrong place. For other applications (\emph{e.g.} identifying pairs
-of interacting species), one may conversely care more about getting as
-many pairs right, even though the mistakes accumulate in the form of a
-slightly worse estimate of network structure. How these two approaches
-can be reconciled is something to evaluate on a case-by-case basis,
-especially since there is no guarantee that an esemble model will always
-be the most precise one. Despite this apparent tension at the heart of
-the predictive exercise, we can use the results presented here to
-suggest a number of guidelines.
+structure (see tbl.~\ref{tbl:comparison} and fig.~\ref{fig:ecovalid}),
+it ultimately remains a poor classifier, as evidenced by its low PR-AUC.
+This suggests that the goal of predicting \emph{interactions} and
+predicting \emph{networks} may not always be solvable in the same way --
+of course a perfect classifier of interactions would make a perfect
+network prediction; indeed, the best scoring predictor of interactions
+(the ensemble model) had the best prediction of network structure. The
+tasks of predicting networks structure and of predicting interactions
+within networks are essentially two different ones. For some
+applications (\emph{.e.g.} comparison of network structure across
+gradients), one may care more about a robust estimate of the structure,
+at the cost at putting some interactions at the wrong place. For other
+applications (\emph{e.g.} identifying pairs of interacting species), one
+may conversely care more about getting as many pairs right, even though
+the mistakes accumulate in the form of a slightly worse estimate of
+network structure. How these two approaches can be reconciled is
+something to evaluate on a case-by-case basis, especially since there is
+no guarantee that an esemble model will always be the most precise one.
+Despite this apparent tension at the heart of the predictive exercise,
+we can use the results presented here to suggest a number of guidelines.
 
@@
 skill), the overall model performance is better evaluated through the
-use of MCC fig.~\ref{fig:biasmccinf}. Because \(F_1\) is monotonously
-sensitive to classifier bias fig.~\ref{fig:bias} and network connectance
-fig.~\ref{fig:connectance}, MCC should be prefered as a measure of model
-evaluation and comparison. When dealing with multiple models, we
-therefore suggest to find the optimal threshold using informedness, and
-to pick the best model using MCC (assuming one does not want to use an
-ensemble model).
+use of MCC (figs.~\ref{fig:optimbias}, \ref{fig:optimvalue}). Because
+\(F_1\) is monotonously sensitive to classifier bias fig.~\ref{fig:bias}
+and network connectance fig.~\ref{fig:connectance}, MCC should be
+prefered as a measure of model evaluation and comparison. When dealing
+with multiple models, we therefore suggest to find the optimal threshold
+using informedness, and to pick the best model using MCC (assuming one
+does not want to use an ensemble model).
 
@@
 Third, because the PR-AUC responds more to network connectance
-fig.~\ref{fig:optimvalue} and training set imbalance
-fig.~\ref{fig:biasrocpr} than ROC-AUC, it should be used as a measure of
-model performance over the ROC-AUC. This is not to say that ROC-AUC
+(fig.~\ref{fig:optimvalue}) and training set imbalance
+(fig.~\ref{fig:optimbias}) than ROC-AUC, it should be used as a measure
+of model performance over the ROC-AUC. This is not to say that ROC-AUC
 should be discarded (in fact, a low ROC-AUC is undoubtedly a sign of an
@@
 directly compared to the values for the actual prediction. A striking
-result fig.~\ref{fig:optimbias} is that Informedness was almost always
-maximal at 50/50 balance (regardless of connectance), whereas MCC
+result from fig.~\ref{fig:optimbias} is that Informedness was almost
+always maximal at 50/50 balance (regardless of connectance), whereas MCC
 required \emph{more} positives to be maximized when connectance
@@
 
+\leavevmode\hypertarget{ref-Barber2007ModCom}{}%
+Barber, M. J. (2007). Modularity and community detection in bipartite
+networks. \emph{Physical Review E}, \emph{76}(6), 066102.
+\url{https://doi.org/10.1103/PhysRevE.76.066102}
+
+\leavevmode\hypertarget{ref-Bastolla2009ArcMut}{}%
+Bastolla, U., Fortuna, M. A., Pascual-García, A., Ferrera, A., Luque,
+B., \& Bascompte, J. (2009). The architecture of mutualistic networks
+minimizes competition and increases biodiversity. \emph{Nature},
+\emph{458}(7241), 1018--1020. \url{https://doi.org/10.1038/nature07950}
+
 \leavevmode\hypertarget{ref-Beauchesne2016ThiOut}{}%
@@
 e0222916. \url{https://doi.org/10.1371/journal.pone.0222916}
+
+\leavevmode\hypertarget{ref-Delmas2018AnaEco}{}%
+Delmas, E., Besson, M., Brice, M.-H., Burkle, L. A., Dalla Riva, G. V.,
+Fortin, M.-J., Gravel, D., Guimarães, P. R., Hembry, D. H., Newman, E.
+A., Olesen, J. M., Pires, M. M., Yeakel, J. D., \& Poisot, T. (2018).
+Analysing ecological networks of species interactions. \emph{Biological
+Reviews}, 112540. \url{https://doi.org/10.1111/brv.12433}
 

### 2022-06-09 draft.tex
Deploying to gh-pages from @ PoisotLab/ms_networkary_classifiers@ae2c54fa2068f9f57b1e56217327ebc87637c577 🚀

@@
 and \(\kappa\). Accuracy increases with skill, but also increases when
-the bias tends towards estimating \emph{fewer} interactions. The \(F_1\)
-score increases with skill but also increases when the bias tends
-towards estimating \emph{more} interactions. Interestingly, \(\kappa\)
-responds as expected to skill (being negative whenever \(s < 0.5\)), and
-peaks for values of \(b \approx 0.5\); nevertheless, the value of bias
-for which \(\kappa\) is maximized in \emph{not} \(b=0.5\), but instead
-increases with classifier skill. In other words, at equal skill,
-maximizing \(\kappa\) would lead to select a \emph{more} biased
-classifier.}\label{fig:bias}
+the bias tends towards estimating \emph{fewer} interactions (this
+follows from the derivations in the text, not shown in the figure).
+Interestingly, \(\kappa\) responds as expected to skill (being negative
+whenever \(s < 0.5\)), and peaks for values of \(b \approx 0.5\);
+nevertheless, the value of bias for which \(\kappa\) is maximized in
+\emph{not} \(b=0.5\), but instead increases with classifier skill. In
+other words, at equal skill, maximizing \(\kappa\) would lead to select
+a \emph{more} biased classifier.}\label{fig:bias}
 }
@@
 
-In order to pick the best adjacency matrix for a given trained machine,
+In order to pick the best confusion matrix for a given trained machine,
 we performed a thresholding approach using 500 steps on predictions from

### 2022-06-10 draft.tex
Deploying to gh-pages from @ PoisotLab/ms_networkary_classifiers@c70bb0c32fc9f7ee1cd0c8eef075914e1493ff52 🚀

@@
 unweighted, directed network) is a column vector
-\(\mathbf{o}^T = [\rho, \qquad (1-\rho)]\) (we can safely drop the
-\(S^2\) terms, as we will work on the confusion matrix, which ends up
-expressing \emph{relative} values). We will apply skill and bias to this
-matrix, and measure how a selection of performance metrics respond to
-changes in these values, in order to assess their suitability for model
-evaluation.
+\(\mathbf{o}^T = [\rho, (1-\rho)]\) (we can safely drop the \(S^2\)
+terms, as we will work on the confusion matrix, which ends up expressing
+\emph{relative} values). We will apply skill and bias to this matrix,
+and measure how a selection of performance metrics respond to changes in
+these values, in order to assess their suitability for model evaluation.
 
@@
 \(\rho\). The predictions of a no-skill classifier can be expressed as a
-row vector \(\mathbf{p} = [\rho (1-\rho)]\). The confusion matrix
+row vector \(\mathbf{p}^T = [\rho, (1-\rho)]\). The confusion matrix
 \(\mathbf{M}\) for a no-skill classifier is given by the element-wise
@@
 matrix \(\mathbf{S}\) with diagonal elements equal to \(s\), and
-off-diagonal elements equal to \((1-s)\), and re-express the
-skill-adjusted confusion matrix as \(\mathbf{M} \odot \mathbf{S}\),
-\emph{i.e.}
+off-diagonal elements equal to \((1-s)\), which allows to regulate how
+many predictions are wrong, under the assumption that the bias is the
+same (\emph{i.e.} the classifier is as likely to make a false positive
+or a false negative). The skill-adjusted confusion matrix is
+\(\mathbf{M} \odot \mathbf{S}\), \emph{i.e.}
 

### 2022-09-11 draft.tex
Deploying to gh-pages from @ PoisotLab/ms_networkary_classifiers@ae26545e508f95c05f6e91f39f9b29085a150e2a 🚀

@@
 on simulated networks, we develop a series of guidelines related to the
-choice of measures to use for model selection, and the degree of
-unbiasing to apply to the training dataset.%
-    \item Neither classifier accuracy nor the ROC-AUC are informative
-measures for the performance of interaction prediction. PR-AUC is a
-fairer assessment of performance. In some cases, even standard measures
-can lead to selecting a more biased classifier because the effect of
-connectance is strong. The amount of correction to apply to the training
-dataset depends on network connectance, on the measure to be optimized,
-and only weakly on the classifier.%
+choice of measures to use for model selection, and the ways to assemble
+the training dataset.%
+    \item Neither classifier accuracy nor the area under the receiver
+operating characteristic curve (ROC-AUC) are informative measures for
+the performance of interaction prediction. The area under the
+precision-recall curve (PR-AUC) is a fairer assessment of performance.
+In some cases, even standard measures can lead to selecting a more
+biased classifier because the effect of connectance is strong. The
+amount of correction to apply to the training dataset depends on network
+connectance, on the measure to be optimized, and only weakly on the
+classifier.%
     \item These results reveal that training machines to predict
@@
 
-Ecological networks are a backbone for key ecological and evolutionary
-processes; yet enumerating all of the interactions between \(S\) species
-is a daunting task, as it scales with \(S^2\), \emph{i.e.} the squared
-species richness (Martinez, 1992). Recent contributions to the field of
-ecological network prediction (Becker et al., 2022; Pichler et al.,
-2020; Strydom et al., 2021) highlight that although interactions can be
-predicted by adding ecologically relevant information (in the form of,
-\emph{e.g.} traits), we do not have robust guidelines as to how the
-predictive ability of these models should be evaluated, nor about how
-the models should be trained. Here, by relying on simple derivations and
-a series of simulations, we formulate a number of such guidelines,
-specifically for the case of binary classifiers derived from thresholded
-values. Specifically, we conduct an investigation of the models in terms
-of their skill (ability to make the right prediction), bias (trends
-towards systematically over-predicting one class), class imbalance (the
-relative number of cases representing interactions), and show how these
-effects interact. We conclude on the fact that models with the best
-interaction-scale predictive score do not necessarily result in the most
-accurate representation of the network.
+Species interactions, forming ecological networks, are a backbone for
+key ecological and evolutionary processes; yet enumerating all of the
+interactions between \(S\) species is a daunting task, as it scales with
+\(S^2\), \emph{i.e.} the squared species richness (Martinez, 1992).
+Recent contributions to the field of ecological network prediction
+(Becker et al., 2022; Pichler et al., 2020; Strydom et al., 2021)
+highlight that although interactions can be predicted by adding
+ecologically relevant information (in the form of, \emph{e.g.} traits),
+we do not have robust guidelines as to how the predictive ability of
+models recommending species interactions should be evaluated, nor about
+how these models should be trained. Here, by relying on simple
+derivations and a series of simulations, we formulate a number of such
+guidelines, specifically for the case of binary classifiers derived from
+thresholded values. Specifically, we conduct an investigation of the
+models in terms of their skill (ability to make the right prediction),
+bias (trends towards systematically over-predicting one class), class
+imbalance (the relative number of cases representing interactions), and
+show how these effects interact. We conclude on the fact that models
+with the best interaction-scale predictive score do not necessarily
+result in the most accurate representation of the true network.
 
@@
 methodological issues with two fields in biology: species distribution
-models (SDMs), and genomics. SDMs suffers from issues affecting
+modelling (SDMs), and genomics. SDMs suffers from issues affecting
 interactions prediction, namely low prevalence (due to sparsity of
 observations/interactions) and data aggregation (due to bias in sampling
-some locations/species). In previous work, Allouche et al. (2006)
-suggested that \(\kappa\) was a better test of model performance than
-the True Skill Statistic (TSS; which we refer to as Youden's
-informedness); these conclusions were later criticized by Somodi et al.
-(2017), who emphasized that informedness' is affected both by prevalence
-and bias. Although this work offers recommendations about the comparison
-of models, it doesn't establishes baselines or good practices for
-training on imbalanced ecological data, or ways to remedy the imbalance.
-Steen et al. (2021) show that, when applying spatial thinning (a process
-that has no analogue in networks), the best approach to train ML-based
-SDMs varies according to the balancing of the dataset, and the
-evaluation measures used. This suggests that there is no single
-``recipe'' that is guaranteed to give the best model. By contrast to
-networks, SDMs have the advantage of being able to both thin datasets to
-remove some of the sampling bias (\emph{e.g.} Inman et al., 2021), but
-also to create pseudo-absences to inflate the number of supposed
-negatives in the dataset (\emph{e.g.} Iturbide et al., 2015).
+some locations/species). An important challenge lies in the fact that
+the best measure to quantify the performance of a model is not
+necessarilly a point of consensus (these methods, their interpretation,
+and the way they are measured, are covered in depth in the next
+section). In previous work, Allouche et al. (2006) suggested that
+Cohen's \(\kappa\) agreement score (\(\kappa\) thereafter) was a better
+test of model performance than the True Skill Statistic (TSS; which we
+refer to as Youden's informedness thereafter); these conclusions were
+later criticized by Somodi et al. (2017), who emphasized that
+informedness is affected both by prevalence and bias. Although this work
+offers recommendations about the comparison of models, it doesn't
+establishes baselines or good practices for training on imbalanced
+ecological data, or ways to remedy the imbalance. Steen et al. (2021)
+show that, when applying spatial thinning (artificially re-balancing
+observation data in space to avoid artifacts due to auto-correlation),
+the best approach to train ML-based SDMs varies according to the
+balancing of the dataset, and the evaluation measures used; there is no
+single ``recipe'' that is guaranteed to give the best model. By contrast
+to networks, SDMs have the advantage of being able to both thin datasets
+to remove some of the sampling bias (\emph{e.g.} Inman et al., 2021),
+but also to create pseudo-absences to inflate the number of supposed
+negatives in the dataset (\emph{e.g.} Iturbide et al., 2015). These
+powerful ways to remove data bias often have no analogue in networks,
+removing one potential tool from our methodological toolkit, and making
+the task of network prediction through classification potentially more
+demanding, and more prone to underlying data biases.
 
@@
 bookmaker's accuracy, and the True-Skill Statistic) when the imbalance
-in the dataset may not be representative of the actual imbalance.
+in the dataset may not be representative of the actual imbalance. In a
+way, the measures themselves need not be a strong focus for network
+prediction, as they are routinely used in other field; the
+discipline-specific question we seek to address is: `which metric should
+be employed when predicting networks, and how to optimize it?'
 
@@
 data on interaction makes a lot of imbalance correction methods (see
-\emph{e.g.} Branco et al., 2015) hard to apply, which indicators can be
-optimized by sacrificing least amount of positive interaction data?
-
-It may sound counter-intuitive to care so deeply about how good a
-classifier with no-skill is, as by definition, is has no skill. The
-necessity of this exercise has its roots in the paradox of accuracy:
-when the desired class (``two species interact'') is rare, a model that
-gets less ecologically performant by only predicting the opposite class
-(``these two species do not interact'') sees its accuracy increase;
-because most of the guesses have ``these two species do not interact''
-as a correct answer, a model that never predicts interactions would be
-right an overwhelming majority of the time; it would also be utterly
-useless. Herein lies the core challenge of predicting species
-interactions: the extreme imbalance between classes makes the training
-of predictive models difficult, and their validation even more so as we
-do not reliably know which negatives are true. The connectance (the
-proportion of realized interactions, usually the number of interactions
-divided by the number of species pairs) of empirical networks is usually
-well under 20\%, with larger networks having a lower connectance
-(MacDonald et al., 2020), and therefore being increasingly difficult to
-predict.
+\emph{e.g.} Branco et al., 2015) hard to apply, which measures of model
+performance can be optimized by sacrificing least amount of positive
+interaction data?
+
+A preliminary question is to examin the baseline performance of these
+measures, \emph{i.e.} the values they would take on hypothetical
+networks based on a classifier that has no-skill. It may sound
+counter-intuitive to care so deeply about how good a classifier with
+no-skill is, as by definition, is has no skill. The necessity of this
+exercise has its roots in the paradox of accuracy: when the desired
+class (``two species interact'') is rare, a model that gets less
+ecologically performant by only predicting the opposite class (``these
+two species do not interact'') sees its accuracy increase; because most
+of the guesses have ``these two species do not interact'' as a correct
+answer, a model that never predicts interactions would be right an
+overwhelming majority of the time; it would also be utterly useless.
+Herein lies the core challenge of predicting species interactions: the
+extreme imbalance between classes makes the training of predictive
+models difficult, and their validation even more so as we do not
+reliably know which negatives are true. The connectance (the proportion
+of realized interactions, usually the number of interactions divided by
+the number of species pairs) of empirical networks is usually well under
+20\%, with larger networks having a lower connectance (MacDonald et al.,
+2020), and therefore being increasingly difficult to predict.
 
@@
 Binary classifiers, which it to say, machine learning algorithms whose
-answer is a categorical value, are usually assessed by measuring
-properties of their confusion matrix, \emph{i.e.} the contingency table
-reporting true/false positive/negative hits. A confusion matrix is laid
-out as
+answer is a binary value, are usually assessed by measuring properties
+of their confusion matrix, \emph{i.e.} the contingency table reporting
+true/false positive/negative hits. A confusion matrix is laid out as
 
@@
 other measures informing us on how the predictions fail.
-
-There is an immense diversity of measures to evaluate the performance of
-classification tasks (Ferri et al., 2009). Here we will focus on five of
-them with high relevance for imbalanced learning (He \& Ma, 2013). The
-choice of metrics with relevance to class-imbalanced problems is
-fundamental, because as Japkowicz (2013) unambiguously concluded,
-``relatively robust procedures used for unskewed data can break down
-miserably when the data is skewed.'' Following Japkowicz (2013), we
-focus on two ranking metrics (the areas under the Receiver Operating
-Characteristic and Precision Recall curves), and three threshold metrics
-(\(\kappa\), informedness, and MCC; we will briefly discuss \(F_1\) but
-show early on that it has undesirable properties).
-
-The \(\kappa\) measure (Landis \& Koch, 1977) establishes the extent to
-which two observers (the network and the prediction) agree, and is
-measured as
-
-\[
-2\frac{tp\times tn - fn\times fp}{(tp+fp)\times (fp+tn)+(tn+fp)\times (tn+fn)} \,.
-\]
-
-Informedness (Youden, 1950) (also known as bookmaker informedness or the
-True Skill Statistic) is \(\text{TPR}+\text{TNR}-1\), where
-\(\text{TPR}= tp/(tp+fn)\) and \(\text{TNR} = tn/(tn+fp)\). Informedness
-can be used to find the optimal cutpoint in thresholding analyses
-(Schisterman et al., 2005); indeed, the maximal informedness corresponds
-to the point on the ROC curve that is closest to the perfect classifier
-point. The formula for informedness is
-
-\[\frac{tp}{tp+fn}+\frac{tn}{tn+fp}-1\,.\]
-
-The MCC is defined as
-
-\[
-\frac{tp\times tn - fn\times fp}{\sqrt{(tp+fp)\times (tp+fn)\times (tn+fp)\times (tn+fn)}} \,.
-\]
-
-Finally, \(F_1\) is the harmonic mean of precision (the chance that
-interaction was correctly detected as such) and sensitivity (the ability
-to correctly classify interactions), and is defined as
-
-\[
-2\frac{tp}{2\times tp + fp + fn}\,.
-\]
 
@@
 on the \(x\) axis, and the positive predictive value on the \(y\) axis.
-By comparison with the previous paragraph, it is obvious that \(F_1\)
-and MCC have ties to the PR curve (being close to the expected PR-AUC),
-and that informedness has ties to the ROC curve (whereby the threshold
-maximizing informedness is also the point of maximal inflection on the
-ROC curve). One important difference between ROC and PR is that the
-later does not prominently account for the size of the true negative
-compartments: in short, it is more sensitive to the correct positive
-predictions. In a context of strong imbalance, PR-AUC is therefore a
-more stringent test of model performance.
+
+There is an immense diversity of measures to evaluate the performance of
+classification tasks (Ferri et al., 2009). Here we will focus on five of
+them with high relevance for imbalanced learning (He \& Ma, 2013). The
+choice of metrics with relevance to class-imbalanced problems is
+fundamental, because as Japkowicz (2013) unambiguously concluded,
+``relatively robust procedures used for unskewed data can break down
+miserably when the data is skewed.'' Following Japkowicz (2013), we
+focus on two ranking metrics (the areas under the Receiver Operating
+Characteristic and Precision Recall curves), and three threshold metrics
+(\(\kappa\), informedness, and MCC; we will briefly discuss \(F_1\) but
+show early on that it has undesirable properties).
+
+The \(\kappa\) measure (Landis \& Koch, 1977) establishes the extent to
+which two observers (the network and the prediction) agree, and is
+measured as
+
+\[
+2\frac{tp\times tn - fn\times fp}{(tp+fp)\times (fp+tn)+(tn+fp)\times (tn+fn)} \,.
+\]
+
+Informedness (Youden, 1950) (also known as bookmaker informedness or the
+True Skill Statistic) is \(\text{TPR}+\text{TNR}-1\), where
+\(\text{TPR}= tp/(tp+fn)\) and \(\text{TNR} = tn/(tn+fp)\). Informedness
+can be used to find the optimal cutpoint in thresholding analyses
+(Schisterman et al., 2005); indeed, the maximal informedness corresponds
+to the point on the ROC curve that is closest to the perfect classifier
+point. The formula for informedness is
+
+\[\frac{tp}{tp+fn}+\frac{tn}{tn+fp}-1\,.\]
+
+The MCC is defined as
+
+\[
+\frac{tp\times tn - fn\times fp}{\sqrt{(tp+fp)\times (tp+fn)\times (tn+fp)\times (tn+fn)}} \,.
+\]
+
+Finally, \(F_1\) is the harmonic mean of precision (the chance that
+interaction was correctly detected as such) and sensitivity (the ability
+to correctly classify interactions), and is defined as
+
+\[
+2\frac{tp}{2\times tp + fp + fn}\,.
+\]
+
+One noteworthy fact is that \(F_1\) and MCC have ties to the PR curve
+(being close to the expected PR-AUC), and that informedness has ties to
+the ROC curve (whereby the threshold maximizing informedness is also the
+point of maximal inflection on the ROC curve). One important difference
+between ROC and PR is that the later does not prominently account for
+the size of the true negative compartments: in short, it is more
+sensitive to the correct positive predictions. In a context of strong
+imbalance, PR-AUC is therefore a more stringent test of model
+performance.
 
@@
 
-In this section, we will assume a network of connectance \(\rho\),
-\emph{i.e.} having \(\rho S^2\) interactions (where \(S\) is the species
-richness), and \((1-\rho) S^2\) non-interactions. Therefore, the vector
-describing the \emph{true} state of the network (assumed to be an
-unweighted, directed network) is a column vector
+In this section, we will assume a network with connectance equal to a
+scalar \(\rho\), \emph{i.e.} having \(\rho S^2\) interactions (where
+\(S\) is the species richness), and \((1-\rho) S^2\) non-interactions.
+Therefore, the vector describing the \emph{true} state of the network
+(assumed to be an unweighted, directed network) is a column vector
 \(\mathbf{o}^T = [\rho, (1-\rho)]\) (we can safely drop the \(S^2\)
@@
 
-The training sample is composed of 50\% of the \(10^4\) possible entries
-in the network, \emph{i.e.} \(n=5000\). Out of these interactions, we
-pick a proportion \(\nu\) (the training set balance) to be positive, so
-that the training set has \(\nu n\) interactions, and \((1-\nu) n\)
-non-interactions. We vary \(\nu\) uniformly in \(]0,1[\). This allows to
-evaluate how the measures of binary classification performance respond
-to artificially rebalanced dataset for a given network connectance. The
-rest of the dataset is used as a testing set, on which all further
-measures are calculated. Note that although the training set is balanced
-arbitrarily, the testing set is assembled so that it has the exact
-connectance of the entire network; this ensures that the model is
-evaluated under the class imbalance where the predictions will be made,
-which represents a more meaningful evaluation. Note also that although
-the simulated networks are bipartite, the algorithms have no
-``knowledge'' of the network structure, and simply look at pairs of
-species; therefore, the approach outlined here would also work for
-unipartite networks.
+The training sample is composed of a random pick of up to 50\% of the
+\(10^4\) possible entries in the network, \emph{i.e.} \(n=5000\). Out of
+these interactions, we pick a proportion \(\nu\) (the training set
+balance) to be positive, so that the training set has \(\nu n\)
+interactions, and \((1-\nu) n\) non-interactions. We vary \(\nu\)
+uniformly in \(]0,1[\). This allows to evaluate how the measures of
+binary classification performance respond to artificially rebalanced
+dataset for a given network connectance. The rest of the dataset is used
+as a testing set, on which all further measures are calculated. Note
+that although the training set is balanced arbitrarily, the testing set
+is assembled so that it has the exact connectance of the entire network;
+this ensures that the model is evaluated under the class imbalance where
+the predictions will be made, which represents a more meaningful
+evaluation. Furthermore, to avoid artifacts due to different sizes of
+the training and testing set within a single network, the number of
+entries in both sets are equal. Note also that although the simulated
+networks are bipartite, the algorithms have no ``knowledge'' of the
+network structure, and simply look at pairs of species; therefore, the
+approach outlined here would also work for unipartite networks.
 
@@
 (\emph{e.g.} Desjardins-Proulx et al., 2017), a regression tree, a
-regression random forest, and a boosted regression tree. Following
-results from Pichler et al. (2020), linear models have not been
-considered (in any way, the relationship in the simulated networks is
-non-linear). The point of these numerical experiments is \emph{not} to
-recommend the best model (this is likely problem-specific), but to
-highlight a series of recommendations that would work for supervised
-learning tasks. All models were taken from the \texttt{MLJ.jl} package
-(Blaom et al., 2020; Blaom \& Vollmer, 2020) in Julia 1.7 (Bezanson et
-al., 2017). All machines use the default parameterization; this is an
-obvious deviation from best practices, as the hyperparameters of any
-machine require training before its application on a real dataset. As we
-use 612500 such datasets, this would require over 2 millions unique
-instances of tweaking the hyperparameters, which is prohibitive from a
-computing time point of view. An important thing to keep in mind is that
-the problem we simulate has been designed to be simple to solve: we
-expect all machines with sensible default parameters to fare well ---
-the results presented in the later sections show that this assumption is
-warranted, and we further checked that the models do not overfit by
-ensuring that there is never more than 5\% of difference between the
-accuracy on the training and testing sets. All machines return a
-quantitative prediction, usually (but not necessarily) in \([0,1]\),
-which is proportional (but not necessarily linearly) to the probability
-of an interaction between \(i\) and \(j\). The ROC-AUC and PR-AUC (and
-therefore the thresholds) can be measured by integrating over the domain
-of the values return by each machine, but in order to make the
-average-based ensemble model more meaningful, all predictions are
-expressed in \([0,1]\).
+regression random forest, and a boosted regression tree; the later three
+methods are turned into classifiers using thresholding, which oftentimes
+provides better results than classification when faced with class
+imbalance (Hong et al., 2016). Following results from Pichler et al.
+(2020), linear models have not been considered (in any way, the
+relationship in the simulated networks is non-linear). The point of
+these numerical experiments is \emph{not} to recommend the best model
+(this is likely problem-specific), but to highlight a series of
+recommendations that would work for supervised learning tasks. All
+models were taken from the \texttt{MLJ.jl} package (Blaom et al., 2020;
+Blaom \& Vollmer, 2020) in Julia 1.7 (Bezanson et al., 2017). All
+machines use the default parameterization; this is an obvious deviation
+from best practices, as the hyperparameters of any machine require
+training before its application on a real dataset. As we use 612500 such
+datasets, this would require over 2 millions unique instances of
+tweaking the hyperparameters, which is prohibitive from a computing time
+point of view. An important thing to keep in mind is that the problem we
+simulate has been designed to be simple to solve: we expect all machines
+with sensible default parameters to fare well --- the results presented
+in the later sections show that this assumption is warranted, and we
+further checked that the models do not overfit by ensuring that there is
+never more than 5\% of difference between the accuracy on the training
+and testing sets. All machines return a quantitative prediction, usually
+(but not necessarily) in \([0,1]\), which is proportional (but not
+necessarily linearly) to the probability of an interaction between \(i\)
+and \(j\). The ROC-AUC and PR-AUC (and therefore the thresholds) can be
+measured by integrating over the domain of the values return by each
+machine, but in order to make the average-based ensemble model more
+meaningful, all predictions are expressed in \([0,1]\).
 
@@
 biological traits. This may be because the dataset of Becker et al.
-(2022) was known to be under-sampled, and so the network itself
-contained less information than the network and species traits. There is
-no general conclusion to draw from either these results or ours, besides
-reinforcing the need to be pragmatic about which models should be
-included in the ensemble, and whether to use an ensemble at all. In a
-sense, the surprising performance of the ensemble model should form the
-basis of the first broad recommendation: optimal training set balance
-and its interaction with connectance and the specific binary classifier
-used is, in a sense, an hyperparameter that should be assessed. The
+(2022) was known to be under-sampled, and so the network alone contained
+less information than the combination of the network and species traits.
+There is no general conclusion to draw from either these results or
+ours, besides reinforcing the need to be pragmatic about which models
+should be included in the ensemble, and whether to use an ensemble at
+all. In a sense, the surprising performance of the ensemble model should
+form the basis of the first broad recommendation: optimal training set
+balance and its interaction with connectance and the specific binary
+classifier used is, in a sense, an hyperparameter that should be
+assessed following the approach outlined in this manuscript. The
 distribution of results in fig.~\ref{fig:optimbias} and
@@
 within networks are essentially two different ones. For some
-applications (\emph{.e.g.} comparison of network structure across
+applications (\emph{e.g.} comparison of network structure across
 gradients), one may care more about a robust estimate of the structure,
@@
 something to evaluate on a case-by-case basis, especially since there is
-no guarantee that an esemble model will always be the most precise one.
+no guarantee that an ensemble model will always be the most precise one.
 Despite this apparent tension at the heart of the predictive exercise,
@@
 use of MCC (figs.~\ref{fig:optimbias}, \ref{fig:optimvalue}). Because
-\(F_1\) is monotonously sensitive to classifier bias fig.~\ref{fig:bias}
-and network connectance fig.~\ref{fig:connectance}, MCC should be
-prefered as a measure of model evaluation and comparison. When dealing
-with multiple models, we therefore suggest to find the optimal threshold
-using informedness, and to pick the best model using MCC (assuming one
-does not want to use an ensemble model).
+\(F_1\) is monotonously sensitive to classifier bias
+(fig.~\ref{fig:bias}) and network connectance
+(fig.~\ref{fig:connectance}), MCC should be prefered as a measure of
+model evaluation and comparison. When dealing with multiple models, we
+therefore suggest to find the optimal threshold using informedness, and
+to pick the best model using MCC (assuming one does not want to use an
+ensemble model).
 
@@
 more tractable is that some interactions can safely be assumed to be
-impossible; indeed, a lot of networks admit a stochastic block model as
-a good approximation (\emph{e.g.} Xie et al., 2017). In ecological
-networks, this can be due to spatial constrains (Valdovinos, 2019), or
-to the long-standing knowledge that some links are ``forbidden'' due to
-traits (Olesen et al., 2011) or abundances (Canard et al., 2014). The
-matching rules (Olito \& Fox, 2015; Strona \& Veech, 2017) can be
-incorporated in the model either by adding compatibility traits, or by
-\emph{only} training the model on pairs of species that are not likely
-to be forbidden links; having this information would allow to assemble
-training/testing sets that have true negatives, and in this situation,
-it may be possible to use the more usual 70/30 split. Besides forbidden
-links, a real-life case that may arise is multi-interaction or
-multi-layer networks (Pilosof et al., 2017). These can be studied using
-the same general approach outlined here, either by assuming that pairs
-of species can interact in more than one way (wherein one would train a
-model for each type of interaction, based on the relevant predictors),
-or by assuming that pairs of species can only have one type of
-interaction (wherein this becomes a multi-label classification problem).
+impossible; indeed, a lot of networks can be reasonably well described
+using a stochastic block model (\emph{e.g.} Xie et al., 2017). In
+ecological networks, this can be due to spatial constraints (Valdovinos,
+2019), or to the long-standing knowledge that some links are
+``forbidden'' due to traits (Olesen et al., 2011) or abundances (Canard
+et al., 2014). The matching rules (Olito \& Fox, 2015; Strona \& Veech,
+2017) can be incorporated in the model either by adding compatibility
+traits, or by \emph{only} training the model on pairs of species that
+are not likely to be forbidden links. Knowledge of true negative
+interactions could be propagated in training/testing sets that have true
+negatives, and in this situation, it may be possible to use the more
+usual 70/30 split for training/testing folds as the need to protect
+against potential unbalance is lowered. Besides forbidden links, a
+real-life case that may arise is multi-interaction or multi-layer
+networks (Pilosof et al., 2017). These can be studied using the same
+general approach outlined here, either by assuming that pairs of species
+can interact in more than one way (wherein one would train a model for
+each type of interaction, based on the relevant predictors), or by
+assuming that pairs of species can only have one type of interaction
+(wherein this becomes a multi-label classification problem).
 
@@
 
+\leavevmode\hypertarget{ref-Hong2016DeaCla}{}%
+Hong, C., Ghosh, R., \& Srinivasan, S. (2016). \emph{Dealing with Class
+Imbalance using Thresholding}.
+\url{https://doi.org/10.48550/arXiv.1607.02705}
+
 \leavevmode\hypertarget{ref-Inman2021ComSam}{}%

### 2023-01-18 Poisot_2023_guidelines_prediction_draft.tex
Deploying to gh-pages from @ PoisotLab/ms_networkary_classifiers@08949b36220530a13348d4076144bc29f375baf9 🚀

--- 2022-09-11 draft.tex
@@
 \usepackage{lastpage}
+
+\usepackage{gensymb}
 
@@
 \setlength{\csllabelwidth}{3em}
-\newenvironment{CSLReferences}[3] % #1 hanging-ident, #2 entry spacing
+\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
  {% don't indent paragraphs
@@
 \vfill
-This work is released by its authors under a CC-BY 4.0 license\hfill\ccby\\
-Last revision: \emph{\today}
-
-\clearpage
-\thispagestyle{empty}
-
-\vfill
 
 \begin{enumerate}
-    \item The prediction of species interactions is gaining momentum as
-a way to circumvent limitations in data volume. Yet, ecological networks
+    \itemThe prediction of species interactions is gaining momentum as a
+way to circumvent limitations in data volume. Yet, ecological networks
 are challenging to predict because they are typically small and sparse.
@@
 models can be trained for this specific problem.%
-    \item Using simple mathematical arguments and numerical experiments
+    \itemUsing simple mathematical arguments and numerical experiments
 in which a variety of classifiers (for supervised learning) are trained
@@
 the training dataset.%
-    \item Neither classifier accuracy nor the area under the receiver
+    \itemNeither classifier accuracy nor the area under the receiver
 operating characteristic curve (ROC-AUC) are informative measures for
@@
 classifier.%
-    \item These results reveal that training machines to predict
-networks is a challenging task, and that in virtually all cases, the
-composition of the training set needs to be fine-tuned before performing
-the actual training. We discuss these consequences in the context of the
-low volume of data.%
+    \itemThese results reveal that training machines to predict networks
+is a challenging task, and that in virtually all cases, the composition
+of the training set needs to be fine-tuned before performing the actual
+training. We discuss these consequences in the context of the low volume
+of data.%
 \end{enumerate}
+
+\vfill
+This work is released by its authors under a CC-BY 4.0 license\hfill\ccby\\
+Last revision: \emph{\today}
+
+\clearpage
+\thispagestyle{empty}
+
+\vfill
 
@@
 discipline-specific question we seek to address is: `which metric should
-be employed when predicting networks, and how to optimize it?'
+be employed when predicting networks, and how to optimize it?'.
 
@@
 ``relatively robust procedures used for unskewed data can break down
-miserably when the data is skewed.'' Following Japkowicz (2013), we
+miserably when the data is skewed''. Following Japkowicz (2013), we
 focus on two ranking metrics (the areas under the Receiver Operating
@@
 ``Dataset'' panel). Increasing the value of the \(\xi\) parameter would
-make the diagonal structure ``broader,'' leading to more interactions. A
+make the diagonal structure ``broader'', leading to more interactions. A
 visual inspection of the results is important, as it highlights how some
@@
 analysis; specifically, we measured the connectance, nestedness
-(\(\eta\); Bastolla et al., 2009), modularity (\(Q\); Barber, 2007),
-asymmetry (\(A\); Delmas et al., 2018), and Jaccard network
-dissimilarity (Canard et al., 2014). This process was repeated 250
-times, and the results are presented in tbl.~\ref{tbl:comparison}. The
-k-NN model is an interesting instance here: it produces the network that
-looks the most like the original dataset, despite having the lowest
+{[}\(\eta\); Bastolla et al. (2009){]}, modularity {[}\(Q\); Barber
+(2007){]}, asymmetry {[}\(A\); Delmas et al. (2018){]}, and Jaccard
+network dissimilarity (Canard et al., 2014). This process was repeated
+250 times, and the results are presented in tbl.~\ref{tbl:comparison}.
+The k-NN model is an interesting instance here: it produces the network
+that looks the most like the original dataset, despite having the lowest
 PR-AUC, suggesting it hits high recall at the cost of low precision. The
@@
 \hypertarget{tbl:comparison}{}
-\begin{longtable}[]{@{}rccccccccc@{}}
+\begin{longtable}[]{@{}
+  >{\raggedleft\arraybackslash}p{(\columnwidth - 18\tabcolsep) * \real{0.1548}}
+  >{\centering\arraybackslash}p{(\columnwidth - 18\tabcolsep) * \real{0.0952}}
+  >{\centering\arraybackslash}p{(\columnwidth - 18\tabcolsep) * \real{0.0952}}
+  >{\centering\arraybackslash}p{(\columnwidth - 18\tabcolsep) * \real{0.0833}}
+  >{\centering\arraybackslash}p{(\columnwidth - 18\tabcolsep) * \real{0.0952}}
+  >{\centering\arraybackslash}p{(\columnwidth - 18\tabcolsep) * \real{0.0952}}
+  >{\centering\arraybackslash}p{(\columnwidth - 18\tabcolsep) * \real{0.0952}}
+  >{\centering\arraybackslash}p{(\columnwidth - 18\tabcolsep) * \real{0.0952}}
+  >{\centering\arraybackslash}p{(\columnwidth - 18\tabcolsep) * \real{0.0952}}
+  >{\centering\arraybackslash}p{(\columnwidth - 18\tabcolsep) * \real{0.0952}}@{}}
 \caption{\label{tbl:comparison}Values of four performance metrics, and
@@
 column indicate an average \(\ge 0.99\).}\tabularnewline
-\toprule
-Model & MCC & Inf. & ROC-AUC & PR-AUC & Conn. & \(\eta\) & \(Q\) & \(A\)
-& Jaccard\tabularnewline
-\midrule
+\toprule()
+\begin{minipage}[b]{\linewidth}\raggedleft
+Model
+\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
+MCC
+\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
+Inf.
+\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
+ROC-AUC
+\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
+PR-AUC
+\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
+Conn.
+\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
+\(\eta\)
+\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
+\(Q\)
+\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
+\(A\)
+\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
+Jaccard
+\end{minipage} \\
+\midrule()
 \endfirsthead
-\toprule
-Model & MCC & Inf. & ROC-AUC & PR-AUC & Conn. & \(\eta\) & \(Q\) & \(A\)
-& Jaccard\tabularnewline
-\midrule
+\toprule()
+\begin{minipage}[b]{\linewidth}\raggedleft
+Model
+\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
+MCC
+\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
+Inf.
+\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
+ROC-AUC
+\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
+PR-AUC
+\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
+Conn.
+\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
+\(\eta\)
+\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
+\(Q\)
+\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
+\(A\)
+\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
+Jaccard
+\end{minipage} \\
+\midrule()
 \endhead
 Decision tree & 0.59 & 0.94 & 0.97 & 0.04 & 0.17 & 0.64 & 0.37 &
-\textbf{0.42} & 0.1\tabularnewline
-BRT & 0.46 & 0.91 & 0.97 & 0.36 & 0.2 & 0.78 & 0.29 & 0.41 &
-0.19\tabularnewline
+\textbf{0.42} & 0.1 \\
+BRT & 0.46 & 0.91 & 0.97 & 0.36 & 0.2 & 0.78 & 0.29 & 0.41 & 0.19 \\
 Random Forest & 0.72 & \textbf{0.98} & 0.99 & 0.1 & \textbf{0.16} &
-\textbf{0.61} & 0.38 & \textbf{0.42} & \textbf{0.06}\tabularnewline
+\textbf{0.61} & 0.38 & \textbf{0.42} & \textbf{0.06} \\
 k-NN & 0.71 & \textbf{0.98} & 0.99 & 0.02 & \textbf{0.16} &
-\textbf{0.61} & \textbf{0.39} & \textbf{0.42} &
-\textbf{0.06}\tabularnewline
+\textbf{0.61} & \textbf{0.39} & \textbf{0.42} & \textbf{0.06} \\
 \emph{Ensemble} & \textbf{0.74} & \textbf{0.98} & \textbf{1.0} &
 \textbf{0.79} & \textbf{0.16} & \textbf{0.61} & 0.38 & \textbf{0.42} &
-\textbf{0.06}\tabularnewline
-\emph{Data} & & & & & 0.16 & 0.56 & 0.41 & 0.42 & 0.0\tabularnewline
-\bottomrule
+\textbf{0.06} \\
+\emph{Data} & & & & & 0.16 & 0.56 & 0.41 & 0.42 & 0.0 \\
+\bottomrule()
 \end{longtable}
@@
 model) will give the best validation potential, and in this light is
-very likely informedness (maximizing informedness is, in fact, the
+very likely informedness {[}maximizing informedness is, in fact, the
 generally accepted default for imbalanced classification regardless of
-the problem domain; Schisterman et al., 2005). This last result further
-strengthens the assumption that the amount of bias \emph{is} an
+the problem domain; Schisterman et al. (2005){]}. This last result
+further strengthens the assumption that the amount of bias \emph{is} an
 hyper-parameter that must be fine-tuned, as using the wrong bias can
@@
 \begin{CSLReferences}{1}{0}
-\leavevmode\hypertarget{ref-Allouche2006AssAcc}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-Allouche2006AssAcc}{}}%
 Allouche, O., Tsoar, A., \& Kadmon, R. (2006). Assessing the accuracy of
@@
 
-\leavevmode\hypertarget{ref-Barber2007ModCom}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-Barber2007ModCom}{}}%
 Barber, M. J. (2007). Modularity and community detection in bipartite
@@
 
-\leavevmode\hypertarget{ref-Bastolla2009ArcMut}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-Bastolla2009ArcMut}{}}%
 Bastolla, U., Fortuna, M. A., Pascual-García, A., Ferrera, A., Luque,
@@
 
-\leavevmode\hypertarget{ref-Beauchesne2016ThiOut}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-Beauchesne2016ThiOut}{}}%
 Beauchesne, D., Desjardins-Proulx, Archambault, P., \& Gravel, D.
@@
 
-\leavevmode\hypertarget{ref-Becker2022OptPre}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-Becker2022OptPre}{}}%
 Becker, D. J., Albery, G. F., Sjodin, A. R., Poisot, T., Bergner, L. M.,
@@
 
-\leavevmode\hypertarget{ref-Bezanson2017JulFre}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-Bezanson2017JulFre}{}}%
 Bezanson, J., Edelman, A., Karpinski, S., \& Shah, V. (2017). Julia: A
@@
 
-\leavevmode\hypertarget{ref-Blaom2020MljJul}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-Blaom2020MljJul}{}}%
 Blaom, A. D., Kiraly, F., Lienart, T., Simillides, Y., Arenas, D., \&
@@
 
-\leavevmode\hypertarget{ref-Blaom2020FleMod}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-Blaom2020FleMod}{}}%
 Blaom, A. D., \& Vollmer, S. J. (2020). \emph{Flexible model composition
@@
 
-\leavevmode\hypertarget{ref-Boughorbel2017OptCla}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-Boughorbel2017OptCla}{}}%
 Boughorbel, S., Jarray, F., \& El-Anbari, M. (2017). Optimal classifier
@@
 
-\leavevmode\hypertarget{ref-Branco2015SurPre}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-Branco2015SurPre}{}}%
 Branco, P., Torgo, L., \& Ribeiro, R. (2015). \emph{A Survey of
@@
 
-\leavevmode\hypertarget{ref-Canard2014EmpEva}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-Canard2014EmpEva}{}}%
 Canard, E. F., Mouquet, N., Mouillot, D., Stanko, M., Miklisova, D., \&
@@
 
-\leavevmode\hypertarget{ref-Caron2022AddElt}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-Caron2022AddElt}{}}%
 Caron, D., Maiorano, L., Thuiller, W., \& Pollock, L. J. (2022).
@@
 
-\leavevmode\hypertarget{ref-Chicco2020AdvMat}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-Chicco2020AdvMat}{}}%
 Chicco, D., \& Jurman, G. (2020). The advantages of the Matthews
@@
 
-\leavevmode\hypertarget{ref-Chicco2021MatCor}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-Chicco2021MatCor}{}}%
 Chicco, D., Tötsch, N., \& Jurman, G. (2021). The Matthews correlation
@@
 
-\leavevmode\hypertarget{ref-deAguiar2019RevBia}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-deAguiar2019RevBia}{}}%
 de Aguiar, M. A. M., Newman, E. A., Pires, M. M., Yeakel, J. D.,
@@
 
-\leavevmode\hypertarget{ref-Delgado2019WhyCoh}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-Delgado2019WhyCoh}{}}%
 Delgado, R., \& Tibau, X.-A. (2019). Why Cohen's Kappa should be avoided
@@
 
-\leavevmode\hypertarget{ref-Delmas2018AnaEco}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-Delmas2018AnaEco}{}}%
 Delmas, E., Besson, M., Brice, M.-H., Burkle, L. A., Dalla Riva, G. V.,
@@
 
-\leavevmode\hypertarget{ref-Desjardins-Proulx2017EcoInt}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-Desjardins-Proulx2017EcoInt}{}}%
 Desjardins-Proulx, P., Laigle, I., Poisot, T., \& Gravel, D. (2017).
@@
 
-\leavevmode\hypertarget{ref-Ferri2009ExpCom}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-Ferri2009ExpCom}{}}%
 Ferri, C., Hernández-Orallo, J., \& Modroiu, R. (2009). An experimental
@@
 
-\leavevmode\hypertarget{ref-He2013ImbLea}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-He2013ImbLea}{}}%
 He, H., \& Ma, Y. (Eds.). (2013). \emph{Imbalanced Learning:
@@
 
-\leavevmode\hypertarget{ref-Hong2016DeaCla}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-Hong2016DeaCla}{}}%
 Hong, C., Ghosh, R., \& Srinivasan, S. (2016). \emph{Dealing with Class
@@
 
-\leavevmode\hypertarget{ref-Inman2021ComSam}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-Inman2021ComSam}{}}%
 Inman, R., Franklin, J., Esque, T., \& Nussear, K. (2021). Comparing
@@
 
-\leavevmode\hypertarget{ref-Iturbide2015FraSpe}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-Iturbide2015FraSpe}{}}%
 Iturbide, M., Bedia, J., Herrera, S., del Hierro, O., Pinto, M., \&
@@
 
-\leavevmode\hypertarget{ref-Japkowicz2013AssMet}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-Japkowicz2013AssMet}{}}%
 Japkowicz, N. (2013). Assessment Metrics for Imbalanced Learning. In
@@
 
-\leavevmode\hypertarget{ref-Jeni2013FacImb}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-Jeni2013FacImb}{}}%
 Jeni, L. A., Cohn, J. F., \& De La Torre, F. (2013). Facing Imbalanced
@@
 
-\leavevmode\hypertarget{ref-Jordano2016ChaEco}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-Jordano2016ChaEco}{}}%
 Jordano, P. (2016a). Chasing Ecological Interactions. \emph{PLOS Biol},
@@
 
-\leavevmode\hypertarget{ref-Jordano2016SamNet}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-Jordano2016SamNet}{}}%
 Jordano, P. (2016b). Sampling networks of ecological interactions.
@@
 
-\leavevmode\hypertarget{ref-Landis1977MeaObs}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-Landis1977MeaObs}{}}%
 Landis, J. R., \& Koch, G. G. (1977). The Measurement of Observer
@@
 
-\leavevmode\hypertarget{ref-MacDonald2020RevLin}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-MacDonald2020RevLin}{}}%
 MacDonald, A. A. M., Banville, F., \& Poisot, T. (2020). Revisiting the
@@
 
-\leavevmode\hypertarget{ref-Martinez1992ConCon}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-Martinez1992ConCon}{}}%
 Martinez, N. D. (1992). Constant Connectance in Community Food Webs.
 \emph{The American Naturalist}, \emph{139}(6), 1208--1218.
-\url{http://www.jstor.org/stable/2462337}
-
-\leavevmode\hypertarget{ref-McLeod2021SamAsy}{}%
+\url{https://www.jstor.org/stable/2462337}
+
+\leavevmode\vadjust pre{\hypertarget{ref-McLeod2021SamAsy}{}}%
 McLeod, A., Leroux, S. J., Gravel, D., Chu, C., Cirtwill, A. R., Fortin,
@@
 
-\leavevmode\hypertarget{ref-Olesen2011MisFor}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-Olesen2011MisFor}{}}%
 Olesen, J. M., Bascompte, J., Dupont, Y. L., Elberling, H., Rasmussen,
@@
 
-\leavevmode\hypertarget{ref-Olito2015SpeTra}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-Olito2015SpeTra}{}}%
 Olito, C., \& Fox, J. W. (2015). Species traits and abundances predict
@@
 
-\leavevmode\hypertarget{ref-Pichler2020MacLea}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-Pichler2020MacLea}{}}%
 Pichler, M., Boreux, V., Klein, A., Schleuning, M., \& Hartig, F.
@@
 
-\leavevmode\hypertarget{ref-Pilosof2017MulNat}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-Pilosof2017MulNat}{}}%
 Pilosof, S., Porter, M. A., Pascual, M., \& Kéfi, S. (2017). The
@@
 
-\leavevmode\hypertarget{ref-Poisot2021GloKno}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-Poisot2021GloKno}{}}%
 Poisot, T., Bergeron, G., Cazelles, K., Dallas, T., Gravel, D.,
@@
 
-\leavevmode\hypertarget{ref-Poisot2021ImpMam}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-Poisot2021ImpMam}{}}%
 Poisot, T., Ouellet, M.-A., Mollentze, N., Farrell, M. J., Becker, D.
@@
 
-\leavevmode\hypertarget{ref-Saito2015PrePlo}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-Saito2015PrePlo}{}}%
 Saito, T., \& Rehmsmeier, M. (2015). The Precision-Recall Plot Is More
@@
 
-\leavevmode\hypertarget{ref-Schisterman2005OptCut}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-Schisterman2005OptCut}{}}%
 Schisterman, E. F., Perkins, N. J., Liu, A., \& Bondell, H. (2005).
@@
 
-\leavevmode\hypertarget{ref-Somodi2017PreDep}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-Somodi2017PreDep}{}}%
 Somodi, I., Lepesi, N., \& Botta‐Dukát, Z. (2017). Prevalence dependence
@@
 
-\leavevmode\hypertarget{ref-Steen2021SpaThi}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-Steen2021SpaThi}{}}%
 Steen, V. A., Tingley, M. W., Paton, P. W. C., \& Elphick, C. S. (2021).
@@
 
-\leavevmode\hypertarget{ref-Strona2017ForPer}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-Strona2017ForPer}{}}%
 Strona, G., \& Veech, J. A. (2017). Forbidden versus permitted
@@
 
-\leavevmode\hypertarget{ref-Strydom2021RoaPre}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-Strydom2021RoaPre}{}}%
 Strydom, T., Catchen, M. D., Banville, F., Caron, D., Dansereau, G.,
@@
 
-\leavevmode\hypertarget{ref-Valdovinos2019MutNet}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-Valdovinos2019MutNet}{}}%
 Valdovinos, F. S. (2019). Mutualistic networks: Moving closer to a
@@
 
-\leavevmode\hypertarget{ref-Weitz2005CoeArm}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-Weitz2005CoeArm}{}}%
 Weitz, J. S., Hartman, H., \& Levin, S. A. (2005). Coevolutionary arms
@@
 
-\leavevmode\hypertarget{ref-Whalen2021NavPit}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-Whalen2021NavPit}{}}%
 Whalen, S., Schreiber, J., Noble, W. S., \& Pollard, K. S. (2021).
@@
 
-\leavevmode\hypertarget{ref-Wood2015EffSpa}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-Wood2015EffSpa}{}}%
 Wood, S. A., Russell, R., Hanson, D., Williams, R. J., \& Dunne, J. A.
@@
 
-\leavevmode\hypertarget{ref-Xie2017ComCom}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-Xie2017ComCom}{}}%
 Xie, J.-R., Zhang, P., Zhang, H.-F., \& Wang, B.-H. (2017). Completeness
@@
 
-\leavevmode\hypertarget{ref-Youden1950IndRat}{}%
+\leavevmode\vadjust pre{\hypertarget{ref-Youden1950IndRat}{}}%
 Youden, W. J. (1950). Index for rating diagnostic tests. \emph{Cancer},

### 2023-01-18 Poisot_2023_guidelines_prediction_preprint.tex
Deploying to gh-pages from @ PoisotLab/ms_networkary_classifiers@08949b36220530a13348d4076144bc29f375baf9 🚀

--- 2023-01-18 Poisot_2023_guidelines_prediction_draft.tex
@@
 \newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth}{#1}}
-\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}\geometry{verbose,letterpaper,tmargin=2.2cm,bmargin=2.2cm,lmargin=2.2cm,rmargin=2.2cm}
-
-\usepackage{lineno}
-\usepackage[nolists,noheads]{endfloat}
+\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}\usepackage[table,dvipsnames]{xcolor}
+
+\geometry{includemp,
+            letterpaper,
+            top=2.4cm,
+            bottom=2.4cm,
+            left=1.0cm,
+            right=1.0cm,
+            marginparwidth=5cm,
+            marginparsep=1.0cm}
+
+\usepackage[singlelinecheck=off]{caption}
+
+\captionsetup{
+  font={small},
+  labelfont={bf},
+  format=plain,
+  labelsep=quad
+}
+
+\usepackage{floatrow}
+
+\floatsetup[figure]{margins=hangright,
+              facing=no,
+              capposition=beside,
+              capbesideposition={center,outside},
+              floatwidth=\textwidth}
+
+% \floatsetup[table]{margins=hangright,
+%              facing=no,
+%              capposition=beside,
+%              capbesideposition={center,outside},
+%              floatwidth=\textwidth}
 
@@
 
-\tolerance=1
-\emergencystretch=\maxdimen
-\hyphenpenalty=10000
-\hbadness=10000
-
-\doublespacing
+\setcounter{secnumdepth}{5}
+
+\usepackage{titlesec}
+
+\titleformat{\section}[block]
+{\normalfont\large\sffamily}
+{\thesection}{.5em}{\titlerule\\[.8ex]\bfseries}
+
+\titleformat{\subsection}[runin]
+{\normalfont\fontseries{b}\selectfont\filright\sffamily}
+{\thesubsection.}{.5em}{}
+
+\titleformat{\subsubsection}[runin]
+{\normalfont\itshape\rmfamily\bfseries}{\thesubsubsection}{1em}{}
+
+\fancypagestyle{firstpage}
+{
+   \fancyhf{}
+   \renewcommand{\headrulewidth}{0pt}
+   \fancyfoot[R]{\footnotesize\ccby}
+   \fancyfoot[L]{\footnotesize\sffamily\today}
+}
 
@@
 }
+
+\usepackage{tikz}
 \begin{document}
-\raggedright
-\thispagestyle{empty}
-{\Large\bfseries\sffamily Guidelines for the prediction of species
-interactions through binary classification}
-\vskip 5em
-
+\tikz [remember picture, overlay] %
+\node [shift={(-0.6in,1.1cm)},scale=0.2,opacity=0.4] at (current page.south east)[anchor=south east]{\includegraphics{logo}};%
+\pagestyle{normal}
+\thispagestyle{firstpage}
+
+\newcommand{\colorRule}[3][black]{\textcolor[HTML]{#1}{\rule{#2}{#3}}}
+
+\noindent {\LARGE \textbf{\textsf{Guidelines for the prediction of
+species interactions through binary classification}}}
+
+\medskip
+\begin{flushleft}
+{\small
 %
@@
 \,\textsuperscript{1,2}
-
-\textsuperscript{1}\,Université de
-Montréal\quad \textsuperscript{2}\,Québec Centre for Biodiversity
-Sciences
-
+\vskip 1em
+\textsuperscript{1}\,Université de Montréal; \textsuperscript{2}\,Québec
+Centre for Biodiversity Sciences\\
+\vskip 1em
 \textbf{Correspondance to:}\\
 Timothée Poisot --- \texttt{timothee.poisot@umontreal.ca}\\
-
-\vfill
+}
+\end{flushleft}
+
+\vskip 2em
+\makebox[0pt][l]{\colorRule[CCCCCC]{2.0\textwidth}{0.5pt}}
+\vskip 2em
+\noindent
+
+\marginpar{\vskip 1em\flushright
+{\small{\bfseries Keywords}:\par
+species interaction networks\\binary classifiers\\machine
+learning\\regression\\supervised learning\\}
+}
 
@@
 
-\vfill
-This work is released by its authors under a CC-BY 4.0 license\hfill\ccby\\
-Last revision: \emph{\today}
-
-\clearpage
-\thispagestyle{empty}
-
-\vfill
-
-\vfill
-
-\clearpage
-\linenumbers
-\pagestyle{normal}
+\vskip 2em
+\makebox[0pt][l]{\colorRule[CCCCCC]{2.0\textwidth}{0.5pt}}
+\vskip 2em
 

### 2023-01-27 Poisot_2023_guidelines_prediction_draft.tex
Deploying to gh-pages from @ PoisotLab/ms_networkary_classifiers@03944f45cb3ceab9378f415b05e746b5f347c090 🚀

@@
 \begin{enumerate}
-    \itemThe prediction of species interactions is gaining momentum as a
-way to circumvent limitations in data volume. Yet, ecological networks
+    \item The prediction of species interactions is gaining momentum as
+a way to circumvent limitations in data volume. Yet, ecological networks
 are challenging to predict because they are typically small and sparse.
@@
 models can be trained for this specific problem.%
-    \itemUsing simple mathematical arguments and numerical experiments
+    \item Using simple mathematical arguments and numerical experiments
 in which a variety of classifiers (for supervised learning) are trained
@@
 the training dataset.%
-    \itemNeither classifier accuracy nor the area under the receiver
+    \item Neither classifier accuracy nor the area under the receiver
 operating characteristic curve (ROC-AUC) are informative measures for
@@
 classifier.%
-    \itemThese results reveal that training machines to predict networks
-is a challenging task, and that in virtually all cases, the composition
-of the training set needs to be fine-tuned before performing the actual
-training. We discuss these consequences in the context of the low volume
-of data.%
+    \item These results reveal that training machines to predict
+networks is a challenging task, and that in virtually all cases, the
+composition of the training set needs to be fine-tuned before performing
+the actual training. We discuss these consequences in the context of the
+low volume of data.%
 \end{enumerate}
